---
title: "coiaf"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose

`coiaf` is a package for estimating complexity of infection (COI) from bi-allelic SNP data. In comparison to other available methods, `coiaf` is fast, accurate, and can be used to estimate COI for a large number of samples.

## Existing resources

The [coiaf website](https://bailey-lab.github.io/coiaf/) contains a detailed tutorial on how to use the package, as well as a vignette with worked examples.

The paper describing the package is available [here](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1010247).

## Citation
```{r, eval=FALSE}
 @article{
  Paschalidis_Watson_Aydemir_Verity_Bailey_2023, 
  title={coiaf: Directly estimating complexity of infection with allele frequencies}, 
  volume={19},
  ISSN={1553-7358},
  DOI={10.1371/journal.pcbi.1010247},
  number={6},
  journal={PLOS Computational Biology},
  publisher={Public Library of Science},
  author={Paschalidis, Aris and Watson, Oliver J. and Aydemir, Ozkan and Verity, Robert and Bailey, Jeffrey A.},
  year={2023},
  month=jun,
  pages={e1010247},
  language={en} 
}
```