---
title: IDM Benchmarking Results
author: Bing Guo
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load in the Benchmarking Outputs

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(jsonlite)
library(magrittr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(tidyr)

input_config <- list(
  sim_info_dir = "simulations/SpotMalaria",
  tool_res_dir = "tool_results/IDM",
  tool_params = "tool_results/idm.json"
)
```

First, all of the tool results, tool parameters, expected simulated values, and
simulation parameters are read into a tibble.

```{r}
# simulation truth data frame
df_st <- tibble(
  sim_truth_path = list.files(
    input_config$sim_info_dir,
    pattern = "population_aa_variants_freqs_prevs.tsv.gz",
    full.names = TRUE,
    recursive = TRUE
  )
) %>%
  mutate(
    # read all files, each as a row
    sim_truth = map(
      sim_truth_path,
      read_tsv,
      col_types = cols(
        population_name = col_character(),
        country = col_character(),
        pop = col_character(),
        GeneName = col_character(),
        transcript_id = col_character(),
        aa_position = col_integer(),
        aa = col_character(),
        is_ref = col_character(),
        allele_count = col_integer(),
        sample_count = col_integer(),
        total_pop_alleles = col_integer(),
        total_pop_samples = col_integer(),
        unique_alleles = col_integer(),
        freq = col_double(),
        prev = col_double(),
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>%
  unite(variant, transcript_id, aa_position, aa, sep = ":") %>%
  rename(
    sim_id = population_name,
    exp_freq = freq,
  ) %>%
  select(
    sim_id,
    variant,
    exp_freq
  )
str(df_st)
```


```{r}
# Read in all simulation parameters into one tibble --------------------
df_sp <- tibble(
  sim_params_path = list.files(
    input_config$sim_info_dir,
    pattern = "simulation_parameters.json",
    full.names = TRUE,
    recursive = TRUE
  )
) %>%
  mutate(
    sim_params = map(
      sim_params_path,
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>%
  rename(
    sim_id = population_name,
    true_mean_coi = mean_coi
  ) %>%
  select(sim_id, country, pop_alpha, coi_r, coi_p, k_s, true_mean_coi) %>%
  mutate(coi_bin = case_when(
    true_mean_coi < 2 ~ 0,
    true_mean_coi >= 2 ~ 1,
  )) %>%
  mutate(
    sim_id = factor(sim_id, levels = sort(unique(sim_id))),
    country = factor(country, levels = sort(unique(country))),
    pop_alpha = factor(
      pop_alpha, c(99, 9, 2.33),
      labels = c("low_rel", "med_rel", "high_rel")
    ),
    coi_bin = factor(
      coi_bin,
      levels = c(0, 1), labels = c("coi_lt_2", "coi_ge_2")
    ),
    coi_r = factor(
      coi_r,
      c("0.25", "0.5"),
      labels = c("low_coi_param", "high_coi_param")
    ),
    k_s = factor(
      k_s,
      levels = c(0.5, 0.2),
      labels = c("low_recomb_param", "high_recomb_param")
    ),
  ) %>%
  select(-coi_p)
str(df_sp)
```

```{r}
# Read in all IDM results into one tibble ------------------------------
df_tr <- tibble(
  tool_res_path = list.files(
    input_config$tool_res_dir,
    pattern = "*.tsv.gz",
    full.names = TRUE,
    recursive = TRUE
  )
) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(sim_id = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path,
      read_tsv,
      col_types = cols(
        variant = col_character(),
        freq = col_double(),
      ),
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>%
  rename(
    obs_freq = freq,
  ) %>%
  mutate(
    sim_id = factor(sim_id, levels = sort(unique(sim_id))),
    paramset_id = factor(
      paramset_id,
      levels = sort(unique(paramset_id))
    ),
  )
str(df_tr)
```

```{r}
# Read in tool parameters, and join ------------------------------------
df_tp <- jsonlite::fromJSON(input_config$tool_params) %>%
  as_tibble() %>%
  unnest(paramset) %>%
  mutate(
    paramset_id = factor(paramset_id),
    model = factor(model, levels = c("OM", "IDM")),
    eps_initial = factor(eps_initial),
    lambda_initial = factor(lambda_initial)
  )
str(df_tp)
```





### Join tables

```{r}
df_st_tr_sp_tp <- df_st %>%
  # may contain NAs
  full_join(df_tr, by = c("sim_id", "variant")) %>%
  select(sim_id, variant, paramset_id, obs_freq, exp_freq) %>%
  left_join(df_sp, by = "sim_id") %>%
  left_join(df_tp, by = "paramset_id") %>%
  mutate(freq_bin = cut(replace(exp_freq, is.na(exp_freq), 0),
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, Inf),
    labels = c("<0.2", "<0.4", "<0.6", "<0.8", ">=0.8"),
  ))
str(df_st_tr_sp_tp)
```

## Analysis

### Allele Frequency

Calculating RMSE

```{r}
freq_rmse <- df_st_tr_sp_tp %>%
  drop_na() %>%
  group_by(
    # sim_id,
    paramset_id,
    country,
    pop_alpha,
    coi_r,
    coi_bin,
    k_s,
    model,
    lambda_initial,
    eps_initial,
  ) %>%
  summarize(
    freq_rmse = sqrt(mean((obs_freq - exp_freq)^2)),
    num_rows = n(),
    .groups = "drop"
  )
str(freq_rmse)
```

Calculating RMSE by frequency bins
```{r}
freq_rmse_by_freq <- df_st_tr_sp_tp %>%
  replace_na(list(obs_freq = 0, true_freq = 0)) %>%
  group_by(
    # sim_id,
    paramset_id,
    country,
    pop_alpha,
    coi_r,
    coi_bin,
    k_s,
    model,
    lambda_initial,
    eps_initial,
    freq_bin, # <---
  ) %>%
  summarize(
    freq_rmse = sqrt(mean((obs_freq - exp_freq)^2)),
    num_rows = n(),
    .groups = "drop"
  )
str(freq_rmse_by_freq)
```

### Explore how tool parameters affect RMSE

```{r}
freq_rmse %>%
  filter(
    country == "Cambodia",
    pop_alpha == "high_rel",
    coi_r == "low_coi_param",
  ) %>%
  ggplot(
    mapping = aes(x = factor(lambda_initial), y = freq_rmse, color = model)
  ) +
  geom_boxplot() +
  facet_wrap(vars(eps_initial), labeller = label_both) +
  labs(
    y = "RMSE of Frequency Estimate",
  ) +
  theme(legend.position = "bottom")
```
Different IDM tool parameters, such as `model`, `eps_intial`, and `lambda_initial`, do
not seem to change the RMSE value significantly.


## Explore how simulation parameters affect RMSE

```{r}
freq_rmse %>%
  filter(lambda_initial == 0.1, eps_initial == 0.1, model == "IDM") %>%
  data.frame() %>%
  ggplot(mapping = aes(x = country, y = freq_rmse)) +
  geom_boxplot()
```

With default tools parameters, RMSE is higher in Cambodian than in Ghana.



```{r}
freq_rmse %>%
  filter(lambda_initial == 0.1, eps_initial == 0.1, model == "IDM") %>%
  data.frame() %>%
  ggplot(mapping = aes(x = pop_alpha, y = freq_rmse)) +
  geom_boxplot()
```

RMSE is slightly higher in high relatedness simulations.


```{r}
freq_rmse %>%
  filter(lambda_initial == 0.1, eps_initial == 0.1, model == "IDM") %>%
  data.frame() %>%
  ggplot(mapping = aes(x = k_s, y = freq_rmse)) +
  geom_boxplot()
```

RMSE does not differ too much between high and low recombination simulations.

```{r}
freq_rmse %>%
  filter(lambda_initial == 0.1, eps_initial == 0.1, model == "IDM") %>%
  ggplot(mapping = aes(x = factor(coi_bin), y = freq_rmse)) +
  geom_boxplot()
```

There is a clear trend of higher RMSE in simulations with higher true mean COI.



```{r}
freq_rmse %>%
  filter(lambda_initial == 0.1, eps_initial == 0.1, model == "IDM") %>%
  ggplot(mapping = aes(x = factor(coi_r), y = freq_rmse)) +
  geom_boxplot()
```

Similarly, higher RMSE in high COI parameter simulations.

We can look at combinations of simulation parameters to see how they affect RMSE.

```{r}
freq_rmse %>%
  filter(lambda_initial == 0.1, eps_initial == 0.1, model == "IDM") %>%
  ggplot(mapping = aes(x = country, y = freq_rmse, color = coi_r)) +
  geom_boxplot() +
  facet_grid(rows = vars(k_s), cols = vars(pop_alpha)) +
  labs(
    x = "Simulated Relatedness",
    y = "RMSE of Frequency Estimate",
  ) +
  theme(legend.position = "bottom")
```

Again, we see that:

- RMSE is higher in Cambodian simulations, and in high COI. 
- The effect seems to be more pronounced in low recombination simulations for high relatedness simulations. 
- Errors are slightly higher in high relatedness simulations.


We can also look at RMSE by frequency bins.
```{r}
freq_rmse_by_freq %>% 
  replace_na(list(freq_rmse=0))%>% # DOUBLE CHECK THIS REPLACEMENT
  filter(lambda_initial == 0.1, eps_initial == 0.1, model == "IDM") %>%
  ggplot(mapping = aes(x = country, y = freq_rmse, color = coi_r)) +
  geom_boxplot() +
  facet_grid(rows = vars(k_s), cols = vars(freq_bin), labeller = label_both) +
  labs(
    y = "RMSE of Frequency Estimate",
  ) +
  theme(legend.position = "bottom")
```

Here, we see that RMSE is much higher for low frequency alleles. 

We may want to consider normalize the RMSE by allele frequencies?

## Calculate false negative and fale discovery rates

Using whether an allele was observed (population frequency > 0) or not to
determine whehter each observation is a true positive (TP), true negative (TN),
false positive (FP), or false negative (FN).

```{r}
ftnp <- df_st_tr_sp_tp %>%
  select(-paramset_id, -model, -lambda_initial, -eps_initial) %>%
  replace_na(list(obs_freq = 0, exp_freq = 0)) %>%
  mutate(
    ftnp = case_when(
      obs_freq == 0 & exp_freq != 0 ~ "FN",
      obs_freq == 0 & exp_freq == 0 ~ "TN",
      obs_freq != 0 & exp_freq != 0 ~ "TP",
      obs_freq != 0 & exp_freq == 0 ~ "FP"
    )
  )
str(ftnp)
```

Grouping by `simulation` and `ftnp` category to calculate False negative rate
(FNR) and False Discovery Rate (FDR).

FDR is used instead of False positive rate (FPR) because the number of true
negatives is always zero, and FPR would be 1 and not informative.

```{r}
fnr_fdr <- ftnp %>%
  group_by(sim_id, ftnp) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_wider(
    names_from = ftnp,
    values_from = count,
    values_fill = 0
  ) %>%
  mutate(FNR = FN / (FN + TP), FDR = FP / (FP + TP)) %>%
  select(-FN, -FP, -TN, -TP)
str(fnr_fdr)
```

Calculating FNR and FDR by allele frequency bins

```{r}
ggplot(ftnp, aes(x = exp_freq, fill = country)) +
  geom_histogram(alpha = 0.6, position = "identity") +
  labs(
    title = "Histogram of Value by Group (Overlapping)",
    x = "allele frequency",
    y = "Frequency"
  ) +
  scale_fill_manual(values = c("blue", "red")) # Customize colors
```

```{r}
fnr_fdr_by_freq <- ftnp %>%
  mutate(freq_bin = cut(exp_freq,
    breaks = c(-Inf, 0.2, 0.4, 0.6, 0.8, Inf),
    labels = c("<0.2", "<0.4", "<0.6", "<0.8", ">=0.8"),
  )) %>%
  group_by(sim_id, freq_bin, ftnp) %>%
  summarise(count = n(), .groups = "drop") %>%
  pivot_wider(
    names_from = ftnp,
    values_from = count,
    values_fill = 0
  ) %>%
  mutate(FNR = FN / (FN + TP), FDR = FP / (FP + TP)) %>%
  select(-FN, -FP, -TN, -TP)

# combine simulation parameter information
sim_info <- df_sp %>%
  select(sim_id, country, pop_alpha, coi_r, k_s) %>%
  distinct() %>%
  drop_na()

fnr_fdr <- fnr_fdr %>% left_join(sim_info)
fnr_fdr_by_freq <- fnr_fdr_by_freq %>% left_join(sim_info)
```

### FNR

```{r}
fnr_fdr %>%
  ggplot(mapping = aes(x = country, y = FNR, color = factor(coi_r))) +
  geom_boxplot(outlier.shape = NA) +
  facet_grid(rows = vars(k_s), cols = vars(pop_alpha), labeller = label_both) +
  labs(
    x = "Country",
    y = "False negative rate",
  ) +
  coord_cartesian(ylim = c(0, 0.25)) +
  theme(legend.position = "bottom")
```

High FNR in Ghana-like simulations with high COI parameter and high relatedness.


### FNR by frequency bins
```{r, fig.width = 7, fig.height = 7}
fnr_fdr_by_freq %>%
  ggplot(mapping = aes(x = country, y = FNR, color = factor(coi_r))) +
  geom_boxplot() +
  facet_grid(
    rows = vars(freq_bin),
    cols = vars(pop_alpha),
    labeller = label_both
  ) +
  labs(
    x = "Country",
    y = "False negative rate",
  ) +
  coord_cartesian(ylim = c(0, 0.6)) +
  theme(legend.position = "bottom")
```

It is clear that alleles with very low frequencies or high frequencies have much higher FNR, especially in Ghana-like simulations.

## FDR

```{r}
fnr_fdr %>%
  ggplot(mapping = aes(x = country, y = FDR, color = factor(coi_r))) +
  geom_boxplot(outlier.shape = NA) +
  facet_grid(rows = vars(k_s), cols = vars(pop_alpha), labeller = label_both) +
  labs(
    x = "Country",
    y = "False Discovery rate",
  ) +
  theme(legend.position = "bottom")
```

Low FDR overall, but slightly higher in Ghana-like simulations with high COI parameter and high relatedness.

## FDR by frequency bins
```{r, fig.width = 7, fig.height = 7}
fnr_fdr_by_freq %>%
  ggplot(mapping = aes(x = country, y = FDR, color = factor(coi_r))) +
  geom_boxplot() +
  facet_grid(
    rows = vars(freq_bin),
    cols = vars(pop_alpha),
    labeller = label_both
  ) +
  labs(
    x = "Country",
    y = "False Discovery Rate",
  ) +
  coord_cartesian(ylim = c(0, 0.2)) +
  theme(legend.position = "bottom")
```

When examining frequency bins, FDR is very high for alleles with lower frequencies. 



## Summary and Next Steps

### Summary

- IDM tool parameters do not seem to affect RMSE. 

  - We may want to explore how missing data rates and IDM model (OM, IDM) affect RMSE in future analyses.

- Simulation parameters does affect RMSE.

  - RMSE is higher in Cambodian-like simulations and in high COI parameter simulations
  - RMSE is slightly higher in high relatedness simulations.
  - RMSE is much higher for high frequency alleles.
    - suggesting RMSE being normalized by allele frequency?
  
- Simulation parameters also affect FNR and FDR.

  - FNR is higher in Ghana-like simulations with high COI parameter and high relatedness.
  - FDR is slightly higher in Ghana-like simulations with high COI parameter and high relatedness.
  - Both FNR and FDR are much higher for low frequency alleles, especially in Ghana-like simulations.
  
### Next steps

- Consider normalizing RMSE by allele frequency.
- Statistical testing
- Correlation between simulation parameters, such country and COI, and relatedness.
- Explore how missing data rates and IDM model (OM, IDM) affect RMSE, FNR, and FDR.
- Compare with other tools.s
