---
title: IDM Benchmarking Results
author: Bing Guo
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)
```


## Set up 

### Libraries used in the analysis below 

```{r, warning=FALSE, message=FALSE}
# data manipulation libraries 
library(dplyr)
library(readr)
library(stringr)
library(tibble)
library(tidyr) 
library(magrittr)
library(jsonlite)
library(purrr)
library(rlang) # for as_name

# plotting 
library(ggplot2)
library(scales) # help with axes 
library(GGally) # ggpairs for pairwise comparisons 

# interactive html outputs
library(plotly) # ggplotly to turn any ggplot object into an interactive plot 
library(DT) # data tables for displaying tables 

# stats package
library(DescTools) # CCC, Calculates Lin's concordance correlation coefficient
```

### Some helpful R utilities 

```{r}
source("../scripts/benchmarking_functions.R")


```


## Load in the Benchmarking Outputs

```{r}
input_config <- list(
  sim_info_dir = "../../data/benchmarking/simulations/SpotMalaria",
  tool_res_dir = "../../data/benchmarking/tool_results/IDM",
  tool_params = "../../data/benchmarking/tool_results/idm.json"
)
```


First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.

### Reading simulation truths (expected allele frequencies)


```{r}
# simulation truth data frame
df_st <- tibble(
  sim_truth_path = list.files(
    input_config$sim_info_dir,
    pattern = "population_aa_variants_freqs_prevs.tsv.gz",
    
    full.names = TRUE,
    recursive = TRUE
  )
) %>%
  mutate(
    # read all files, each as a row
    sim_truth = map(
      sim_truth_path,
      read_tsv,
      col_types = cols(
        population_name = col_character(),
        country = col_character(),
        pop = col_character(),
        GeneName = col_character(),
        transcript_id = col_character(),
        aa_position = col_integer(),
        aa = col_character(),
        is_ref = col_character(),
        allele_count = col_integer(),
        sample_count = col_integer(),
        total_pop_alleles = col_integer(),
        total_pop_samples = col_integer(),
        unique_alleles = col_integer(),
        freq = col_double(),
        prev = col_double(),
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>%
  unite(variant, transcript_id, aa_position, aa, sep = ":") %>%
  rename(
    sim_id = population_name,
    exp_freq = freq,
  ) %>%
  select(
    sim_id,
    variant,
    exp_freq
  )
# str(df_st)
#  $ sim_id  : chr [1:69053] "simulated_Cambodia_pop_001" "simulated_Cambodia_pop_001" "simulated_Cambodia_pop_001" "simulated_Cambodia_pop_001" ...
#  $ variant : chr [1:69053] "PF3D7_0508900.1:935:Y" "PF3D7_0728200.1:543:N" "PF3D7_0728200.1:543:S" "PF3D7_1133400.1:121:E" ...
#  $ exp_freq: num [1:69053] 1 0.379 0.621 1 1 ...
```

### Read simulation parameters

```{r}
# Read in all simulation parameters into one tibble --------------------
df_sp <- tibble(
  sim_params_path = list.files(
    input_config$sim_info_dir,
    pattern = "simulation_parameters.json",
    full.names = TRUE,
    recursive = TRUE
  )
) %>%
  mutate(
    sim_params = map(
      sim_params_path,
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>%
  rename(
    sim_id = population_name,
    true_mean_coi = mean_coi
  ) %>%
  select(sim_id, country, pop_alpha, coi_r, coi_p, k_s, true_mean_coi) %>%
  mutate(
    sim_id = factor(sim_id, levels = sort(unique(sim_id))),
    country = factor(country, levels = sort(unique(country))),
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>%
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  )) %>% 
  mutate(pop_cat = factor(pop_cat, levels = sort(unique((.$pop_cat)))), 
         coi_cat = factor(coi_cat, levels = sort(unique((.$coi_cat)))), 
         between_sample_related_cat = factor(between_sample_related_cat, levels = sort(unique((.$between_sample_related_cat)))), 
         nonrecombinant_cat = factor(nonrecombinant_cat, levels = sort(unique((.$nonrecombinant_cat)))), 
         meiosis_cat = factor(meiosis_cat, levels = sort(unique((.$meiosis_cat))))) %>%
  select(-coi_p, -coi_r, -k_s, -pop_alpha)
# str(df_sp)
#  $ sim_id                    : Factor w/ 240 levels "simulated_Cambodia_pop_001",..: 1 2 3 4 5 6 7 8 9 10 ...
#  $ country                   : Factor w/ 2 levels "Cambodia","Ghana": 1 1 1 1 1 1 1 1 1 1 ...
#  $ true_mean_coi             : num [1:240] 1.33 1.28 3.82 3.9 1.36 1.44 4.1 3.43 1.33 1.31 ...
#  $ coi_cat                   : Factor w/ 2 levels "MeanCOI=1.26",..: 1 1 2 2 1 1 2 2 1 1 ...
#  $ between_sample_related_cat: Factor w/ 3 levels "_1%_related",..: 1 1 1 1 2 2 2 2 3 3 ...
#  $ nonrecombinant_cat        : Factor w/ 2 levels "20%_non-recombinant",..: 2 1 2 1 2 1 2 1 2 1 ...
#  $ meiosis_cat               : Factor w/ 2 levels "mean_meiosis=1.0",..: 1 2 1 2 1 2 1 2 1 2 ...
#  $ pop_cat                   : Factor w/ 12 levels "_1%_related__MeanCOI=1.26__20%_non-recombinant",..: 2 1 4 3 6 5 8 7 10 9 ...
```

### Read tool (IDM) results

```{r}
# Read in all IDM results into one tibble ------------------------------
df_tr <- tibble(
  tool_res_path = list.files(
    input_config$tool_res_dir,
    pattern = "*.tsv.gz",
    full.names = TRUE,
    recursive = TRUE
  )
) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(sim_id = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path,
      read_tsv,
      col_types = cols(
        variant = col_character(),
        freq = col_double(),
      ),
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>%
  rename(
    obs_freq = freq,
  ) %>%
  mutate(
    sim_id = factor(sim_id, levels = sort(unique(sim_id))),
    paramset_id = factor(
      paramset_id,
      levels = sort(unique(paramset_id))
    ),
  ) %>%
  # remove placeholder NAs
  drop_na(obs_freq) 
str(df_tr)
#  $ sim_id     : Factor w/ 240 levels "simulated_Cambodia_pop_001",..: 1 1 1 1 1 1 1 1 1 1 ...
#  $ paramset_id: Factor w/ 24 levels "idm_paramset1",..: 1 1 1 1 1 1 1 1 1 1 ...
#  $ variant    : chr [1:1548408] "PF3D7_0104100.1:304:R" "PF3D7_0104100.1:304:S" "PF3D7_0104100.1:311:E" "PF3D7_0104100.1:311:G" ...
#  $ obs_freq   : num [1:1548408] NA NA 0.576 0.424 1 ...
```

### Read tool (IDM) parameters
```{r}
# Read in tool parameters, and join ------------------------------------
df_tp <- jsonlite::fromJSON(input_config$tool_params) %>%
  as_tibble() %>%
  unnest(paramset) %>%
  mutate(
    paramset_id = factor(paramset_id),
    model = factor(model, levels = c("OM", "IDM")),
    eps_initial = factor(eps_initial),
    lambda_initial = factor(lambda_initial)
  )
# str(df_tp)
#  $ paramset_id   : Factor w/ 24 levels "idm_paramset1",..: 1 12 18 19 20 21 22 23 24 2 ...
#  $ model         : Factor w/ 2 levels "OM","IDM": 1 2 1 2 1 2 1 2 1 2 ...
#  $ lambda_initial: Factor w/ 4 levels "0.1","0.3","1",..: 1 1 1 1 1 1 2 2 2 2 ...
#  $ eps_initial   : Factor w/ 3 levels "0.015","0.1",..: 1 1 2 2 3 3 1 1 2 2 ...
```

### Join tables

```{r}
df_tr_wide <- df_tr%>% pivot_wider(names_from = paramset_id, values_from = obs_freq) 
# A tibble: 64,517 × 26
#    sim_id     variant idm_paramset1 idm_paramset10 idm_paramset11 idm_paramset12
#    <fct>      <chr>           <dbl>          <dbl>          <dbl>          <dbl>
#  1 simulated… PF3D7_…        NA             NA             NA             NA    
#  3 simulated… PF3D7_…         0.576          0.576          0.576          0.576
#  5 simulated… PF3D7_…         1              1              1              1    

# NOTE: here, pivot_wider and then pivot_long were called to ensure that there
# is a record for any variants that were true positives or false positives in any
# parameter sets  
df_st_tr_sp_tp <- df_st %>% 
  full_join(df_tr_wide) %>% 
  mutate(across(starts_with("idm_paramset"), ~ replace_na(.x, 0))) %>% 
  mutate(exp_freq = replace_na(exp_freq, 0)) %>% 
  pivot_longer(cols = starts_with("idm_paramset"), names_to = "paramset_id", values_to = "obs_freq")  %>%
  select(sim_id, variant, paramset_id, obs_freq, exp_freq) %>%
  left_join(df_sp, by = "sim_id") %>%
  left_join(df_tp, by = "paramset_id") 
str(df_st_tr_sp_tp)
# tibble [1,692,552 × 15] (S3: tbl_df/tbl/data.frame)
#  $ sim_id                    : chr [1:1692552] "simulated_Cambodia_pop_001" "simulated_Cambodia_pop_001" 
#  $ variant                   : chr [1:1692552] "PF3D7_0508900.1:935:Y" "PF3D7_0508900.1:935:Y" "PF3D7_0508900.1:935:Y" "PF3D7_0508900.1:935:Y" ...
#  $ paramset_id               : chr [1:1692552] "idm_paramset1" "idm_paramset10" "idm_paramset11" "idm_paramset12" ...
#  $ obs_freq                  : num [1:1692552] 1 1 1 1 1 1 1 1 1 1 ...
#  $ exp_freq                  : num [1:1692552] 1 1 1 1 1 1 1 1 1 1 ...
#  $ country                   : Factor w/ 2 levels "Cambodia","Ghana": 1 1 1 1 1 1 1 1 1 1 ...
#  $ true_mean_coi             : num [1:1692552] 1.33 1.33 1.33 1.33 1.33 1.33 1.33 1.33 1.33 1.33 ...
#  $ coi_cat                   : Factor w/ 2 levels "MeanCOI=1.26",..: 1 1 1 1 1 1 1 1 1 1 ...
#  $ between_sample_related_cat: Factor w/ 3 levels "_1%_related",..: 1 1 1 1 1 1 1 1 1 1 ...
#  $ nonrecombinant_cat        : Factor w/ 2 levels "20%_non-recombinant",..: 2 2 2 2 2 2 2 2 2 2 ...
#  $ meiosis_cat               : Factor w/ 2 levels "mean_meiosis=1.0",..: 1 1 1 1 1 1 1 1 1 1 ...
#  $ pop_cat                   : Factor w/ 12 levels "_1%_related__MeanCOI=1.26__20%_non-recombinant",..: 2 2 2 2 2 2 2 2 2 2 ...
#  $ model                     : Factor w/ 2 levels "OM","IDM": 1 2 1 2 1 2 1 2 1 2 ...
#  $ lambda_initial            : Factor w/ 4 levels "0.1","0.3","1",..: 1 2 2 2 3 3 3 3 3 3 ...
#  $ eps_initial               : Factor w/ 3 levels "0.015","0.1",..: 1 2 3 3 1 1 2 2 3 3 ...

```



## Benchmarking 


### Allele frequencies

#### Metrics 

*  Plotting Bland-Altman plots (plotting the mean of expected and observed frequencies vs their difference) 
*  Calculating the RMSE (taking the square root of the mean squared difference between expected and observed)  
*  Calculating [Lin's concordance correlation coefficient](https://www.rdocumentation.org/packages/DescTools/versions/0.99.60/topics/CCC) for agreement on a continuous measure 
*  Calculating contingency table metrics for false positive and false negatives  



#### Bland-Altman Plots 
```{r}
df_st_tr_sp_tp  %>% mutate (is_fp = (exp_freq == 0), is_fn = (obs_freq == 0), is_tp = (exp_freq>0) & (obs_freq>0)) %>% select (is_fp, is_fn, is_tp) %>% count(is_fp, is_fn, is_tp)
```


```{r}
df_ba <-  df_st_tr_sp_tp %>% 
    filter(obs_freq>0, exp_freq > 0) %>% # focusing on true positive allele (excluding false positive and false negative)
    add_diff_mean(expected_col = exp_freq, observed_col = obs_freq)
df_ba

```

First, all of the differences between predicted and expected frequencies were 
pooled together and visualized in a Bland-Altman plot. A Bland-Altman plot 
shows the mean of the observed (in this case, the value estimated by the tool) 
and the expected (in this case, the simulated truth) values for a point 
estimate plotted against the difference in those observed and expected values. 
The horizontal lines indicate two standard deviations in this difference above 
and below the mean. In the plots below, the point estimates were binned into 
hexes, with the color of the hex indicating the density of points.

```{r}
make_bland_altman_plot_by_label(df_ba, "obs_freq")
```

In a Bland-Altman plot, points clustered along y = 0 indicate a good 
prediction, whereas deviations from this line indicate ranges of values that 
have worse performance. We can see that overall IDM has estimated allele 
frequencies quite accurately. Estimates are generally more accurate for high or 
low allele frequencies, which is expected. There are some substantial 
overestimates in the middle of the range, but these are the minority.

Note that false negatives and false positives were excluded from the analysis 
in this section - those are analyzed further below.

2. Showing the different parameter sets 
  - Results are not so sensitive to variations in tool parameters used.

```{r}
#| fig-width: 20
#| fig-height: 12
make_bland_altman_plot_by_label(df_ba, "obs_freq") + 
  facet_wrap(~paramset_id, ncol = 8)
```


3. Showing the different population characteristics 
  - Larger errors occur in simulations with higher COIs and higher relatedness
  - Recombination parameters seem to have little impact 

```{r}
#| fig-width: 20
#| fig-height: 12
make_bland_altman_plot_by_label(df_ba, "obs_freq") + 
  facet_grid(between_sample_related_cat ~interaction(coi_cat, nonrecombinant_cat, sep = "\n"))
```


#### Concordance 

1. Overall 

```{r}
ggpairs(df_ba, columns = c("exp_freq", "obs_freq"), lower = list(continuous = wrap(my_hex)),
        diag  = list(continuous = wrap("barDiag", binwidth = 0.05)), 
        upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)), 
        title = "all")

```


2. Plotting difference parameter sets  
  - Results are not so sensitive to variations in tool parameters used.

```{r}
uniq_parmset_ids <- df_ba %>% distinct(paramset_id) %>% pull(paramset_id)

res <- lapply(uniq_parmset_ids, function(ps_id){
  ggpairs(df_ba %>% filter(paramset_id == ps_id), 
                        columns = c("exp_freq", "obs_freq"), 
                        lower = list(continuous = wrap(my_hex)), 
                        diag  = list(continuous = wrap("barDiag", binwidth = 0.05)), 
                        upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)), 
                        title = ps_id)
  }
) 

names(res) <- uniq_parmset_ids
```


```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(res))
```

3. Different population metrics 


3.1. COI categories 
- Higher COI corresponds to lower concordance.

```{r}
tmp1 = ggpairs(df_ba %>% filter(coi_cat == "MeanCOI=1.26"),
        columns = c("exp_freq", "obs_freq"),
        lower = list(continuous = wrap(my_hex)),
        diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
        upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
        title = "MeanCOI=1.26")
tmp2 = ggpairs(df_ba %>% filter(coi_cat == "MeanCOI=3.41"),
        columns = c("exp_freq", "obs_freq"),
        lower = list(continuous = wrap(my_hex)),
        diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
        upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
        title = "MeanCOI=3.41")

```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("MeanCOI_1.26" = tmp1, "MeanCOI_3.41" = tmp2)))
```


3.2. between sample related categories 
- CCC values:
  - 0.1% related: 0.808
  - 10%  related: 0.816
  - 30%  related: 0.794

```{r}
tmp1 = ggpairs(df_ba %>%
         filter(between_sample_related_cat == "_1%_related"),
       columns = c("exp_freq", "obs_freq"),
       lower = list(continuous = wrap(my_hex)),
       diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
       upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
       title = "_1%_related")
tmp2 = ggpairs(df_ba %>%
         filter(between_sample_related_cat == "10%_related"),
       columns = c("exp_freq", "obs_freq"),
       lower = list(continuous = wrap(my_hex)),
       diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
       upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
       title = "10%_related")
tmp3 = ggpairs(df_ba %>%
         filter(between_sample_related_cat == "30%_related"),
       columns = c("exp_freq", "obs_freq"),
       lower = list(continuous = wrap(my_hex)),
       diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
       upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
       title = "30%_related")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(
  list("_1perc_between_sample_related" = tmp1,
       "10perc_between_sample_related" = tmp2, 
       "30perc_between_sample_related" = tmp3 
       )))
```

3.3. Recombination categories 
- CCC value:
  - 50% non-recombinant: 0.802
  - 20% non-recombinant: 0.812
```{r}
tmp1 = ggpairs(df_ba %>% filter(nonrecombinant_cat == "50%_non-recombinant"),
              columns = c("exp_freq", "obs_freq"),
              lower = list(continuous = wrap(my_hex)),
              diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
              upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
              title = "50%_non-recombinant")
tmp2 = ggpairs(df_ba %>% filter(nonrecombinant_cat == "20%_non-recombinant"),
              columns = c("exp_freq", "obs_freq"),
              lower = list(continuous = wrap(my_hex)),
              diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
              upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
              title = "20%_non-recombinant")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("20perc_non_recomb" = tmp2,
                                      "50perc_non_recomb" = tmp1
                                      )))
```

##### Concordance Correlation Coefficient (CCC) 

```{r}
# calculate Concordance Correlation Coefficient (CCC) 
df_ccc <-  compute_ccc( df_st_tr_sp_tp,
    expected_col  = exp_freq,
    observed_cols = obs_freq,
    group_cols    = c(sim_id, paramset_id)
  )
  
# Gather into long format, left join in the simulation parameters
df_ccc_gat = 
  df_ccc %>%
  select(sim_id, paramset_id, ends_with("_ccc")) %>% 
  gather(allele_freq_estimate, ccc, 3:ncol(.)) %>%
  mutate(allele_freq_estimate = gsub("_ccc", "", allele_freq_estimate)) %>%
  left_join(df_sp)

```

1. CCC over all comparing parameter sets
  - no big difference

```{r}
ggplot(df_ccc_gat %>% mutate(paramset_id = factor(paramset_id, level = uniq_parmset_ids))) +
  geom_boxplot( aes(fill = paramset_id, y = ccc, x = allele_freq_estimate) ) +
  theme_minimal() +
  # scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")


```

```{r}
#| results: asis
#| echo: false
create_tabsetOfHtmlWidgets <- function(htmlObjectsList) {
  out <- capture.output({
    cat("::: {.panel-tabset}\n")
    purrr::iwalk(htmlObjectsList, ~{
      cat('## ', .y, '\n\n', sep = "")
      tempList <- list(item = .x)
      print(htmltools::tagList(tempList))
      cat('\n\n')
    })
    cat(":::\n")
  })
  paste0(out, collapse = "\n")
}


```

2. CCC on the different simulation population characteristics 
   - Higher COI corresponds to lower concordance.

```{r}
tmp1 =  ggplot(df_ccc_gat) +
  geom_boxplot( aes(x = pop_cat, y = ccc, fill = country) ) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

tmp2 = ggplot(df_ccc_gat) + 
  geom_boxplot( aes(x = between_sample_related_cat, y = ccc, fill = country) ) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

tmp3 = ggplot(df_ccc_gat) +
  geom_boxplot( aes(x = nonrecombinant_cat, y = ccc, fill = country) ) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

tmp4 = ggplot(df_ccc_gat) +
  geom_boxplot( aes(x = coi_cat, y = ccc, fill = country) ) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| echo: false
#| results: asis
#| fig-width: 8
#| fig-height: 8

cat(create_tabsetOfGgplotObjects(list(
  "relatedness x coi x recombination x country" = tmp1,
  "relatedness x country" = tmp2,
  "recombinantion x country" = tmp3,
  "coi x country" = tmp4
)))
```

#### RMSE   
```{r}
#' @rdname vectorized_metrics
#' @export
compute_rmse <- function(df, expected_col, observed_cols, group_cols = NULL) {
  stopifnot(is.data.frame(df))
  expected_col <- rlang::enquo(expected_col)

  obs_expr <- rlang::enexpr(observed_cols)
  obs_quos <- if (rlang::is_call(obs_expr, "c")) {
    purrr::map(rlang::call_args(obs_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(obs_expr, env = rlang::caller_env()))
  }

  grp_expr <- rlang::enexpr(group_cols)
  grp_quos <- if (missing(group_cols) || is.null(grp_expr)) {
    list()
  } else if (rlang::is_call(grp_expr, "c")) {
    purrr::map(rlang::call_args(grp_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(grp_expr, env = rlang::caller_env()))
  }

  metric_fun <- function(q) {
    obs_name <- rlang::as_name(q)
    if (length(grp_quos) > 0) {
      df %>%
        dplyr::group_by(!!!grp_quos) %>%
        dplyr::summarise(
          !!paste0(obs_name, "_rmse") := {
            x <- !!q; y <- !!expected_col
            ok <- !is.na(x) & !is.na(y) & (x>0) & (x<1) & (y>0) & (y<1)
            n  <- sum(ok)
            if (n == 0) NA_real_ else sqrt(sum((x[ok] - y[ok])^2) / n)
          },
          .groups = "drop"
        )
    } else {
      dplyr::summarise(df,
        !!paste0(obs_name, "_rmse") := {
          x <- !!q; y <- !!expected_col
          ok <- !is.na(x) & !is.na(y)
          n  <- sum(ok)
          if (n == 0) NA_real_ else sqrt(sum((x[ok] - y[ok])^2) / n)
        }
      )
    }
  }

  if (length(grp_quos) > 0) {
    out <- df %>% dplyr::distinct(!!!grp_quos)
    join_by_cols <- names(out)
    for (q in obs_quos) {
      out <- dplyr::left_join(out, metric_fun(q), by = join_by_cols)
    }
  } else {
    out <- dplyr::bind_cols(purrr::map(obs_quos, metric_fun))
  }

  out
}

```

```{r}
# calculate RMSE 
df_rmse = df_st_tr_sp_tp %>%
  compute_rmse(
    expected_col  = exp_freq,
    observed_cols = obs_freq,
    group_cols    = c(sim_id, paramset_id)
  )

# Gather into long format, left join in the simulation parameters
df_rmse_gat = df_rmse %>%
  gather(allele_freq_estimate, rmse, 3:ncol(.)) %>%
  mutate(allele_freq_estimate = gsub("_rmse", "", allele_freq_estimate))%>%
  left_join(df_sp)
```


1. RMSE over all  parameter sets
  - no differences

```{r}
#| echo: false
#| results: asis
#| fig-width: 8
#| fig-height: 8
ggplot(df_rmse_gat %>% mutate(paramset_id = factor(paramset_id, level = uniq_parmset_ids))) +
  geom_boxplot( aes( y = rmse, x = paramset_id) ) +
  custom_ggplot_theme_xRotate +
  theme_minimal() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


2. RMSE on the different simulation population characteristics 
  - Higher COI corresponds to higher RMSE

```{r}

tmp1 = ggplot(df_rmse_gat) +
  geom_boxplot( aes(x = pop_cat, y = rmse, fill = country) ) +
  custom_ggplot_theme_xRotate

tmp2 = ggplot(df_rmse_gat) +
  geom_boxplot( aes(x = between_sample_related_cat, y = rmse, fill = country) ) +
  custom_ggplot_theme_xRotate

tmp3 = ggplot(df_rmse_gat) +
  geom_boxplot( aes(x = nonrecombinant_cat, y = rmse, fill = country) ) +
  custom_ggplot_theme_xRotate

tmp4 = ggplot(df_rmse_gat) +
  geom_boxplot( aes(x = coi_cat, y = rmse, fill = country) ) +
  custom_ggplot_theme_xRotate

```

```{r}
#| echo: false
#| results: asis
#| fig-width: 8
#| fig-height: 8

cat(create_tabsetOfGgplotObjects(list(
  "relatedness x coi x recombination x country" = tmp1,
  "relatedness x country" = tmp2,
  "recombination x country" = tmp3,
  "coi x country" = tmp4

)))
```



<!-- #### False positive and false negatives -->

```{r}
# True Positive is both estimator and expected is great than 0 
# False Negative is that expected > 0 and estimator is 0 
# False Positive is that expected is 0 and estimator is > 0

df_fnfp <- df_st_tr_sp_tp %>% 
  mutate(TP = as.numeric(exp_freq > 0 & obs_freq > 0), 
         FN = as.numeric(exp_freq > 0 & obs_freq == 0), 
         FP = as.numeric(exp_freq == 0 & obs_freq > 0))

hap_metric_map <- c(
  "Recall"               = "recall",
  "Precision"            = "precision",
  "NPV"                  = "tpr",
  "F1-score"             = "f1_score",
  "False Discovery Rate" = "fdr",
  "Overlap Coefficient"  = "overlap_coefficient"
)
```


1. all

```{r}
#| echo: false
#| results: asis
#| fig-width: 8
#| fig-height: 8
tmp <-bind_rows(
  df_fnfp %>%  add_contigency_metrics_no_TN(summarize_first = T) %>% mutate(label='all'),
)
radar_by_metrics(tmp,  metric_map = hap_metric_map, color_col = "label")

```



2. Different coi, relatedness, recombination, and country
  - low coi and cambodia is associated with better accuracy metrics

```{r}
#| echo: false
#| results: asis
#| fig-width: 8
#| fig-height: 8

# coi_cat
tmp <-bind_rows(
  df_fnfp %>% filter(coi_cat == "MeanCOI=1.26")%>%  add_contigency_metrics_no_TN(summarize_first = T) %>% mutate(label='MeanCOI=1.26'),
  df_fnfp %>% filter(coi_cat == "MeanCOI=3.41")%>%  add_contigency_metrics_no_TN(summarize_first = T) %>% mutate(label='MeanCOI=3.41'),

)
tmp1 <- radar_by_metrics(tmp,  metric_map = hap_metric_map, color_col = "label")

# between_sample_related_cat
tmp <-bind_rows(
  df_fnfp %>% filter(between_sample_related_cat == "_1%_related")%>%  add_contigency_metrics_no_TN(summarize_first = T) %>% mutate(label='_1%_related'),
  df_fnfp %>% filter(between_sample_related_cat == "10%_related")%>%  add_contigency_metrics_no_TN(summarize_first = T) %>% mutate(label='10%_related'),
  df_fnfp %>% filter(between_sample_related_cat == "30%_related")%>%  add_contigency_metrics_no_TN(summarize_first = T) %>% mutate(label='30%_related')
)
tmp2 <- radar_by_metrics(tmp,  metric_map = hap_metric_map, color_col = "label")

# nonrecombinant_cat
tmp <-bind_rows(
  df_fnfp %>% filter(nonrecombinant_cat == "50%_non-recombinant")%>%  add_contigency_metrics_no_TN(summarize_first = T) %>% mutate(label='50%_non-recombinant'),
  df_fnfp %>% filter(nonrecombinant_cat == "20%_non-recombinant")%>%  add_contigency_metrics_no_TN(summarize_first = T) %>% mutate(label='20%_non-recombinant')
)
tmp3 <- radar_by_metrics(tmp,  metric_map = hap_metric_map, color_col = "label")

# country
tmp <-bind_rows(
  df_fnfp %>% filter(country == "Cambodia")%>%  add_contigency_metrics_no_TN(summarize_first = T) %>% mutate(label='Cambodia'),
  df_fnfp %>% filter(country == "Ghana")%>%  add_contigency_metrics_no_TN(summarize_first = T) %>% mutate(label='Ghana')
)
tmp4 <- radar_by_metrics(tmp,  metric_map = hap_metric_map, color_col = "label")

cat(create_tabsetOfGgplotObjects(list(
  'coi' = tmp1,
  "relatedness" = tmp2,
  "recombination" = tmp3,
  "country" = tmp4
)))

```





