---
title: "Using FreqEstimationModel to Calculate Multi-Locus Allele Frequency"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(kableExtra)
```

# The data
We are going to use data from here [main Data tab](https://mrc-ide.github.io/PGEforge/website_docs/data_description.html).
```{r, warning=FALSE, message=FALSE, error=FALSE}
library(tibble)
library(dplyr)
library(FreqEstimationModel)
library(foreach)
library(doMC)
library(iterators)
library(parallel)
library(rngtools)
library(validate)
library(magrittr)
library(stringr)
library(optparse)
library(readr)
library(purrr)
library(variantstring)
```

## Transferring data into the PlasmoGenEpi format
For now, I'm reading in a sample amino_acid_calls table. Working on creating this from the vcf file!
```{r formatting_1}
setwd("~/Life/Collaboration with Dr. Wesolowski/for_alfred/test_PGE/PGEcore/scripts/FreqEstimationModel_wrapper/")
input_data <- read.csv("example_amino_acid_calls.tsv", sep="\t")
```

## Formatting the data from PGE to FEM
We need to transform the data from PlasmoGenEpi format to FEM format, which is a matrix indicating whether targets in a given sample are heterozygous, match the reference sequence, or are exclusively mutated.

The sample matrix is of the form n x m where n = the samples in the dataset and m = targets. Each position will indicate if there is alignment at the given target.
```{r formatting_2}
input_data$unique_targets <- paste(input_data$gene_id, input_data$aa_position, sep=":")
unique_targets <- unique(input_data$unique_targets)
unique_sample_ids <- unique(input_data$specimen_id)

sample_matrix <- matrix(99,  nrow=length(unique_sample_ids), ncol=length(unique_targets))
#99 is the "no data" identifier
colnames(sample_matrix) <- unique_targets
rownames(sample_matrix) <- unique_sample_ids
```


For this tutorial, I'm iterating through the dataset first by sample and then by target. We filter the PGE data to only look at *unique* reads at a given target, and then count how many unique calls there are. Currently, FEM does not support >2 alleles at a given target.
```{r sample_matrix population}
# populate the input matrix with 0 (only reference), 0.5 ("het" cal), or 1 (only non-reference)
for(sample in input_data$specimen_id){
  cut_df <- input_data[input_data$specimen_id==sample,]
  for(unique_target in unique_targets){
    cut_cut_df <- cut_df[cut_df$unique_targets==unique_target,]
    nvals <- length(unique(cut_cut_df$aa)) 
    if (nvals > 2){
      print("too many alleles for FEM")
    }
    if (nvals == 2){
      sample_matrix[sample, unique_target] <- 0.5 # "heterozygous" call
    }
    if (nvals == 1){
      if (cut_cut_df[1, "ref_aa"]==cut_cut_df[1, "aa"]){
        val <- 0
      }
      else {val <- 1}
      sample_matrix[sample, unique_target] <- val
    }
  
  }
}

print(sample_matrix)
```

##Running FEM

I took this directly from Aimee. I need to understand it.
```{r Running FEM}
#alt_alleles <- sample_matrix_list[[2]]
COI <- 3
seed <- 1

data_summary <- list()
data_summary$Data <- sample_matrix
#data_summary <- data.frame(data_summary)
runtime <- system.time({
    print('point -1')
    thinning_interval <- 1 # Number of iterations per chain that are not saved
    no_traces_preburnin <- 10000 # For more traces but manageable pdf plots, don't exceed 10k and increase thinning interval instead
    no_mcmc_chains <- 3 # Number of MCMC chains to run
    parallel <- FALSE # Set to true if running code in parallel (this option isn't active yet)
    log_like_zero <- FALSE # QC check: sets log(p(yi|ai)) to zero (only impacts NGS)
    NGS <- FALSE
    mcmc_variable_list <- list(
        no_mcmc_chains = no_mcmc_chains,
        no_traces_preburnin = no_traces_preburnin,
        thinning_interval = thinning_interval,
        NGS = NGS,
        log_like_zero = log_like_zero
    )
    if (!NGS | log_like_zero) { # Set to true to keep partial observations rather than discard them
        augment_missing_data <- TRUE
    } else {
        augment_missing_data <- FALSE
    }
    if (log_like_zero) { # When log_like_zero == TRUE, code should return the prior
        moi_prior <- "Uniform" # Use a prior that is easy to eye-ball
    } else {
        moi_prior <- "Poisson" # Choose between 'Uniform', 'Poisson', 'Geometric' or 'nBinomial'
    }
    moi_max <- 8 # Maximum MOI regarded as possible by the model (I haven't tested beyond 20)
    print('point 0')
    
    moi_hyperparameter <- COI 
    moi_size_hyperparameter <- 0.5 # Only applies if moi_prior == 'nBinomial' (hyperparameter for the prior on the MOI if the prior is the negative Binomial)
    moi_prior_min2 <- NULL # Specify the lower bound for the MOI per individual
    moi_initial <- NULL # If null, the initial vector of MOIs is set internally, otherwise set to input moi_initial
    moi_list <- list(
        moi_hyperparameter = moi_hyperparameter,
        moi_size_hyperparameter = moi_size_hyperparameter,
        moi_prior = moi_prior,
        moi_max = moi_max,
        moi_prior_min2 = moi_prior_min2,
        moi_initial = moi_initial
    )
    processed_data_list <- preprocess_data(
        data_summary,
        log_like_zero,
        NGS,
        augment_missing_data,
        moi_prior_min2
    )
    
    print("pointA")
    test_process <- preprocess_data(data_summary, FALSE, FALSE, TRUE, FALSE)
    print("pointB")
    frequency_hyperparameter <- rep(1, processed_data_list$no_haplotypes) # The Parameter vector for the Dirichlet prior on the frequency vector
    frequency_initial <- NULL # If null, external_frequency_initial set internally, otherwise set to input external_frequency_initial
    frequency_list <- list(
        frequency_hyperparameter = frequency_hyperparameter,
        frequency_initial = frequency_initial
    )
    
    set.seed(1)

    # Run MCMC - works up to here
    results <- mcmc_sampling_parallel(processed_data_list,
        moi_list,
        frequency_list,
        mcmc_variable_list,
        cores_max = 3
    )
    print('point 1')
    # Generate numerical approximations of posteriors by removing burnin
    burnin <- 1:(0.5 * mcmc_variable_list$no_traces_preburnin) # remove first half
    if (mcmc_variable_list$no_mcmc_chains > 1) {
        alply_genotype_freq_store_chains_burnin <- plyr::alply(results$genotype_freq_store_chains[-burnin, , ], 3)
    } else {
        alply_genotype_freq_store_chains_burnin <- results$genotype_freq_store_chains[-burnin, , ]
    }
    print('point 2')
    # Use mcmc.list to represent frequency chains across parallel runs (see ?coda::mcmc.list)
    mcmc_frequency_chains <- coda::mcmc.list(lapply(alply_genotype_freq_store_chains_burnin, # 3 for splitting by third dimension (nothing to do with no. of chains)
        coda::mcmc,
        start = (max(burnin) + 1) * mcmc_variable_list$thinning_interval,
        end = mcmc_variable_list$no_traces_preburnin * mcmc_variable_list$thinning_interval,
        thin = mcmc_variable_list$thinning_interval
    ))
    print('point 3')
    mcmc_As <- abind::abind(plyr::alply(results$genotype_count_store_chains[-burnin, , , ], 4), along = 1) # Haplotype counts for all chains excluding burnin
    mcmc_mois <- apply(mcmc_As, c(1, 2), sum)

    # Extract population-level haplotype frequency estimates
    pop_freq <- cbind(
        freq = summary(mcmc_frequency_chains)$statistics[, "Mean"],
        median_freq = summary(mcmc_frequency_chains)$quantiles[, 3],
        "CI_2.5" = summary(mcmc_frequency_chains)$quantiles[, 1],
        "CI_97.5" = CI_upper <- summary(mcmc_frequency_chains)$quantiles[, 5]
    )
    print("point 4")
    pop_prev <- 1-(1-pop_freq)^median(mcmc_mois)
    inf_prev <- t(apply(mcmc_As, 2, function(x) colMeans(x > 0)))
    prev <- colMeans(inf_prev)
    pop_freq <- cbind(pop_freq, prev)
})
# Convert row names to a new column named "sequence"
sequence_column <- data.frame(sequence = rownames(pop_freq), stringsAsFactors = FALSE)
pop_freq <- cbind(sequence_column, pop_freq)
rownames(pop_freq) <- NULL
```


##Interpreting the Output


Make as many sections as you need. Try to break long documents out over multiple pages.

## Summary

It's a good idea to round off with a summary of what we did.

PS, note that the name of the tutorial on the PGEforge website can be whatever you want - it does not have to match the name of this tutorial.