---
title: "FreqEstimationModel"
output: html_document
author: "Shazia Ruybal-Pesántez"
date: "2025-07-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose

The `FreqEstimationModel` R package estimates population-level frequencies of genetic markers of antimalarial resistance where

- genetic markers of antimalarial resistance are either alleles at single biallelic SNPs or sequences of alleles (haplotypes) over a small number of biallelic SNPs (at most seven) in the malaria parasite genome;
- population-level parasite frequencies (vs within-infection parasite frequencies) are estimated by jointly modelling parasite genetic data on many infections, including those that contain genetically distinct parasites (i.e., those that are polyclonal);
- parasite genetic data are prevalence data (i.e., they describe the presence of alleles). A non-default version of the model does exploit information on read-depths (see note on chapter 6 below), but it has not been peer-reviewed.

## Existing resources

The [`FreqEstimationModel` GitHub repository](https://github.com/aimeertaylor/FreqEstimationModel) includes scripts for implementing FreqEstimationModel as well as vignettes demonstrating usage. 

## Citation

The publications associated with the `FreqEstimationModel` can be found at: 

- [Taylor et al 2014 Malaria Journal](https://malariajournal.biomedcentral.com/articles/10.1186/1475-2875-13-102)
- [Taylor et al 2016 Open Forum Infectious Diseases](https://doi.org/10.1093/ofid/ofw229)