---
title: "Freq Estimation Model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose

Here is the purpose of FEM as taken from the github repository: 
"An R package developed to estimate population-level frequencies of genetic markers of antimalarial resistance where

* genetic markers of antimalarial resistance are either alleles at single biallelic SNPs or sequences of alleles (haplotypes) over a small number of biallelic SNPs (at most seven) in the malaria parasite genome;
* population-level parasite frequencies (vs within-infection parasite frequencies) are estimated by jointly modelling parasite genetic data on many infections, including those that contain genetically distinct parasites (i.e., those that are polyclonal);
* parasite genetic data are prevalence data (i.e., they describe the presence of alleles)." 

FEM allows for the calculation of multi-locus allele frequencies for biallelic SNPs.

## Existing resources

- All information is [here](https://github.com/aimeertaylor/FreqEstimationModel)

## Citation
```
@Manual{,
  title = {FreqEstimationModel},
  author = {Aimee Taylor)}},
  year = {2015},
  note = {R package version 0.1, commit 8c010f47579d947e0ff1d90a7f96893555435462},
  url = {https://github.com/aimeertaylor/FreqEstimationModel},
}
```