---
title: "benchmarking_functions"
author: "JNN"
date: "2025-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)



```
## Load in the Benchmarking Outputs
```{r}
library(dplyr)
library(ggplot2)
library(jsonlite)
library(magrittr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(tidyr)

inp <- list(
  sim_info_dir = "C:/Users/Josphat.Nyataya/OneDrive - USAMRU Kenya/Desktop/Desktop/R_analyses_2024/naive_slaf_results/simulations/SpotMalaria", 
  tool_res_dir = "C:/Users/Josphat.Nyataya/OneDrive - USAMRU Kenya/Desktop/Desktop/R_analyses_2024/naive_slaf_results/naive_slaf", 
  tool_params = "C:/Users/Josphat.Nyataya/OneDrive - USAMRU Kenya/Desktop/Desktop/R_analyses_2024/naive_slaf_results/naive_slaf.json"
)
```
##First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.
```{r}
# Read all expected frequencies into one tibble ------------------------
sim_truth <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_aa_variants_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_integer(), 
        population_name = col_character(), 
        country = col_character(), 
        GeneName = col_character(), 
        transcript_id = col_character(), 
        aa = col_character(), 
        is_ref = col_logical(), 
        freq = col_double(), 
        prev = col_double()
      ), 
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>%
  rename(
    sim_id = population_name, 
    exp_freq = freq, 
    exp_prev = prev, 
    exp_n_samp = sample_count
  ) %>%
  select(
    -pop, 
    -country, 
    -allele_count, 
    -GeneName, 
    -is_ref, 
    -total_pop_samples, 
    -total_pop_alleles, 
    -unique_alleles
  )

# Read in all simulation parameters into one tibble --------------------
sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>%
  select(
    -collection_year, 
    -obs_freq_monoclonal, 
    -obs_non_recombinant, 
    -mean_k, 
    -sample_count, 
    -est_between_sample_relatedness
  ) %>%
  rename(
    sim_id = population_name, 
    true_mean_coi = mean_coi
  )
sim_truth <- sim_truth %>%
  left_join(sim_params, by = "sim_id")

# Read in all FEM results into one tibble ------------------------------
tool_results <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "*.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(sim_id = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(), 
        variant = col_character()
      ), 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>%
  rename(
    obs_freq = freq
  )

# Read in tool parameters, and join ------------------------------------
tool_params <- jsonlite::fromJSON(inp$tool_params) %>%
  as_tibble() %>%
  mutate(slaf_method = paramset$method) %>%
  select(-paramset)
tool_results <- tool_results %>%
  left_join(tool_params, by = "paramset_id")

# Join everything into one tibble --------------------------------------
sim_truth <- sim_truth %>% mutate(variant = paste(transcript_id, aa_position, aa, sep = ":"))
tool_results_wtruth <- tool_results %>%
  # Use an inner join because FEM also estimates frequencies for 
  # subsets of the requested multilocus genotypes, and these 
  # frequencies are not benchmarked
  inner_join(sim_truth, by = c("sim_id", "variant"))
```
## Loaded it into R session
```{r}
getwd()
source("hackathon_helper.R")
```
##selected the requred columns (observed and expected from the original data) and rename them accordingly
```{r}
combined_s <- full_join(sim_truth, tool_results, by = c("sim_id", "variant"))
df <- combined_s %>% select(obs_freq, exp_freq, slaf_method)
```
##run the means and difference of the observed and expected columns
```{r}
df <- add_diff_mean(df, expected_col = exp_freq, observed_col = obs_freq)
```
##Now we can make the bland altman plot
```{r}
g1 <- make_bland_altman_plot_by_label(df, label = "obs_freq", bins = 30) + transparentBackground
g1
```
## To evaluate the performance of each method, we split the BA plot by slaf_method
```{r}
df_rc <- df %>% filter(slaf_method == "read_count_prop")
df_pa <- df %>% filter(slaf_method == "presence_absence")
```
##Make BA for slaf_method read_count
```{r}
s1 <- make_bland_altman_plot_by_label(df_rc, label = "obs_freq", bins = 30) + transparentBackground 
s1
```
## Do the same for presence_absence
```{r}
s2 <- make_bland_altman_plot_by_label(df_pa, label = "obs_freq", bins = 30) + transparentBackground
s2
```
## We now proceed to make the general concordance plot
```{r}
g2 <- my_hex(df, aes(x = exp_freq, y = obs_freq, bin = 30)) + transparentBackground
g2
```
## create a correlation plot by read count method
```{r}
c1 <- my_hex(df_rc, aes(x = exp_freq, y = obs_freq, bin = 30)) + transparentBackground
c1
```
##make the same plot by presence_absence
```{r}
c2 <- my_hex(df_pa, aes(x = exp_freq, y = obs_freq, bin = 30)) + transparentBackground
c2
```
## generate the overal correlation coefficient and lin's ccc
```{r}
d2 <- my_corr_ccc(df, aes(x = exp_freq, y = obs_freq, method = "pearson", digits = 3, ccc_conf = 0.95, size = 4))
d2
```
## We can aslo estimate the correlation coefficients of each methodd
## start with read_count
```{r}
d3_rc <- my_corr_ccc(df_rc, aes(x = exp_freq, y = obs_freq, method = "pearson", digits = 3, ccc_conf = 0.95, size = 4))
d3_rc
```
## Do the same correlation for presence_absence
```{r}
d3_pa <- my_corr_ccc(df_pa, aes(x = exp_freq, y = obs_freq, method = "pearson", digits = 3, ccc_conf = 0.95, size = 4))
d3_pa
```
##To generate the contingency tables, we extract the TP, FN and FP 
```{r}
#classify the variants as "False positive" or as "False Negative"
combined_2 <- combined_s %>% mutate(status = case_when(!is.na(obs_freq) & is.na(exp_freq) ~ "FP", is.na(obs_freq) & !is.na(exp_freq) ~ "FN", TRUE ~ "TP"))

#summarize the counts
count_summarize <- combined_2 %>% group_by(status) %>% summarise(count = n(), .groups = "drop")

count_summarize
```
##convert the count_summary tibble to a dataframe and ensure the three columns are well labelled
```{r}
count_summarize_df <- as.data.frame(count_summarize)
input <- as.data.frame(t(count_summarize_df))
colnames(input)[1] <- "FN"
colnames(input)[2] <- "FP"
colnames(input)[3] <- "TP"
input <- input[-1, ]
input <- input %>% mutate(TP = as.numeric(TP),                                                                     FP = as.numeric(FP),                                                                     FN = as.numeric(FN)                                            )

input
```
## Generate the contingency table
```{r}
d1 <- input2 <- add_contigency_metrics_no_TN(input = input, summarize_first = F)
d1
```
##generate a radar plot but we first generate the metric map
```{r}
metric_map1 <- c("Sensitivity" = "sensitivity", "Precision" = "precision", "F1-score" = "f1_score", "FDR" = "fdr")
g3 <- radar_by_metrics(input2, color_col = NULL, facet_col = NULL, metric_map = metric_map1, palette = NULL, title = "Perfomance radar", outer_pad = 0.12, ring_breaks = seq(0,1,0.25))
g3
```


