---
title: "MultiLociBiallelicModel"
output: html_document
author: "Nicholas Hathaway"
date: "2023-12-11"
bibliography: reference.bib
code-link: true
code-copy: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose  

The code supplied by this paper does a maximum-likelihood (MLE) method to estimate: 

1)  haplotype frequencies and prevalence
2)  multiplicity of infection (MOI/COI) from SNP data. 

The functions here provide functionality to predict possible haplotype prevalence within the population that lead to the current set of data. Also estimates MOI/COI based on these estimates. Has to take only biallelic SNPs and compuationally can be limited by the number of loci supplied (in publication used 10 loci). 

## Existing resources

- Example file can be found [here](https://github.com/Maths-against-Malaria/MultiLociBiallelicModel/blob/main/src/SNP_MLE.R)

## Citation

[@Tsoungui_Obama2022-gz]


