---
title: "Pixelate"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose

We often need to make maps of prevalence, but these are estimates and should ideally be presented with their associated uncertainty. Displaying uncertainty is challenging because the scale can be difficult to interpret and the information also tends to be presented in a separate map. Pixelate solve this problem by simultaneously presenting predictions and uncertainty in a single map by varying pixel size. Large pixels (and hence "blocky" areas of the map) represent high uncertainty, while small pixels (more "high resolution" areas of the map) represent low uncertainty. This provides an intuitive way of putting all the information in once place.

## Existing resources

The Pixelate package contains vignettes with both quick and detailed examples. You can access these vignettes by loading the package and running `vignette("pixelate")`.

## Citation

The method is described in [this arXiv preprint](https://doi.org/10.48550/arXiv.2005.11993):

Taylor, Aimee R., James A. Watson, and Caroline O. Buckee. "Pixelate to communicate: visualising uncertainty in maps of disease risk and other spatial continua." arXiv preprint arXiv:2005.11993 (2020).
