---
title: MOIRE Benchmarking Results
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)
```


## Set up 

### Libraries used in the analysis below 

```{r, warning=FALSE, message=FALSE}
# data manipulation libraries 
library(dplyr)
library(readr)
library(stringr)
library(tibble)
library(tidyr) 
library(magrittr)
library(jsonlite)
library(purrr)
library(rlang) # for as_name

# plotting 
library(ggplot2)
library(scales) # help with axises 
library(GGally) # ggpairs for pairwise comparisons 

# interactive html outputs
library(plotly) # ggplotly to turn any ggplot object into an interactive plot 
library(DT) # data tables for displaying tables 

# stats package
library(DescTools) # CCC, Calculates Lin's concordance correlation coefficient
```

### Some helpful R utilities 

```{r}
#| code-fold: true

# commonly used functions  
`%!in%` <- Negate(`%in%`)
is.notna <-function(x){
  return(!is.na(x))
}

# ggplot themes 
transparentBackground = theme(
  panel.background = element_rect(fill='transparent'), #transparent panel bg
  plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
  # panel.grid.major = element_blank(), #remove major gridlines
  # panel.grid.minor = element_blank(), #remove minor gridlines
  legend.background = element_rect(color = NA, fill='transparent'), #transparent legend bg
  legend.box.background = element_rect(color = NA, fill='transparent') #transparent legend panel
)
custom_ggplot_theme_noTransparentBackground = theme_bw() +
  # theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank() )+
  theme(axis.line.x = element_line(color="black", linewidth = 0.3),axis.line.y =
          element_line(color="black", linewidth = 0.3))+
  theme(text=element_text(size=12, family="Helvetica"))+
  theme(axis.text.y = element_text(size=12))+
  theme(axis.text.x = element_text(size=12)) +
  theme(legend.position = "bottom") + 
  theme(plot.title = element_text(hjust = 0.5))

custom_ggplot_theme = custom_ggplot_theme_noTransparentBackground +
  transparentBackground
custom_ggplot_theme_backgroundTransparent = custom_ggplot_theme

custom_ggplot_theme_xRotate_noBackgroundTransparent = theme_bw() +
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank() )+
  theme(axis.line.x = element_line(color="black", linewidth = 0.3),axis.line.y =
          element_line(color="black", linewidth = 0.3))+
  theme(text=element_text(size=12, family="Helvetica"))+
  theme(axis.text.y = element_text(size=12))+
  theme(axis.text.x = element_text(size=12)) +
  theme(legend.position = "bottom") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.text.x = element_text(size=12, angle = -90, vjust = 0.5, hjust = 0)) 

custom_ggplot_theme_xRotate = custom_ggplot_theme_xRotate_noBackgroundTransparent + 
  transparentBackground

custom_ggplot_theme_xRotate_backgroundTransparent = custom_ggplot_theme_xRotate

# color blind safe distinct color palettes 
colorPalette_08 = c("#2271B2","#F748A5","#359B73","#F0E442","#D55E00","#3DB7E9","#E69F00","#000000")
colorPalette_12 = c("#E20134","#FF6E3A","#008DF9","#8400CD","#FFC33B","#9F0162","#009F81","#FF5AAF","#00FCCF","#00C2F9","#FFB2FD","#A40122")
colorPalette_15 = c("#F60239","#003C86","#EF0096","#9400E6","#009FFA","#008169","#68023F","#00DCB5","#FFCFE2","#FF71FD","#7CFFFA","#6A0213","#008607","#00E307","#FFDC3D")


# creating a datatable html widget
create_dt <- function(x){
  DT::datatable(x,
                extensions = 'Buttons',
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All"))), 
                filter = "top")
}

# for creating tabsets html outputs
create_tabsetOfHtmlWidgets <- function(htmlObjectsList) {
  zz <- textConnection("foo", "w")
  sink(zz)
  cat("::: {.panel-tabset}\n")
  iwalk(htmlObjectsList, ~ {
    cat('## ', .y, '\n\n')
    tempList = list()
    tempList[["item"]] = .x
    print(htmltools::tagList(tempList))
    cat('\n\n')
  })
  cat(":::\n")
  sink()
  close(zz)
  paste0(foo, collapse = "\n")
}

create_tabsetOfGgplotObjects <- function(ggplotObjectsList) {
  cat("\n\n::: {.panel-tabset}\n\n")
  iwalk(ggplotObjectsList, ~ {
    cat('## ', .y, '\n\n')
    print(.x)
    cat('\n\n')
  })
  cat(":::\n")
}

```

```{r}
#' Add various contingency table metrics, input must have TP, FP, TN, FN columns defined already
#'
#' @param input A data table with at least TP, FP, TN, FN
#' @param summarize_first Whether or not to sum up the columns first, will lose any non-grouping columns though
#'
#' @returns the input table plus various performance metrics 
#' @export
add_contigency_metrics <- function(input, summarize_first = F){
  output = input
  if(summarize_first){
    output = output  %>%
      summarise(TP = sum(TP), 
                FN = sum(FN), 
                FP = sum(FP), 
                TN = sum(TN))
  }
 output = output %>%
  mutate(
    sensitivity           = if_else(TP + FN > 0, TP / (TP + FN), NA_real_),
    recall                = sensitivity,
    specificity           = if_else(TN + FP > 0, TN / (TN + FP), NA_real_),
    ppv                   = if_else(TP + FP > 0, TP / (TP + FP), NA_real_),
    precision             = ppv,
    npv                   = if_else(TN + FN > 0, TN / (TN + FN), NA_real_),
    accuracy              = (TP + TN) / (TP + TN + FP + FN),
    prevalence            = (TP + FN) / (TP + TN + FP + FN),
    overlap_coefficient   = ifelse(TP + FP + FN > 0, TP / (TP + FP + FN), NA_real_),
    tpr                   = sensitivity,
    fpr                   = if_else(FP + TN > 0, FP / (FP + TN), NA_real_),
    fdr                   = if_else(TP + FP > 0, FP / (TP + FP), NA_real_),
    f1_score              = if_else((2 * TP + FP + FN) > 0, (2 * TP) / (2 * TP + FP + FN), NA_real_),
    balanced_accuracy     = (sensitivity + specificity) / 2
  )
  return(output)
}


#' Add various contingency table metrics but only if they don't depend on TN, input must have TP, FP, FN columns defined already
#'
#' @param input A data table with at least TP, FP, FN
#' @param summarize_first Whether or not to sum up the columns first, will lose any non-grouping columns though
#' @returns the input table plus various performance metrics 
#' @export
add_contigency_metrics_no_TN <- function(input, summarize_first = F){
  output = input
  if(summarize_first){
    output = output  %>%
      summarise(TP = sum(TP), 
                FN = sum(FN), 
                FP = sum(FP))
  }
 output = output %>%
  mutate(
    sensitivity           = if_else(TP + FN > 0, TP / (TP + FN), NA_real_),
    recall                = sensitivity,
    ppv                   = if_else(TP + FP > 0, TP / (TP + FP), NA_real_),
    precision             = ppv,
    overlap_coefficient   = ifelse(TP + FP + FN > 0, TP / (TP + FP + FN), NA_real_),
    tpr                   = sensitivity,
    fdr                   = if_else(TP + FP > 0, FP / (TP + FP), NA_real_),
    f1_score              = if_else((2 * TP + FP + FN) > 0, (2 * TP) / (2 * TP + FP + FN), NA_real_)
  )
  return(output)
}

```

```{r}

#' Plot metrics that range from 0 to 1 in radar perform with optional coloring and faceting, all input must have 1 obs per df (+/- groupings)
#'
#' @param df The summarized metrics table 
#' @param color_col an optional coloring if there are multiple groups in data 
#' @param facet_col an optional faceting if there are multiple groups in data
#' @param metric_map a map to turn the metrics into nice lables, can also be a 1:1 table, is also used to indicate what metrics are being plotted 
#' @param palette the coloring palette being used if coloring, will be passed to scale_color_manual
#' @param title the title for the plot
#' @param outer_pad the padding used to make sure labels aren't clipped off 
#' @param ring_breaks where to put the breaks in the rignt 
#'
#' @returns a ggplot object of the plot which can be further modified 
#' @export
radar_by_metrics <- function(df,
                             color_col = NULL,          
                             facet_col = NULL,          
                             metric_map,                # named chr: "Nice_Col_Name_Bro" = "df_col"
                             palette = NULL,
                             title = "Performance radar",
                             outer_pad = 0.12,
                             ring_breaks = seq(0, 1, 0.25)) {
  # Requires: ggplot2, dplyr, tidyr, tibble, scales
  is_named <- function(x) !is.null(names(x)) && all(nzchar(names(x)))
  stopifnot(is.character(metric_map), is_named(metric_map))
  stopifnot(all(unname(metric_map) %in% names(df)))

  has_color <- is.character(color_col) && length(color_col) == 1L && nzchar(color_col)
  has_facet <- is.character(facet_col) && length(facet_col) == 1L && nzchar(facet_col)

  sel_cols <- c(if (has_color) color_col,
                if (has_facet) facet_col,
                unname(metric_map))

  metrics <- names(metric_map); k <- length(metrics)

  # Long format
  radar_long_ix <- df %>%
    dplyr::select(dplyr::all_of(sel_cols)) %>%
    {
      if (has_color && has_facet) dplyr::rename(., .color = dplyr::all_of(color_col),
                                                   .facet = dplyr::all_of(facet_col))
      else if (has_color && !has_facet) dplyr::rename(., .color = dplyr::all_of(color_col)) %>%
             dplyr::mutate(.facet = "(all)")
      else if (!has_color && has_facet) dplyr::rename(., .facet = dplyr::all_of(facet_col)) %>%
             dplyr::mutate(.color = "(all)")
      else dplyr::mutate(., .color = "(all)", .facet = "(all)")
    } %>%
    tidyr::pivot_longer(cols = dplyr::all_of(unname(metric_map)),
                        names_to = "metric_key", values_to = "value") %>%
    dplyr::mutate(
      metric = factor(metrics[match(metric_key, metric_map)], levels = metrics),
      value  = as.numeric(value),
      x_idx  = as.integer(metric)
    ) %>%
    dplyr::select(-metric_key)

  # Strict checks 
  ## Strict range check: finite & in [0,1]
  bad <- radar_long_ix %>% dplyr::filter(!is.na(value) & (!is.finite(value) | value < 0 | value > 1))
  if (nrow(bad) > 0) {
    summary_msg <- bad %>%
      dplyr::group_by(metric) %>%
      dplyr::summarise(n = dplyr::n(),
                       min_val = min(value, na.rm = TRUE),
                       max_val = max(value, na.rm = TRUE),
                       .groups = "drop") %>%
      dplyr::mutate(msg = sprintf("%s (n=%d, range=[%g, %g])", metric, n, min_val, max_val)) %>%
      dplyr::pull(msg) %>%
      paste(collapse = "; ")
    stop(sprintf("Out-of-bounds metric values (expected finite in [0,1]). Offenders: %s", summary_msg),
         call. = FALSE)
  }

  ## Uniqueness check: exactly 0/1 non-NA per point (facet × color × metric), throw if not 
  g_cols <- c(".facet", ".color", "metric", "x_idx")
  dup_check <- radar_long_ix %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(g_cols))) %>%
    dplyr::summarise(n_non_na = sum(!is.na(value)), .groups = "drop") %>%
    dplyr::filter(n_non_na > 1)

  if (nrow(dup_check) > 0) {
    offenders <- dup_check %>%
      dplyr::mutate(msg = sprintf("[facet=%s, color=%s, metric=%s, count=%d]",
                                  .facet, .color, as.character(metric), n_non_na)) %>%
      dplyr::pull(msg)
    stop(
      paste0(
        "Multiple observations per point detected. The input must be pre-summarized to one row ",
        "per (facet, color, metric). Offenders (showing up to first 10): ",
        paste(head(offenders, 10), collapse = "; "),
        if (nrow(dup_check) > 10) sprintf(" ... and %d more.", nrow(dup_check) - 10) else ""
      ),
      call. = FALSE
    )
  }

  # Keep at most one non-NA row per point (ok after uniqueness check)
  radar_long_ix <- radar_long_ix %>%
    dplyr::filter(!is.na(value)) %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(g_cols))) %>%
    dplyr::slice(1) %>%
    dplyr::ungroup()

  # Close each line: append first vertex at x = k+1
  close_lines <- function(d) {
    d <- dplyr::arrange(d, x_idx)
    if (nrow(d) == 0) return(d)
    first <- d[1, ]; first$x_idx <- k + 1
    dplyr::bind_rows(d, first)
  }

  group_vars <- c(".facet", ".color")
  radar_closed_ix <- radar_long_ix %>%
    dplyr::group_by(dplyr::across(dplyr::all_of(group_vars))) %>%
    dplyr::group_modify(~ close_lines(.x)) %>%
    dplyr::ungroup()

  # Background bands
  x_seq <- c(seq_len(k), k + 1)
  bands <- tibble::tibble(
    band_id = factor(seq_len(length(ring_breaks) - 1)),
    ymin = head(ring_breaks, -1),
    ymax = tail(ring_breaks, -1)
  ) %>% tidyr::crossing(tibble::tibble(x = x_seq))

  # Build plot
  p <- ggplot() +
    geom_ribbon(
      data = bands,
      aes(x = x, ymin = ymin, ymax = ymax, group = band_id),
      inherit.aes = FALSE, alpha = 0.08, fill = "grey50", color = NA
    ) +
    {
      if (has_color) {
        list(
          geom_line(
            data = dplyr::arrange(radar_closed_ix, .facet, .color, x_idx),
            aes(x = x_idx, y = value, color = .color, group = .color),
            linewidth = 0.8, show.legend = TRUE
          ),
          geom_point(
            data = radar_long_ix,
            aes(x = x_idx, y = value, color = .color, group = .color),
            size = 2, show.legend = TRUE
          )
        )
      } else {
        list(
          geom_line(
            data = dplyr::arrange(radar_closed_ix, .facet, x_idx),
            aes(x = x_idx, y = value, group = 1),
            linewidth = 0.9, color = "black", show.legend = FALSE
          ),
          geom_point(
            data = radar_long_ix,
            aes(x = x_idx, y = value),
            size = 2, color = "black", show.legend = FALSE
          )
        )
      }
    } +
    coord_polar(clip = "off") +
    scale_x_continuous(limits = c(1, k + 1), breaks = seq_len(k), labels = metrics) +
    scale_y_continuous(limits = c(0, 1 + outer_pad),
                                breaks = ring_breaks,
                                labels = scales::percent_format(accuracy = 1),
                                expand = expansion(mult = c(0, 0))) +
    labs(title = title, x = NULL, y = NULL) +
    theme_minimal(base_size = 12) +
    theme(
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(linewidth = 0.3),
      axis.text.x = element_text(size = 11, face = "bold"),
      axis.text.y = element_text(size = 9),
      legend.position = if (has_color) "bottom" else "none",
      strip.text = element_text(face = "bold"),
      plot.margin = margin(10, 30, 10, 30)
    )

  # Facet if requested
  if (has_facet) p <- p + facet_wrap(~ .facet)

  # Palette / legend title
  if (has_color && !is.null(palette)) {
    p <- p + scale_color_manual(values = palette, name = color_col)
  } else if (has_color) {
    p <- p + labs(color = color_col)
  }

  return(p)
}

```

```{r}
# custom hex plot function for GGally
my_hex <- function(data, mapping, ...) {
  ggplot(data, mapping) +
    geom_hex(aes(fill = after_stat(count)), ...) +  # map fill here
    scale_fill_viridis_c(trans = "log10", name = "count") +  # log scale
    theme_minimal() +
    geom_smooth() + 
    # geom_smooth(method = 'gam', formula = 'y ~ s(x, bs = "cs")') +
    geom_abline(intercept = 0,
                slope = 1,
                color = "red")
}

# custom correlation and CCC 
my_corr_ccc <- function(data, mapping,
                        method = "pearson",
                        digits = 3,
                        ccc_conf = 0.95,
                        size = 4, ...) {
  # extract columns from mapping
  x <- GGally::eval_data_col(data, mapping$x)
  y <- GGally::eval_data_col(data, mapping$y)

  ok <- is.finite(x) & is.finite(y)
  x <- x[ok]; y <- y[ok]

  label <- "Not enough data"
  if (length(x) >= 2 && stats::sd(x) > 0 && stats::sd(y) > 0) {
    r <- suppressWarnings(stats::cor(x, y, method = method, use = "complete.obs"))
    ccc <- try(DescTools::CCC(x, y, ci = "z-transform", conf.level = ccc_conf), silent = TRUE)

    if (inherits(ccc, "try-error")) {
      label <- sprintf("cor = %s\nCCC = NA", signif(r, digits))
    } else {
      label <- paste0(
        method, " cor = ", signif(r, digits), "\n",
        "CCC = ",
        paste0(
          signif(ccc$rho.c$est, digits), " (95% CI =",
          signif(ccc$rho.c$lwr.ci, digits), "–",
          signif(ccc$rho.c$upr.ci, digits), ")"
        )
      )
    }
  }
  
  # draw centered text in the panel
  ggplot(data.frame(x = 0.5, y = 0.5, label = label), aes(x, y, label = label)) +
    geom_text(size = size, lineheight = 1.05, ...) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE) +
    theme_void()
}

```

```{r}
#' Add columns of mean of observed and expected and the observed minus the expected 
#'
#' @param df the dataframe to add to, needs an observation column and an expected column 
#' @param expected_col character string of the expected column
#' @param observed_col character string of the observed column
#'
#' @returns the input with the difference from expected and mean of expected and observed
#' @export
add_diff_mean <- function(df, expected_col, observed_col) {
  expected_col <- enquo(expected_col)
  observed_col <- enquo(observed_col)

  obs_name <- as_name(observed_col)

  df %>%
    rowwise() %>%
    mutate(
      !!paste0(obs_name, "_diff_from_exp") := !!observed_col - !!expected_col,
      !!paste0(obs_name, "_mean_with_exp") := mean(c(!!observed_col, !!expected_col))
    ) %>%
    ungroup()
}
```

```{r}
#' Create a Bland–Altman style plot from observed vs expected COI measures
#'
#' This function generates a hex-binned Bland–Altman plot for a given COI measure
#' compared against an expected COI. It automatically looks up the mean and diff
#' columns based on a provided label (e.g., `"ecoi"`, `"coi"`, `"naive_coi"`, 
#' `"offset_naive_coi"`).
#'
#' @param df A data frame containing the mean and diff columns.
#' @param label A character string specifying the observed measure name. 
#'   The function will look for columns named `paste0(label, mean_suffix)` 
#'   and `paste0(label, diff_suffix)`.
#' @param mean_suffix A character string appended to the label to form the
#'   mean column name. Default is `"_mean_with_exp"`.
#' @param diff_suffix A character string appended to the label to form the
#'   diff column name. Default is `"_diff_from_exp"`.
#' @param expected_label A character string used in axis labels to describe
#'   the expected measure. Default is `"Expected COI"`.
#' @param bins Integer, number of bins for the hex plot. Default is `30`.
#'
#' @returns A `ggplot` object representing the Bland–Altman style plot.
#' @export
#'
#' @examples
#' \dontrun{
#' # Example with ecoi
#' p_ecoi <- make_bland_altman_plot_by_label(moire_all_coi_summary, "ecoi")
#' 
#' # Example with naive_coi
#' p_naive <- make_bland_altman_plot_by_label(moire_all_coi_summary, "naive_coi")
#' 
#' # Example with custom suffixes (if column names differ)
#' p_custom <- make_bland_altman_plot_by_label(
#'   df = moire_all_coi_summary,
#'   label = "ecoi",
#'   mean_suffix = "_mean",
#'   diff_suffix = "_diff"
#' )
#' }
make_bland_altman_plot_by_label <- function(
  df,
  label,
  mean_suffix = "_mean_with_exp",
  diff_suffix = "_diff_from_exp",
  expected_label = "Expected",
  bins = 30 
) {
  stopifnot(is.data.frame(df), is.character(label), length(label) == 1)

  mean_col <- paste0(label, mean_suffix)
  diff_col <- paste0(label, diff_suffix)

  if (!(mean_col %in% names(df))) {
    stop(sprintf("Column '%s' not found in df.", mean_col))
  }
  if (!(diff_col %in% names(df))) {
    stop(sprintf("Column '%s' not found in df.", diff_col))
  }

  # limits of agreement (±2 SD) computed once
  sd_val <- stats::sd(df[[diff_col]], na.rm = TRUE)

  ggplot(df) +
    geom_hex(aes(x = .data[[mean_col]], y = .data[[diff_col]], fill = log10(after_stat(count))),
             bins = bins) +
    scale_fill_viridis_c(name = "log10(count)") +
    theme_minimal() +
    geom_smooth(aes(x = .data[[mean_col]], y = .data[[diff_col]]),
                method = "gam", se = FALSE) +
    geom_hline(yintercept =  2 * sd_val) +
    geom_hline(yintercept = -2 * sd_val) +
    labs(
      title = label,
      y = paste0(label, " - ", expected_label),
      x = paste0("mean (", label, ", ", expected_label, ")")
    ) +
    coord_equal()
}


```

```{r}
#' @title Concordance / Error Metrics (single or multiple observed columns)
#' @keywords CCC RMSE MAE COI
#'
#' @description
#' Compute metrics between an expected column and one or more observed columns.
#' Each function accepts an optional set of grouping columns. If omitted, an
#' overall (single-row) summary is returned.
#'
#' @param df A data frame.
#' @param expected_col Unquoted column name of expected values.
#' @param observed_cols One symbol or \code{c(sym1, sym2, ...)} of observed columns.
#' @param group_cols Optional grouping columns supplied like
#'   \code{c(population_name, paramset_id)}. If omitted or \code{NULL},
#'   computes overall metrics (no grouping).
#'
#' @return A tibble with group columns (if any) and metric columns per observed variable.
#'
#' @importFrom dplyr group_by summarise distinct left_join bind_cols
#' @importFrom rlang enquo enexpr is_call call_args as_quosure caller_env as_name
#' @importFrom purrr map map_dbl
#' @name vectorized_metrics


# ---- CCC (DescTools::CCC) ----------------------------------------------------

#' @rdname vectorized_metrics
#' @export
compute_ccc <- function(df, expected_col, observed_cols, group_cols = NULL) {
  stopifnot(is.data.frame(df))
  expected_col <- rlang::enquo(expected_col)

  # capture observed columns (symbol or c(...))
  obs_expr <- rlang::enexpr(observed_cols)
  obs_quos <- if (rlang::is_call(obs_expr, "c")) {
    purrr::map(rlang::call_args(obs_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(obs_expr, env = rlang::caller_env()))
  }

  # capture grouping columns (NULL/omitted => no groups)
  grp_expr <- rlang::enexpr(group_cols)
  grp_quos <- if (missing(group_cols) || is.null(grp_expr)) {
    list()
  } else if (rlang::is_call(grp_expr, "c")) {
    purrr::map(rlang::call_args(grp_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(grp_expr, env = rlang::caller_env()))
  }

  if (length(grp_quos) > 0) {
    out <- df %>% dplyr::distinct(!!!grp_quos)
    join_by_cols <- names(out)
    for (q in obs_quos) {
      obs_name <- rlang::as_name(q)
      one <- df %>%
        dplyr::group_by(!!!grp_quos) %>%
        dplyr::summarise(.ccc = list(DescTools::CCC(!!expected_col, !!q)), .groups = "drop") %>%
        dplyr::mutate(
          !!paste0(obs_name, "_ccc")       := purrr::map_dbl(.ccc, ~ .x$rho.c$est),
          !!paste0(obs_name, "_ccc_lower") := purrr::map_dbl(.ccc, ~ .x$rho.c$lwr.ci),
          !!paste0(obs_name, "_ccc_upper") := purrr::map_dbl(.ccc, ~ .x$rho.c$upr.ci)
        ) %>%
        dplyr::select(-.ccc)
      out <- dplyr::left_join(out, one, by = join_by_cols)
    }
  } else {
    pieces <- purrr::map(obs_quos, function(q) {
      obs_name <- rlang::as_name(q)
      tmp <- df %>% dplyr::summarise(.ccc = list(DescTools::CCC(!!expected_col, !!q)))
      tibble::tibble(
        !!paste0(obs_name, "_ccc")       := purrr::map_dbl(tmp$.ccc, ~ .x$rho.c$est),
        !!paste0(obs_name, "_ccc_lower") := purrr::map_dbl(tmp$.ccc, ~ .x$rho.c$lwr.ci),
        !!paste0(obs_name, "_ccc_upper") := purrr::map_dbl(tmp$.ccc, ~ .x$rho.c$upr.ci)
      )
    })
    out <- dplyr::bind_cols(pieces)
  }

  out
}

# ---- RMSE --------------------------------------------------------------------

#' @rdname vectorized_metrics
#' @export
compute_rmse <- function(df, expected_col, observed_cols, group_cols = NULL) {
  stopifnot(is.data.frame(df))
  expected_col <- rlang::enquo(expected_col)

  obs_expr <- rlang::enexpr(observed_cols)
  obs_quos <- if (rlang::is_call(obs_expr, "c")) {
    purrr::map(rlang::call_args(obs_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(obs_expr, env = rlang::caller_env()))
  }

  grp_expr <- rlang::enexpr(group_cols)
  grp_quos <- if (missing(group_cols) || is.null(grp_expr)) {
    list()
  } else if (rlang::is_call(grp_expr, "c")) {
    purrr::map(rlang::call_args(grp_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(grp_expr, env = rlang::caller_env()))
  }

  metric_fun <- function(q) {
    obs_name <- rlang::as_name(q)
    if (length(grp_quos) > 0) {
      df %>%
        dplyr::group_by(!!!grp_quos) %>%
        dplyr::summarise(
          !!paste0(obs_name, "_rmse") := {
            x <- !!q; y <- !!expected_col
            ok <- !is.na(x) & !is.na(y)
            n  <- sum(ok)
            if (n == 0) NA_real_ else sqrt(sum((x[ok] - y[ok])^2) / n)
          },
          .groups = "drop"
        )
    } else {
      dplyr::summarise(df,
        !!paste0(obs_name, "_rmse") := {
          x <- !!q; y <- !!expected_col
          ok <- !is.na(x) & !is.na(y)
          n  <- sum(ok)
          if (n == 0) NA_real_ else sqrt(sum((x[ok] - y[ok])^2) / n)
        }
      )
    }
  }

  if (length(grp_quos) > 0) {
    out <- df %>% dplyr::distinct(!!!grp_quos)
    join_by_cols <- names(out)
    for (q in obs_quos) {
      out <- dplyr::left_join(out, metric_fun(q), by = join_by_cols)
    }
  } else {
    out <- dplyr::bind_cols(purrr::map(obs_quos, metric_fun))
  }

  out
}

# ---- MAE ---------------------------------------------------------------------

#' @rdname vectorized_metrics
#' @export
compute_mae <- function(df, expected_col, observed_cols, group_cols = NULL) {
  stopifnot(is.data.frame(df))
  expected_col <- rlang::enquo(expected_col)

  obs_expr <- rlang::enexpr(observed_cols)
  obs_quos <- if (rlang::is_call(obs_expr, "c")) {
    purrr::map(rlang::call_args(obs_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(obs_expr, env = rlang::caller_env()))
  }

  grp_expr <- rlang::enexpr(group_cols)
  grp_quos <- if (missing(group_cols) || is.null(grp_expr)) {
    list()
  } else if (rlang::is_call(grp_expr, "c")) {
    purrr::map(rlang::call_args(grp_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(grp_expr, env = rlang::caller_env()))
  }

  metric_fun <- function(q) {
    obs_name <- rlang::as_name(q)
    if (length(grp_quos) > 0) {
      df %>%
        dplyr::group_by(!!!grp_quos) %>%
        dplyr::summarise(
          !!paste0(obs_name, "_mean_abs_error") := {
            x <- !!q; y <- !!expected_col
            ok <- !is.na(x) & !is.na(y)
            if (!any(ok)) NA_real_ else mean(abs(x[ok] - y[ok]))
          },
          .groups = "drop"
        )
    } else {
      dplyr::summarise(df,
        !!paste0(obs_name, "_mean_abs_error") := {
          x <- !!q; y <- !!expected_col
          ok <- !is.na(x) & !is.na(y)
          if (!any(ok)) NA_real_ else mean(abs(x[ok] - y[ok]))
        }
      )
    }
  }

  if (length(grp_quos) > 0) {
    out <- df %>% dplyr::distinct(!!!grp_quos)
    join_by_cols <- names(out)
    for (q in obs_quos) {
      out <- dplyr::left_join(out, metric_fun(q), by = join_by_cols)
    }
  } else {
    out <- dplyr::bind_cols(purrr::map(obs_quos, metric_fun))
  }

  out
}

```


## Load in the Benchmarking Outputs

```{r}
inp <- list(
  sim_info_dir = "~/Downloads/hackathon_2025_07_data/simulations/SpotMalaria", 
  tool_res_dir = "~/Downloads/hackathon_2025_07_data/tool_results/MOIRE/", 
  madhatter_res_dir = "~/Downloads/hackathon_2025_07_data/tool_results/madhatter/",
  tool_params = "~/Downloads/hackathon_2025_07_data/tool_results/moire.json"
)
```


First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.

### Reading in expected truths 

Read per in expected allele frequencies per sample tables to get true COIs

```{r}
expected_simulated_frequencies_per_sample <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "trimmed_expected_simulated_frequencies_per_sample.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(library_sample_name = libary_sample_name)

expected_simulated_frequencies_per_sample_COIs = expected_simulated_frequencies_per_sample %>% 
  select(panel, population_name, country, library_sample_name, COI) %>% 
  distinct() %>% 
  dplyr::rename(expected_COI = COI) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Read per in expected allele frequencies tables population level 
```{r}
population_trimmed_mhap_freqs_prevs <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_trimmed_mhap_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(),
        population_name = col_character(), 
        pop = col_character(), 
        country = col_character(), 
        target = col_character(), 
        seq = col_character()
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(target_id = target)
```


Read in all simulation parameters into one tibble. Encode different population parameters as descriptive factors. 

```{r}

sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>% 
  mutate(
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>% 
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  )) %>% 
  mutate(pop_cat = factor(pop_cat, levels = sort(unique((.$pop_cat)))), 
         coi_cat = factor(coi_cat, levels = sort(unique((.$coi_cat)))), 
         between_sample_related_cat = factor(between_sample_related_cat, levels = sort(unique((.$between_sample_related_cat)))), 
         nonrecombinant_cat = factor(nonrecombinant_cat, levels = sort(unique((.$nonrecombinant_cat)))), 
         meiosis_cat = factor(meiosis_cat, levels = sort(unique((.$meiosis_cat)))))

create_dt(sim_params)
```

Read in madhatter pipeline allele tables 

```{r}
madhatter_allele_data <- tibble(
    allele_data_path = list.files(
      inp$madhatter_res_dir, 
      pattern = "allele_data.txt.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    allele_data = map(
      allele_data_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  mutate(tool_res_dir = dirname(allele_data_path)) %>%
  mutate(population_name = gsub("_mh_results", "", basename(tool_res_dir))) %>%
  select(-allele_data_path, -tool_res_dir) %>%
  unnest(allele_data) %>% 
  mutate(SampleID = gsub("_S[0-9]+$", "", SampleID)) %>% 
  dplyr::rename(library_sample_name = SampleID, 
                target_id = Locus, 
                seq = ASV, 
                reads = Reads) 
```

### Read in MOIRE results

MOIRE outputs information on complexity of infection (COI) and on population allele frequencies and several different outputs need to be read in. 

Getting the allele frequencies data in the `allele_freq_summary.tsv.gz` files 

```{r}
moire_allele_freq_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "allele_freq_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 
```

Getting the complexity of infection (COI) from the `coi_summary.tsv.gz` files 

```{r}
moire_coi_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "^coi_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id)) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Getting the effective COI from the `effective_coi_summary.tsv.gz` files. 

```{r}
moire_effective_coi_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "effective_coi_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id))%>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Getting the expected heterozygosity (he) estimates `he_summary.tsv.gz` files 

```{r}
moire_he_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "he_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res)
```

Getting the within host relatedness from the `relatedness_summary.tsv.gz` files 

```{r}
moire_relatedness_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "relatedness_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id))%>% 
  dplyr::rename(library_sample_name = specimen_id)

```

Getting what tool parameters were used. For MOIRE benchmarking we set the priors on the assumed amount of relatedness with moire_paramset1 assuming uniform relatedness and then each moire_paramset 2 and 3 increasing the amount of assumed relatedness. moire_paramsets3-6 are the same prior relatedness as 1-3 but have more parallel burin to explore more pace for convergence.   

```{r}
# Read in tool parameters
tool_params <- jsonlite::fromJSON(inp$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    r_alpha = paramset.r_alpha,
    r_beta = paramset.r_beta
  )

create_dt(tool_params)
```

## Benchmarking 

### COI  

#### Metrics 

*  Plotting Bland-Altman plots (plotting the mean of expected and obs vs difference) 
*  Calculating the RMSE (taking the square root of the mean squared difference between expected and observed)  
*  Calculating [Lin's concordance correlation coefficient](https://www.rdocumentation.org/packages/DescTools/versions/0.99.60/topics/CCC) for agreement on a continuous measure 
*  Calculating contingency table of determining a sample as monoclonal vs polyclonal 

#### Results 

MOIRE reports several COIs: 

*  naive_coi - The max alleles called for any locus for a sample 
*  offset_naive_coi - The 2nd max alleles called for any locus for a sample 
*  coi - COI reported by MOIRE 
*  ecoi - ecoi is the MOIRE COI adjusted for the amount of within sample relatedness 

```{r}
# combine the COI summaries and ecoi summaries and combine with the expected outputs for COI
moire_all_coi_summary = moire_coi_summary %>% 
  left_join(moire_effective_coi_summary) %>% 
  left_join(expected_simulated_frequencies_per_sample_COIs %>% 
              mutate(library_sample_name = as.character(library_sample_name))) %>% 
                                                        left_join(sim_params)

# calculate the means and differences for Bland-Altman plots

moire_all_coi_summary = moire_all_coi_summary %>% 
  add_diff_mean(expected_COI, naive_coi) %>% 
  add_diff_mean(expected_COI, offset_naive_coi)%>% 
  add_diff_mean(expected_COI, coi)%>% 
  add_diff_mean(expected_COI, ecoi)

```

##### Bland-Altman COI plots 

Overall 



```{r}
moire_all_coi_summary_overall_ecoi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "ecoi", expected_label = "Expected COI")
moire_all_coi_summary_overall_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "coi", expected_label = "Expected COI")
moire_all_coi_summary_overall_naive_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "naive_coi", expected_label = "Expected COI")
moire_all_coi_summary_overall_offset_naive_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "offset_naive_coi", expected_label = "Expected COI")

```


```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive COI" = moire_all_coi_summary_overall_naive_coi_bland_altman_plot, 
    "offset naive COI" = moire_all_coi_summary_overall_offset_naive_coi_bland_altman_plot, 
    "COI" = moire_all_coi_summary_overall_coi_bland_altman_plot, 
    "eCOI" = moire_all_coi_summary_overall_ecoi_bland_altman_plot
       )
))
```


Showing the different parameter sets 

```{r}
moire_all_coi_summary_ecoi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "ecoi") + 
  facet_wrap(~paramset_id, ncol = 3)
moire_all_coi_summary_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "coi") + 
  facet_wrap(~paramset_id, ncol = 3)
moire_all_coi_summary_offset_naive_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "offset_naive_coi") + 
  facet_wrap(~paramset_id, ncol = 3) 
moire_all_coi_summary_naive_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "naive_coi") + 
  facet_wrap(~paramset_id, ncol = 3)

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive COI" = moire_all_coi_summary_naive_coi_bland_altman_plot, 
    "offset naive COI" = moire_all_coi_summary_offset_naive_coi_bland_altman_plot, 
    "COI" = moire_all_coi_summary_coi_bland_altman_plot, 
    "eCOI" = moire_all_coi_summary_ecoi_bland_altman_plot
       )
))
```

Write to pdf  
```{r}
#| results: hide
cairo_pdf("moire_all_cois_bland_altman_plot.pdf", width = 10, height = 10)
print(moire_all_coi_summary_ecoi_bland_altman_plot)
print(moire_all_coi_summary_coi_bland_altman_plot)
print(moire_all_coi_summary_offset_naive_coi_bland_altman_plot)
print(moire_all_coi_summary_naive_coi_bland_altman_plot)
dev.off()
```

Showing the different population characteristics 

```{r}
moire_all_coi_summary_ecoi_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "ecoi") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)

moire_all_coi_summary_coi_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "coi") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)

moire_all_coi_summary_naive_coi_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "naive_coi") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)

moire_all_coi_summary_offset_naive_coi_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "offset_naive_coi") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)
```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive COI" = moire_all_coi_summary_naive_coi_by_pop_bland_altman_plot, 
    "offset naive COI" = moire_all_coi_summary_offset_naive_coi_by_pop_bland_altman_plot, 
    "COI" = moire_all_coi_summary_coi_by_pop_bland_altman_plot, 
    "eCOI" = moire_all_coi_summary_ecoi_by_pop_bland_altman_plot
       )
))
```

Write to pdf  
```{r}
#| results: hide
cairo_pdf("moire_all_cois_by_pop_bland_altman_plot.pdf", width = 10, height = 10)
print(moire_all_coi_summary_ecoi_by_pop_bland_altman_plot)
print(moire_all_coi_summary_coi_by_pop_bland_altman_plot)
print(moire_all_coi_summary_offset_naive_coi_by_pop_bland_altman_plot)
print(moire_all_coi_summary_naive_coi_by_pop_bland_altman_plot)
dev.off()
```


##### Concordance 

Plotting concordance between all COI metrics reported by MOIRE.

Overall 

```{r}
moire_all_coi_summary_coi_measures_plot_moire_all = ggpairs(moire_all_coi_summary,
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "all")

```

```{r}
print(moire_all_coi_summary_coi_measures_plot_moire_all)
```


Plotting difference parameter sets  

```{r}
moire_all_coi_summary_coi_measures_plot_moire_paramset1 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset1"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset1")

moire_all_coi_summary_coi_measures_plot_moire_paramset2 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset2"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset2")

moire_all_coi_summary_coi_measures_plot_moire_paramset3 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset3"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset3")

moire_all_coi_summary_coi_measures_plot_moire_paramset4 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset4"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset4")

moire_all_coi_summary_coi_measures_plot_moire_paramset5 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset5"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset5")

moire_all_coi_summary_coi_measures_plot_moire_paramset6 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset6"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset6")

```

```{r}
#| results: hide
cairo_pdf("moire_coi_pairwise_comparison_parametersets.pdf", width = 15, height = 15)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset1)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset4)

print(moire_all_coi_summary_coi_measures_plot_moire_paramset2)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset5)

print(moire_all_coi_summary_coi_measures_plot_moire_paramset3)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset6)
dev.off()
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("paramset1" = moire_all_coi_summary_coi_measures_plot_moire_paramset1,
                                      "paramset4" = moire_all_coi_summary_coi_measures_plot_moire_paramset4,
                                      "paramset2" = moire_all_coi_summary_coi_measures_plot_moire_paramset2,
                                      "paramset5" = moire_all_coi_summary_coi_measures_plot_moire_paramset5,
                                      "paramset3" = moire_all_coi_summary_coi_measures_plot_moire_paramset3, 
                                      "paramset6" = moire_all_coi_summary_coi_measures_plot_moire_paramset6 
                                      )))
```

Different population metrics 


COI categories 

```{r}
moire_all_coi_summary_coi_measures_plot_moire_MeanCOI1.26 = ggpairs(moire_all_coi_summary %>%
                                                    filter(coi_cat == "MeanCOI=1.26"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=1.26")

moire_all_coi_summary_coi_measures_plot_moire_MeanCOI3.41 = ggpairs(moire_all_coi_summary %>%
                                                    filter(coi_cat == "MeanCOI=3.41"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=3.41")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("MeanCOI_1.26" = moire_all_coi_summary_coi_measures_plot_moire_MeanCOI1.26,
                                      "MeanCOI_3.41" = moire_all_coi_summary_coi_measures_plot_moire_MeanCOI3.41
                                      )))
```


between sample related categories 

```{r}
moire_all_coi_summary_coi_measures_plot_moire_1perc_related = ggpairs(moire_all_coi_summary %>%
                                                    filter(between_sample_related_cat == "_1%_related"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "_1%_related")

moire_all_coi_summary_coi_measures_plot_moire_10perc_related = ggpairs(moire_all_coi_summary %>%
                                                    filter(between_sample_related_cat == "10%_related"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "10%_related")

moire_all_coi_summary_coi_measures_plot_moire_30perc_related = ggpairs(moire_all_coi_summary %>%
                                                    filter(between_sample_related_cat == "30%_related"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "30%_related")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("_1perc_between_sample_related" = moire_all_coi_summary_coi_measures_plot_moire_1perc_related,
                                      "10perc_between_sample_related" = moire_all_coi_summary_coi_measures_plot_moire_10perc_related, 
                                      "30perc_between_sample_related" = moire_all_coi_summary_coi_measures_plot_moire_30perc_related
                                      )))
```

Recombination categories 
```{r}
moire_all_coi_summary_coi_measures_plot_moire_50perc_non_recomb = ggpairs(moire_all_coi_summary %>%
                                                    filter(nonrecombinant_cat == "50%_non-recombinant"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "50%_non-recombinant")

moire_all_coi_summary_coi_measures_plot_moire_20perc_non_recomb = ggpairs(moire_all_coi_summary %>%
                                                    filter(nonrecombinant_cat == "20%_non-recombinant"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "20%_non-recombinant")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("20perc_non_recomb" = moire_all_coi_summary_coi_measures_plot_moire_20perc_non_recomb,
                                      "50perc_non_recomb" = moire_all_coi_summary_coi_measures_plot_moire_50perc_non_recomb
                                      )))
```


```{r}
#| results: hide
cairo_pdf("moire_coi_pairwise_comparison_pop_parameters.pdf", width = 15, height = 15)
print(moire_all_coi_summary_coi_measures_plot_moire_MeanCOI1.26)
print(moire_all_coi_summary_coi_measures_plot_moire_MeanCOI3.41)

print(moire_all_coi_summary_coi_measures_plot_moire_1perc_related)
print(moire_all_coi_summary_coi_measures_plot_moire_10perc_related)
print(moire_all_coi_summary_coi_measures_plot_moire_30perc_related)

print(moire_all_coi_summary_coi_measures_plot_moire_20perc_non_recomb)
print(moire_all_coi_summary_coi_measures_plot_moire_50perc_non_recomb)
dev.off()
```




```{r}
# calculate Concordance Correlation Coefficient (CCC) 
moire_all_coi_summary_ccc_by_pop <- moire_all_coi_summary %>%
  compute_ccc(
    expected_col  = expected_COI,
    observed_cols = c(coi, ecoi, naive_coi, offset_naive_coi),
    group_cols    = c(population_name, paramset_id)
  )

# Gather into long format, left join in the simulation parameters
moire_all_coi_summary_ccc_by_pop_gat = moire_all_coi_summary_ccc_by_pop %>%
  select(population_name, paramset_id, ends_with("_ccc")) %>% 
  gather(coi_estimate, ccc, 3:ncol(.)) %>%
  mutate(coi_estimate = gsub("_ccc", "", coi_estimate))%>%
  left_join(sim_params)
```




ccc across all parameter sets for all 4 estimators 

```{r}
moire_all_coi_summary_ccc_plot_just_estimators = ggplot(moire_all_coi_summary_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(y = ccc, x = coi_estimate, fill = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

ggplotly(moire_all_coi_summary_ccc_plot_just_estimators)
```

ccc over all comparing parameter sets for each estimate

```{r}
moire_all_coi_summary_ccc_plot_fill_paramset_id = ggplot(moire_all_coi_summary_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_id, y = ccc, x = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_plot_fill_coi_estimate = ggplot(moire_all_coi_summary_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = coi_estimate, y = ccc, x = paramset_id), position = "dodge"
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfHtmlWidgets(
  list(
  "By Parameter Set" = ggplotly(moire_all_coi_summary_ccc_plot_fill_paramset_id) %>% layout(boxmode = "group"), 
  "By COI Estimator" = ggplotly(moire_all_coi_summary_ccc_plot_fill_coi_estimate) %>% layout(boxmode = "group")
  )
))

```

ccc on the different simulation population characteristics 

```{r}
moire_all_coi_summary_ccc_by_pop_gat_large_plot_by_country = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat_by_country = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = moire_all_coi_summary_ccc_by_pop_gat_large_plot,
  "est_between_sample_related" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat,
  "coi_cat" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat,
  "nonrecombinant" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country
)))
```

##### RMSE   

```{r}
# calculate RMSE 
moire_all_coi_summary_rmse <- moire_all_coi_summary %>%
  compute_rmse(
    expected_col  = expected_COI,
    observed_cols = c(coi, ecoi, naive_coi, offset_naive_coi),
    group_cols    = c(population_name, paramset_id)
  )

# Gather into long format, left join in the simulation parameters
moire_all_coi_summary_rmse_gat = moire_all_coi_summary_rmse %>%
  gather(coi_estimate, rmse, 3:ncol(.)) %>%
  mutate(coi_estimate = gsub("_rmse", "", coi_estimate))%>%
  left_join(sim_params)

```

RMSE across all parameter sets for all 4 estimators 

```{r}
moire_all_coi_summary_rmse_plot_just_estimators = ggplot(moire_all_coi_summary_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(y = rmse, x = coi_estimate, fill = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

ggplotly(moire_all_coi_summary_rmse_plot_just_estimators)
```

RMSE over all comparing parameter sets for each estimate

```{r}
moire_all_coi_summary_rmse_plot_fill_paramset_id = ggplot(moire_all_coi_summary_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_id, y = rmse, x = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

moire_all_coi_summary_rmse_plot_fill_coi_estimate = ggplot(moire_all_coi_summary_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = coi_estimate, y = rmse, x = paramset_id), position = "dodge"
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfHtmlWidgets(
  list(
  "By Parameter Set" = ggplotly(moire_all_coi_summary_rmse_plot_fill_paramset_id) %>% layout(boxmode = "group"), 
  "By COI Estimator" = ggplotly(moire_all_coi_summary_rmse_plot_fill_coi_estimate) %>% layout(boxmode = "group")
  )
))

```

RMSE on the different simulation population characteristics 

```{r}
moire_all_coi_summary_rmse_gat_large_plot_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_all_coi_summary_rmse_gat_large_plot_coi_cat_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_all_coi_summary_rmse_gat_large_plot = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate

moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate

moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate

moire_all_coi_summary_rmse_gat_large_plot_coi_cat = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = moire_all_coi_summary_rmse_gat_large_plot,
  "est_between_sample_related" = moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat,
  "coi_cat" = moire_all_coi_summary_rmse_gat_large_plot_coi_cat,
  "nonrecombinant" = moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = moire_all_coi_summary_rmse_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = moire_all_coi_summary_rmse_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat_by_country
)))
```

<!-- ###### Testing for significant affects of population  -->

<!-- ####### COI  -->

<!-- ```{r} -->
<!-- library(finalfit) -->
<!-- explanatory = c("coi_cat", "nonrecombinant_cat", "between_sample_related_cat", "country", "paramset_id") -->
<!-- dependent = 'coi_rmse' -->
<!-- moire_all_coi_summary_rmse %>% -->
<!--   left_join(sim_params) %>% -->
<!--   mutate(coi_cat = factor(coi_cat), -->
<!--          nonrecombinant_cat = factor(nonrecombinant_cat), -->
<!--          between_sample_related_cat = factor(between_sample_related_cat), -->
<!--          country = factor(country), -->
<!--          paramset_id = factor(paramset_id)) %>% -->
<!--   finalfit("coi_rmse", explanatory, metrics=TRUE) -> t2 -->
<!-- knitr::kable(t2[[1]], row.names=FALSE, align=c("l", "l", "r", "r", "r", "r")) -->
<!-- knitr::kable(t2[[2]], row.names=FALSE, col.names="") -->

<!-- ``` -->

<!-- ```{r, echo = T, eval=T} -->
<!-- #| fig-width: 15 -->
<!-- moire_all_coi_summary_rmse %>% -->
<!--   left_join(sim_params) %>% -->
<!--   mutate(coi_cat = factor(coi_cat), -->
<!--          nonrecombinant_cat = factor(nonrecombinant_cat), -->
<!--          between_sample_related_cat = factor(between_sample_related_cat), -->
<!--          country = factor(country), -->
<!--          paramset_id = factor(paramset_id))  %>% -->
<!--   # filter(coi_rmse < 1) %>% -->
<!--   coefficient_plot("coi_rmse", -->
<!--           # "country" -->
<!--           explanatory -->
<!--           ) -->

<!-- ``` -->

<!-- ####### eCOI  -->
<!-- ```{r} -->
<!-- library(finalfit) -->
<!-- explanatory = c("coi_cat", "nonrecombinant_cat", "between_sample_related_cat", "country", "paramset_id") -->
<!-- dependent = 'ecoi_rmse' -->
<!-- moire_all_coi_summary_rmse %>% -->
<!--   left_join(sim_params) %>% -->
<!--   mutate(coi_cat = factor(coi_cat), -->
<!--          nonrecombinant_cat = factor(nonrecombinant_cat), -->
<!--          between_sample_related_cat = factor(between_sample_related_cat), -->
<!--          country = factor(country), -->
<!--          paramset_id = factor(paramset_id)) %>% -->
<!--   finalfit("ecoi_rmse", explanatory, metrics=TRUE) -> t2 -->
<!-- knitr::kable(t2[[1]], row.names=FALSE, align=c("l", "l", "r", "r", "r", "r")) -->
<!-- knitr::kable(t2[[2]], row.names=FALSE, col.names="") -->

<!-- ``` -->

<!-- ```{r, echo = T, eval=T} -->
<!-- #| fig-width: 15 -->
<!-- moire_all_coi_summary_rmse %>% -->
<!--   left_join(sim_params) %>% -->
<!--   mutate(coi_cat = factor(coi_cat), -->
<!--          nonrecombinant_cat = factor(nonrecombinant_cat), -->
<!--          between_sample_related_cat = factor(between_sample_related_cat), -->
<!--          country = factor(country), -->
<!--          paramset_id = factor(paramset_id))  %>% -->
<!--   # filter(ecoi_rmse < 1) %>% -->
<!--   coefficient_plot("ecoi_rmse", -->
<!--           # "country" -->
<!--           explanatory -->
<!--           ) -->

<!-- ``` -->


#### Monoclonal vs polyclonal determination 



```{r}
# Round eCOI and COI for testing 

moire_all_coi_summary_mono_vs_poly = moire_all_coi_summary %>%
  mutate(ecoi  = round(ecoi), 
         coi = round(coi)) %>% 
  select(population_name, paramset_id, library_sample_name,expected_COI, coi, ecoi, naive_coi, offset_naive_coi) %>% 
  gather(coi_estimator, determined_coi, 5:ncol(.)) %>%
  mutate(
    test_classification = case_when(
      determined_coi == 1  & expected_COI == 1 ~ "TruePositive",
      determined_coi == 1  & expected_COI != 1 ~ "FalsePositive",
      determined_coi != 1  & expected_COI == 1 ~ "FalseNegative",
      determined_coi != 1  & expected_COI != 1 ~ "TrueNegative"
    )
  ) %>%
  left_join(sim_params) 

mono_vs_poly_by_parameterset <- moire_all_coi_summary_mono_vs_poly %>%
  group_by(coi_estimator, paramset_id) %>%
  summarise(
    TP = sum(determined_coi == 1  & expected_COI == 1, na.rm = TRUE),
    FP = sum(determined_coi == 1  & expected_COI != 1, na.rm = TRUE),
    FN = sum(determined_coi != 1  & expected_COI == 1, na.rm = TRUE),
    TN = sum(determined_coi != 1  & expected_COI != 1, na.rm = TRUE),
  ) %>% 
  add_contigency_metrics()

```

Performance on detecting monoclonal for each coi estimator across the different parameter sets 

```{r}
#| fig-width: 20
#| fig-height: 20
metric_map <- c(
  "Sensitivity"        = "sensitivity",
  "Specificity"        = "specificity",
  "Precision (PPV)"    = "ppv",
  "NPV"                = "npv",
  "F1-score"           = "f1_score",
  "Balanced Accuracy"  = "balanced_accuracy"
)

mono_vs_poly_by_parameterset_radar = radar_by_metrics(
  mono_vs_poly_by_parameterset,
  facet_col = "paramset_id",
  color_col = "coi_estimator",
  metric_map = metric_map,
  palette = colorPalette_08
)

print(mono_vs_poly_by_parameterset_radar)
```


### Allele frequencies

Since the allele calling pipeline might have baseline inaccuracies, calculating naive estimates and false positive / false negative rates for the pipeline data itself as a bunch mark can be a helpful comparator. 

#### Naive allele frequencies 

```{r}
madhatter_allele_data_naive_allele_freq_totals = madhatter_allele_data %>%
  group_by(population_name, target_id) %>%
  summarise(allele_total = n(),
            sample_total = n_distinct(library_sample_name))

madhatter_allele_data_naive_allele_freq_prev = madhatter_allele_data %>%
  group_by(population_name, target_id, seq) %>%
  summarise(
    allele_count = n(),
    sample_count = n_distinct(library_sample_name)
  ) %>%
  left_join(madhatter_allele_data_naive_allele_freq_totals) %>%
  mutate(allele_freq  = allele_count/allele_total,
         sample_prev = sample_count/sample_total) %>%
  mutate(population_name_target_id_seq = paste0(population_name, "-", target_id, "-", seq))


madhatter_allele_data_naive_allele_freq_prev_out = madhatter_allele_data_naive_allele_freq_prev %>% 
  select(population_name, target_id, seq, allele_freq, sample_prev) %>% 
  dplyr::rename(naive_freq = allele_freq, 
                naive_prev = sample_prev)


population_trimmed_mhap_freqs_prevs_sel = population_trimmed_mhap_freqs_prevs %>% 
  select(population_name, target_id, seq, freq, prev) %>% 
  dplyr::rename(expected_freq = freq, 
                expected_prev = prev) %>% 
  full_join(madhatter_allele_data_naive_allele_freq_prev_out)

# fill in zeros for missing or false alleles 
population_trimmed_mhap_freqs_prevs_sel = population_trimmed_mhap_freqs_prevs_sel %>% 
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq), 
         expected_prev = ifelse(is.na(expected_prev), 0, expected_prev), 
         naive_freq = ifelse(is.na(naive_freq), 0, naive_freq), 
         naive_prev = ifelse(is.na(naive_prev), 0, naive_prev)
         )

# moire doesn't calculate prevalence, so don't include 
population_trimmed_mhap_freqs_prevs_sel_out = population_trimmed_mhap_freqs_prevs_sel %>% 
  select(-naive_prev, -expected_prev)

```

#### Adding expected and naive estimates to MOIRE 

```{r}

moire_param_map <- moire_allele_freq_summary %>%
  distinct(population_name, target_id, paramset_id)

population_trimmed_mhap_freqs_prevs_sel_out_expanded <- population_trimmed_mhap_freqs_prevs_sel_out %>%
  inner_join(moire_param_map, by = c("population_name", "target_id"))

moire_allele_freq_summary_sel = moire_allele_freq_summary %>%
  select(population_name, paramset_id, target_id, seq, freq) %>%
  dplyr::rename(moire_freq = freq) %>%
  full_join(population_trimmed_mhap_freqs_prevs_sel_out_expanded) %>%
  mutate(
    moire_freq = ifelse(is.na(moire_freq), 0, moire_freq),
    expected_freq = ifelse(is.na(expected_freq), 0, expected_freq),
    naive_freq = ifelse(is.na(naive_freq), 0, naive_freq)
  ) %>% 
  left_join(sim_params)

```

#### Metrics 

*  Plotting Bland-Altman plots (plotting the mean of expected and obs vs difference) 
*  Calculating the RMSE (taking the square root of the mean squared difference between expected and observed)  
*  Calculating [Lin's concordance correlation coefficient](https://www.rdocumentation.org/packages/DescTools/versions/0.99.60/topics/CCC) for agreement on a continuous measure 
*  Calculating contingency table metrics for false positive and false negatives  



##### Bland-Altman Plots 


```{r}
moire_allele_freq_summary_sel = moire_allele_freq_summary_sel %>% 
    add_diff_mean(expected_col = expected_freq,
                  observed_col = moire_freq)%>% 
    add_diff_mean(expected_col = expected_freq,
                  observed_col = naive_freq)

```



Overall 

```{r}
moire_allele_freq_summary_sel_overall_moire_freq_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "moire_freq")
moire_allele_freq_summary_sel_overall_naive_freq_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "naive_freq")

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive_freq" = moire_allele_freq_summary_sel_overall_naive_freq_bland_altman_plot, 
    "moire_freq" = moire_allele_freq_summary_sel_overall_moire_freq_bland_altman_plot
       )
))
```


Showing the different parameter sets 

```{r}
moire_allele_freq_summary_sel_moire_freq_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "moire_freq") + 
  facet_wrap(~paramset_id, ncol = 3)


moire_allele_freq_summary_sel_naive_freq_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "naive_freq") + 
  facet_wrap(~paramset_id, ncol = 3) 

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive_freq" = moire_allele_freq_summary_sel_naive_freq_bland_altman_plot, 
    "moire_freq" = moire_allele_freq_summary_sel_moire_freq_bland_altman_plot
       )
))
```

Write to pdf  
```{r}
#| results: hide
cairo_pdf("moire_all_allele_freqs_bland_altman_plot.pdf", width = 10, height = 10)
print(moire_allele_freq_summary_sel_moire_freq_bland_altman_plot)
print(moire_allele_freq_summary_sel_naive_freq_bland_altman_plot)
dev.off()
```

Showing the different population characteristics 

```{r}
moire_allele_freq_summary_sel_moire_freq_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "moire_freq") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)

moire_allele_freq_summary_sel_naive_freq_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "naive_freq") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)
```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive_freq" = moire_allele_freq_summary_sel_naive_freq_by_pop_bland_altman_plot, 
    "moire_freq" = moire_allele_freq_summary_sel_moire_freq_by_pop_bland_altman_plot
       )
))
```

Write to pdf  
```{r}
#| results: hide
cairo_pdf("moire_all_allele_freqs_by_pop_bland_altman_plot.pdf", width = 10, height = 10)
print(moire_allele_freq_summary_sel_moire_freq_by_pop_bland_altman_plot)
print(moire_allele_freq_summary_sel_naive_freq_by_pop_bland_altman_plot)
dev.off()
```



##### Concordance 

Plotting concordance between allele frequencies metrics reported by MOIRE.

Overall 

```{r}
moire_all_summary_allele_freq_measures_plot_moire_all = ggpairs(moire_allele_freq_summary_sel,
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "all")

```

```{r}
print(moire_all_summary_allele_freq_measures_plot_moire_all)
```


Plotting difference parameter sets  

```{r}
moire_all_summary_allele_freq_measures_plot_moire_paramset1 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset1"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset1")

moire_all_summary_allele_freq_measures_plot_moire_paramset2 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset2"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset2")

moire_all_summary_allele_freq_measures_plot_moire_paramset3 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset3"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset3")

moire_all_summary_allele_freq_measures_plot_moire_paramset4 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset4"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset4")

moire_all_summary_allele_freq_measures_plot_moire_paramset5 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset5"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset5")

moire_all_summary_allele_freq_measures_plot_moire_paramset6 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset6"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset6")

```

```{r}
#| results: hide
cairo_pdf("moire_allele_freq_pairwise_comparison_parametersets.pdf", width = 15, height = 15)
print(moire_all_summary_allele_freq_measures_plot_moire_paramset1)
print(moire_all_summary_allele_freq_measures_plot_moire_paramset4)

print(moire_all_summary_allele_freq_measures_plot_moire_paramset2)
print(moire_all_summary_allele_freq_measures_plot_moire_paramset5)

print(moire_all_summary_allele_freq_measures_plot_moire_paramset3)
print(moire_all_summary_allele_freq_measures_plot_moire_paramset6)
dev.off()
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("paramset1" = moire_all_summary_allele_freq_measures_plot_moire_paramset1,
                                      "paramset4" = moire_all_summary_allele_freq_measures_plot_moire_paramset4,
                                      "paramset2" = moire_all_summary_allele_freq_measures_plot_moire_paramset2,
                                      "paramset5" = moire_all_summary_allele_freq_measures_plot_moire_paramset5,
                                      "paramset3" = moire_all_summary_allele_freq_measures_plot_moire_paramset3, 
                                      "paramset6" = moire_all_summary_allele_freq_measures_plot_moire_paramset6 
                                      )))
```

Different population metrics 


COI categories 

```{r}
moire_all_summary_allele_freq_measures_plot_moire_MeanCOI1.26 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(coi_cat == "MeanCOI=1.26"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=1.26")

moire_all_summary_allele_freq_measures_plot_moire_MeanCOI3.41 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(coi_cat == "MeanCOI=3.41"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=3.41")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("MeanCOI_1.26" = moire_all_summary_allele_freq_measures_plot_moire_MeanCOI1.26,
                                      "MeanCOI_3.41" = moire_all_summary_allele_freq_measures_plot_moire_MeanCOI3.41
                                      )))
```


between sample related categories 

```{r}
moire_all_summary_allele_freq_measures_plot_moire_1perc_related = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(between_sample_related_cat == "_1%_related"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "_1%_related")

moire_all_summary_allele_freq_measures_plot_moire_10perc_related = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(between_sample_related_cat == "10%_related"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "10%_related")

moire_all_summary_allele_freq_measures_plot_moire_30perc_related = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(between_sample_related_cat == "30%_related"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "30%_related")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("_1perc_between_sample_related" = moire_all_summary_allele_freq_measures_plot_moire_1perc_related,
                                      "10perc_between_sample_related" = moire_all_summary_allele_freq_measures_plot_moire_10perc_related, 
                                      "30perc_between_sample_related" = moire_all_summary_allele_freq_measures_plot_moire_30perc_related
                                      )))
```

Recombination categories 
```{r}
moire_all_summary_allele_freq_measures_plot_moire_50perc_non_recomb = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(nonrecombinant_cat == "50%_non-recombinant"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "50%_non-recombinant")

moire_all_summary_allele_freq_measures_plot_moire_20perc_non_recomb = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(nonrecombinant_cat == "20%_non-recombinant"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "20%_non-recombinant")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("20perc_non_recomb" = moire_all_summary_allele_freq_measures_plot_moire_20perc_non_recomb,
                                      "50perc_non_recomb" = moire_all_summary_allele_freq_measures_plot_moire_50perc_non_recomb
                                      )))
```


```{r}
#| results: hide
cairo_pdf("moire_allele_freq_pairwise_comparison_pop_parameters.pdf", width = 15, height = 15)
print(moire_all_summary_allele_freq_measures_plot_moire_MeanCOI1.26)
print(moire_all_summary_allele_freq_measures_plot_moire_MeanCOI3.41)

print(moire_all_summary_allele_freq_measures_plot_moire_1perc_related)
print(moire_all_summary_allele_freq_measures_plot_moire_10perc_related)
print(moire_all_summary_allele_freq_measures_plot_moire_30perc_related)

print(moire_all_summary_allele_freq_measures_plot_moire_20perc_non_recomb)
print(moire_all_summary_allele_freq_measures_plot_moire_50perc_non_recomb)
dev.off()
```



```{r}
# calculate Concordance Correlation Coefficient (CCC) 
moire_allele_freq_summary_sel_ccc_by_pop = moire_allele_freq_summary_sel %>%
  compute_ccc(
    expected_col  = expected_freq,
    observed_cols = c(moire_freq, naive_freq),
    group_cols    = c(population_name, paramset_id)
  )
  
# Gather into long format, left join in the simulation parameters
moire_allele_freq_summary_sel_ccc_by_pop_gat = moire_allele_freq_summary_sel_ccc_by_pop %>%
  select(population_name, paramset_id, ends_with("_ccc")) %>% 
  gather(allele_freq_estimate, ccc, 3:ncol(.)) %>%
  mutate(allele_freq_estimate = gsub("_ccc", "", allele_freq_estimate))%>%
  left_join(sim_params)

```

ccc across all parameter sets for all 4 estimators 

```{r}
moire_allele_freq_summary_sel_ccc_plot_just_estimators = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(y = ccc, x = allele_freq_estimate, fill = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

ggplotly(moire_allele_freq_summary_sel_ccc_plot_just_estimators)
```

ccc over all comparing parameter sets for each estimate

```{r}
moire_allele_freq_summary_sel_ccc_plot_fill_paramset_id = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_id, y = ccc, x = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_plot_fill_allele_freq_estimate = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = allele_freq_estimate, y = ccc, x = paramset_id), position = "dodge"
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfHtmlWidgets(
  list(
  "By Parameter Set" = ggplotly(moire_allele_freq_summary_sel_ccc_plot_fill_paramset_id) %>% layout(boxmode = "group"), 
  "By allele freq Estimator" = ggplotly(moire_allele_freq_summary_sel_ccc_plot_fill_allele_freq_estimate) %>% layout(boxmode = "group")
  )
))

```

ccc on the different simulation population characteristics 

```{r}
moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_by_country = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_coi_cat_by_country = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_between_sample_related_cat = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_nonrecombinant_cat = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_coi_cat = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot,
  "est_between_sample_related" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_between_sample_related_cat,
  "coi_cat" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_coi_cat,
  "nonrecombinant" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country
)))
```


##### RMSE   

```{r}
# calculate RMSE 
moire_allele_freq_summary_sel_rmse = moire_allele_freq_summary_sel %>%
  compute_rmse(
    expected_col  = expected_freq,
    observed_cols = c(moire_freq, naive_freq),
    group_cols    = c(population_name, paramset_id)
  )

# Gather into long format, left join in the simulation parameters
moire_allele_freq_summary_sel_rmse_gat = moire_allele_freq_summary_sel_rmse %>%
  gather(allele_freq_estimate, rmse, 3:ncol(.)) %>%
  mutate(allele_freq_estimate = gsub("_rmse", "", allele_freq_estimate))%>%
  left_join(sim_params)

```

RMSE across all parameter sets for all 4 estimators 

```{r}
moire_allele_freq_summary_sel_rmse_plot_just_estimators = ggplot(moire_allele_freq_summary_sel_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(y = rmse, x = allele_freq_estimate, fill = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

ggplotly(moire_allele_freq_summary_sel_rmse_plot_just_estimators)
```

RMSE over all comparing parameter sets for each estimate

```{r}
moire_allele_freq_summary_sel_rmse_plot_fill_paramset_id = ggplot(moire_allele_freq_summary_sel_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_id, y = rmse, x = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

moire_allele_freq_summary_sel_rmse_plot_fill_allele_freq_estimate = ggplot(moire_allele_freq_summary_sel_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = allele_freq_estimate, y = rmse, x = paramset_id), position = "dodge"
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfHtmlWidgets(
  list(
  "By Parameter Set" = ggplotly(moire_allele_freq_summary_sel_rmse_plot_fill_paramset_id) %>% layout(boxmode = "group"), 
  "By COI Estimator" = ggplotly(moire_allele_freq_summary_sel_rmse_plot_fill_allele_freq_estimate) %>% layout(boxmode = "group")
  )
))

```

RMSE on the different simulation population characteristics 

```{r}
moire_allele_freq_summary_sel_rmse_gat_large_plot_by_country = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_allele_freq_summary_sel_rmse_gat_large_plot_between_sample_related_cat_by_country = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_allele_freq_summary_sel_rmse_gat_large_plot_nonrecombinant_cat_by_country = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_allele_freq_summary_sel_rmse_gat_large_plot_coi_cat_by_country = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_allele_freq_summary_sel_rmse_gat_large_plot = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate

moire_allele_freq_summary_sel_rmse_gat_large_plot_between_sample_related_cat = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate

moire_allele_freq_summary_sel_rmse_gat_large_plot_nonrecombinant_cat = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate

moire_allele_freq_summary_sel_rmse_gat_large_plot_coi_cat = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = moire_allele_freq_summary_sel_rmse_gat_large_plot,
  "est_between_sample_related" = moire_allele_freq_summary_sel_rmse_gat_large_plot_between_sample_related_cat,
  "coi_cat" = moire_allele_freq_summary_sel_rmse_gat_large_plot_coi_cat,
  "nonrecombinant" = moire_allele_freq_summary_sel_rmse_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = moire_allele_freq_summary_sel_rmse_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = moire_allele_freq_summary_sel_rmse_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = moire_allele_freq_summary_sel_rmse_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = moire_allele_freq_summary_sel_rmse_gat_large_plot_nonrecombinant_cat_by_country
)))
```



##### False positive and false negatives

```{r}
moire_allele_freq_summary_sel_contignency = moire_allele_freq_summary_sel %>% 
  select(population_name, paramset_id, target_id, seq, expected_freq, moire_freq, naive_freq) %>% 
  gather(estimator, freq, 6:7) %>% 
  mutate(estimator = gsub("_freq", "", estimator))%>% 
  mutate(estimator = gsub("naive", "mad4hatter", estimator))

# moire appears to have some allele freq estimates that are extremely low even for some expected frequencies close to 1000, to count these 
moire_allele_freq_summary_sel_contignency = moire_allele_freq_summary_sel_contignency %>% 
  mutate(freq = ifelse(estimator == "moire" & expected_freq/freq > 1000 & expected_freq > 0.05, 0, freq))

# True Positive is both estimator and expected is great than 0 
# False Negative is that expected > 0 and estimator is 0 
# False Positive is that expected is 0 and estimator is > 0
# (could also consider like above if the frequenices are extremely off from each other, than make it as False Positive or false negative)

moire_allele_freq_summary_sel_contignency = moire_allele_freq_summary_sel_contignency %>% 
  mutate(TP = as.numeric(expected_freq > 0 & freq > 0), 
         FN = as.numeric(expected_freq > 0 & freq == 0), 
         FP = as.numeric(expected_freq == 0 & freq > 0))

```


Has the same Haplotype performance as input data, no deviations from the data that it works with. 

```{r}
moire_allele_freq_summary_sel_contignency_overall = moire_allele_freq_summary_sel_contignency %>% 
  group_by(estimator) %>%
  add_contigency_metrics_no_TN(summarize_first = T)

create_dt(moire_allele_freq_summary_sel_contignency_overall)

hap_metric_map <- c(
  "Recall"               = "recall",
  "Precision"            = "precision",
  "NPV"                  = "tpr",
  "F1-score"             = "f1_score",
  "False Discovery Rate" = "fdr",
  "Overlap Coefficient"  = "overlap_coefficient"
)

radar_by_metrics(moire_allele_freq_summary_sel_contignency_overall,  metric_map = hap_metric_map, color_col = "estimator", palette = colorPalette_08)
```


### he

MOIRE also reports on he for each target 

```{r}
moire_allele_freq_summary_sel_hes = moire_allele_freq_summary_sel %>% 
  group_by(population_name, paramset_id, target_id) %>% 
  summarise(
    expected_he = 1 - sum(expected_freq^2), 
    naive_he    = 1 - sum(naive_freq^2), 
    moire_recalc_he = 1 - sum(moire_freq^2)
  ) %>% 
  full_join(moire_he_summary %>% 
              select(population_name, paramset_id, target_id, he) %>% 
              dplyr::rename(moire_he = he)) %>% 
  left_join(sim_params)

```



##### Concordance 

Plotting concordance between all he reported. The moire_recalc_he is the allele frequencies re-calculated (1 - sum(moire_freq^2))

Overall 

```{r}
moire_all_summary_allele_freq_measures_plot_moire_all = ggpairs(moire_allele_freq_summary_sel_hes,
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "all")

```

```{r}
print(moire_all_summary_allele_freq_measures_plot_moire_all)
```


Plotting difference parameter sets  

```{r}
moire_all_summary_he_measures_plot_moire_paramset1 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset1"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset1")

moire_all_summary_he_measures_plot_moire_paramset2 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset2"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset2")

moire_all_summary_he_measures_plot_moire_paramset3 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset3"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset3")

moire_all_summary_he_measures_plot_moire_paramset4 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset4"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset4")

moire_all_summary_he_measures_plot_moire_paramset5 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset5"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset5")

moire_all_summary_he_measures_plot_moire_paramset6 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset6"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset6")

```

```{r}
#| results: hide
cairo_pdf("moire_he_pairwise_comparison_parametersets.pdf", width = 15, height = 15)
print(moire_all_summary_he_measures_plot_moire_paramset1)
print(moire_all_summary_he_measures_plot_moire_paramset4)

print(moire_all_summary_he_measures_plot_moire_paramset2)
print(moire_all_summary_he_measures_plot_moire_paramset5)

print(moire_all_summary_he_measures_plot_moire_paramset3)
print(moire_all_summary_he_measures_plot_moire_paramset6)
dev.off()
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("paramset1" = moire_all_summary_he_measures_plot_moire_paramset1,
                                      "paramset4" = moire_all_summary_he_measures_plot_moire_paramset4,
                                      "paramset2" = moire_all_summary_he_measures_plot_moire_paramset2,
                                      "paramset5" = moire_all_summary_he_measures_plot_moire_paramset5,
                                      "paramset3" = moire_all_summary_he_measures_plot_moire_paramset3, 
                                      "paramset6" = moire_all_summary_he_measures_plot_moire_paramset6 
                                      )))
```

Different population metrics 


COI categories 

```{r}
moire_all_summary_he_measures_plot_moire_MeanCOI1.26 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(coi_cat == "MeanCOI=1.26"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=1.26")

moire_all_summary_he_measures_plot_moire_MeanCOI3.41 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(coi_cat == "MeanCOI=3.41"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=3.41")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("MeanCOI_1.26" = moire_all_summary_he_measures_plot_moire_MeanCOI1.26,
                                      "MeanCOI_3.41" = moire_all_summary_he_measures_plot_moire_MeanCOI3.41
                                      )))
```


between sample related categories 

```{r}
moire_all_summary_he_measures_plot_moire_1perc_related = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(between_sample_related_cat == "_1%_related"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "_1%_related")

moire_all_summary_he_measures_plot_moire_10perc_related = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(between_sample_related_cat == "10%_related"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "10%_related")

moire_all_summary_he_measures_plot_moire_30perc_related = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(between_sample_related_cat == "30%_related"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "30%_related")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("_1perc_between_sample_related" = moire_all_summary_he_measures_plot_moire_1perc_related,
                                      "10perc_between_sample_related" = moire_all_summary_he_measures_plot_moire_10perc_related, 
                                      "30perc_between_sample_related" = moire_all_summary_he_measures_plot_moire_30perc_related
                                      )))
```

Recombination categories 
```{r}
moire_all_summary_he_measures_plot_moire_50perc_non_recomb = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(nonrecombinant_cat == "50%_non-recombinant"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "50%_non-recombinant")

moire_all_summary_he_measures_plot_moire_20perc_non_recomb = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(nonrecombinant_cat == "20%_non-recombinant"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "20%_non-recombinant")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("20perc_non_recomb" = moire_all_summary_he_measures_plot_moire_20perc_non_recomb,
                                      "50perc_non_recomb" = moire_all_summary_he_measures_plot_moire_50perc_non_recomb
                                      )))
```


```{r}
#| results: hide
cairo_pdf("moire_he_pairwise_comparison_pop_parameters.pdf", width = 15, height = 15)
print(moire_all_summary_he_measures_plot_moire_MeanCOI1.26)
print(moire_all_summary_he_measures_plot_moire_MeanCOI3.41)

print(moire_all_summary_he_measures_plot_moire_1perc_related)
print(moire_all_summary_he_measures_plot_moire_10perc_related)
print(moire_all_summary_he_measures_plot_moire_30perc_related)

print(moire_all_summary_he_measures_plot_moire_20perc_non_recomb)
print(moire_all_summary_he_measures_plot_moire_50perc_non_recomb)
dev.off()
```



```{r}
# calculate Concordance Correlation Coefficient (CCC) 
moire_he_summary_sel_hes_ccc_by_pop = moire_allele_freq_summary_sel_hes %>%
  compute_ccc(
    expected_col = expected_he, 
    observed_cols = c(moire_he, moire_recalc_he, naive_he), 
    group_cols = c(population_name, paramset_id)
  )
# Gather into long format, left join in the simulation parameters
moire_he_summary_sel_hes_ccc_by_pop_gat = moire_he_summary_sel_hes_ccc_by_pop %>%
  select(population_name, paramset_id, ends_with("_ccc")) %>% 
  gather(allele_freq_estimate, ccc, 3:ncol(.)) %>%
  mutate(allele_freq_estimate = gsub("_ccc", "", allele_freq_estimate))%>%
  left_join(sim_params)

```

ccc across all parameter sets for all 4 estimators 

```{r}
moire_allele_freq_summary_sel_hes_ccc_plot_just_estimators = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(y = ccc, x = allele_freq_estimate, fill = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

ggplotly(moire_allele_freq_summary_sel_hes_ccc_plot_just_estimators)
```

ccc over all comparing parameter sets for each estimate

```{r}
moire_allele_freq_summary_sel_hes_ccc_plot_fill_paramset_id = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_id, y = ccc, x = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_hes_ccc_plot_fill_allele_freq_estimate = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = allele_freq_estimate, y = ccc, x = paramset_id), position = "dodge"
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfHtmlWidgets(
  list(
  "By Parameter Set" = ggplotly(moire_allele_freq_summary_sel_hes_ccc_plot_fill_paramset_id) %>% layout(boxmode = "group"), 
  "By allele freq Estimator" = ggplotly(moire_allele_freq_summary_sel_hes_ccc_plot_fill_allele_freq_estimate) %>% layout(boxmode = "group")
  )
))

```

ccc on the different simulation population characteristics 

```{r}
moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_by_country = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_coi_cat_by_country = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_between_sample_related_cat = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_nonrecombinant_cat = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_coi_cat = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot,
  "est_between_sample_related" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_between_sample_related_cat,
  "coi_cat" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_coi_cat,
  "nonrecombinant" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country
)))
```





