---
title: moire Benchmarking Results
author: Nicholas Hathaway
output: html_document
toc-depth: 5
code-fold: true
fig-width: 10
fig-height: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)
```


## Set up 

Libraries used in the analysis below 

```{r, warning=FALSE, message=FALSE}
# data manipulation libraries 
library(dplyr)
library(readr)
library(stringr)
library(tibble)
library(tidyr) 
library(magrittr)
library(jsonlite)
library(purrr)

# plotting 
library(ggplot2)
library(GGally) # ggpairs for pairwise comparisons 

# interactive html outputs
library(plotly) # ggplotly to turn any ggplot object into an interactive plot 
library(DT) # data tables for displaying tables 

# stats package
library(DescTools) # CCC, Calculates Lin's concordance correlation coefficient
```

Some helpful R utilities 

```{r}
#| code-fold: true

# commonly used functions  
`%!in%` <- Negate(`%in%`)
is.notna <-function(x){
  return(!is.na(x))
}
# ggplot themes 
transparentBackground = theme(
  panel.background = element_rect(fill='transparent'), #transparent panel bg
  plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
  panel.grid.major = element_blank(), #remove major gridlines
  panel.grid.minor = element_blank(), #remove minor gridlines
  legend.background = element_rect(color = NA, fill='transparent'), #transparent legend bg
  legend.box.background = element_rect(color = NA, fill='transparent') #transparent legend panel
)
custom_theme_noTransparentBackground = theme_bw() +
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank() )+
  theme(axis.line.x = element_line(color="black", linewidth = 0.3),axis.line.y =
          element_line(color="black", linewidth = 0.3))+
  theme(text=element_text(size=12, family="Helvetica"))+
  theme(axis.text.y = element_text(size=12))+
  theme(axis.text.x = element_text(size=12)) +
  theme(legend.position = "bottom") + 
  theme(plot.title = element_text(hjust = 0.5))

custom_theme = custom_theme_noTransparentBackground +
  transparentBackground
custom_theme_backgroundTransparent = custom_theme

custom_theme_xRotate_noBackgroundTransparent = theme_bw() +
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank() )+
  theme(axis.line.x = element_line(color="black", linewidth = 0.3),axis.line.y =
          element_line(color="black", linewidth = 0.3))+
  theme(text=element_text(size=12, family="Helvetica"))+
  theme(axis.text.y = element_text(size=12))+
  theme(axis.text.x = element_text(size=12)) +
  theme(legend.position = "bottom") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.text.x = element_text(size=12, angle = -90, vjust = 0.5, hjust = 0)) 

custom_theme_xRotate = custom_theme_xRotate_noBackgroundTransparent + 
  transparentBackground

custom_theme_xRotate_backgroundTransparent = custom_theme_xRotate

# color blind safe distinct color palettes 
colorPalette_08 = c("#2271B2","#F748A5","#359B73","#F0E442","#D55E00","#3DB7E9","#E69F00","#000000")
colorPalette_12 = c("#E20134","#FF6E3A","#008DF9","#8400CD","#FFC33B","#9F0162","#009F81","#FF5AAF","#00FCCF","#00C2F9","#FFB2FD","#A40122")
colorPalette_15 = c("#F60239","#003C86","#EF0096","#9400E6","#009FFA","#008169","#68023F","#00DCB5","#FFCFE2","#FF71FD","#7CFFFA","#6A0213","#008607","#00E307","#FFDC3D")


# creating a datatable html widget
create_dt <- function(x){
  DT::datatable(x,
                extensions = 'Buttons',
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All"))), 
                filter = "top")
}

# for creating tabsets html outputs
create_tabsetOfHtmlWidgets <- function(htmlObjectsList) {
  zz <- textConnection("foo", "w")
  sink(zz)
  cat("::: {.panel-tabset}\n")
  iwalk(htmlObjectsList, ~ {
    cat('## ', .y, '\n\n')
    tempList = list()
    tempList[["item"]] = .x
    print(htmltools::tagList(tempList))
    cat('\n\n')
  })
  cat(":::\n")
  sink()
  close(zz)
  paste0(foo, collapse = "\n")
}

create_tabsetOfGgplotObjects <- function(ggplotObjectsList) {
  cat("\n\n::: {.panel-tabset}\n\n")
  iwalk(ggplotObjectsList, ~ {
    cat('## ', .y, '\n\n')
    print(.x)
    cat('\n\n')
  })
  cat(":::\n")
}

```


## Load in the Benchmarking Outputs

```{r}
inp <- list(
  sim_info_dir = "~/Downloads/hackathon_2025_07_data/simulations/SpotMalaria", 
  tool_res_dir = "~/Downloads/hackathon_2025_07_data/tool_results/MOIRE/", 
  madhatter_res_dir = "~/Downloads/hackathon_2025_07_data/tool_results/madhatter/",
  tool_params = "~/Downloads/hackathon_2025_07_data/tool_results/moire.json"
)
```


First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.

### Reading in expected truths 

Read per in expected allele frequencies per sample tables to get true COIs

```{r}
expected_simulated_frequencies_per_sample <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "trimmed_expected_simulated_frequencies_per_sample.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(library_sample_name = libary_sample_name)

expected_simulated_frequencies_per_sample_COIs = expected_simulated_frequencies_per_sample %>% 
  select(panel, population_name, country, library_sample_name, COI) %>% 
  distinct() %>% 
  dplyr::rename(expected_COI = COI) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Read per in expected allele frequencies tables population level 
```{r}
population_trimmed_mhap_freqs_prevs <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_trimmed_mhap_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(),
        population_name = col_character(), 
        pop = col_character(), 
        country = col_character(), 
        target = col_character(), 
        seq = col_character()
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth)
```


Read in all simulation parameters into one tibble

```{r}
sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>% 
  mutate(
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>% 
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  ))

create_dt(sim_params)
```

Read in madhatter pipeline allele tables 

```{r}
madhatter_allele_data <- tibble(
    allele_data_path = list.files(
      inp$madhatter_res_dir, 
      pattern = "allele_data.txt.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    allele_data = map(
      allele_data_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  mutate(tool_res_dir = dirname(allele_data_path)) %>%
  mutate(population_name = gsub("_mh_results", "", basename(tool_res_dir))) %>%
  select(-allele_data_path, -tool_res_dir) %>%
  unnest(allele_data) %>% 
  mutate(SampleID = gsub("_S[0-9]+$", "", SampleID)) %>% 
  dplyr::rename(library_sample_name = SampleID, 
                target_id = Locus, 
                seq = ASV, 
                reads = Reads) 
```

### Read in moire results

moire outputs information on complexity of infection (COI) and on population allele frequencies and several different outputs need to be read in. 

Getting the allele frequencies data in the `allele_freq_summary.tsv.gz` files 

```{r}
moire_allele_freq_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "allele_freq_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 
```

Getting the complexity of infection (COI) from the `coi_summary.tsv.gz` files 

```{r}
moire_coi_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "^coi_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id)) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Getting the effective COI from the `effective_coi_summary.tsv.gz` files. 

```{r}
moire_effective_coi_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "effective_coi_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id))%>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Getting the expected heterozygosity (he) estimates `he_summary.tsv.gz` files 

```{r}
moire_he_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "he_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res)
```

Getting the within host relatedness from the `relatedness_summary.tsv.gz` files 

```{r}
moire_relatedness_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "relatedness_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id))%>% 
  dplyr::rename(library_sample_name = specimen_id)

```

Getting what tool parameters were used. For moire benchmarking we set the priors on the assumed amount of relatedness with moire_paramset1 assuming uniform relatedness and then each moire_paramset 2 and 3 increasing the amount of assumed relatedness. moire_paramsets3-6 are the same prior relatedness as 1-3 but have more parallel burin to explore more pace for convergence.   

```{r}
# Read in tool parameters
tool_params <- jsonlite::fromJSON(inp$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    r_alpha = paramset.r_alpha,
    r_beta = paramset.r_beta
  )

create_dt(tool_params)
```


## Benchmarking 



### COI  

#### Metrics 

*  Plotting Bland-Altman plots (plotting the mean of expected and obs vs difference) 
*  Calculating the root mean square error (taking the squart root of the mean squared difference between epxected and observed)  
*  Calculating Lin's concordance correlation coefficient for agreement on a continuous measure 
*  Calculating contingency table of determining a sample as monoclonal vs polyclonal 

#### Results 

moire reports several COIs: 

*  naive_coi - The max alleles called for any locus for a sample 
*  offset_naive_coi - The 2nd max alleles called for any locus for a sample 
*  coi - COI reported by moire 
*  ecoi - ecoi is the COI adjusted for the amount of within sample relatedness 

```{r}
# combine the COI summaries and ecoi summaries and combine with the expected outputs for COI
moire_all_coi_summary = moire_coi_summary %>% 
  left_join(moire_effective_coi_summary) %>% 
  left_join(expected_simulated_frequencies_per_sample_COIs %>% 
              mutate(library_sample_name = as.character(library_sample_name))) %>% 
                                                        left_join(sim_params)

# calculate the means and differences for Bland-Altman plots 
moire_all_coi_summary = moire_all_coi_summary %>% 
  rowwise() %>% 
  mutate(naive_coi_diff = naive_coi - expected_COI, 
         offset_naive_coi_diff = offset_naive_coi - expected_COI, 
         coi_diff = coi - expected_COI, 
         ecoi_diff = ecoi - expected_COI, 
         
         naive_coi_mean = mean(naive_coi + expected_COI), 
         offset_naive_coi_mean = mean(offset_naive_coi + expected_COI), 
         coi_mean = mean(coi + expected_COI), 
         ecoi_mean = mean(ecoi + expected_COI)) %>% 
  ungroup() 
```

##### Bland-Altman COI plots 

Showing the different parameter sets 


```{r}
moire_all_coi_summary_ecoi_bland_altman_plot = ggplot(moire_all_coi_summary) + 
  geom_hex(
    aes(x = ecoi_mean, y = ecoi_diff, fill = log10(after_stat(count)))) +
  scale_fill_viridis_c(name = "log10(count)") +
  theme_minimal() + 
  geom_smooth(aes(x = ecoi_mean, y = ecoi_diff), method = "gam", se = F) + 
  facet_wrap(~paramset_id, ncol = 3)+ 
  geom_hline(aes(yintercept = 2 * sd(ecoi_diff)))  + 
  geom_hline(aes(yintercept = -2 * sd(ecoi_diff))) + 
  labs(title = "ecoi") + 
  coord_equal()

moire_all_coi_summary_coi_bland_altman_plot = ggplot(moire_all_coi_summary) + 
  geom_hex(
    aes(x = coi_mean, y = coi_diff, fill = log10(after_stat(count)))) +
  scale_fill_viridis_c(name = "log10(count)") +
  theme_minimal() + 
  geom_smooth(aes(x = coi_mean, y = coi_diff), method = "gam", se = F) + 
  facet_wrap(~paramset_id, ncol = 3)+ 
  geom_hline(aes(yintercept = 2 * sd(ecoi_diff)))  + 
  geom_hline(aes(yintercept = -2 * sd(ecoi_diff))) + 
  labs(title = "coi") + 
  coord_equal()

moire_all_coi_summary_offset_naive_coi_bland_altman_plot = ggplot(moire_all_coi_summary) + 
  geom_hex(
    aes(x = offset_naive_coi_mean, y = offset_naive_coi_diff, fill = log10(after_stat(count)))) +
  scale_fill_viridis_c(name = "log10(count)") +
  theme_minimal() + 
  coord_equal() + 
  geom_smooth(aes(x = offset_naive_coi_mean, y = offset_naive_coi_diff), method = "gam", se = F) + 
  geom_hline(aes(yintercept = 2 * sd(offset_naive_coi_diff)))  + 
  geom_hline(aes(yintercept = -2 * sd(offset_naive_coi_diff))) + 
  facet_wrap(~paramset_id, ncol = 3) + 
  labs(title = "offset_naive_coi")

moire_all_coi_summary_naive_coi_bland_altman_plot = ggplot(moire_all_coi_summary) + 
  geom_hex(
    aes(x = naive_coi_mean, y = naive_coi_diff, fill = log10(after_stat(count)))) +
  scale_fill_viridis_c(name = "log10(count)") +
  theme_minimal() + 
  coord_equal() + 
  geom_smooth(aes(x = naive_coi_mean, y = naive_coi_diff), method = "gam", se = F) + 
  geom_hline(aes(yintercept = 2 * sd(naive_coi_diff)))  + 
  geom_hline(aes(yintercept = -2 * sd(naive_coi_diff))) + 
  facet_wrap(~paramset_id, ncol = 3) + 
  labs(title = "naive_coi")

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive COI" = moire_all_coi_summary_naive_coi_bland_altman_plot, 
    "offset naive COI" = moire_all_coi_summary_offset_naive_coi_bland_altman_plot, 
    "COI" = moire_all_coi_summary_coi_bland_altman_plot, 
    "eCOI" = moire_all_coi_summary_ecoi_bland_altman_plot
       )
))
```

Write to pdf  
```{r}
#| results: hide
cairo_pdf("moire_all_cois_bland_altman_plot.pdf", width = 10, height = 10)
print(moire_all_coi_summary_ecoi_bland_altman_plot)
print(moire_all_coi_summary_coi_bland_altman_plot)
print(moire_all_coi_summary_offset_naive_coi_bland_altman_plot)
print(moire_all_coi_summary_naive_coi_bland_altman_plot)
dev.off()
```

Showing the different population characteristics 

```{r}
moire_all_coi_summary_ecoi_by_pop_bland_altman_plot = ggplot(moire_all_coi_summary) + 
  geom_hex(
    aes(x = ecoi_mean, y = ecoi_diff, fill = log10(after_stat(count)))) +
  scale_fill_viridis_c(name = "log10(count)") +
  theme_minimal() + 
  geom_smooth(aes(x = ecoi_mean, y = ecoi_diff), method = "gam", se = F) + 
  facet_grid(coi_cat~between_sample_related_cat)+ 
  geom_hline(aes(yintercept = 2 * sd(ecoi_diff)))  + 
  geom_hline(aes(yintercept = -2 * sd(ecoi_diff))) + 
  labs(title = "ecoi") + 
  coord_equal()

moire_all_coi_summary_coi_by_pop_bland_altman_plot = ggplot(moire_all_coi_summary) + 
  geom_hex(
    aes(x = coi_mean, y = coi_diff, fill = log10(after_stat(count)))) +
  scale_fill_viridis_c(name = "log10(count)") +
  theme_minimal() + 
  geom_smooth(aes(x = coi_mean, y = coi_diff), method = "gam", se = F) + 
  facet_grid(coi_cat~between_sample_related_cat)+ 
  geom_hline(aes(yintercept = 2 * sd(coi_diff)))  + 
  geom_hline(aes(yintercept = -2 * sd(coi_diff))) + 
  labs(title = "coi") + 
  coord_equal()

moire_all_coi_summary_naive_coi_by_pop_bland_altman_plot = ggplot(moire_all_coi_summary) + 
  geom_hex(
    aes(x = naive_coi_mean, y = naive_coi_diff, fill = log10(after_stat(count)))) +
  scale_fill_viridis_c(name = "log10(count)") +
  theme_minimal() + 
  geom_smooth(aes(x = naive_coi_mean, y = naive_coi_diff), method = "gam", se = F) + 
  facet_grid(coi_cat~between_sample_related_cat)+ 
  geom_hline(aes(yintercept = 2 * sd(naive_coi_diff)))  + 
  geom_hline(aes(yintercept = -2 * sd(naive_coi_diff))) + 
  labs(title = "naive_coi") + 
  coord_equal()

moire_all_coi_summary_offset_naive_coi_by_pop_bland_altman_plot = ggplot(moire_all_coi_summary) + 
  geom_hex(
    aes(x = offset_naive_coi_mean, y = offset_naive_coi_diff, fill = log10(after_stat(count)))) +
  scale_fill_viridis_c(name = "log10(count)") +
  theme_minimal() + 
  geom_smooth(aes(x = offset_naive_coi_mean, y = offset_naive_coi_diff), method = "gam", se = F) + 
  facet_grid(coi_cat~between_sample_related_cat)+ 
  geom_hline(aes(yintercept = 2 * sd(offset_naive_coi_diff)))  + 
  geom_hline(aes(yintercept = -2 * sd(offset_naive_coi_diff))) + 
  labs(title = "offset_naive_coi") + 
  coord_equal()
```


```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive COI" = moire_all_coi_summary_naive_coi_by_pop_bland_altman_plot, 
    "offset naive COI" = moire_all_coi_summary_offset_naive_coi_by_pop_bland_altman_plot, 
    "COI" = moire_all_coi_summary_coi_by_pop_bland_altman_plot, 
    "eCOI" = moire_all_coi_summary_ecoi_by_pop_bland_altman_plot
       )
))
```

Write to pdf  
```{r}
#| results: hide
cairo_pdf("moire_all_cois_by_pop_bland_altman_plot.pdf", width = 10, height = 10)
print(moire_all_coi_summary_ecoi_by_pop_bland_altman_plot)
print(moire_all_coi_summary_coi_by_pop_bland_altman_plot)
print(moire_all_coi_summary_offset_naive_coi_by_pop_bland_altman_plot)
print(moire_all_coi_summary_naive_coi_by_pop_bland_altman_plot)
dev.off()
```

##### RMSE   

```{r}

moire_all_coi_summary_rmse = moire_all_coi_summary %>%
  group_by(population_name, paramset_id) %>%
  summarise(
    coi_rmse = sqrt(sum(coi_diff^2)/length(coi_diff)),
    ecoi_rmse = sqrt(sum(ecoi_diff^2)/length(ecoi_diff)),
    naive_coi_rmse = sqrt(sum(naive_coi_diff^2)/length(naive_coi_diff)),
    offset_naive_coi_rmse = sqrt(sum(offset_naive_coi_diff^2)/length(offset_naive_coi_diff))
  )

moire_all_coi_summary_rmse_gat = moire_all_coi_summary_rmse %>%
  gather(coi_estimate, rmse, 3:ncol(.)) %>%
  mutate(coi_estimate = gsub("_rmse", "", coi_estimate))


moire_all_coi_summary_rmse_monoclonal = moire_all_coi_summary %>%
  filter(expected_COI == 1) %>%
  group_by(population_name, paramset_id) %>%
  summarise(
    coi_rmse = sqrt(sum(coi_diff^2)/length(coi_diff)),
    ecoi_rmse = sqrt(sum(ecoi_diff^2)/length(ecoi_diff)),
    naive_coi_rmse = sqrt(sum(naive_coi_diff^2)/length(naive_coi_diff)),
    offset_naive_coi_rmse = sqrt(sum(offset_naive_coi_diff^2)/length(offset_naive_coi_diff))
  )

moire_all_coi_summary_rmse_monoclonal_gat = moire_all_coi_summary_rmse_monoclonal %>%
  gather(coi_estimate, rmse, 3:ncol(.)) %>%
  mutate(coi_estimate = gsub("_rmse", "", coi_estimate))

```

RMSE across all parameter sets for all 4 estimators 

```{r}
moire_all_coi_summary_rmse_plot_just_estimators = ggplot(moire_all_coi_summary_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(y = rmse, x = coi_estimate, fill = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

ggplotly(moire_all_coi_summary_rmse_plot_just_estimators)
```


RMSE over all comparing parameter sets for each estimate

```{r}
moire_all_coi_summary_rmse_plot_fill_paramset_id = ggplot(moire_all_coi_summary_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_id, y = rmse, x = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

moire_all_coi_summary_rmse_plot_fill_coi_estimate = ggplot(moire_all_coi_summary_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = coi_estimate, y = rmse, x = paramset_id), position = "dodge"
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfHtmlWidgets(
  list(
  "By Parameter Set" = ggplotly(moire_all_coi_summary_rmse_plot_fill_paramset_id) %>% layout(boxmode = "group"), 
  "By COI Estimator" = ggplotly(moire_all_coi_summary_rmse_plot_fill_coi_estimate) %>% layout(boxmode = "group")
  )
))

```

Monoclonal RMSE

```{r}
moire_all_coi_summary_rmse_plot_fill_paramset_id_coi_estimate_monoclonal = ggplot(moire_all_coi_summary_rmse_monoclonal_gat) +
  geom_boxplot(
    aes(x = paramset_id, y = rmse, fill = coi_estimate, group = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)
```

```{r}
ggplotly(moire_all_coi_summary_rmse_plot_fill_paramset_id_coi_estimate_monoclonal) %>% layout(boxmode = "group")
```

```{r}

moire_all_coi_summary_rmse_gat = moire_all_coi_summary_rmse_gat %>%
  left_join(sim_params)

moire_all_coi_summary_rmse_gat_large_plot_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_theme_xRotate

cairo_pdf("moire_coi_summary_rmse_large_plot.pdf", width = 15, height = 10)
print(moire_all_coi_summary_rmse_gat_large_plot_by_country)
dev.off()

moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_theme_xRotate

moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_theme_xRotate

moire_all_coi_summary_rmse_gat_large_plot_coi_cat_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_theme_xRotate


moire_all_coi_summary_rmse_gat_large_plot = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_theme_xRotate

moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_theme_xRotate

moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_theme_xRotate

moire_all_coi_summary_rmse_gat_large_plot_coi_cat = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_theme_xRotate
```


### RMSE breakdown over different simulation parameters

```{r}
#| echo: false
#| results: asis


cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = moire_all_coi_summary_rmse_gat_large_plot,
  "est_between_sample_related" = moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat,
  "coi" = moire_all_coi_summary_rmse_gat_large_plot_coi_cat,
  "nonrecombinant" = moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = moire_all_coi_summary_rmse_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = moire_all_coi_summary_rmse_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat_by_country
)))
```

#### COI 

```{r}
library(finalfit)
explanatory = c("coi_cat", "nonrecombinant_cat", "between_sample_related_cat", "country", "paramset_id")
dependent = 'coi_rmse'
moire_all_coi_summary_rmse %>%
  left_join(sim_params) %>%
  mutate(coi_cat = factor(coi_cat),
         nonrecombinant_cat = factor(nonrecombinant_cat),
         between_sample_related_cat = factor(between_sample_related_cat),
         country = factor(country),
         paramset_id = factor(paramset_id)) %>%
  finalfit("coi_rmse", explanatory, metrics=TRUE) -> t2
knitr::kable(t2[[1]], row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
knitr::kable(t2[[2]], row.names=FALSE, col.names="")

```

```{r, echo = T, eval=T}
#| fig-width: 15
moire_all_coi_summary_rmse %>%
  left_join(sim_params) %>%
  mutate(coi_cat = factor(coi_cat),
         nonrecombinant_cat = factor(nonrecombinant_cat),
         between_sample_related_cat = factor(between_sample_related_cat),
         country = factor(country),
         paramset_id = factor(paramset_id))  %>%
  # filter(coi_rmse < 1) %>%
  coefficient_plot("coi_rmse",
          # "country"
          explanatory
          )

```


#### eCOI 
```{r}
library(finalfit)
explanatory = c("coi_cat", "nonrecombinant_cat", "between_sample_related_cat", "country", "paramset_id")
dependent = 'ecoi_rmse'
moire_all_coi_summary_rmse %>%
  left_join(sim_params) %>%
  mutate(coi_cat = factor(coi_cat),
         nonrecombinant_cat = factor(nonrecombinant_cat),
         between_sample_related_cat = factor(between_sample_related_cat),
         country = factor(country),
         paramset_id = factor(paramset_id)) %>%
  finalfit("ecoi_rmse", explanatory, metrics=TRUE) -> t2
knitr::kable(t2[[1]], row.names=FALSE, align=c("l", "l", "r", "r", "r", "r"))
knitr::kable(t2[[2]], row.names=FALSE, col.names="")

```

```{r, echo = T, eval=T}
#| fig-width: 15
moire_all_coi_summary_rmse %>%
  left_join(sim_params) %>%
  mutate(coi_cat = factor(coi_cat),
         nonrecombinant_cat = factor(nonrecombinant_cat),
         between_sample_related_cat = factor(between_sample_related_cat),
         country = factor(country),
         paramset_id = factor(paramset_id))  %>%
  # filter(ecoi_rmse < 1) %>%
  coefficient_plot("ecoi_rmse",
          # "country"
          explanatory
          )

```


```{r}
# custom hex plot function for GGally
my_hex <- function(data, mapping, ...) {
  ggplot(data, mapping) +
    geom_hex(aes(fill = after_stat(count)), ...) +  # map fill here
    scale_fill_viridis_c(trans = "log10", name = "count") +  # log scale
    theme_minimal() +
    geom_smooth(method = 'gam', formula = 'y ~ s(x, bs = "cs")') +
    geom_abline(intercept = 0,
                slope = 1,
                color = "red")
}

# custom correlation and CCC 
my_corr_ccc <- function(data, mapping,
                        method = "pearson",
                        digits = 3,
                        ccc_conf = 0.95,
                        size = 4, ...) {
  # extract columns from mapping
  x <- GGally::eval_data_col(data, mapping$x)
  y <- GGally::eval_data_col(data, mapping$y)

  ok <- is.finite(x) & is.finite(y)
  x <- x[ok]; y <- y[ok]

  label <- "Not enough data"
  if (length(x) >= 2 && stats::sd(x) > 0 && stats::sd(y) > 0) {
    r <- suppressWarnings(stats::cor(x, y, method = method, use = "complete.obs"))
    ccc <- try(DescTools::CCC(x, y, ci = "z-transform", conf.level = ccc_conf), silent = TRUE)

    if (inherits(ccc, "try-error")) {
      label <- sprintf("cor = %s\nCCC = NA", signif(r, digits))
    } else {
      label <- paste0(
        method, " cor = ", signif(r, digits), "\n",
        "CCC = ",
        paste0(
          signif(ccc$rho.c$est, digits), " (",
          signif(ccc$rho.c$lwr.ci, digits), "–",
          signif(ccc$rho.c$upr.ci, digits), ")"
        )
      )
    }
  }
    # draw centered text in the panel
  ggplot(data.frame(x = 0.5, y = 0.5, label = label), aes(x, y, label = label)) +
    geom_text(size = size, lineheight = 1.05, ...) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1), expand = FALSE) +
    theme_void()
}

moire_all_coi_summary_coi_measures_plot_moire_paramset1 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset1"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset1")
print(moire_all_coi_summary_coi_measures_plot_moire_paramset1)

moire_all_coi_summary_coi_measures_plot_moire_paramset2 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset2"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag = list(continuous = "barDiag"),
                                                  title = "moire_paramset2")

moire_all_coi_summary_coi_measures_plot_moire_paramset3 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset3"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag = list(continuous = "barDiag"),
                                                  title = "moire_paramset3")

moire_all_coi_summary_coi_measures_plot_moire_paramset4 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset4"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag = list(continuous = "barDiag"),
                                                  title = "moire_paramset4")

moire_all_coi_summary_coi_measures_plot_moire_paramset5 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset5"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag = list(continuous = "barDiag"),
                                                  title = "moire_paramset5")

moire_all_coi_summary_coi_measures_plot_moire_paramset6 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset6"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag = list(continuous = "barDiag"),
                                                  title = "moire_paramset6")

```

```{r}
#| results: hide
cairo_pdf("moire_coi_pairwise_comparison.pdf", width = 15, height = 15)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset1)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset4)

print(moire_all_coi_summary_coi_measures_plot_moire_paramset2)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset5)

print(moire_all_coi_summary_coi_measures_plot_moire_paramset3)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset6)
dev.off()
```


Plotting correlation between the different COI estimates compared to expected COI over each parameter set

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("paramset1" = moire_all_coi_summary_coi_measures_plot_moire_paramset1,
                                      "paramset4" = moire_all_coi_summary_coi_measures_plot_moire_paramset4,
                                      "paramset2" = moire_all_coi_summary_coi_measures_plot_moire_paramset2,
                                      "paramset5" = moire_all_coi_summary_coi_measures_plot_moire_paramset5,
                                      "paramset3" = moire_all_coi_summary_coi_measures_plot_moire_paramset3, 
                                      "paramset6" = moire_all_coi_summary_coi_measures_plot_moire_paramset6 
                                      )))
```




```{r}
ecoi_ccc = CCC(moire_all_coi_summary$coi, moire_all_coi_summary$expected_COI)

concordance_expectedCOI_ecoi = survival::concordance(expected_COI~ecoi, data = moire_all_coi_summary)
concordance_expectedCOI_ecoi

concordance_expectedCOI_coi = survival::concordance(expected_COI~coi, data = moire_all_coi_summary)
concordance_expectedCOI_coi

concordance_expectedCOI_naive_coi = survival::concordance(expected_COI~naive_coi, data = moire_all_coi_summary)
concordance_expectedCOI_naive_coi

concordance_expectedCOI_offset_naive_coi = survival::concordance(expected_COI~offset_naive_coi, data = moire_all_coi_summary)
concordance_expectedCOI_offset_naive_coi

```

### Monoclonal Contignency tables for e-coi cut off 

With e-coi being a continuous variable rather than an integer, will test how rounding would affect measurement

```{r}
moire_all_coi_summary_ecoi_cut_off_metrics = tibble()
for (ecoi_cut_off in seq(1, 2, by = 0.05)) {
  moire_all_coi_summary_mod = moire_all_coi_summary %>%
    mutate(ecoi_below_cut_off  = ecoi <= ecoi_cut_off,
           monoclonal = expected_COI == 1)
  moire_all_coi_summary_mod_summary = moire_all_coi_summary_mod %>%
    group_by(population_name,
             paramset_id,
             ecoi_below_cut_off,
             monoclonal) %>%
    count() %>%
    mutate(
      classification = case_when(
        ecoi_below_cut_off & monoclonal ~ "TruePositive",
        ecoi_below_cut_off & !monoclonal ~ "FalsePositive",!ecoi_below_cut_off &
          monoclonal ~ "FalseNegative",!ecoi_below_cut_off &
          !monoclonal ~ "TrueNegative"
      )
    )
  metrics <- moire_all_coi_summary_mod_summary %>%
    group_by(population_name, paramset_id) %>%
    summarise(
      TP = sum(n[ecoi_below_cut_off &  monoclonal], na.rm = TRUE),
      FP = sum(n[ecoi_below_cut_off & !monoclonal], na.rm = TRUE),
      TN = sum(n[!ecoi_below_cut_off & !monoclonal], na.rm = TRUE),
      FN = sum(n[!ecoi_below_cut_off &  monoclonal], na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      sensitivity = if_else(TP + FN > 0, TP / (TP + FN), NA_real_),
      specificity = if_else(TN + FP > 0, TN / (TN + FP), NA_real_),
      ppv        = if_else(TP + FP > 0, TP / (TP + FP), NA_real_),
      npv        = if_else(TN + FN > 0, TN / (TN + FN), NA_real_),
      accuracy   = (TP + TN) / (TP + TN + FP + FN),
      prevalence = (TP + FN) / (TP + TN + FP + FN),
      tpr         = sensitivity,
      fpr         = if_else(FP + TN > 0, FP / (FP + TN), NA_real_),
      f1_score    = if_else((2 * TP + FP + FN) > 0, (2 * TP) / (2 * TP + FP + FN), NA_real_),
      balanced_accuracy = (sensitivity + specificity) / 2
    ) %>%
    mutate(ecoi_cut_off = ecoi_cut_off)
  moire_all_coi_summary_ecoi_cut_off_metrics = bind_rows(moire_all_coi_summary_ecoi_cut_off_metrics, metrics)
}

moire_all_coi_summary_ecoi_cut_off_metrics = moire_all_coi_summary_ecoi_cut_off_metrics %>%
  left_join(sim_params)

ggplot(moire_all_coi_summary_ecoi_cut_off_metrics) +
  geom_boxplot(aes(x = paramset_id, y = sensitivity, fill = factor(ecoi_cut_off))) +
  facet_grid(coi_cat~.)

ggplot(moire_all_coi_summary_ecoi_cut_off_metrics) +
  geom_boxplot(aes(x = paramset_id, y = f1_score, fill = factor(ecoi_cut_off))) +
  facet_grid(coi_cat~.)

```



<!-- ### Allele frequencies  -->

<!-- moire_allele_freq_summary -->
<!-- moire_he_summary -->

<!-- ```{r} -->
madhatter_allele_data_naive_allele_freq_totals = madhatter_allele_data %>%
  group_by(population_name, target_id) %>%
  summarise(allele_total = n(),
            sample_total = n_distinct(library_sample_name))

madhatter_allele_data_naive_allele_freq_prev = madhatter_allele_data %>%
  group_by(population_name, target_id, seq) %>%
  summarise(
    allele_count = n(),
    sample_count = n_distinct(library_sample_name)
  ) %>%
  left_join(madhatter_allele_data_naive_allele_freq_totals) %>%
  mutate(allele_freq  = allele_count/allele_total,
         sample_prev = sample_count/sample_total) %>%
  mutate(population_name_target_id_seq = paste0(population_name, "-", target_id, "-", seq))


<!-- population_trimmed_mhap_freqs_prevs_sel = population_trimmed_mhap_freqs_prevs %>%  -->
<!--   dplyr::rename(target_id = target) %>%  -->
<!--   mutate(population_name_target_id_seq = paste0(population_name, "-", target_id, "-", seq)) -->
<!-- colnames(population_trimmed_mhap_freqs_prevs_sel)[6:12] = paste0("expected_", colnames(population_trimmed_mhap_freqs_prevs_sel)[6:12]) -->


<!-- population_trimmed_mhap_freqs_prevs_sel_inMadhatterRes = population_trimmed_mhap_freqs_prevs_sel %>%  -->
<!--   filter(population_name_target_id_seq %in% madhatter_allele_data_naive_allele_freq_prev$population_name_target_id_seq) -->

<!-- population_trimmed_mhap_freqs_prevs_sel_inMadhatterRes_he = population_trimmed_mhap_freqs_prevs_sel_inMadhatterRes %>%  -->
<!--   group_by(population_name, country, pop, target_id) %>%  -->
<!--   summarise(he = 1 - sum(expected_freq^2)) -->

<!-- population_trimmed_mhap_freqs_prevs_sel_notInMadhatterRes = population_trimmed_mhap_freqs_prevs_sel %>%  -->
<!--   filter(population_name_target_id_seq %!in% madhatter_allele_data_naive_allele_freq_prev$population_name_target_id_seq) -->

<!-- moire_allele_freq_summary_mod = moire_allele_freq_summary %>%  -->
<!--   full_join(population_trimmed_mhap_freqs_prevs_sel_inMadhatterRes %>%  -->
<!--               select(-pop, -country)) %>%  -->
<!--   mutate(population_name_target_id_seq = paste0(population_name, "-", target_id, "-", seq)) %>%  -->
<!--   mutate(freq = ifelse(is.na(freq), 0, freq),  -->
<!--          expected_freq = ifelse(is.na(expected_freq), 0, expected_freq)) -->

<!-- population_trimmed_mhap_freqs_prevs_sel_inMadhatterRes_notInMoire = population_trimmed_mhap_freqs_prevs_sel_inMadhatterRes %>%  -->
<!--   filter(population_name_target_id_seq %!in% moire_allele_freq_summary_mod$population_name_target_id_seq) -->

<!-- moire_allele_freq_summary_mod_notInMadhatterRes = moire_allele_freq_summary_mod %>% -->
<!--   filter(!is.na(freq)) %>% -->
<!--   filter(population_name_target_id_seq %!in% madhatter_allele_data_naive_allele_freq_prev$population_name_target_id_seq) -->

<!-- madhatter_allele_data_naive_allele_freq_prev_notInMoireRes = madhatter_allele_data_naive_allele_freq_prev %>% -->
<!--   filter(population_name_target_id_seq %!in% moire_allele_freq_summary_mod$population_name_target_id_seq ) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- #| fig-width: 10 -->
<!-- #| fig-height: 10 -->
<!-- #|  -->
<!-- moire_allele_freq_summary_mod_diff = moire_allele_freq_summary_mod %>%  -->
<!--   mutate(freq_diff = freq - expected_freq) %>%  -->
<!--   mutate(abs_freq_diff = abs(freq_diff)) %>%  -->
<!--   left_join(sim_params) -->

<!-- ggplot(moire_allele_freq_summary_mod_diff) +  -->
<!--   geom_boxplot(aes(x = pop_cat, y = abs(freq_diff), fill = paramset_id)) +  -->
<!--   custom_theme_xRotate -->


<!-- library(finalfit) -->
<!-- explanatory = c("coi_cat", "nonrecombinant_cat", "between_sample_related_cat", "country", "paramset_id") -->
<!-- dependent = 'abs_freq_diff' -->
<!-- moire_allele_freq_summary_mod_diff %>% -->
<!--   mutate(coi_cat = factor(coi_cat),  -->
<!--          nonrecombinant_cat = factor(nonrecombinant_cat),  -->
<!--          between_sample_related_cat = factor(between_sample_related_cat),  -->
<!--          country = factor(country),  -->
<!--          paramset_id = factor(paramset_id)) %>%  -->
<!--   finalfit(dependent, explanatory, metrics=TRUE) -> t2 -->
<!-- knitr::kable(t2[[1]], row.names=FALSE, align=c("l", "l", "r", "r", "r", "r")) -->
<!-- knitr::kable(t2[[2]], row.names=FALSE, col.names="") -->

<!-- moire_allele_freq_summary_mod_diff_rmse = moire_allele_freq_summary_mod_diff %>%  -->
<!--   ungroup() %>%  -->
<!--   group_by(population_name, paramset_id) %>%  -->
<!--   reframe(rmse = sqrt(sum(freq_diff^2)/length(freq_diff))) %>%  -->
<!--   left_join(sim_params) -->

<!-- moire_allele_freq_summary_mod_diff_rmse %>% -->
<!--   mutate(coi_cat = factor(coi_cat),  -->
<!--          nonrecombinant_cat = factor(nonrecombinant_cat),  -->
<!--          between_sample_related_cat = factor(between_sample_related_cat),  -->
<!--          country = factor(country),  -->
<!--          paramset_id = factor(paramset_id)) %>%  -->
<!--   finalfit("rmse", explanatory, metrics=TRUE) -> t2 -->
<!-- knitr::kable(t2[[1]], row.names=FALSE, align=c("l", "l", "r", "r", "r", "r")) -->
<!-- knitr::kable(t2[[2]], row.names=FALSE, col.names="") -->


<!-- moire_allele_freq_summary_mod_diff_rmse_plot_all = ggplot(moire_allele_freq_summary_mod_diff_rmse) +  -->
<!--   geom_boxplot( -->
<!--     aes(x = paramset_id,  -->
<!--         y = rmse) -->
<!--   ) +  -->
<!--   custom_theme_xRotate -->

<!-- print(moire_allele_freq_summary_mod_diff_rmse_plot_all) -->

<!-- moire_allele_freq_summary_mod_diff_rmse_plot = ggplot(moire_allele_freq_summary_mod_diff_rmse) +  -->
<!--   geom_boxplot( -->
<!--     aes(x = pop_cat,  -->
<!--         y = rmse,  -->
<!--         fill = country) -->
<!--   ) +  -->
<!--   facet_grid(paramset_id~.) +  -->
<!--   custom_theme_xRotate -->

<!-- print(moire_allele_freq_summary_mod_diff_rmse_plot) -->

<!-- moire_allele_freq_summary_mod_diff_rmse_plot_between_sample_related_cat = ggplot(moire_allele_freq_summary_mod_diff_rmse %>%  -->
<!--                                                                                    filter(rmse < 0.2)) +  -->
<!--   geom_boxplot( -->
<!--     aes(x = between_sample_related_cat,  -->
<!--         y = rmse,  -->
<!--         fill = country) -->
<!--   ) +  -->
<!--   facet_grid(paramset_id~.) +  -->
<!--   custom_theme_xRotate -->

<!-- print(moire_allele_freq_summary_mod_diff_rmse_plot_between_sample_related_cat) -->
<!-- ``` -->

<!-- ```{r} -->

<!-- madhatter_allele_data_naive_allele_freq_prev_mod = madhatter_allele_data_naive_allele_freq_prev -->
<!-- colnames(madhatter_allele_data_naive_allele_freq_prev_mod)[4:9] = paste0("naive_", colnames(madhatter_allele_data_naive_allele_freq_prev)[4:9]) -->

<!-- moire_allele_freq_summary_mod_diff_with_naive = moire_allele_freq_summary_mod_diff %>%  -->
<!--   left_join(madhatter_allele_data_naive_allele_freq_prev_mod) -->



<!-- ``` -->



<!-- ```{r} -->
<!-- moire_allele_freq_summary_mod_with_naive_paramset1 = ggpairs(moire_allele_freq_summary_mod_diff_with_naive %>%  -->
<!--                                                     filter(paramset_id == "moire_paramset1"), -->
<!--                                                   columns = c("freq", "naive_allele_freq", "expected_freq"),  -->
<!--                                                   diag = list(continuous = "barDiag"), -->
<!--                                                   title = "moire_paramset1")  -->

<!-- moire_allele_freq_summary_mod_with_naive_paramset2 = ggpairs(moire_allele_freq_summary_mod_diff_with_naive %>%  -->
<!--                                                     filter(paramset_id == "moire_paramset2"), -->
<!--                                                   columns = c("freq", "naive_allele_freq", "expected_freq"),  -->
<!--                                                   diag = list(continuous = "barDiag"),  -->
<!--                                                   title = "moire_paramset2") -->

<!-- moire_allele_freq_summary_mod_with_naive_paramset3 = ggpairs(moire_allele_freq_summary_mod_diff_with_naive %>%  -->
<!--                                                     filter(paramset_id == "moire_paramset3"), -->
<!--                                                   columns = c("freq", "naive_allele_freq", "expected_freq"),  -->
<!--                                                   diag = list(continuous = "barDiag"),  -->
<!--                                                   title = "moire_paramset3") -->


<!-- ``` -->


<!-- Plotting correlation between the different COI estimates compared to expected COI over each parameter set  -->

<!-- ```{r} -->
<!-- #| results: asis -->
<!-- #| echo: false -->
<!-- #| fig-column: screen-inset -->

<!-- cat(create_tabsetOfGgplotObjects(list("paramset1" = moire_allele_freq_summary_mod_with_naive_paramset1,  -->
<!--                                       "paramset2" = moire_allele_freq_summary_mod_with_naive_paramset2,  -->
<!--                                       "paramset3" = moire_allele_freq_summary_mod_with_naive_paramset3))) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- moire_allele_freq_summary_mod_full = moire_allele_freq_summary %>%  -->
<!--   full_join(population_trimmed_mhap_freqs_prevs_sel_inMadhatterRes %>%  -->
<!--               select(-pop, -country)) %>%  -->
<!--   mutate(population_name_target_id_seq = paste0(population_name, "-", target_id, "-", seq)) %>%  -->
<!--   group_by(population_name, target_id) %>%  -->
<!--   mutate(n_obs = sum(!is.na(freq)),  -->
<!--          n_exp = sum(!is.na(expected_freq))) %>%  -->
<!--   mutate(freq_mod = ifelse(is.na(freq), 0, freq)) %>%  -->

<!--   mutate(freq_diff = freq_mod - expected_freq) -->

<!-- moire_allele_freq_summary_mod_full_summary = moire_allele_freq_summary_mod_full %>%  -->
<!--   mutate( -->
<!--     acc_shared = (freq_diff)/sqrt(expected_freq * (1 - expected_freq)) -->
<!--     sophie_metric = (freq_diff) / sqrt( -->
<!--       (freq_mod * (1 - freq_mod)) -->
<!--     ) -->
<!--   ) -->


<!-- ``` -->

<!-- #### he  -->

<!-- ```{r} -->
<!-- #| fig-width: 10 -->
<!-- #| fig-height: 10 -->
<!-- moire_he_summary_mod = moire_he_summary %>%  -->
<!--   left_join(population_trimmed_mhap_freqs_prevs_sel_inMadhatterRes_he %>%  -->
<!--               dplyr::rename(expected_he = he)) %>%  -->
<!--   mutate(he_diff = he - expected_he) -->

<!-- moire_he_summary_mod_rmse = moire_he_summary_mod %>%  -->
<!--   group_by(population_name, paramset_id) %>%  -->
<!--   summarise( -->
<!--     rmse = sqrt(sum(he_diff^2)/length(he_diff))) %>%  -->
<!--   left_join(sim_params) -->


<!-- moire_he_summary_mod_rmse_plot = ggplot(moire_he_summary_mod_rmse) +  -->
<!--   geom_boxplot( -->
<!--     aes(x = pop_cat,  -->
<!--         y = rmse,  -->
<!--         fill = country) -->
<!--   ) +  -->
<!--   facet_grid(paramset_id~.) +  -->
<!--   custom_theme_xRotate -->

<!-- print(moire_he_summary_mod_rmse_plot) -->
<!-- ``` -->






