---
title: MOIRE Benchmarking Results
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)
```


## Set up 

### Libraries used in the analysis below and Some helpful R utilities


```{r, warning=FALSE, message=FALSE}
source("../scripts/benchmarking_functions.R")
```


## Load in the Benchmarking Outputs

```{r}
inp <- list(
  sim_info_dir = "../../data/benchmarking/simulations/SpotMalaria", 
  tool_res_dir = "../../data/benchmarking/tool_results/MOIRE/", 
  madhatter_res_dir = "../../data/benchmarking/tool_results/madhatter/",
  tool_params = "../../data/benchmarking/tool_results/moire.json"
)
```


First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.

### Reading in expected truths 

Read per in expected allele frequencies per sample tables to get true COIs

```{r}
expected_simulated_frequencies_per_sample <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "trimmed_expected_simulated_frequencies_per_sample.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(library_sample_name = libary_sample_name)

expected_simulated_frequencies_per_sample_COIs = expected_simulated_frequencies_per_sample %>% 
  select(panel, population_name, country, library_sample_name, COI) %>% 
  distinct() %>% 
  dplyr::rename(expected_COI = COI) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Read per in expected allele frequencies tables population level 
```{r}
population_trimmed_mhap_freqs_prevs <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_trimmed_mhap_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(),
        population_name = col_character(), 
        pop = col_character(), 
        country = col_character(), 
        target = col_character(), 
        seq = col_character()
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(target_id = target)
```


Read in all simulation parameters into one tibble. Encode different population parameters as descriptive factors. 

```{r}

sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>% 
  mutate(
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>% 
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  )) %>% 
  mutate(pop_cat = factor(pop_cat, levels = sort(unique((.$pop_cat)))), 
         coi_cat = factor(coi_cat, levels = sort(unique((.$coi_cat)))), 
         between_sample_related_cat = factor(between_sample_related_cat, levels = sort(unique((.$between_sample_related_cat)))), 
         nonrecombinant_cat = factor(nonrecombinant_cat, levels = sort(unique((.$nonrecombinant_cat)))), 
         meiosis_cat = factor(meiosis_cat, levels = sort(unique((.$meiosis_cat)))))

create_dt(sim_params)
```

Read in madhatter pipeline allele tables 

```{r}
madhatter_allele_data <- tibble(
    allele_data_path = list.files(
      inp$madhatter_res_dir, 
      pattern = "allele_data.txt.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    allele_data = map(
      allele_data_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  mutate(tool_res_dir = dirname(allele_data_path)) %>%
  mutate(population_name = gsub("_mh_results", "", basename(tool_res_dir))) %>%
  select(-allele_data_path, -tool_res_dir) %>%
  unnest(allele_data) %>% 
  mutate(SampleID = gsub("_S[0-9]+$", "", SampleID)) %>% 
  dplyr::rename(library_sample_name = SampleID, 
                target_id = Locus, 
                seq = ASV, 
                reads = Reads) 
```

### Read in MOIRE results

MOIRE outputs information on complexity of infection (COI) and on population allele frequencies and several different outputs need to be read in. 

Getting the allele frequencies data in the `allele_freq_summary.tsv.gz` files 

```{r}
moire_allele_freq_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "allele_freq_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 
```

Getting the complexity of infection (COI) from the `coi_summary.tsv.gz` files 

```{r}
moire_coi_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "^coi_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id)) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Getting the effective COI from the `effective_coi_summary.tsv.gz` files. 

```{r}
moire_effective_coi_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "effective_coi_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id))%>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Getting the expected heterozygosity (he) estimates `he_summary.tsv.gz` files 

```{r}
moire_he_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "he_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res)
```

Getting the within host relatedness from the `relatedness_summary.tsv.gz` files 

```{r}
moire_relatedness_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "relatedness_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id))%>% 
  dplyr::rename(library_sample_name = specimen_id)

```

Getting what tool parameters were used. For MOIRE benchmarking we set the priors on the assumed amount of relatedness with moire_paramset1 assuming uniform relatedness and then each moire_paramset 2 and 3 increasing the amount of assumed relatedness. moire_paramsets3-6 are the same prior relatedness as 1-3 but have more parallel burin to explore more pace for convergence.   

```{r}
# Read in tool parameters
tool_params <- jsonlite::fromJSON(inp$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    r_alpha = paramset.r_alpha,
    r_beta = paramset.r_beta
  )

create_dt(tool_params)
```

## Benchmarking 

### COI  

#### Metrics 

*  Plotting Bland-Altman plots (plotting the mean of expected and obs vs difference) 
*  Calculating the RMSE (taking the square root of the mean squared difference between expected and observed)  
*  Calculating [Lin's concordance correlation coefficient](https://www.rdocumentation.org/packages/DescTools/versions/0.99.60/topics/CCC) for agreement on a continuous measure 
*  Calculating contingency table of determining a sample as monoclonal vs polyclonal 

#### Results 

MOIRE reports several COIs: 

*  naive_coi - The max alleles called for any locus for a sample 
*  offset_naive_coi - The 2nd max alleles called for any locus for a sample 
*  coi - COI reported by MOIRE 
*  ecoi - ecoi is the MOIRE COI adjusted for the amount of within sample relatedness 

```{r}
# combine the COI summaries and ecoi summaries and combine with the expected outputs for COI
moire_all_coi_summary = moire_coi_summary %>% 
  left_join(moire_effective_coi_summary) %>% 
  left_join(expected_simulated_frequencies_per_sample_COIs %>% 
              mutate(library_sample_name = as.character(library_sample_name))) %>% 
                                                        left_join(sim_params)

# calculate the means and differences for Bland-Altman plots

moire_all_coi_summary = moire_all_coi_summary %>% 
  add_diff_mean(expected_COI, naive_coi) %>% 
  add_diff_mean(expected_COI, offset_naive_coi)%>% 
  add_diff_mean(expected_COI, coi)%>% 
  add_diff_mean(expected_COI, ecoi)

```

##### Bland-Altman COI plots 

Overall 



```{r}
moire_all_coi_summary_overall_ecoi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "ecoi", expected_label = "Expected COI")
moire_all_coi_summary_overall_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "coi", expected_label = "Expected COI")
moire_all_coi_summary_overall_naive_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "naive_coi", expected_label = "Expected COI")
moire_all_coi_summary_overall_offset_naive_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "offset_naive_coi", expected_label = "Expected COI")

```


```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive COI" = moire_all_coi_summary_overall_naive_coi_bland_altman_plot, 
    "offset naive COI" = moire_all_coi_summary_overall_offset_naive_coi_bland_altman_plot, 
    "COI" = moire_all_coi_summary_overall_coi_bland_altman_plot, 
    "eCOI" = moire_all_coi_summary_overall_ecoi_bland_altman_plot
       )
))
```


Showing the different parameter sets 

```{r}
moire_all_coi_summary_ecoi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "ecoi") + 
  facet_wrap(~paramset_id, ncol = 3)
moire_all_coi_summary_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "coi") + 
  facet_wrap(~paramset_id, ncol = 3)
moire_all_coi_summary_offset_naive_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "offset_naive_coi") + 
  facet_wrap(~paramset_id, ncol = 3) 
moire_all_coi_summary_naive_coi_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "naive_coi") + 
  facet_wrap(~paramset_id, ncol = 3)

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive COI" = moire_all_coi_summary_naive_coi_bland_altman_plot, 
    "offset naive COI" = moire_all_coi_summary_offset_naive_coi_bland_altman_plot, 
    "COI" = moire_all_coi_summary_coi_bland_altman_plot, 
    "eCOI" = moire_all_coi_summary_ecoi_bland_altman_plot
       )
))
```

Write to pdf  
```{r}
#| results: hide
cairo_pdf("moire_all_cois_bland_altman_plot.pdf", width = 10, height = 10)
print(moire_all_coi_summary_ecoi_bland_altman_plot)
print(moire_all_coi_summary_coi_bland_altman_plot)
print(moire_all_coi_summary_offset_naive_coi_bland_altman_plot)
print(moire_all_coi_summary_naive_coi_bland_altman_plot)
dev.off()
```

Showing the different population characteristics 

```{r}
moire_all_coi_summary_ecoi_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "ecoi") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)

moire_all_coi_summary_coi_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "coi") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)

moire_all_coi_summary_naive_coi_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "naive_coi") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)

moire_all_coi_summary_offset_naive_coi_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_all_coi_summary, "offset_naive_coi") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)
```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive COI" = moire_all_coi_summary_naive_coi_by_pop_bland_altman_plot, 
    "offset naive COI" = moire_all_coi_summary_offset_naive_coi_by_pop_bland_altman_plot, 
    "COI" = moire_all_coi_summary_coi_by_pop_bland_altman_plot, 
    "eCOI" = moire_all_coi_summary_ecoi_by_pop_bland_altman_plot
       )
))
```

Write to pdf  
```{r}
#| results: hide
cairo_pdf("moire_all_cois_by_pop_bland_altman_plot.pdf", width = 10, height = 10)
print(moire_all_coi_summary_ecoi_by_pop_bland_altman_plot)
print(moire_all_coi_summary_coi_by_pop_bland_altman_plot)
print(moire_all_coi_summary_offset_naive_coi_by_pop_bland_altman_plot)
print(moire_all_coi_summary_naive_coi_by_pop_bland_altman_plot)
dev.off()
```


##### Concordance 

Plotting concordance between all COI metrics reported by MOIRE.

Overall 

```{r}
moire_all_coi_summary_coi_measures_plot_moire_all = ggpairs(moire_all_coi_summary,
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "all")

```

```{r}
print(moire_all_coi_summary_coi_measures_plot_moire_all)
```


Plotting difference parameter sets  

```{r}
moire_all_coi_summary_coi_measures_plot_moire_paramset1 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset1"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset1")

moire_all_coi_summary_coi_measures_plot_moire_paramset2 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset2"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset2")

moire_all_coi_summary_coi_measures_plot_moire_paramset3 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset3"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset3")

moire_all_coi_summary_coi_measures_plot_moire_paramset4 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset4"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset4")

moire_all_coi_summary_coi_measures_plot_moire_paramset5 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset5"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset5")

moire_all_coi_summary_coi_measures_plot_moire_paramset6 = ggpairs(moire_all_coi_summary %>%
                                                    filter(paramset_id == "moire_paramset6"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset6")

```

```{r}
#| results: hide
cairo_pdf("moire_coi_pairwise_comparison_parametersets.pdf", width = 15, height = 15)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset1)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset4)

print(moire_all_coi_summary_coi_measures_plot_moire_paramset2)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset5)

print(moire_all_coi_summary_coi_measures_plot_moire_paramset3)
print(moire_all_coi_summary_coi_measures_plot_moire_paramset6)
dev.off()
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("paramset1" = moire_all_coi_summary_coi_measures_plot_moire_paramset1,
                                      "paramset4" = moire_all_coi_summary_coi_measures_plot_moire_paramset4,
                                      "paramset2" = moire_all_coi_summary_coi_measures_plot_moire_paramset2,
                                      "paramset5" = moire_all_coi_summary_coi_measures_plot_moire_paramset5,
                                      "paramset3" = moire_all_coi_summary_coi_measures_plot_moire_paramset3, 
                                      "paramset6" = moire_all_coi_summary_coi_measures_plot_moire_paramset6 
                                      )))
```

Different population metrics 


COI categories 

```{r}
moire_all_coi_summary_coi_measures_plot_moire_MeanCOI1.26 = ggpairs(moire_all_coi_summary %>%
                                                    filter(coi_cat == "MeanCOI=1.26"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=1.26")

moire_all_coi_summary_coi_measures_plot_moire_MeanCOI3.41 = ggpairs(moire_all_coi_summary %>%
                                                    filter(coi_cat == "MeanCOI=3.41"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=3.41")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("MeanCOI_1.26" = moire_all_coi_summary_coi_measures_plot_moire_MeanCOI1.26,
                                      "MeanCOI_3.41" = moire_all_coi_summary_coi_measures_plot_moire_MeanCOI3.41
                                      )))
```


between sample related categories 

```{r}
moire_all_coi_summary_coi_measures_plot_moire_1perc_related = ggpairs(moire_all_coi_summary %>%
                                                    filter(between_sample_related_cat == "_1%_related"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "_1%_related")

moire_all_coi_summary_coi_measures_plot_moire_10perc_related = ggpairs(moire_all_coi_summary %>%
                                                    filter(between_sample_related_cat == "10%_related"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "10%_related")

moire_all_coi_summary_coi_measures_plot_moire_30perc_related = ggpairs(moire_all_coi_summary %>%
                                                    filter(between_sample_related_cat == "30%_related"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "30%_related")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("_1perc_between_sample_related" = moire_all_coi_summary_coi_measures_plot_moire_1perc_related,
                                      "10perc_between_sample_related" = moire_all_coi_summary_coi_measures_plot_moire_10perc_related, 
                                      "30perc_between_sample_related" = moire_all_coi_summary_coi_measures_plot_moire_30perc_related
                                      )))
```

Recombination categories 
```{r}
moire_all_coi_summary_coi_measures_plot_moire_50perc_non_recomb = ggpairs(moire_all_coi_summary %>%
                                                    filter(nonrecombinant_cat == "50%_non-recombinant"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "50%_non-recombinant")

moire_all_coi_summary_coi_measures_plot_moire_20perc_non_recomb = ggpairs(moire_all_coi_summary %>%
                                                    filter(nonrecombinant_cat == "20%_non-recombinant"),
                                                  columns = c("expected_COI", "naive_coi", "offset_naive_coi", "coi", "ecoi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "20%_non-recombinant")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("20perc_non_recomb" = moire_all_coi_summary_coi_measures_plot_moire_20perc_non_recomb,
                                      "50perc_non_recomb" = moire_all_coi_summary_coi_measures_plot_moire_50perc_non_recomb
                                      )))
```


```{r}
#| results: hide
cairo_pdf("moire_coi_pairwise_comparison_pop_parameters.pdf", width = 15, height = 15)
print(moire_all_coi_summary_coi_measures_plot_moire_MeanCOI1.26)
print(moire_all_coi_summary_coi_measures_plot_moire_MeanCOI3.41)

print(moire_all_coi_summary_coi_measures_plot_moire_1perc_related)
print(moire_all_coi_summary_coi_measures_plot_moire_10perc_related)
print(moire_all_coi_summary_coi_measures_plot_moire_30perc_related)

print(moire_all_coi_summary_coi_measures_plot_moire_20perc_non_recomb)
print(moire_all_coi_summary_coi_measures_plot_moire_50perc_non_recomb)
dev.off()
```




```{r}
# calculate Concordance Correlation Coefficient (CCC) 
moire_all_coi_summary_ccc_by_pop <- moire_all_coi_summary %>%
  compute_ccc(
    expected_col  = expected_COI,
    observed_cols = c(coi, ecoi, naive_coi, offset_naive_coi),
    group_cols    = c(population_name, paramset_id)
  )

# Gather into long format, left join in the simulation parameters
moire_all_coi_summary_ccc_by_pop_gat = moire_all_coi_summary_ccc_by_pop %>%
  select(population_name, paramset_id, ends_with("_ccc")) %>% 
  gather(coi_estimate, ccc, 3:ncol(.)) %>%
  mutate(coi_estimate = gsub("_ccc", "", coi_estimate))%>%
  left_join(sim_params)
```




ccc across all parameter sets for all 4 estimators 

```{r}
moire_all_coi_summary_ccc_plot_just_estimators = ggplot(moire_all_coi_summary_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(y = ccc, x = coi_estimate, fill = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

ggplotly(moire_all_coi_summary_ccc_plot_just_estimators)
```

ccc over all comparing parameter sets for each estimate

```{r}
moire_all_coi_summary_ccc_plot_fill_paramset_id = ggplot(moire_all_coi_summary_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_id, y = ccc, x = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_plot_fill_coi_estimate = ggplot(moire_all_coi_summary_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = coi_estimate, y = ccc, x = paramset_id), position = "dodge"
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfHtmlWidgets(
  list(
  "By Parameter Set" = ggplotly(moire_all_coi_summary_ccc_plot_fill_paramset_id) %>% layout(boxmode = "group"), 
  "By COI Estimator" = ggplotly(moire_all_coi_summary_ccc_plot_fill_coi_estimate) %>% layout(boxmode = "group")
  )
))

```

ccc on the different simulation population characteristics 

```{r}
moire_all_coi_summary_ccc_by_pop_gat_large_plot_by_country = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat_by_country = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat = ggplot(moire_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = moire_all_coi_summary_ccc_by_pop_gat_large_plot,
  "est_between_sample_related" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat,
  "coi_cat" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat,
  "nonrecombinant" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = moire_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country
)))
```

##### RMSE   

```{r}
# calculate RMSE 
moire_all_coi_summary_rmse <- moire_all_coi_summary %>%
  compute_rmse(
    expected_col  = expected_COI,
    observed_cols = c(coi, ecoi, naive_coi, offset_naive_coi),
    group_cols    = c(population_name, paramset_id)
  )

# Gather into long format, left join in the simulation parameters
moire_all_coi_summary_rmse_gat = moire_all_coi_summary_rmse %>%
  gather(coi_estimate, rmse, 3:ncol(.)) %>%
  mutate(coi_estimate = gsub("_rmse", "", coi_estimate))%>%
  left_join(sim_params)

```

RMSE across all parameter sets for all 4 estimators 

```{r}
moire_all_coi_summary_rmse_plot_just_estimators = ggplot(moire_all_coi_summary_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(y = rmse, x = coi_estimate, fill = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

ggplotly(moire_all_coi_summary_rmse_plot_just_estimators)
```

RMSE over all comparing parameter sets for each estimate

```{r}
moire_all_coi_summary_rmse_plot_fill_paramset_id = ggplot(moire_all_coi_summary_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_id, y = rmse, x = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

moire_all_coi_summary_rmse_plot_fill_coi_estimate = ggplot(moire_all_coi_summary_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = coi_estimate, y = rmse, x = paramset_id), position = "dodge"
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfHtmlWidgets(
  list(
  "By Parameter Set" = ggplotly(moire_all_coi_summary_rmse_plot_fill_paramset_id) %>% layout(boxmode = "group"), 
  "By COI Estimator" = ggplotly(moire_all_coi_summary_rmse_plot_fill_coi_estimate) %>% layout(boxmode = "group")
  )
))

```

RMSE on the different simulation population characteristics 

```{r}
moire_all_coi_summary_rmse_gat_large_plot_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_all_coi_summary_rmse_gat_large_plot_coi_cat_by_country = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_all_coi_summary_rmse_gat_large_plot = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate

moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate

moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate

moire_all_coi_summary_rmse_gat_large_plot_coi_cat = ggplot(moire_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~coi_estimate) +
  custom_ggplot_theme_xRotate

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = moire_all_coi_summary_rmse_gat_large_plot,
  "est_between_sample_related" = moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat,
  "coi_cat" = moire_all_coi_summary_rmse_gat_large_plot_coi_cat,
  "nonrecombinant" = moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = moire_all_coi_summary_rmse_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = moire_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = moire_all_coi_summary_rmse_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = moire_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat_by_country
)))
```

<!-- ###### Testing for significant affects of population  -->

<!-- ####### COI  -->

<!-- ```{r} -->
<!-- library(finalfit) -->
<!-- explanatory = c("coi_cat", "nonrecombinant_cat", "between_sample_related_cat", "country", "paramset_id") -->
<!-- dependent = 'coi_rmse' -->
<!-- moire_all_coi_summary_rmse %>% -->
<!--   left_join(sim_params) %>% -->
<!--   mutate(coi_cat = factor(coi_cat), -->
<!--          nonrecombinant_cat = factor(nonrecombinant_cat), -->
<!--          between_sample_related_cat = factor(between_sample_related_cat), -->
<!--          country = factor(country), -->
<!--          paramset_id = factor(paramset_id)) %>% -->
<!--   finalfit("coi_rmse", explanatory, metrics=TRUE) -> t2 -->
<!-- knitr::kable(t2[[1]], row.names=FALSE, align=c("l", "l", "r", "r", "r", "r")) -->
<!-- knitr::kable(t2[[2]], row.names=FALSE, col.names="") -->

<!-- ``` -->

<!-- ```{r, echo = T, eval=T} -->
<!-- #| fig-width: 15 -->
<!-- moire_all_coi_summary_rmse %>% -->
<!--   left_join(sim_params) %>% -->
<!--   mutate(coi_cat = factor(coi_cat), -->
<!--          nonrecombinant_cat = factor(nonrecombinant_cat), -->
<!--          between_sample_related_cat = factor(between_sample_related_cat), -->
<!--          country = factor(country), -->
<!--          paramset_id = factor(paramset_id))  %>% -->
<!--   # filter(coi_rmse < 1) %>% -->
<!--   coefficient_plot("coi_rmse", -->
<!--           # "country" -->
<!--           explanatory -->
<!--           ) -->

<!-- ``` -->

<!-- ####### eCOI  -->
<!-- ```{r} -->
<!-- library(finalfit) -->
<!-- explanatory = c("coi_cat", "nonrecombinant_cat", "between_sample_related_cat", "country", "paramset_id") -->
<!-- dependent = 'ecoi_rmse' -->
<!-- moire_all_coi_summary_rmse %>% -->
<!--   left_join(sim_params) %>% -->
<!--   mutate(coi_cat = factor(coi_cat), -->
<!--          nonrecombinant_cat = factor(nonrecombinant_cat), -->
<!--          between_sample_related_cat = factor(between_sample_related_cat), -->
<!--          country = factor(country), -->
<!--          paramset_id = factor(paramset_id)) %>% -->
<!--   finalfit("ecoi_rmse", explanatory, metrics=TRUE) -> t2 -->
<!-- knitr::kable(t2[[1]], row.names=FALSE, align=c("l", "l", "r", "r", "r", "r")) -->
<!-- knitr::kable(t2[[2]], row.names=FALSE, col.names="") -->

<!-- ``` -->

<!-- ```{r, echo = T, eval=T} -->
<!-- #| fig-width: 15 -->
<!-- moire_all_coi_summary_rmse %>% -->
<!--   left_join(sim_params) %>% -->
<!--   mutate(coi_cat = factor(coi_cat), -->
<!--          nonrecombinant_cat = factor(nonrecombinant_cat), -->
<!--          between_sample_related_cat = factor(between_sample_related_cat), -->
<!--          country = factor(country), -->
<!--          paramset_id = factor(paramset_id))  %>% -->
<!--   # filter(ecoi_rmse < 1) %>% -->
<!--   coefficient_plot("ecoi_rmse", -->
<!--           # "country" -->
<!--           explanatory -->
<!--           ) -->

<!-- ``` -->


#### Monoclonal vs polyclonal determination 



```{r}
# Round eCOI and COI for testing 

moire_all_coi_summary_mono_vs_poly = moire_all_coi_summary %>%
  mutate(ecoi  = round(ecoi), 
         coi = round(coi)) %>% 
  select(population_name, paramset_id, library_sample_name,expected_COI, coi, ecoi, naive_coi, offset_naive_coi) %>% 
  gather(coi_estimator, determined_coi, 5:ncol(.)) %>%
  mutate(
    test_classification = case_when(
      determined_coi == 1  & expected_COI == 1 ~ "TruePositive",
      determined_coi == 1  & expected_COI != 1 ~ "FalsePositive",
      determined_coi != 1  & expected_COI == 1 ~ "FalseNegative",
      determined_coi != 1  & expected_COI != 1 ~ "TrueNegative"
    )
  ) %>%
  left_join(sim_params) 

mono_vs_poly_by_parameterset <- moire_all_coi_summary_mono_vs_poly %>%
  group_by(coi_estimator, paramset_id) %>%
  summarise(
    TP = sum(determined_coi == 1  & expected_COI == 1, na.rm = TRUE),
    FP = sum(determined_coi == 1  & expected_COI != 1, na.rm = TRUE),
    FN = sum(determined_coi != 1  & expected_COI == 1, na.rm = TRUE),
    TN = sum(determined_coi != 1  & expected_COI != 1, na.rm = TRUE),
  ) %>% 
  add_contigency_metrics()

```

Performance on detecting monoclonal for each coi estimator across the different parameter sets 

```{r}
#| fig-width: 20
#| fig-height: 20
metric_map <- c(
  "Sensitivity"        = "sensitivity",
  "Specificity"        = "specificity",
  "Precision (PPV)"    = "ppv",
  "NPV"                = "npv",
  "F1-score"           = "f1_score",
  "Balanced Accuracy"  = "balanced_accuracy"
)

mono_vs_poly_by_parameterset_radar = radar_by_metrics(
  mono_vs_poly_by_parameterset,
  facet_col = "paramset_id",
  color_col = "coi_estimator",
  metric_map = metric_map,
  palette = colorPalette_08
)

print(mono_vs_poly_by_parameterset_radar)
```


### Allele frequencies

Since the allele calling pipeline might have baseline inaccuracies, calculating naive estimates and false positive / false negative rates for the pipeline data itself as a bunch mark can be a helpful comparator. 

#### Naive allele frequencies 

```{r}
madhatter_allele_data_naive_allele_freq_totals = madhatter_allele_data %>%
  group_by(population_name, target_id) %>%
  summarise(allele_total = n(),
            sample_total = n_distinct(library_sample_name))

madhatter_allele_data_naive_allele_freq_prev = madhatter_allele_data %>%
  group_by(population_name, target_id, seq) %>%
  summarise(
    allele_count = n(),
    sample_count = n_distinct(library_sample_name)
  ) %>%
  left_join(madhatter_allele_data_naive_allele_freq_totals) %>%
  mutate(allele_freq  = allele_count/allele_total,
         sample_prev = sample_count/sample_total) %>%
  mutate(population_name_target_id_seq = paste0(population_name, "-", target_id, "-", seq))


madhatter_allele_data_naive_allele_freq_prev_out = madhatter_allele_data_naive_allele_freq_prev %>% 
  select(population_name, target_id, seq, allele_freq, sample_prev) %>% 
  dplyr::rename(naive_freq = allele_freq, 
                naive_prev = sample_prev)


population_trimmed_mhap_freqs_prevs_sel = population_trimmed_mhap_freqs_prevs %>% 
  select(population_name, target_id, seq, freq, prev) %>% 
  dplyr::rename(expected_freq = freq, 
                expected_prev = prev) %>% 
  full_join(madhatter_allele_data_naive_allele_freq_prev_out)

# fill in zeros for missing or false alleles 
population_trimmed_mhap_freqs_prevs_sel = population_trimmed_mhap_freqs_prevs_sel %>% 
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq), 
         expected_prev = ifelse(is.na(expected_prev), 0, expected_prev), 
         naive_freq = ifelse(is.na(naive_freq), 0, naive_freq), 
         naive_prev = ifelse(is.na(naive_prev), 0, naive_prev)
         )

# moire doesn't calculate prevalence, so don't include 
population_trimmed_mhap_freqs_prevs_sel_out = population_trimmed_mhap_freqs_prevs_sel %>% 
  select(-naive_prev, -expected_prev)

```

#### Adding expected and naive estimates to MOIRE 

```{r}

moire_param_map <- moire_allele_freq_summary %>%
  distinct(population_name, target_id, paramset_id)

population_trimmed_mhap_freqs_prevs_sel_out_expanded <- population_trimmed_mhap_freqs_prevs_sel_out %>%
  inner_join(moire_param_map, by = c("population_name", "target_id"))

moire_allele_freq_summary_sel = moire_allele_freq_summary %>%
  select(population_name, paramset_id, target_id, seq, freq) %>%
  dplyr::rename(moire_freq = freq) %>%
  full_join(population_trimmed_mhap_freqs_prevs_sel_out_expanded) %>%
  mutate(
    moire_freq = ifelse(is.na(moire_freq), 0, moire_freq),
    expected_freq = ifelse(is.na(expected_freq), 0, expected_freq),
    naive_freq = ifelse(is.na(naive_freq), 0, naive_freq)
  ) %>% 
  left_join(sim_params)

```

#### Metrics 

*  Plotting Bland-Altman plots (plotting the mean of expected and obs vs difference) 
*  Calculating the RMSE (taking the square root of the mean squared difference between expected and observed)  
*  Calculating [Lin's concordance correlation coefficient](https://www.rdocumentation.org/packages/DescTools/versions/0.99.60/topics/CCC) for agreement on a continuous measure 
*  Calculating contingency table metrics for false positive and false negatives  



##### Bland-Altman Plots 


```{r}
moire_allele_freq_summary_sel = moire_allele_freq_summary_sel %>% 
    add_diff_mean(expected_col = expected_freq,
                  observed_col = moire_freq)%>% 
    add_diff_mean(expected_col = expected_freq,
                  observed_col = naive_freq)

```



Overall 

```{r}
moire_allele_freq_summary_sel_overall_moire_freq_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "moire_freq")
moire_allele_freq_summary_sel_overall_naive_freq_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "naive_freq")

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive_freq" = moire_allele_freq_summary_sel_overall_naive_freq_bland_altman_plot, 
    "moire_freq" = moire_allele_freq_summary_sel_overall_moire_freq_bland_altman_plot
       )
))
```


Showing the different parameter sets 

```{r}
moire_allele_freq_summary_sel_moire_freq_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "moire_freq") + 
  facet_wrap(~paramset_id, ncol = 3)


moire_allele_freq_summary_sel_naive_freq_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "naive_freq") + 
  facet_wrap(~paramset_id, ncol = 3) 

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive_freq" = moire_allele_freq_summary_sel_naive_freq_bland_altman_plot, 
    "moire_freq" = moire_allele_freq_summary_sel_moire_freq_bland_altman_plot
       )
))
```

Write to pdf  
```{r}
#| results: hide
cairo_pdf("moire_all_allele_freqs_bland_altman_plot.pdf", width = 10, height = 10)
print(moire_allele_freq_summary_sel_moire_freq_bland_altman_plot)
print(moire_allele_freq_summary_sel_naive_freq_bland_altman_plot)
dev.off()
```

Showing the different population characteristics 

```{r}
moire_allele_freq_summary_sel_moire_freq_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "moire_freq") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)

moire_allele_freq_summary_sel_naive_freq_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(moire_allele_freq_summary_sel, "naive_freq") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)
```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list(
    "naive_freq" = moire_allele_freq_summary_sel_naive_freq_by_pop_bland_altman_plot, 
    "moire_freq" = moire_allele_freq_summary_sel_moire_freq_by_pop_bland_altman_plot
       )
))
```

Write to pdf  
```{r}
#| results: hide
cairo_pdf("moire_all_allele_freqs_by_pop_bland_altman_plot.pdf", width = 10, height = 10)
print(moire_allele_freq_summary_sel_moire_freq_by_pop_bland_altman_plot)
print(moire_allele_freq_summary_sel_naive_freq_by_pop_bland_altman_plot)
dev.off()
```



##### Concordance 

Plotting concordance between allele frequencies metrics reported by MOIRE.

Overall 

```{r}
moire_all_summary_allele_freq_measures_plot_moire_all = ggpairs(moire_allele_freq_summary_sel,
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.01)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "all")

```

```{r}
print(moire_all_summary_allele_freq_measures_plot_moire_all)
```


```{r}
moire_all_summary_allele_freq_measures_plot_moire_all_filt = ggpairs(moire_allele_freq_summary_sel %>% 
                                                                       filter(expected_freq > 0.05, 
                                                                              expected_freq < 0.95),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.01)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "all")

```

```{r}
print(moire_all_summary_allele_freq_measures_plot_moire_all_filt)
```


Plotting difference parameter sets  

```{r}
moire_all_summary_allele_freq_measures_plot_moire_paramset1 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset1"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset1")

moire_all_summary_allele_freq_measures_plot_moire_paramset2 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset2"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset2")

moire_all_summary_allele_freq_measures_plot_moire_paramset3 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset3"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset3")

moire_all_summary_allele_freq_measures_plot_moire_paramset4 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset4"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset4")

moire_all_summary_allele_freq_measures_plot_moire_paramset5 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset5"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset5")

moire_all_summary_allele_freq_measures_plot_moire_paramset6 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(paramset_id == "moire_paramset6"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset6")

```

```{r}
#| results: hide
cairo_pdf("moire_allele_freq_pairwise_comparison_parametersets.pdf", width = 15, height = 15)
print(moire_all_summary_allele_freq_measures_plot_moire_paramset1)
print(moire_all_summary_allele_freq_measures_plot_moire_paramset4)

print(moire_all_summary_allele_freq_measures_plot_moire_paramset2)
print(moire_all_summary_allele_freq_measures_plot_moire_paramset5)

print(moire_all_summary_allele_freq_measures_plot_moire_paramset3)
print(moire_all_summary_allele_freq_measures_plot_moire_paramset6)
dev.off()
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("paramset1" = moire_all_summary_allele_freq_measures_plot_moire_paramset1,
                                      "paramset4" = moire_all_summary_allele_freq_measures_plot_moire_paramset4,
                                      "paramset2" = moire_all_summary_allele_freq_measures_plot_moire_paramset2,
                                      "paramset5" = moire_all_summary_allele_freq_measures_plot_moire_paramset5,
                                      "paramset3" = moire_all_summary_allele_freq_measures_plot_moire_paramset3, 
                                      "paramset6" = moire_all_summary_allele_freq_measures_plot_moire_paramset6 
                                      )))
```

Different population metrics 


COI categories 

```{r}
moire_all_summary_allele_freq_measures_plot_moire_MeanCOI1.26 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(coi_cat == "MeanCOI=1.26"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=1.26")

moire_all_summary_allele_freq_measures_plot_moire_MeanCOI3.41 = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(coi_cat == "MeanCOI=3.41"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=3.41")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("MeanCOI_1.26" = moire_all_summary_allele_freq_measures_plot_moire_MeanCOI1.26,
                                      "MeanCOI_3.41" = moire_all_summary_allele_freq_measures_plot_moire_MeanCOI3.41
                                      )))
```


Looking at estimates between 0.05 and 0.95 so calculations aren't as biases for large number of allele frequencies of 1. 

```{r}
moire_all_summary_allele_freq_measures_plot_moire_MeanCOI1.26_filt = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(coi_cat == "MeanCOI=1.26")%>% 
                                                                       filter(expected_freq > 0.05, 
                                                                              expected_freq < 0.95),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=1.26")

moire_all_summary_allele_freq_measures_plot_moire_MeanCOI3.41_filt = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(coi_cat == "MeanCOI=3.41")%>% 
                                                                       filter(expected_freq > 0.05, 
                                                                              expected_freq < 0.95),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=3.41")
```


Especially at high mean population COIs, MOIRE output performs naive estimates of allele frequencies in this range of 0.05-0.95

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("MeanCOI_1.26" = moire_all_summary_allele_freq_measures_plot_moire_MeanCOI1.26_filt,
                                      "MeanCOI_3.41" = moire_all_summary_allele_freq_measures_plot_moire_MeanCOI3.41_filt
                                      )))
```




between sample related categories 

```{r}
moire_all_summary_allele_freq_measures_plot_moire_1perc_related = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(between_sample_related_cat == "_1%_related"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "_1%_related")

moire_all_summary_allele_freq_measures_plot_moire_10perc_related = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(between_sample_related_cat == "10%_related"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "10%_related")

moire_all_summary_allele_freq_measures_plot_moire_30perc_related = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(between_sample_related_cat == "30%_related"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "30%_related")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("_1perc_between_sample_related" = moire_all_summary_allele_freq_measures_plot_moire_1perc_related,
                                      "10perc_between_sample_related" = moire_all_summary_allele_freq_measures_plot_moire_10perc_related, 
                                      "30perc_between_sample_related" = moire_all_summary_allele_freq_measures_plot_moire_30perc_related
                                      )))
```

Recombination categories 
```{r}
moire_all_summary_allele_freq_measures_plot_moire_50perc_non_recomb = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(nonrecombinant_cat == "50%_non-recombinant"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "50%_non-recombinant")

moire_all_summary_allele_freq_measures_plot_moire_20perc_non_recomb = ggpairs(moire_allele_freq_summary_sel %>%
                                                    filter(nonrecombinant_cat == "20%_non-recombinant"),
                                                  columns = c("expected_freq", "moire_freq", "naive_freq"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "20%_non-recombinant")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("20perc_non_recomb" = moire_all_summary_allele_freq_measures_plot_moire_20perc_non_recomb,
                                      "50perc_non_recomb" = moire_all_summary_allele_freq_measures_plot_moire_50perc_non_recomb
                                      )))
```


```{r}
#| results: hide
cairo_pdf("moire_allele_freq_pairwise_comparison_pop_parameters.pdf", width = 15, height = 15)
print(moire_all_summary_allele_freq_measures_plot_moire_MeanCOI1.26)
print(moire_all_summary_allele_freq_measures_plot_moire_MeanCOI3.41)

print(moire_all_summary_allele_freq_measures_plot_moire_1perc_related)
print(moire_all_summary_allele_freq_measures_plot_moire_10perc_related)
print(moire_all_summary_allele_freq_measures_plot_moire_30perc_related)

print(moire_all_summary_allele_freq_measures_plot_moire_20perc_non_recomb)
print(moire_all_summary_allele_freq_measures_plot_moire_50perc_non_recomb)
dev.off()
```



```{r}
# calculate Concordance Correlation Coefficient (CCC) 
moire_allele_freq_summary_sel_ccc_by_pop = moire_allele_freq_summary_sel %>%
  compute_ccc(
    expected_col  = expected_freq,
    observed_cols = c(moire_freq, naive_freq),
    group_cols    = c(population_name, paramset_id)
  )
  
# Gather into long format, left join in the simulation parameters
moire_allele_freq_summary_sel_ccc_by_pop_gat = moire_allele_freq_summary_sel_ccc_by_pop %>%
  select(population_name, paramset_id, ends_with("_ccc")) %>% 
  gather(allele_freq_estimate, ccc, 3:ncol(.)) %>%
  mutate(allele_freq_estimate = gsub("_ccc", "", allele_freq_estimate))%>%
  left_join(sim_params)

```

ccc across all parameter sets for all 4 estimators 

```{r}
moire_allele_freq_summary_sel_ccc_plot_just_estimators = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(y = ccc, x = allele_freq_estimate, fill = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

ggplotly(moire_allele_freq_summary_sel_ccc_plot_just_estimators)
```

ccc over all comparing parameter sets for each estimate

```{r}
moire_allele_freq_summary_sel_ccc_plot_fill_paramset_id = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_id, y = ccc, x = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_plot_fill_allele_freq_estimate = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = allele_freq_estimate, y = ccc, x = paramset_id), position = "dodge"
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfHtmlWidgets(
  list(
  "By Parameter Set" = ggplotly(moire_allele_freq_summary_sel_ccc_plot_fill_paramset_id) %>% layout(boxmode = "group"), 
  "By allele freq Estimator" = ggplotly(moire_allele_freq_summary_sel_ccc_plot_fill_allele_freq_estimate) %>% layout(boxmode = "group")
  )
))

```

ccc on the different simulation population characteristics 

```{r}
moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_by_country = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_coi_cat_by_country = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_between_sample_related_cat = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_nonrecombinant_cat = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_coi_cat = ggplot(moire_allele_freq_summary_sel_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot,
  "est_between_sample_related" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_between_sample_related_cat,
  "coi_cat" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_coi_cat,
  "nonrecombinant" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = moire_allele_freq_summary_sel_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country
)))
```


##### RMSE   

```{r}
# calculate RMSE 
moire_allele_freq_summary_sel_rmse = moire_allele_freq_summary_sel %>%
  compute_rmse(
    expected_col  = expected_freq,
    observed_cols = c(moire_freq, naive_freq),
    group_cols    = c(population_name, paramset_id)
  )

# Gather into long format, left join in the simulation parameters
moire_allele_freq_summary_sel_rmse_gat = moire_allele_freq_summary_sel_rmse %>%
  gather(allele_freq_estimate, rmse, 3:ncol(.)) %>%
  mutate(allele_freq_estimate = gsub("_rmse", "", allele_freq_estimate))%>%
  left_join(sim_params)

```

RMSE across all parameter sets for all 4 estimators 

```{r}
moire_allele_freq_summary_sel_rmse_plot_just_estimators = ggplot(moire_allele_freq_summary_sel_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(y = rmse, x = allele_freq_estimate, fill = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

ggplotly(moire_allele_freq_summary_sel_rmse_plot_just_estimators)
```

RMSE over all comparing parameter sets for each estimate

```{r}
moire_allele_freq_summary_sel_rmse_plot_fill_paramset_id = ggplot(moire_allele_freq_summary_sel_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_id, y = rmse, x = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

moire_allele_freq_summary_sel_rmse_plot_fill_allele_freq_estimate = ggplot(moire_allele_freq_summary_sel_rmse_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = allele_freq_estimate, y = rmse, x = paramset_id), position = "dodge"
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfHtmlWidgets(
  list(
  "By Parameter Set" = ggplotly(moire_allele_freq_summary_sel_rmse_plot_fill_paramset_id) %>% layout(boxmode = "group"), 
  "By COI Estimator" = ggplotly(moire_allele_freq_summary_sel_rmse_plot_fill_allele_freq_estimate) %>% layout(boxmode = "group")
  )
))

```

RMSE on the different simulation population characteristics 

```{r}
moire_allele_freq_summary_sel_rmse_gat_large_plot_by_country = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_allele_freq_summary_sel_rmse_gat_large_plot_between_sample_related_cat_by_country = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_allele_freq_summary_sel_rmse_gat_large_plot_nonrecombinant_cat_by_country = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_allele_freq_summary_sel_rmse_gat_large_plot_coi_cat_by_country = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

moire_allele_freq_summary_sel_rmse_gat_large_plot = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate

moire_allele_freq_summary_sel_rmse_gat_large_plot_between_sample_related_cat = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate

moire_allele_freq_summary_sel_rmse_gat_large_plot_nonrecombinant_cat = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate

moire_allele_freq_summary_sel_rmse_gat_large_plot_coi_cat = ggplot(moire_allele_freq_summary_sel_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = moire_allele_freq_summary_sel_rmse_gat_large_plot,
  "est_between_sample_related" = moire_allele_freq_summary_sel_rmse_gat_large_plot_between_sample_related_cat,
  "coi_cat" = moire_allele_freq_summary_sel_rmse_gat_large_plot_coi_cat,
  "nonrecombinant" = moire_allele_freq_summary_sel_rmse_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = moire_allele_freq_summary_sel_rmse_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = moire_allele_freq_summary_sel_rmse_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = moire_allele_freq_summary_sel_rmse_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = moire_allele_freq_summary_sel_rmse_gat_large_plot_nonrecombinant_cat_by_country
)))
```



##### False positive and false negatives

```{r}
moire_allele_freq_summary_sel_contignency = moire_allele_freq_summary_sel %>% 
  select(population_name, paramset_id, target_id, seq, expected_freq, moire_freq, naive_freq) %>% 
  gather(estimator, freq, 6:7) %>% 
  mutate(estimator = gsub("_freq", "", estimator))%>% 
  mutate(estimator = gsub("naive", "mad4hatter", estimator))

# moire appears to have some allele freq estimates that are extremely low even for some expected frequencies close to 1000, to count these 
moire_allele_freq_summary_sel_contignency = moire_allele_freq_summary_sel_contignency %>% 
  mutate(freq = ifelse(estimator == "moire" & expected_freq/freq > 1000 & expected_freq > 0.05, 0, freq))

# True Positive is both estimator and expected is great than 0 
# False Negative is that expected > 0 and estimator is 0 
# False Positive is that expected is 0 and estimator is > 0
# (could also consider like above if the frequenices are extremely off from each other, than make it as False Positive or false negative)

moire_allele_freq_summary_sel_contignency = moire_allele_freq_summary_sel_contignency %>% 
  mutate(TP = as.numeric(expected_freq > 0 & freq > 0), 
         FN = as.numeric(expected_freq > 0 & freq == 0), 
         FP = as.numeric(expected_freq == 0 & freq > 0))

```


Has the same Haplotype performance as input data, no deviations from the data that it works with. 

```{r}
moire_allele_freq_summary_sel_contignency_overall = moire_allele_freq_summary_sel_contignency %>% 
  group_by(estimator) %>%
  add_contigency_metrics_no_TN(summarize_first = T)

create_dt(moire_allele_freq_summary_sel_contignency_overall)

hap_metric_map <- c(
  "Recall"               = "recall",
  "Precision"            = "precision",
  "NPV"                  = "tpr",
  "F1-score"             = "f1_score",
  "False Discovery Rate" = "fdr",
  "Overlap Coefficient"  = "overlap_coefficient"
)

radar_by_metrics(moire_allele_freq_summary_sel_contignency_overall,  metric_map = hap_metric_map, color_col = "estimator", palette = colorPalette_08)
```


### he

MOIRE also reports on he for each target 

```{r}
moire_allele_freq_summary_sel_hes = moire_allele_freq_summary_sel %>% 
  group_by(population_name, paramset_id, target_id) %>% 
  summarise(
    expected_he = 1 - sum(expected_freq^2), 
    naive_he    = 1 - sum(naive_freq^2), 
    moire_recalc_he = 1 - sum(moire_freq^2)
  ) %>% 
  full_join(moire_he_summary %>% 
              select(population_name, paramset_id, target_id, he) %>% 
              dplyr::rename(moire_he = he)) %>% 
  left_join(sim_params)

```



##### Concordance 

Plotting concordance between all he reported. The moire_recalc_he is the allele frequencies re-calculated (1 - sum(moire_freq^2))

Overall 

```{r}
moire_all_summary_allele_freq_measures_plot_moire_all = ggpairs(moire_allele_freq_summary_sel_hes,
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "all")

```

```{r}
print(moire_all_summary_allele_freq_measures_plot_moire_all)
```


Plotting difference parameter sets  

```{r}
moire_all_summary_he_measures_plot_moire_paramset1 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset1"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset1")

moire_all_summary_he_measures_plot_moire_paramset2 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset2"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset2")

moire_all_summary_he_measures_plot_moire_paramset3 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset3"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset3")

moire_all_summary_he_measures_plot_moire_paramset4 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset4"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset4")

moire_all_summary_he_measures_plot_moire_paramset5 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset5"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset5")

moire_all_summary_he_measures_plot_moire_paramset6 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(paramset_id == "moire_paramset6"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "moire_paramset6")

```

```{r}
#| results: hide
cairo_pdf("moire_he_pairwise_comparison_parametersets.pdf", width = 15, height = 15)
print(moire_all_summary_he_measures_plot_moire_paramset1)
print(moire_all_summary_he_measures_plot_moire_paramset4)

print(moire_all_summary_he_measures_plot_moire_paramset2)
print(moire_all_summary_he_measures_plot_moire_paramset5)

print(moire_all_summary_he_measures_plot_moire_paramset3)
print(moire_all_summary_he_measures_plot_moire_paramset6)
dev.off()
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("paramset1" = moire_all_summary_he_measures_plot_moire_paramset1,
                                      "paramset4" = moire_all_summary_he_measures_plot_moire_paramset4,
                                      "paramset2" = moire_all_summary_he_measures_plot_moire_paramset2,
                                      "paramset5" = moire_all_summary_he_measures_plot_moire_paramset5,
                                      "paramset3" = moire_all_summary_he_measures_plot_moire_paramset3, 
                                      "paramset6" = moire_all_summary_he_measures_plot_moire_paramset6 
                                      )))
```

Different population metrics 


COI categories 

```{r}
moire_all_summary_he_measures_plot_moire_MeanCOI1.26 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(coi_cat == "MeanCOI=1.26"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=1.26")

moire_all_summary_he_measures_plot_moire_MeanCOI3.41 = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(coi_cat == "MeanCOI=3.41"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=3.41")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("MeanCOI_1.26" = moire_all_summary_he_measures_plot_moire_MeanCOI1.26,
                                      "MeanCOI_3.41" = moire_all_summary_he_measures_plot_moire_MeanCOI3.41
                                      )))
```


between sample related categories 

```{r}
moire_all_summary_he_measures_plot_moire_1perc_related = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(between_sample_related_cat == "_1%_related"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "_1%_related")

moire_all_summary_he_measures_plot_moire_10perc_related = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(between_sample_related_cat == "10%_related"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "10%_related")

moire_all_summary_he_measures_plot_moire_30perc_related = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(between_sample_related_cat == "30%_related"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "30%_related")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("_1perc_between_sample_related" = moire_all_summary_he_measures_plot_moire_1perc_related,
                                      "10perc_between_sample_related" = moire_all_summary_he_measures_plot_moire_10perc_related, 
                                      "30perc_between_sample_related" = moire_all_summary_he_measures_plot_moire_30perc_related
                                      )))
```

Recombination categories 
```{r}
moire_all_summary_he_measures_plot_moire_50perc_non_recomb = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(nonrecombinant_cat == "50%_non-recombinant"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "50%_non-recombinant")

moire_all_summary_he_measures_plot_moire_20perc_non_recomb = ggpairs(moire_allele_freq_summary_sel_hes %>%
                                                    filter(nonrecombinant_cat == "20%_non-recombinant"),
                                                  columns = c("expected_he", "naive_he", "moire_recalc_he", "moire_he"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.05)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "20%_non-recombinant")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("20perc_non_recomb" = moire_all_summary_he_measures_plot_moire_20perc_non_recomb,
                                      "50perc_non_recomb" = moire_all_summary_he_measures_plot_moire_50perc_non_recomb
                                      )))
```


```{r}
#| results: hide
cairo_pdf("moire_he_pairwise_comparison_pop_parameters.pdf", width = 15, height = 15)
print(moire_all_summary_he_measures_plot_moire_MeanCOI1.26)
print(moire_all_summary_he_measures_plot_moire_MeanCOI3.41)

print(moire_all_summary_he_measures_plot_moire_1perc_related)
print(moire_all_summary_he_measures_plot_moire_10perc_related)
print(moire_all_summary_he_measures_plot_moire_30perc_related)

print(moire_all_summary_he_measures_plot_moire_20perc_non_recomb)
print(moire_all_summary_he_measures_plot_moire_50perc_non_recomb)
dev.off()
```



```{r}
# calculate Concordance Correlation Coefficient (CCC) 
moire_he_summary_sel_hes_ccc_by_pop = moire_allele_freq_summary_sel_hes %>%
  compute_ccc(
    expected_col = expected_he, 
    observed_cols = c(moire_he, moire_recalc_he, naive_he), 
    group_cols = c(population_name, paramset_id)
  )
# Gather into long format, left join in the simulation parameters
moire_he_summary_sel_hes_ccc_by_pop_gat = moire_he_summary_sel_hes_ccc_by_pop %>%
  select(population_name, paramset_id, ends_with("_ccc")) %>% 
  gather(allele_freq_estimate, ccc, 3:ncol(.)) %>%
  mutate(allele_freq_estimate = gsub("_ccc", "", allele_freq_estimate))%>%
  left_join(sim_params)

```

ccc across all parameter sets for all 4 estimators 

```{r}
moire_allele_freq_summary_sel_hes_ccc_plot_just_estimators = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(y = ccc, x = allele_freq_estimate, fill = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

ggplotly(moire_allele_freq_summary_sel_hes_ccc_plot_just_estimators)
```

ccc over all comparing parameter sets for each estimate

```{r}
moire_allele_freq_summary_sel_hes_ccc_plot_fill_paramset_id = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_id, y = ccc, x = allele_freq_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_allele_freq_summary_sel_hes_ccc_plot_fill_allele_freq_estimate = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat %>%
                                                            mutate(paramset_id =
                                                                     factor(paramset_id,
                                                                            level = c("moire_paramset1","moire_paramset4",
                                                                                      "moire_paramset2","moire_paramset5",
                                                                                      "moire_paramset3","moire_paramset6"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = allele_freq_estimate, y = ccc, x = paramset_id), position = "dodge"
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfHtmlWidgets(
  list(
  "By Parameter Set" = ggplotly(moire_allele_freq_summary_sel_hes_ccc_plot_fill_paramset_id) %>% layout(boxmode = "group"), 
  "By allele freq Estimator" = ggplotly(moire_allele_freq_summary_sel_hes_ccc_plot_fill_allele_freq_estimate) %>% layout(boxmode = "group")
  )
))

```

ccc on the different simulation population characteristics 

```{r}
moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_by_country = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_coi_cat_by_country = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_between_sample_related_cat = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_nonrecombinant_cat = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_coi_cat = ggplot(moire_he_summary_sel_hes_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc)
  ) +
  facet_grid(paramset_id~allele_freq_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot,
  "est_between_sample_related" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_between_sample_related_cat,
  "coi_cat" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_coi_cat,
  "nonrecombinant" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = moire_he_summary_sel_hes_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country
)))
```





