---
title: "MalHaploFreq About"
output: html_document
author: "Nicholas Hathaway"
date: "2023-12-15"
bibliography: reference.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose

This tool was created to estimate prevalence of haplotypes (linking up to 3 SNPs together) from SNP loci. This tool requires input determined MOI and does 3 loci. Since the creation of this tool other tools have been created that do similar analysis with additional features, for example [MultiLociBiallelicModel](../MultiLociBiallelicModel/MultiLociBiallelicModel_background.qmd) also estimates prevalence but does not require MOI, can do more loci and estimates the MOI from the input data. 

## Citation

[@Hastings2008-mb]
