---
title: "SNP-slice analysis using snp.slicer"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

## Introduction

For this tutorial, we will be using the `snp.slicer` package to estimate the complexity of infection (COI) and infer latent haplotypes from SNP data. Instructions to install this package may be found [here](installation.html)

## The data

`snp.slicer` expects a `data.frame` of SNPs with columns `specimen_id`, `target_id`, `target_value`, and `target_count` representing the specimen identifier, target identifier, target value, and target read count, respectively. We will demonstrate using this package with the simulated SNP barcode data described [here](../../website_docs/data_description.html#simulated-1). We will load the data using the `vcfR` package.

```{r, waning=FALSE, message=FALSE, error=FALSE}
library(tidyverse)
library(here)
library(vcfR)
library(snp.slicer)

vcf_data <- read.vcfR(
  here::here("data/snp_barcode/SpotMalariapfPanel_simData_sanger100.vcf.gz")
)

counts_raw <- t(vcfR::extract.gt(vcf_data, element = "AD"))

read0 <- vcfR::masplit(
  counts_raw, record = 1, sort = FALSE, decreasing = FALSE
) |>
  as.data.frame() |>
  rownames_to_column(var = "specimen_id") |>
  tidyr::pivot_longer(
    cols = -specimen_id,
    names_to = "target_id",
    values_to = "target_count"
  ) |>
  # fill in target value with ref allele
  mutate(target_value = "0")

read1 <- vcfR::masplit(
  counts_raw, record = 2, sort = FALSE, decreasing = FALSE
) |>
  as.data.frame() |>
  rownames_to_column(var = "specimen_id") |>
  tidyr::pivot_longer(
    cols = -specimen_id,
    names_to = "target_id",
    values_to = "target_count"
  ) |>
  # fill in target value with alt allele
  mutate(target_value = "1")

snpslice_data <- rbind(read0, read1) |>
  arrange(specimen_id, target_id)

# run snp.slicer
run <- snp.slicer::snp_slice(
  snpslice_data,
  model = "negative_binomial",
  # set n_mcmc to 100 for quick demo
  n_mcmc = 100,
  store_mcmc = TRUE,
  verbose = FALSE
)

summary(run)
```

`snp.slicer` exposes functionality to calculate multi locus allele frequencies using the posterior distribution of phased strain assignments. Here we will calculate the distribution for the collection `c("Pf3D7_01_v3_145515", "Pf3D7_01_v3_179347", "Pf3D7_01_v3_180554")`

```{r}
af <- snp.slicer::calculate_allele_frequencies(
  run, c("Pf3D7_01_v3_145515", "Pf3D7_01_v3_179347", "Pf3D7_01_v3_180554")
)

af
```

We can also summarize the distribution of MOI using the strain allocation matrix
```{r}
moi_dist <- data.frame(
  specimen_id = run$model_info$processed_data$specimen_ids,
  moi = run$allocation_matrix |> rowSums()
)

ggplot(moi_dist) +
  geom_histogram(aes(x = moi))
```

