---
title: "SNP-slice"
output: html_document
author: "Shazia Ruybal-Pesántez"
date: "2025-07-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose

The `SNP-slice` algorithm enables reconstruction of SNP haplotypes from bi-allelic data, allowing for inference of haplotype/strain frequencies.

## Existing resources

The [`SNP-slice` GitHub repository](https://github.com/nianqiaoju/snp-slice) includes scripts for implementing the SNP-slice algorithm, as well as scripts for analyzing output and generating figures. 

## Citation

The publication associated with the `SNP-slice` algorithm and analysis scripts can be found [here (Ju et al 2024 Bioinformatics)](https://doi.org/10.1093/bioinformatics/btae344).