---
title: Sample Benchmarking for Allele Frequencies
author: Sophie Berube
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=F, message=F)
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)
```

<<<<<<< HEAD
```{r nickFunctions}
library(dplyr)
library(ggplot2)
library(jsonlite)
library(magrittr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(tidyr)
library(flextable)
library(DescTools)
library(hexbin)
#| code-fold: true

# commonly used functions  
`%!in%` <- Negate(`%in%`)
is.notna <-function(x){
  return(!is.na(x))
}
# ggplot themes 
transparentBackground = theme(
  panel.background = element_rect(fill='transparent'), #transparent panel bg
  plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
  panel.grid.major = element_blank(), #remove major gridlines
  panel.grid.minor = element_blank(), #remove minor gridlines
  legend.background = element_rect(color = NA, fill='transparent'), #transparent legend bg
  legend.box.background = element_rect(color = NA, fill='transparent') #transparent legend panel
)
custom_theme_noTransparentBackground = theme_bw() +
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank() )+
  theme(axis.line.x = element_line(color="black", linewidth = 0.3),axis.line.y =
          element_line(color="black", linewidth = 0.3))+
  theme(text=element_text(size=12, family="Helvetica"))+
  theme(axis.text.y = element_text(size=12))+
  theme(axis.text.x = element_text(size=12)) +
  theme(legend.position = "bottom") + 
  theme(plot.title = element_text(hjust = 0.5))

custom_theme = custom_theme_noTransparentBackground +
  transparentBackground
custom_theme_backgroundTransparent = custom_theme

custom_theme_xRotate_noBackgroundTransparent = theme_bw() +
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank() )+
  theme(axis.line.x = element_line(color="black", linewidth = 0.3),axis.line.y =
          element_line(color="black", linewidth = 0.3))+
  theme(text=element_text(size=12, family="Helvetica"))+
  theme(axis.text.y = element_text(size=12))+
  theme(axis.text.x = element_text(size=12)) +
  theme(legend.position = "bottom") + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme(axis.text.x = element_text(size=12, angle = -90, vjust = 0.5, hjust = 0)) 

custom_theme_xRotate = custom_theme_xRotate_noBackgroundTransparent + 
  transparentBackground

custom_theme_xRotate_backgroundTransparent = custom_theme_xRotate


create_dt <- function(x){
  DT::datatable(x,
                extensions = 'Buttons',
                options = list(dom = 'Blfrtip',
                               buttons = c('copy', 'csv', 'excel', 'pdf', 'print'),
                               lengthMenu = list(c(10,25,50,-1),
                                                 c(10,25,50,"All"))), 
                filter = "top")
}

colorPalette_08 = c("#2271B2","#F748A5","#359B73","#F0E442","#D55E00","#3DB7E9","#E69F00","#000000")
colorPalette_12 = c("#E20134","#FF6E3A","#008DF9","#8400CD","#FFC33B","#9F0162","#009F81","#FF5AAF","#00FCCF","#00C2F9","#FFB2FD","#A40122")
colorPalette_15 = c("#F60239","#003C86","#EF0096","#9400E6","#009FFA","#008169","#68023F","#00DCB5","#FFCFE2","#FF71FD","#7CFFFA","#6A0213","#008607","#00E307","#FFDC3D")


create_tabsetOfHtmlWidgets <- function(htmlObjectsList) {
  zz <- textConnection("foo", "w")
  sink(zz)
  cat("::: {.panel-tabset}\n")
  iwalk(htmlObjectsList, ~ {
    cat('## ', .y, '\n\n')
    tempList = list()
    tempList[["item"]] = .x
    print(htmltools::tagList(tempList))
    cat('\n\n')
  })
  cat(":::\n")
  sink()
  close(zz)
  paste0(foo, collapse = "\n")
}


create_tabsetOfGgplotObjects <- function(ggplotObjectsList) {
  # zz <- textConnection("foo", "w")
  # sink(zz)
  cat("\n\n::: {.panel-tabset}\n\n")
  iwalk(ggplotObjectsList, ~ {
    cat('## ', .y, '\n\n')
    print(.x)
    # tempList = list()
    # tempList[["item"]] = .x
    # grid::grid.draw(ggplot2::ggplotGrob(.x))
    cat('\n\n')
  })
  cat(":::\n")
  # sink()
  # close(zz)
  # paste0(foo, collapse = "\n")
}

```



```{r, warning=FALSE, message=FALSE}


inp <- list(
  sim_info_dir = "~/Library/CloudStorage/OneDrive-UniversityofFlorida/Desktop/Research/General_malaria_genomics/PGEForge_benchmarking/SpotMalaria", 
  tool_res_dir = "~/Library/CloudStorage/OneDrive-UniversityofFlorida/Desktop/Research/General_malaria_genomics/PGEForge_benchmarking/naive_mlafp/", 
  tool_params = "~/Library/CloudStorage/OneDrive-UniversityofFlorida/Desktop/Research/General_malaria_genomics/PGEForge_benchmarking/naive_mlafp.json"
=======
```{r, warning=FALSE, message=FALSE}
source("../../scripts/benchmarking_functions.R")
```

```{r, warning=FALSE, message=FALSE}
inp <- list(
  sim_info_dir = "../../../data/benchmarking/simulations/SpotMalaria",
  tool_res_dir = "../../../data/benchmarking/tool_results/naive_mlafp/",
  madhatter_res_dir = "../../../data/benchmarking/tool_results/madhatter/",
  tool_params = "../../../data/benchmarking/tool_results/naive_mlafp.json"
>>>>>>> e7e90d474c6072704988e3ccd452da6363874fa1
)
```

```{r}
# Read all expected frequencies into one tibble ------------------------
sim_truth <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_aa_multiloci_variants_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_integer(), 
        population_name = col_character(), 
        country = col_character(), 
        variant = col_character(), 
        freq = col_double(), 
        prev = col_double(), 
        group_id = col_character()
      ), 
      progress = FALSE
    )
  ) %>%
  dplyr::select(-sim_truth_path) %>%
  unnest(sim_truth) %>%
  rename(
    sim_id = population_name, 
    exp_freq = freq, 
    exp_prev = prev, 
    exp_n_samp = sample_count
  ) %>%
  dplyr::select(
    -pop, 
    -country, 
    -aas_in_variant, 
    -allele_count, 
    -total_samples, 
    -total_alleles
  )

# Read in all simulation parameters into one tibble --------------------
sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  dplyr::select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>%
  dplyr::select(
    -collection_year, 
    -obs_freq_monoclonal, 
    -obs_non_recombinant, 
    -mean_k, 
    -sample_count, 
    -est_between_sample_relatedness
  ) %>%
  rename(
    sim_id = population_name, 
    true_mean_coi = mean_coi
  )
sim_truth <- sim_truth %>%
  left_join(sim_params, by = "sim_id")

# Read in all FEM results into one tibble ------------------------------
tool_results <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "*.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(sim_id = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  dplyr::select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(), 
        variant = col_character(), 
        group_id = col_character(), 
        sample_total= col_character()
      ), 
      progress = FALSE
    )
  ) %>%
  dplyr::select(-tool_res_path) %>%
  unnest(tool_res) %>%
  rename(
    obs_freq = freq, 
    obs_prev = prev,
    obs_n_samp = sample_total
  )

# Read in tool parameters, and join ------------------------------------
tool_params <- jsonlite::fromJSON(inp$tool_params) %>%
  as_tibble()
  #mutate(mean_coi_tool_param = paramset$coi) %>%
  #dplyr::select(-loci_groups, -paramset)
tool_results <- tool_results %>%
  left_join(tool_params, by = "paramset_id")

# Join everything into one tibble --------------------------------------
tool_results_wtruth <- tool_results %>%
  # Use an inner join because FEM also estimates frequencies for 
  # subsets of the requested multilocus genotypes, and these 
  # frequencies are not benchmarked
  inner_join(sim_truth, by = c("sim_id", "variant", "group_id"))

sim_truth<- sim_truth|>
  mutate(exp_n_denom= exp_n_samp/exp_prev)

tool_results<- tool_results|>
  rename(obs_n_denom= obs_n_samp)

tool_results_wtruth|>
  mutate(exp_n_denom = exp_n_samp/exp_prev)|>
  rename(obs_n_denom= obs_n_samp)->tool_results_wtruth 

tool_results_wtruth$obs_n_denom<- as.numeric(tool_results_wtruth$obs_n_denom)

```

# Allele Frequencies 

## Goals for allele frequencies: 


* Report Lin's concordance correlation coefficient with 95% confidence intervals between tool and simulation COIs across different groupings of populations, and parameters. 

* Compare tool allele frequencies to simulated allele frequencies using Bland-Altman plots (mean of tool and simulation vs difference of tool vs simulation) across combinations of loci, and simulation parameters. 
  * Can choose to split this up across different groupings of populations, and simulation parameters. 
  
* For each allele in each locus, report number of samples that contain a false positive (allele detected in sample with tool but not truly present in simulation), and that contain false negatives (allele not detected in sample with tool but is truly present in simulation). Report these as proportions (sensitivity, specificity). 
  
```{r}

# Compute RMSE of frequency and prevalence -----------------------------
tool_results_with_rmse <- tool_results_wtruth %>%
  group_by(
    sim_id, 
    paramset_id, 
    country, 
    group_id, 
    pop_alpha, 
    coi_p, 
    coi_r, 
    k_s
  ) %>%
  summarize(
    freq_rmse = sqrt(mean((obs_freq - exp_freq)^2)), 
    .groups = "drop"
  )

# Show all RMSEs as a boxplot ------------------------------------------
tool_results_with_rmse <- tool_results_with_rmse %>%
  mutate(
    group_id = factor(
      group_id, 
      c(
        "six_med_div", 
        "six_low_div", 
        "four_group1", 
        "two_group1", 
        "pfdhfr_pfdhps", 
        "pfdhps", 
        "pfdhfr"
      ), 
      labels = c(
        "Six, Medium Diversity", 
        "Six, Low Diversity", 
        "Four", 
        "Two", 
        "pfdhfr/pfdhps", 
        "pfdhps", 
        "pfdhfr"
      )
    )
  )%>%
  mutate(
    pop_alpha = factor(
      pop_alpha, 
      c("2.33", "9", "99"), 
      labels = c("High Sim. Rel", "Med. Sim. Rel", "Low Sim. Rel")
    )
  ) %>%
  mutate(
    k_s = factor(
      k_s, 
      c("0.5", "0.2"), 
      labels = c("Med. Recombination", "High Recombination")
    )
  ) %>%
  # coi_p was varied in synchronicity with coi_r to produce different 
  # polyclonality levels, and thus coi_p is implicitly included when 
  # using coi_r as a plotting aesthetic
  mutate(
    coi_r = factor(
      coi_r, 
      c("0.25", "0.5"), 
      labels = c("Low COI/Prop. Polyclonal", "Med. COI/Prop. Polyclonal")
    )
  )

tool_results_with_rmse %>%
  ggplot(mapping = aes(x = pop_alpha, y = freq_rmse, color = group_id)) +
  geom_boxplot() +theme_bw()+
  facet_grid(rows = vars(k_s), cols = vars(coi_r)) +
  labs(
    x = "Simulated Relatedness", 
    y = "RMSE of Allele Frequency Estimate", 
    color = "Loci Set"
  ) +
  theme(legend.position = "bottom")-> RMSE_byScenario_freq

#print(RMSE_byScenario_freq)



```


## Concordance 

```{r}

tool_results_Concordance_Freq <- tool_results_wtruth %>%
  group_by(
    group_id, 
    pop_alpha, 
    coi_r, 
    k_s
  )|>
  summarize(conc_pt_est= CCC(exp_freq, obs_freq)$rho.c$est, 
         conc_lwr_CI= CCC(exp_freq, obs_freq)$rho.c$lwr.ci, 
         conc_upr_CI= CCC(exp_freq, obs_freq)$rho.c$upr.ci)|>
  mutate(
    group_id = factor(
      group_id, 
      c(
        "six_med_div", 
        "six_low_div", 
        "four_group1", 
        "two_group1", 
        "pfdhfr_pfdhps", 
        "pfdhps", 
        "pfdhfr"
      ), 
      labels = c(
        "Six, Medium Diversity", 
        "Six, Low Diversity", 
        "Four", 
        "Two", 
        "pfdhfr/pfdhps", 
        "pfdhps", 
        "pfdhfr"
      )
    )
  )|>
  mutate(
    pop_alpha = factor(
      pop_alpha, 
      c("2.33", "9", "99"), 
      labels = c("High Sim. Rel", "Med. Sim. Rel", "Low Sim. Rel")
    )
  ) %>%
  mutate(
    k_s = factor(
      k_s, 
      c("0.5", "0.2"), 
      labels = c("Med. Recombination", "High Recombination")
    )
  ) %>%
  # coi_p was varied in synchronicity with coi_r to produce different 
  # polyclonality levels, and thus coi_p is implicitly included when 
  # using coi_r as a plotting aesthetic
  mutate(
    coi_r = factor(
      coi_r, 
      c("0.25", "0.5"), 
      labels = c("Low COI/Prop. Polyclonal", "Med. COI/Prop. Polyclonal")
    )
  )




tool_results_Concordance_Freq %>%
  ggplot(mapping = aes(x = pop_alpha, y = conc_pt_est, ymin= conc_lwr_CI, ymax= conc_upr_CI, color = group_id)) +
  geom_pointrange(position = position_jitter(width = 0.2)) +theme_bw()+
  facet_grid(rows = vars(k_s), cols = vars(coi_r)) +
  labs(
    x = "Simulated Relatedness", 
    y = "Concordance Correlation Coeffifient", 
    color = "Loci Set"
  ) + geom_hline(yintercept= 0.95, color='red', lty=2)+
  theme(legend.position = "bottom")-> Concordance_byScenario_prev
```

```{r fig.cap= "Figure 1: Lin's concordance correlation coefficient (with 95%CI) between simulated frequencies and tool estimated frequencies. Estimates compared across simulation scenarios (level of recombination, COI, and relatedness), and marker groups."}
print(Concordance_byScenario_prev)

```

## Bland-Altman Plots 

```{r}

tool_results_wtruth_BA<- tool_results_wtruth|>
  mutate(mean_freq= (obs_freq+exp_freq)/2, 
         dif_freq= obs_freq-exp_freq)|>
   mutate(
    group_id = factor(
      group_id, 
      c(
        "six_med_div", 
        "six_low_div", 
        "four_group1", 
        "two_group1", 
        "pfdhfr_pfdhps", 
        "pfdhps", 
        "pfdhfr"
      ), 
      labels = c(
        "Six, Medium Diversity", 
        "Six, Low Diversity", 
        "Four", 
        "Two", 
        "pfdhfr/pfdhps", 
        "pfdhps", 
        "pfdhfr"
      )
    )
  )|>
  mutate(
    pop_alpha = factor(
      pop_alpha, 
      c("2.33", "9", "99"), 
      labels = c("High Sim. Rel", "Med. Sim. Rel", "Low Sim. Rel")
    )
  ) %>%
  mutate(
    k_s = factor(
      k_s, 
      c("0.5", "0.2"), 
      labels = c("Med. Recombination", "High Recombination")
    )
  ) %>%
  # coi_p was varied in synchronicity with coi_r to produce different 
  # polyclonality levels, and thus coi_p is implicitly included when 
  # using coi_r as a plotting aesthetic
  mutate(
    coi_r = factor(
      coi_r, 
      c("0.25", "0.5"), 
      labels = c("Low COI/Prop. Polyclonal", "Med. COI/Prop. Polyclonal")
    )
  )


LocusSets<- unique(tool_results_wtruth_BA$group_id)

BA_plots<- vector(mode='list', length=length(LocusSets))

for(i in 1:length(LocusSets)){

tool_results_wtruth_BA %>%
    filter(group_id==LocusSets[i])|>
  ggplot() +
  geom_hex(
    aes(x = mean_freq, y = dif_freq, fill = log10(after_stat(count))))+
  scale_fill_viridis_c(name = "log10(count)") +
  theme_minimal() + 
  geom_smooth(aes(x = mean_freq, y = dif_freq), method='loess', se=F) + 
  facet_grid(rows = vars(k_s), cols = vars(coi_r)) +
  labs(
    x = "Mean of Observed and Expected Freq", 
    y = "Difference of Observed and Expected Freq"
  ) +ggtitle(LocusSets[i])-> BA_byScenario_prev
  
  BA_plots[[i]]<- BA_byScenario_prev

}



```





```{r, fig.cap="Figure 2: Mean vs Difference of tool estimated and true alelle frequencies with LOESS line (in blue). Estimates compared across simulation scenarios (level of recombination and COI), and marker groups."}

cat(create_tabsetOfGgplotObjects(
  list(
    "Six, Medium Diversity" = BA_plots[[1]], 
    "Six, Low Diversity" = BA_plots[[2]],
    "Four" = BA_plots[[3]], 
    "Two" = BA_plots[[4]], 
    "pfdhfr/pfdhps" = BA_plots[[5]], 
    "pfdhfr" = BA_plots[[6]], 
    "pfdhps" = BA_plots[[7]])
))
```

## False Positive and False Negative Summaries 

```{r}
tool_results_wtruth %>%
  group_by(
    sim_id, 
    paramset_id, 
    country, 
    group_id, 
    variant,
    pop_alpha, 
    coi_p, 
    coi_r, 
    k_s
  ) %>%
  summarize(
    #this dataset only had false negatives, no false positives
    #if you have false positives you would add them here 
    exp_n_denom= exp_n_denom, 
    obs_n_denom = obs_n_denom,
    false_neg= round(exp_n_denom-obs_n_denom,5),
    sensitivity= round(obs_n_denom/exp_n_denom, 5),
    exp_freq= exp_freq,
    .groups = "drop"
  )|>
   mutate(
    group_id = factor(
      group_id, 
      c(
        "six_med_div", 
        "six_low_div", 
        "four_group1", 
        "two_group1", 
        "pfdhfr_pfdhps", 
        "pfdhps", 
        "pfdhfr"
      ), 
      labels = c(
        "Six, Medium Diversity", 
        "Six, Low Diversity", 
        "Four", 
        "Two", 
        "pfdhfr/pfdhps", 
        "pfdhps", 
        "pfdhfr"
      )
    )
  )|>
  mutate(
    pop_alpha = factor(
      pop_alpha, 
      c("2.33", "9", "99"), 
      labels = c("High Sim. Rel", "Med. Sim. Rel", "Low Sim. Rel")
    )
  ) %>%
  mutate(
    k_s = factor(
      k_s, 
      c("0.5", "0.2"), 
      labels = c("Med. Recombination", "High Recombination")
    )
  ) %>%
  # coi_p was varied in synchronicity with coi_r to produce different 
  # polyclonality levels, and thus coi_p is implicitly included when 
  # using coi_r as a plotting aesthetic
  mutate(
    coi_r = factor(
      coi_r, 
      c("0.25", "0.5"), 
      labels = c("Low COI/Prop. Polyclonal", "Med. COI/Prop. Polyclonal")
    )
  )-> False_negs_byScenario_and_freq



False_negs_byScenario_and_freq %>%
  ggplot(mapping = aes(x = pop_alpha, y = false_neg, color = group_id)) +
  geom_boxplot() +theme_bw()+
  facet_grid(rows = vars(k_s), cols = vars(coi_r)) +
  labs(
    x = "Simulated Relatedness", 
    y = "Number of Samples with False Negative Alleles", 
    color = "Loci Set"
  ) +
  theme(legend.position = "bottom")-> FalseNeg_byScenario
```

```{r, fig.cap="Figure 3: Proportion of samples with false negative detection using tool. Proportions compared across simulation scenarios (level of recombination, COI, and relatedness) and marker groups."}
print(FalseNeg_byScenario)
```

```{r}
False_negs_byScenario_and_freq %>%
  ggplot(mapping = aes(x = pop_alpha, y = sensitivity, color = group_id)) +
  geom_boxplot() +theme_bw()+
  facet_grid(rows = vars(k_s), cols = vars(coi_r)) +
  labs(
    x = "Simulated Relatedness", 
    y = "Sensitivity, Proportion of MadHatter Samples \nwith True Postive Detection", 
    color = "Loci Set"
  ) +
  theme(legend.position = "bottom")-> Sensitivity_byScenario
```

```{r, fig.cap="Figure 4: Proportion of samples with true positive detection using tool. Proportions compared across simulation scenarios (level of recombination, COI, and relatedness) and marker groups."}
print(Sensitivity_byScenario)
```

```{r}
#replicate this plot as necessary for false positives

False_negs_byScenario_and_freq %>%
  mutate(Freq_Quart= ifelse(exp_freq<0.25, "(0,0.25)", 
                            ifelse(exp_freq>=0.25&exp_freq<0.5, "(0.25,0.5)", ifelse(exp_freq>=0.5&exp_freq<0.75, "(0.5,0.75)", "(0.75,1)"))))|>
  ggplot(mapping = aes(x = Freq_Quart, y = false_neg, color = group_id)) +
  geom_boxplot() +theme_bw()+
  facet_grid(rows = vars(k_s), cols = vars(coi_r)) +
  labs(
    x = "Expected Allele Frequencies", 
    y = "Number of Samples with False Negative Alleles", 
    color = "Loci Set"
  ) +
  theme(legend.position = "bottom")-> FalseNeg_byFreq
```

```{r, fig.cap="Figure 5: Proportion of samples with fals negative detection using tool. Proportions compared across simulation scenarios (level of recombination and COI) and marker groups, and relatedness."}
print(FalseNeg_byFreq)
```

```{r}
False_negs_byScenario_and_freq %>%
  mutate(Freq_Quart= ifelse(exp_freq<0.25, "(0,0.25)", 
                            ifelse(exp_freq>=0.25&exp_freq<0.5, "(0.25,0.5)", ifelse(exp_freq>=0.5&exp_freq<0.75, "(0.5,0.75)", "(0.75,1)"))))|>
  ggplot(mapping = aes(x = Freq_Quart, y = sensitivity, color = group_id)) +
  geom_boxplot() +theme_bw()+
  facet_grid(rows = vars(k_s), cols = vars(coi_r)) +
  labs(
    x = "Expected Allele Frequencies", 
    y = "Sensitivity, Proportion of MadHatter Samples \nwith True Postive Detection", 
    color = "Loci Set"
  ) +
  theme(legend.position = "bottom")-> Sensitivity_byFreq
```

```{r, fig.cap="Figure 6: Proportion of samples with true positive detection using tool. Proportions compared across simulation scenarios (level of recombination and COI), marker groups, and true allele frequencies."}
print(Sensitivity_byFreq)

#replicate this plot as necessary for false positives


```

## Preliminary conclusions for allele frequencies: 

* In general the tool performs best with fewer loci of lower diversity among samples with lower COI.  

* There is high concordance between true and estimated allele frequencies for all simulation scenarios and loci sets except those involving six high diversity loci. 

* These patterns are driven by the tool's inability to detect all alleles in the sample (occurance of false negatives), and these are more common for low frequency alleles, and among samples with higher COI. 

# Allele Prevalence

Goals for allele prevalence: 


* Report Lin's concordance correlation coefficient with 95% confidence intervals between tool and simluation COIs across different groupings of populations, and parameters. 

* Compare tool allele prevalence to simulated allele prevalence using Bland-Altman plots (mean of tool and simulation vs difference of tool vs simulation) across combinations of loci, and simulation parameters. 
  * Can choose to split this up across different groupings of populations, and simulation parameters. 
  

  
```{r}

# Compute RMSE of prevalence -----------------------------
tool_results_with_rmse <- tool_results_wtruth %>%
  group_by(
    sim_id, 
    paramset_id, 
    country, 
    group_id, 
    pop_alpha, 
    coi_p, 
    coi_r, 
    k_s
  ) %>%
  summarize(
    prev_rmse = sqrt(mean((obs_prev - exp_prev)^2)), 
    .groups = "drop"
  )

# Show all RMSEs as a boxplot ------------------------------------------
tool_results_with_rmse <- tool_results_with_rmse %>%
  mutate(
    group_id = factor(
      group_id, 
      c(
        "six_med_div", 
        "six_low_div", 
        "four_group1", 
        "two_group1", 
        "pfdhfr_pfdhps", 
        "pfdhps", 
        "pfdhfr"
      ), 
      labels = c(
        "Six, Medium Diversity", 
        "Six, Low Diversity", 
        "Four", 
        "Two", 
        "pfdhfr/pfdhps", 
        "pfdhps", 
        "pfdhfr"
      )
    )
  )%>%
  mutate(
    pop_alpha = factor(
      pop_alpha, 
      c("2.33", "9", "99"), 
      labels = c("High Sim. Rel", "Med. Sim. Rel", "Low Sim. Rel")
    )
  ) %>%
  mutate(
    k_s = factor(
      k_s, 
      c("0.5", "0.2"), 
      labels = c("Med. Recombination", "High Recombination")
    )
  ) %>%
  # coi_p was varied in synchronicity with coi_r to produce different 
  # polyclonality levels, and thus coi_p is implicitly included when 
  # using coi_r as a plotting aesthetic
  mutate(
    coi_r = factor(
      coi_r, 
      c("0.25", "0.5"), 
      labels = c("Low COI/Prop. Polyclonal", "Med. COI/Prop. Polyclonal")
    )
  )

tool_results_with_rmse %>%
  ggplot(mapping = aes(x = pop_alpha, y = prev_rmse, color = group_id)) +
  geom_boxplot() +theme_bw()+
  facet_grid(rows = vars(k_s), cols = vars(coi_r)) +
  labs(
    x = "Simulated Relatedness", 
    y = "RMSE of Allele Prevalence Estimate", 
    color = "Loci Set"
  ) +
  theme(legend.position = "bottom")-> RMSE_byScenario_prev

#print(RMSE_byScenario_prev)



```


## Concordance 

```{r}

tool_results_Concordance_Prev <- tool_results_wtruth %>%
  group_by(
    group_id, 
    pop_alpha, 
    coi_r, 
    k_s
  )|>
  summarize(conc_pt_est= CCC(exp_prev, obs_prev)$rho.c$est, 
         conc_lwr_CI= CCC(exp_prev, obs_prev)$rho.c$lwr.ci, 
         conc_upr_CI= CCC(exp_prev, obs_prev)$rho.c$upr.ci)|>
  mutate(
    group_id = factor(
      group_id, 
      c(
        "six_med_div", 
        "six_low_div", 
        "four_group1", 
        "two_group1", 
        "pfdhfr_pfdhps", 
        "pfdhps", 
        "pfdhfr"
      ), 
      labels = c(
        "Six, Medium Diversity", 
        "Six, Low Diversity", 
        "Four", 
        "Two", 
        "pfdhfr/pfdhps", 
        "pfdhps", 
        "pfdhfr"
      )
    )
  )|>
  mutate(
    pop_alpha = factor(
      pop_alpha, 
      c("2.33", "9", "99"), 
      labels = c("High Sim. Rel", "Med. Sim. Rel", "Low Sim. Rel")
    )
  ) %>%
  mutate(
    k_s = factor(
      k_s, 
      c("0.5", "0.2"), 
      labels = c("Med. Recombination", "High Recombination")
    )
  ) %>%
  # coi_p was varied in synchronicity with coi_r to produce different 
  # polyclonality levels, and thus coi_p is implicitly included when 
  # using coi_r as a plotting aesthetic
  mutate(
    coi_r = factor(
      coi_r, 
      c("0.25", "0.5"), 
      labels = c("Low COI/Prop. Polyclonal", "Med. COI/Prop. Polyclonal")
    )
  )




tool_results_Concordance_Prev %>%
  ggplot(mapping = aes(x = pop_alpha, y = conc_pt_est, ymin= conc_lwr_CI, ymax= conc_upr_CI, color = group_id)) +
  geom_pointrange(position = position_jitter(width = 0.2)) +theme_bw()+
  facet_grid(rows = vars(k_s), cols = vars(coi_r)) +
  labs(
    x = "Simulated Relatedness", 
    y = "Concordance Correlation Coeffifient", 
    color = "Loci Set"
  ) + geom_hline(yintercept= 0.95, color='red', lty=2)+
  theme(legend.position = "bottom")-> Concordance_byScenario_prev
```

```{r, fig.cap ="Figure 7: Concordance correlation coefficient (with 95% CI) between simulated and tool estimated allele prevalence. Concordance is compared across simulation scenarios (level of recombination, COI, and level of relatedness) and marker set."}
print(Concordance_byScenario_prev)

```

## Bland-Altman Plots 

```{r}

tool_results_wtruth_BA<- tool_results_wtruth|>
  mutate(mean_prev= (obs_prev+exp_prev)/2, 
         dif_prev= obs_prev-exp_prev)|>
   mutate(
    group_id = factor(
      group_id, 
      c(
        "six_med_div", 
        "six_low_div", 
        "four_group1", 
        "two_group1", 
        "pfdhfr_pfdhps", 
        "pfdhps", 
        "pfdhfr"
      ), 
      labels = c(
        "Six, Medium Diversity", 
        "Six, Low Diversity", 
        "Four", 
        "Two", 
        "pfdhfr/pfdhps", 
        "pfdhps", 
        "pfdhfr"
      )
    )
  )|>
  mutate(
    pop_alpha = factor(
      pop_alpha, 
      c("2.33", "9", "99"), 
      labels = c("High Sim. Rel", "Med. Sim. Rel", "Low Sim. Rel")
    )
  ) %>%
  mutate(
    k_s = factor(
      k_s, 
      c("0.5", "0.2"), 
      labels = c("Med. Recombination", "High Recombination")
    )
  ) %>%
  # coi_p was varied in synchronicity with coi_r to produce different 
  # polyclonality levels, and thus coi_p is implicitly included when 
  # using coi_r as a plotting aesthetic
  mutate(
    coi_r = factor(
      coi_r, 
      c("0.25", "0.5"), 
      labels = c("Low COI/Prop. Polyclonal", "Med. COI/Prop. Polyclonal")
    )
  )


LocusSets<- unique(tool_results_wtruth_BA$group_id)

BA_plots<- vector(mode='list', length=length(LocusSets))

for(i in 1:length(LocusSets)){

tool_results_wtruth_BA %>%
    filter(group_id==LocusSets[i])|>
  ggplot() +
  geom_hex(
    aes(x = mean_prev, y = dif_prev, fill = log10(after_stat(count))))+
  scale_fill_viridis_c(name = "log10(count)") +
  theme_minimal() + 
  geom_smooth(aes(x = mean_prev, y = dif_prev), method='loess', se=F) + 
  facet_grid(rows = vars(k_s), cols = vars(coi_r)) +
  labs(
    x = "Mean of Observed and Expected Prev", 
    y = "Difference of Observed and Expected Prev"
  ) +ggtitle(LocusSets[i])-> BA_byScenario_prev
  
  BA_plots[[i]]<- BA_byScenario_prev

}



```


```{r, fig.cap= "Figure 8: Mean vs difference of simulated and tool estimated allele prevalence with LOESS line (blue). These relationships are showm across simulation scenarios (level of recombination, and COI) and marker sets."}

cat(create_tabsetOfGgplotObjects(
  list(
    "Six, Medium Diversity" = BA_plots[[1]], 
    "Six, Low Diversity" = BA_plots[[2]],
    "Four" = BA_plots[[3]], 
    "Two" = BA_plots[[4]], 
    "pfdhfr/pfdhps" = BA_plots[[5]], 
    "pfdhfr" = BA_plots[[6]], 
    "pfdhps" = BA_plots[[7]])
))
```

## Preliminary conclusions for allele prevalence: 

* In general the tool performs best with fewer loci of lower diversity among samples with lower COI.  

* There is high concordance between true and estimated allele prevalence for low COI simulation scenarios however for higher COI simulation scenarios multiple loci sets show poor concordance including the set of six high diversity loci and several drug resistance markers. 

* In general, the greatest discrepancy between the true and estimated allele prevalence is found when the prevalence is between 0.25 and 0.5. 
