---
title: "PGEcore"
output: html_document
author: "Shazia Ruybal-Pesántez"
date: "2025-07-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose

The `PGEcore` bespoke scripts perform functionalities not currently fulfilled by any available tool. 

## Existing resources

The [`PGEcore` GitHub repository](https://github.com/nianqiaoju/snp-slice) includes more details on each bespoke script, including usage. 

See:

- [Estimate naive COI](https://github.com/PlasmoGenEpi/PGEcore/scripts/estimate_coi_naive/)
- [Estimate naive single-locus allele frequency](https://github.com/PlasmoGenEpi/PGEcore/scripts/estimate_allele_frequency_naive/)
- [Estimate naive single-locus allele prevalence](https://github.com/PlasmoGenEpi/PGEcore/scripts/estimate_allele_prevalence_naive/)
- [Estimate naive multi-locus genotype frequency and prevalence](https://github.com/PlasmoGenEpi/PGEcore/scripts/multilocus_prevfreq_naive)

(*note to self: these links will be broken until PGEcore merged to main, but should be ready to go once PGEcore is live*)

## Citation

