---
title: "DRpower"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'") |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose

*DRpower* is an R package that is primarily intended for *pfhrp2/3* deletion studies. However, it can also be applied to drug resistance data, or indeed and data that takes the form of a prevalence estimate (numerator and denominator). It can be used in two main ways:

1. **In the design phase of a study** to perform power and sample size calculations. This can tell us how many sites we need to recruit, and how many samples per site.
2. **In the analysis phase of a study** to estimate the prevalence of the marker of interest, and compare this value against a set threshold.

It uses a Bayesian model to estimate the prevalence, which takes into account the potential of intra-cluster correlation in multi-site studies.

## Existing resources

- The [*DRpower* website](https://mrc-ide.github.io/DRpower/) gives details of the statistical arguments behind the tool, installation instructions, and worked tutorials for both design and analysis phases.
- The [*pfhrp2/3* Planner](https://shiny.dide.ic.ac.uk/DRpower-app/) provides a web-based interface to the  *DRpower* tool. It does not have all the functionality of the R package, but can be used to carry out simple design and analysis tasks.

## Citation

Please use the following citation:

```
@Manual{,
    title = {DRpower: Study design and analysis for pfhrp2/3 deletion prevalence studies},
    author = {Bob Verity and Shazia Ruybal},
    note = {R package version 1.0.0},
  }
```