---
title: "Benchmarking of PGEcore Naïve Single Locus Allele Frequency Tool"
author: "Josphat N. Nyataya"
output: html_document
---

```{r setup, include = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(ggplot2)
library(jsonlite)
library(magrittr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(tidyr)

inp <- list(
  sim_info_dir = "../../../data/benchmarking/simulations/SpotMalaria", 
  tool_res_dir = "../../../data/benchmarking/tool_results/naive_slaf", 
  tool_params = "../../../data/benchmarking/tool_results/naive_slaf.json"
)

source("../../scripts/benchmarking_functions.R")
```

```{r}
# Read all expected frequencies into one tibble ------------------------
sim_truth <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_aa_variants_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_integer(), 
        population_name = col_character(), 
        country = col_character(), 
        GeneName = col_character(), 
        transcript_id = col_character(), 
        aa = col_character(), 
        is_ref = col_logical(), 
        freq = col_double(), 
        prev = col_double()
      ), 
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>%
  rename(
    sim_id = population_name, 
    exp_freq = freq, 
    exp_prev = prev, 
    exp_n_samp = sample_count
  ) %>%
  select(
    -pop, 
    -country, 
    -allele_count, 
    -GeneName, 
    -is_ref, 
    -total_pop_samples, 
    -total_pop_alleles, 
    -unique_alleles
  )

# Read in all simulation parameters into one tibble --------------------
sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>%
  select(
    -collection_year, 
    -obs_freq_monoclonal, 
    -obs_non_recombinant, 
    -mean_k, 
    -sample_count, 
    -est_between_sample_relatedness
  ) %>%
  rename(
    sim_id = population_name, 
    true_mean_coi = mean_coi
  )
sim_truth <- sim_truth %>%
  left_join(sim_params, by = "sim_id")

# Read in all FEM results into one tibble ------------------------------
tool_results <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "*.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(sim_id = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(), 
        variant = col_character()
      ), 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>%
  rename(
    obs_freq = freq
  )

# Read in tool parameters, and join ------------------------------------
tool_params <- jsonlite::fromJSON(inp$tool_params) %>%
  as_tibble() %>%
  mutate(slaf_method = paramset$method) %>%
  select(-paramset)
tool_results <- tool_results %>%
  left_join(tool_params, by = "paramset_id")

# Join everything into one tibble --------------------------------------
sim_truth <- sim_truth %>% mutate(variant = paste(transcript_id, aa_position, aa, sep = ":"))
```

```{r}
# Combine the simulated and tool results into a single dataframe
toolres_simtruth_combined <- full_join(sim_truth, tool_results, by = c("sim_id", "variant"))
```

```{r}
# Run the means and difference of the observed and expected columns
toolres_simtruth_combined <- add_diff_mean(toolres_simtruth_combined, expected_col = exp_freq, observed_col = obs_freq)
```

# Concordance

False positives and false negatives were removed from this analysis to avoid non-infinite values in the plots.

```{r}
toolres_simtruth_combined_nomissing <- na.omit(toolres_simtruth_combined)
```

```{r}
 tool_results_Concordance_Freq <- toolres_simtruth_combined_nomissing %>%
         group_by(
         slaf_method, 
         pop_alpha, 
         coi_r, 
         k_s
     )|>
     summarize(conc_pt_est= CCC(exp_freq, obs_freq)$rho.c$est, 
               conc_lwr_CI= CCC(exp_freq, obs_freq)$rho.c$lwr.ci, 
               conc_upr_CI= CCC(exp_freq, obs_freq)$rho.c$upr.ci, .groups = "drop")|>
     mutate(
         group_id = factor(
             slaf_method, 
             c(
                 "read_count", 
                 "presence_absence"
             ), 
             labels = c(
                 "read_count", 
                 "presence_absence"
             )
         )
     )|>
     mutate(
         pop_alpha = factor(
             pop_alpha, 
             c("2.33", "9", "99"), 
             labels = c("High Sim. Rel", "Med. Sim. Rel", "Low Sim. Rel")
         )
     ) %>%
     mutate(
         k_s = factor(
             k_s, 
             c("0.5", "0.2"), 
             labels = c("Med. Recombination", "High Recombination")
         )
     ) %>%
     
     mutate(
         coi_r = factor(
             coi_r, 
             c("0.25", "0.5"), 
             labels = c("Low COI/Prop. Polyclonal", "Med. COI/Prop. Polyclonal")
         )
     )
```

```{r}
tool_results_Concordance_Freq %>%
     ggplot(mapping = aes(x = pop_alpha, y = conc_pt_est, ymin= conc_lwr_CI, ymax= conc_upr_CI, color = slaf_method)) +
     geom_pointrange(position = position_jitter(width = 0.2)) +        theme_bw()+
     facet_grid(rows = vars(k_s), cols = vars(coi_r)) +
     labs(
         x = "Simulated Relatedness", 
         y = "Concordance Correlation Coeffifient", 
         color = "Method"
     ) + geom_hline(yintercept= 0.95, color='red', lty=2)+
     theme(legend.position = "bottom")-> Concordance_byScenario_prev
 
print(Concordance_byScenario_prev)
```

In this figure we see clearly that the read count method of estimating allele frequency was consistent in both low and medium COI, whereas the presence_absence approach yielded lower correlation coefficients in the medium COI simulations.

# Bland-Altman Plots

False positives and false negatives were also removed to create these plots.

```{r}
# We split the data by slaf_method
toolres_simtruth_combined_rc <- toolres_simtruth_combined_nomissing %>% filter(slaf_method == "read_count_prop")
toolres_simtruth_combined_pa <- toolres_simtruth_combined_nomissing %>% filter(slaf_method == "presence_absence")
```

## Presence-Absence Method

```{r message = FALSE}
s2 <- make_bland_altman_plot_by_label(toolres_simtruth_combined_pa, label = "obs_freq", bins = 30) + transparentBackground
s2
```

While the two measures show reasonable overall agreement, the plot reveals evidence of frequency-dependent bias, with overestimation at low frequencies and underestimation at high frequencies.

## Read Count Method

```{r message = FALSE}
s1 <- make_bland_altman_plot_by_label(toolres_simtruth_combined_rc, label = "obs_freq", bins = 30) + transparentBackground 
s1
```

The plot indicates overall good agreement between observed and expected frequencies, with some heteroscedasticity at intermediate frequencies but no strong evidence of consistent bias.

# False Positives and False Negatives

```{r}
# To generate the contingency tables, we extract the TP, FN and FP
combined_2 <- toolres_simtruth_combined %>% mutate(status = case_when(!is.na(obs_freq) & is.na(exp_freq) ~ "FP", is.na(obs_freq) & !is.na(exp_freq) ~ "FN", .default = "TP"))

#summarize the counts
count_summarize <- combined_2 %>% group_by(status) %>% summarise(count = n(), .groups = "drop")
```

```{r}
# Convert the count_summary tibble to a dataframe and ensure the three columns are well labelled
count_summarize_df <- as.data.frame(count_summarize)
input <- as.data.frame(t(count_summarize_df))
colnames(input)[1] <- "FN"
colnames(input)[2] <- "FP"
colnames(input)[3] <- "TP"
input <- input[-1, ]
input <- input %>% mutate(
  TP = as.numeric(TP), 
  FP = as.numeric(FP), 
  FN = as.numeric(FN)
)
```

```{r}
# Generate the contingency table
d1 <- input2 <- add_contigency_metrics_no_TN(input = input, summarize_first = F)
d1
```

```{r}
# Generate a radar plot but we first generate the metric map
metric_map1 <- c("Sensitivity" = "sensitivity", "Precision" = "precision", "F1-score" = "f1_score", "FDR" = "fdr")
g3 <- radar_by_metrics(input2, color_col = NULL, facet_col = NULL, metric_map = metric_map1, palette = NULL, title = "Perfomance radar", outer_pad = 0.12, ring_breaks = seq(0,1,0.25))
g3
```

Overall, this method has excellent sensitivity (98.5%) and is therefore able to detect all positives with few false negatives. However, it also generated many false positives (precision=76.6%).
