---
title: "Naive Single Locus Allele Frequency"
author: "JNN"
date: "2025-08-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load in the Benchmarking Outputs
```{r}
library(dplyr)
library(ggplot2)
library(jsonlite)
library(magrittr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(tidyr)

inp <- list(
  sim_info_dir = "C:/Users/Josphat.Nyataya/OneDrive - USAMRU Kenya/Desktop/Desktop/R_analyses_2024/naive_slaf_results/simulations/SpotMalaria", 
  tool_res_dir = "C:/Users/Josphat.Nyataya/OneDrive - USAMRU Kenya/Desktop/Desktop/R_analyses_2024/naive_slaf_results/naive_slaf", 
  tool_params = "C:/Users/Josphat.Nyataya/OneDrive - USAMRU Kenya/Desktop/Desktop/R_analyses_2024/naive_slaf_results/naive_slaf.json"
)
```
##First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.
```{r}
# Read all expected frequencies into one tibble ------------------------
sim_truth <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_aa_variants_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_integer(), 
        population_name = col_character(), 
        country = col_character(), 
        GeneName = col_character(), 
        transcript_id = col_character(), 
        aa = col_character(), 
        is_ref = col_logical(), 
        freq = col_double(), 
        prev = col_double()
      ), 
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>%
  rename(
    sim_id = population_name, 
    exp_freq = freq, 
    exp_prev = prev, 
    exp_n_samp = sample_count
  ) %>%
  select(
    -pop, 
    -country, 
    -allele_count, 
    -GeneName, 
    -is_ref, 
    -total_pop_samples, 
    -total_pop_alleles, 
    -unique_alleles
  )

# Read in all simulation parameters into one tibble --------------------
sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>%
  select(
    -collection_year, 
    -obs_freq_monoclonal, 
    -obs_non_recombinant, 
    -mean_k, 
    -sample_count, 
    -est_between_sample_relatedness
  ) %>%
  rename(
    sim_id = population_name, 
    true_mean_coi = mean_coi
  )
sim_truth <- sim_truth %>%
  left_join(sim_params, by = "sim_id")

# Read in all FEM results into one tibble ------------------------------
tool_results <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "*.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(sim_id = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(), 
        variant = col_character()
      ), 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>%
  rename(
    obs_freq = freq
  )

# Read in tool parameters, and join ------------------------------------
tool_params <- jsonlite::fromJSON(inp$tool_params) %>%
  as_tibble() %>%
  mutate(slaf_method = paramset$method) %>%
  select(-paramset)
tool_results <- tool_results %>%
  left_join(tool_params, by = "paramset_id")

# Join everything into one tibble --------------------------------------
sim_truth <- sim_truth %>% mutate(variant = paste(transcript_id, aa_position, aa, sep = ":"))
tool_results_wtruth <- tool_results %>%
  # Use an inner join because FEM also estimates frequencies for 
  # subsets of the requested multilocus genotypes, and these 
  # frequencies are not benchmarked
  inner_join(sim_truth, by = c("sim_id", "variant"))
```
### Allele Frequency
```{r}
# Compute RMSE of frequency and prevalence -----------------------------
Allele_freq <- tool_results_wtruth %>%
  group_by(
    sim_id, 
    slaf_method, 
    country,
    pop_alpha,
    coi_p, 
    coi_r, 
    k_s
  ) %>%
  summarize(
    freq_rmse = sqrt(mean((obs_freq - exp_freq)^2)), 
    #prev_rmse = sqrt(mean((obs_prev - exp_prev)^2)), 
    .groups = "drop"
  )

ggplot(Allele_freq, aes(x = country, y = freq_rmse, colour = slaf_method)) + geom_boxplot() + scale_x_discrete() + labs( x = "Population", y = "RMSE of Allele Frequency Estimate", title = "Single Locus Allele frequency estimates") + theme(axis.text.x = element_text(angle = 45, hjust = 1)) + theme_bw()
```
##plotting separately for low and high frequency alleles;in the tool_results_wtruth data frame, i chose to set a threshold of 0.5 for allelic frequency and label anything below as low frequency and anything above 0.5 as high frequency in a separate column.
```{r}
tool_results_wtruth <- tool_results_wtruth %>% mutate(freq_category = ifelse(obs_freq < 0.5, "Low frequency", "High frequency"))
```
## Compute RMSE for low and high frequency alleles
```{r}
Allele_freq2 <- tool_results_wtruth %>%
  group_by(
    sim_id, 
    slaf_method, 
    country,
    freq_category,
    pop_alpha,
    coi_p, 
    coi_r, 
    k_s
  ) %>%
  summarize(
    freq_rmse = sqrt(mean((obs_freq - exp_freq)^2)), 
      .groups = "drop"
  )
ggplot(Allele_freq2, aes(x = country, y = freq_rmse, colour = slaf_method)) + geom_boxplot() + facet_wrap(~freq_category, nrow = 1) + labs(x = "Population", y = "RMSE of Allele Frequency Estimates") + theme_bw() + theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
## To calculate the number of false negatives and false positive, we compared the variant calls by the simulations (sim_truth) table vs the variant calls by the tool (tool_results)table.
```{r}
combined <- full_join(tool_results, sim_truth, by = c("sim_id", "variant"))

#classify the variants as "False positive" or as "False Negative"
combined <- combined %>% mutate(status = case_when(!is.na(obs_freq) & is.na(exp_freq) ~ "False Positive", is.na(obs_freq) & !is.na(exp_freq) ~ "False Negative", TRUE ~ "True Variant"))

#summarize the counts
count_summarize <- combined %>% group_by(status) %>% summarise(count = n(), .groups = "drop")

count_summarize
```
## grouping the false positive by parameterset
```{r}
false_positive <- combined %>% select(sim_id, paramset_id, status, country, slaf_method)
false_positive <- false_positive %>% mutate(status = "False Positive")
false_positive <- false_positive %>% filter(!is.na(slaf_method), !is.na(country))
FP_count <- false_positive %>% group_by(country,slaf_method,status) %>% summarise(count = n(), .groups = "drop")

ggplot(FP_count, aes(x = country, y = count, fill = slaf_method)) + geom_col() + theme_bw() + labs(x = "Country", y = "False positive counts", title = "Propotion of False positive counts by Parameter set")
```
## Using the revised formular for testing the hypothesis that the difference btwn observed and expected allele frequency is zero
```{r}
#created a combined df from tool results table and simulated results table
combined_s <- full_join(sim_truth, tool_results, by = c("sim_id", "variant"))

#create columns based on the formular
combined_s <- combined_s %>% mutate(Pobs_Pexp = obs_freq - exp_freq,
                                    Sum_Pobs_Pexp = obs_freq + exp_freq,
                                    Sqrd_Pobs = obs_freq ^2, 
                                    Sqrd_Pexp = exp_freq ^2,
                                    n = exp_n_samp*true_mean_coi,
                                    denominator = sqrt((Sum_Pobs_Pexp - Sqrd_Pobs- Sqrd_Pexp) / n),
                                    statistic = Pobs_Pexp/denominator)

##clean up the table
combined_s <- combined_s %>% filter(!is.na(statistic), !is.nan(statistic), is.finite(statistic))

##plot a density plot for the distribution of test statitics
ggplot(combined_s, aes(x = statistic)) + geom_density(fill = "lightblue", alpha = 0.6) + theme_bw() + labs(x = "test statistic", y = "density", title = "Distribution of the test statitic")
```
## correlation coeffiecient and plotting
```{r}
cor_coef <- cor(combined_s$exp_freq, combined_s$obs_freq, use = "complete.obs", method = "pearson")
##plotting
ggplot(combined_s, aes(x = exp_freq, y = obs_freq, colour = slaf_method)) + geom_point(alpha = 0.6) + geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") + labs(title = paste0("Correlation: r = ", round(cor_coef, 3)), x = "Simulated Allele Frequency(exp_freq)", y = "Tool-estimated Allele frequency(obs_freq)") + theme_bw()
```