---
title: "Installing paneljudge"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installation from R-Universe
The `paneljudge` R package can be directly installed from [plasmogenepi.r-universe](https://plasmogenepi.r-universe.dev/builds), which greatly simplifies installation. For other tools, please list the steps as needed.


## Installation from Github 
```{r, eval=F}
install.packages("remotes")
remotes::install_github("aimeertaylor/paneljudge")
```

### Dependencies
`paneljudge` relies on the [Rcpp](https://cran.r-project.org/web/packages/Rcpp/index.html) package, which requires certain OS-specific dependencies: 

* Windows
    - Download and install the appropriate version of [Rtools](https://cran.rstudio.com/bin/windows/Rtools/) for your version of R. On installation, ensure you check the box to arrange your system PATH as recommended by Rtools
* Mac OS X
    - Download and install [XCode](http://itunes.apple.com/us/app/xcode/id497799835?mt=12)
    - Within XCode go to Preferences : Downloads and install the Command Line Tools
* Linux (Debian/Ubuntu)
    - Install the core software development utilities required for R package development as well as LaTeX by executing
    ```
    sudo apt-get install r-base-dev texlive-full
    ```

Assuming everything installed correctly, you can now load the package:

```{r, eval=F}
library(paneljudge)
```

