---
title: "paneljudge"
output: html_document
author: "Nick Brazeau"
date: "December 12, 2023"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose
An R package that judges the "informativeness" of a genetic panel (or barcode) in its ability to infer relatedness. The panel is assessed under simulation with an identity-by-descent hidden-markov model backend. Precision in relatedness estimates indicate the superior panel. 


## Existing resources
Taylor et al. Genetics 212.4 (2019): 1337-1351.  
Schaffner et al. Malaria journal 17.1 (2018): 196.

## Citation  
LaVerriere et al., Molecular Ecology Resources 22.6 (2022): 2285-2303   

BibTeX style citation. For an R package, you can get this using `citation(package = "name")`:

Here is an example for DRpower, using `citation(package = "paneljudge")`:

```
@Manual{,
    title = {paneljudge: Judge the performance of a panel of genetic markers using
simulated data},
    author = {Aimee Taylor and Pierre Jacob},
    year = {2023},
    note = {R package version 0.0.0.9000},
  }
```
