---
title: "isoRelate tutorial for generating Identity-by-decent networks: vcf to network"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

## Useful tutorial examples

If you are starting from ped and map files, there is a very clear tutorial `introduction.Rmd` created by isoRelate's author, Lyndal Henden, that uses isoRelates in built data to work through the package functions.
https://github.com/bahlolab/isoRelate/blob/master/vignettes/introduction.Rmd

The data used in the tutorial is a _Plasmodium falciparum_ dataset from Papua New Guinea (PNG), generated by the MalariaGEN Consortium (https://www.malariagen.net/projects/pf3k).

## New tutorial

This tutorial focuses on data from the vcf stage and works through the process to generate appropriate map and ped files from vcf data within R.

## Known errors

**R memory limit: The number of bytes in a character string is limited to 2^31 - 1 ~ 2*10^9, which is also the limit on each dimension of an array.
Error message reads: `cannot allocate vector of length`**

## The data

These headings are placeholders, but you may want to start by giving a brief description of which data you will use. Remember you can point to the [main Data tab](https://mrc-ide.github.io/PGEforge/website_docs/data_description.html) for further details.


Remember to load any packages, including the tool that you are working on:
```{r, warning=FALSE, message=FALSE, error=FALSE}
library(tidyverse)
library(isoRelate)
```


## First analysis section

Make as many sections as you need. Try to break long documents out over multiple pages.

## Summary

It's a good idea to round off with a summary of what we did.

PS, note that the name of the tutorial on the PGEforge website can be whatever you want - it does not have to match the name of this tutorial.
