---
title: MalariaEM Benchmarking Results
author: Zoe Gotthold
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load in the Benchmarking Outputs

```{r, warning=FALSE, message=FALSE, echo = FALSE}
library(dplyr)
library(ggplot2)
library(jsonlite)
library(magrittr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(tidyr)
library(DT)
library(patchwork)
library(ggpmisc)
library(GGally)

inp <- list(
  sim_info_param_dir = "../../data/benchmarking/simulations/SpotMalaria/",
    #"C:/Users/zoeor/OneDrive/Documents/Life/Collaboration with Dr. Wesolowski/malaria.em.benchmark/SpotMalaria",
  sim_info_dir = "../../data/benchmarking/simulations/all_phased_pop_freqs_prevs_out.tsv",
  sim_info_dir_ind = "../../data/benchmarking/simulations/all_phased_sample_out.tsv",
  tool_res_dir = "../../data/benchmarking/tool_results/malariaem/",
  tool_params = "../../data/benchmarking/tool_results/malariaem.json"
)

source("../scripts/benchmarking_functions.R")
```

First, all of the tool results, tool parameters, expected simulated values, and
simulation parameters are read into a tibble.

```{r reading in results, echo = F, warning=F}
# Read all expected frequencies into one tibble ------------------------
sim_truth <- read_table(inp$sim_info_dir) %>% distinct()


# Read in all simulation parameters into one tibble --------------------
sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_param_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  dplyr::select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>%
  dplyr::select(
    -collection_year, 
    -obs_freq_monoclonal, 
    -obs_non_recombinant, 
    -mean_k, 
    -sample_count, 
    -est_between_sample_relatedness,
    -country
  ) %>%
  dplyr::rename(
    sim_id = population_name, 
    true_mean_coi = mean_coi
  )
sim_truth <- sim_truth %>%
  left_join(sim_params, by = join_by(population_name == sim_id))

country_diversity <- sim_truth %>% group_by(population_name, group_id, target) %>%
  dplyr::summarize(n_alleles = n_distinct(seq)) %>% group_by(population_name, group_id) %>%
  dplyr::summarize(max_n_alleles = max(n_alleles))

sim_truth_individual <- read_table(inp$sim_info_dir_ind)

tool_results <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "mlaf_results.tsv", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(sim_id = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  dplyr::select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(), 
        seq = col_character(), 
        target_id = col_character(), 
        gt_id = col_integer(), 
        group_id = col_character()
      ), 
      progress = FALSE
    )
  ) %>%
  dplyr::select(-tool_res_path) %>%
  unnest(tool_res)



tool_results_ind <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "phasing_results.tsv", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(sim_id = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  dplyr::select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(), 
        seq = col_character(), 
        target_id = col_character(), 
        gt_id = col_integer(), 
        group_id = col_character(),
        specimen_id = col_character()
      ), 
      progress = FALSE
    )
  ) %>%
  dplyr::select(-tool_res_path) %>%
  unnest(tool_res)

unique_seqs <- data.frame(seq = unique(c(sim_truth$seq, sim_truth_individual$seq, tool_results$seq, tool_results_ind$seq)))
unique_seqs$key <- rownames(unique_seqs)

sim_truth <- merge(sim_truth, unique_seqs, by="seq")
sim_truth_individual <- merge(sim_truth_individual, unique_seqs, by="seq")
tool_results <- merge(tool_results, unique_seqs, by="seq") %>% distinct()
tool_results_ind <- merge(tool_results_ind, unique_seqs, by="seq")

sim_truth_phased <- sim_truth %>% group_by(population_name, country, pop_simulated_gt_id, allele_freq, group_id) %>%
  arrange(target) %>% dplyr::summarize(seq = paste(seq, collapse=""),
                                target = paste(target, collapse=";"),
                                key = paste(key, collapse = ","),
                                nseqs = n())


sim_truth_phased_ind <- sim_truth_individual %>% 
  group_by(population_name, country, pop_simulated_gt_id, group_id, sample_name, simulated_within_sample_phase_hap_id) %>%
  arrange(target) %>% distinct(seq, target, key) %>% dplyr::summarize(seq = paste(seq, collapse=""),
                                target = paste(target, collapse=";"),
                                key = paste(key, collapse = ","),
                                nseqs = n()) %>%
  group_by(population_name, group_id, sample_name, target) %>%
  dplyr::summarize(key_TRUE = paste(str_sort(unique(key)), collapse="_"))


tool_results_phased <- tool_results %>% group_by(sim_id, group_id, gt_id, freq) %>% arrange(target_id) %>%
  dplyr::summarize(seq = paste(seq, collapse = ""),
            key = paste(key, collapse = ","),
            target = paste(target_id, collapse = ";")) %>% distinct()

tool_results_ind_phased <- tool_results_ind %>% group_by(sim_id, group_id, gt_id, phase_id, specimen_id) %>% 
  arrange(target_id) %>%
  dplyr::summarize(seq = paste(seq, collapse = ""),
            key = paste(key, collapse = ","),
            target = paste(target_id, collapse = ";"),
            posterior_est = min(posterior_est)) %>% distinct() %>%
  mutate(specimen_id = as.numeric(str_split_i(specimen_id, "_", 1))) %>%
  group_by(sim_id, group_id, specimen_id, target, posterior_est) %>%
  dplyr::summarize(key_TOOL = paste(str_sort(unique(key)), collapse="_"))

results_ind <- full_join(sim_truth_phased_ind, tool_results_ind_phased,
                         by = join_by(target==target,
                                      sample_name == specimen_id,
                                      group_id == group_id,
                                      population_name == sim_id)) %>%
  mutate(match = key_TRUE==key_TOOL,
         posterior_est = as.numeric(posterior_est))

results_ind <- left_join(results_ind, sim_params, by = join_by(population_name == sim_id))

# Read in tool parameters
tool_params <- jsonlite::fromJSON(inp$tool_params, flatten = TRUE) %>%
  as_tibble() 

#create_dt(tool_params)

```
## Defining Plot Functions

To quantify accuracy across a variety of simulation parameters, we have plotted data on two axes: one for the COI/proportion polyclonal, and another for the amount of recombination. Each loci group (two loci (set A), two loci (set B), and four loci) is separated out.

```{r defining plot functions}

make_boxplots <- function(df, response_var, ylab, title){
  df <- df %>% left_join(sim_params, by = join_by(population_name==sim_id))
  
  plot <- ggplot(df %>% mutate(pop_alpha = 
                                               case_match(pop_alpha, 2.33 ~ "High \nRelatedness",
                                                          9 ~ "Med. \nRelatedness",
                                                          99 ~ "Low \nRelatedness"
                                                          ),
                                             coi_r = case_match(coi_r, 0.25 ~ "Low COI/Prop. Polyclonal",
                                                                0.5 ~ "Med COI/Prop. Polyclonal"),
                                             k_s = case_match(k_s, 0.2 ~ "High Recombination",
                                                              0.5 ~ "Med Recombination"),
                                             `Loci Group` = case_match(group_id, "two_lowdiv" ~ "Two (Set A)",
                                                                       "two_meddiv" ~ "Two (Set B)", "four" ~ "Four")), 
         aes(x = `Loci Group`, y=!!sym(response_var), fill = `Loci Group`, color = `Loci Group`)) +
    geom_boxplot(outliers = F, alpha = 0.3) + geom_jitter() + theme_bw() + 
    ylab(ylab) + 
    ylim(c(0, 1)) + xlab(NULL) + ggtitle(title) + 
    scale_fill_manual(values = c( "goldenrod", "thistle", "orchid")) +
  scale_color_manual(values = c( "goldenrod", "thistle", "orchid")) +
  #geom_errorbar(aes(ymin = freq_ccc_lower, ymax = freq_ccc_upper), 
    #            position=position_dodge(width=0.01)) + 
  facet_grid(cols = vars(coi_r), rows = vars(k_s)) + xlab(NULL)
  return(plot)
}



make_grouped_plot <- function(df, response_var, ylab, title){
  plot <- ggplot(df %>% mutate(pop_alpha = 
                                               case_match(pop_alpha, 2.33 ~ "High \nRelatedness",
                                                          9 ~ "Med. \nRelatedness",
                                                          99 ~ "Low \nRelatedness"
                                                          ),
                                             coi_r = case_match(coi_r, 0.25 ~ "Low COI/Prop. Polyclonal",
                                                                0.5 ~ "Med COI/Prop. Polyclonal"),
                                             k_s = case_match(k_s, 0.2 ~ "High Recombination",
                                                              0.5 ~ "Med Recombination"),
                                             `Loci Group` = case_match(group_id, "two_lowdiv" ~ "Two, Low Diversity",
                                                                       "two_meddiv" ~ "Two, Med Diversity", "four" ~ "Four")), 
                      aes(x = pop_alpha, y = !!sym(response_var), 
                                       fill = `Loci Group`, color = `Loci Group`)) +
  #geom_jitter() +
  geom_point(alpha = 1, position = position_dodge(width = 0.5),
             size = 3) +
  geom_errorbar(aes(ymin = freq_ccc_lower, ymax = freq_ccc_upper) , 
                position = position_dodge(width = 0.5),
                width = 0,
                linewidth = 1) + 
  scale_fill_manual(values = c( "goldenrod", "thistle", "orchid")) +
  scale_color_manual(values = c( "goldenrod", "thistle", "orchid")) +
  #geom_errorbar(aes(ymin = freq_ccc_lower, ymax = freq_ccc_upper), 
    #            position=position_dodge(width=0.01)) + 
  facet_grid(cols = vars(coi_r), rows = vars(k_s)) + xlab(NULL) + ylab(ylab) + ggtitle(title)
  return(plot)
}
  

```


# Individual-Level Analyses

Malaria.em provides both population-level and individual results. Here, we present the individual-level results (i.e. the phasing within each individual). The posterior estimates generated for each individual were generally higher when the matching was correct. We chose a posterior estimate of > 0.55 as a cutoff for being more likely to be accurate than not in this dataset, but this value will require further validation.

```{r plotting individual results}
conflicted::conflicts_prefer(dplyr::filter)
results_ind <- results_ind %>% mutate(`Loci Group` = 
                                              case_match(group_id, 
                                                         "two_lowdiv" ~ "Two, Low Diversity",
                                                         "two_meddiv" ~ "Two, Med Diversity",
                                                         "four" ~ "Four"))

a <- ggplot(results_ind %>% filter(!is.na(posterior_est)), aes(x = match, y=posterior_est)) + 
  geom_violin(alpha = 0.3, aes(fill = match)) + 
  theme_bw() + scale_fill_manual(values = c("maroon", "darkblue")) + facet_wrap(~`Loci Group`) + ggtitle("Figure 1a")

b <- ggplot(results_ind %>% filter(!is.na(posterior_est)), aes(fill = match, x = posterior_est)) + 
  geom_density(alpha = 0.3)  + theme_bw() + scale_fill_manual(values = c("maroon", "darkblue")) + 
  facet_wrap(~`Loci Group`) + ggtitle("Figure 1b")

a/b


results_ind_summary <- results_ind %>% group_by(population_name, group_id) %>%
  dplyr::summarize(prop_perfect = sum(match)/ n(), .groups = "keep") %>% 
  mutate(country = str_split_i(population_name, "_", 2),
         id = str_split_i(population_name, "_", 4),
         short_name = paste(country, id, sep="_")) 


results_ind_summary_55 <- results_ind %>%
  filter(posterior_est > 0.55) %>% 
  group_by(population_name, group_id,) %>%
  dplyr::summarize(prop_perfect = sum(match)/ n(), .groups = "keep") %>% 
  mutate(country = str_split_i(population_name, "_", 2),
         id = str_split_i(population_name, "_", 4),
         short_name = paste(country, id, sep="_"))

samples_lost_summary_55 <- results_ind %>% 
  group_by(population_name, group_id) %>%
  summarize(Lost = (sum(posterior_est < 0.55) / sum(posterior_est>=0)))

all_accuracy <- make_boxplots(results_ind_summary, "prop_perfect", 
              "Proportion of Samples Correctly Phased", "Fig. 2. Efficacy of Sample-Level Phasing")

post_55 <- make_boxplots(results_ind_summary_55, "prop_perfect", 
              "Proportion of Samples Correctly Phased", "Fig. 3. Efficacy of Sample-Level Phasing (Post. Est > 0.55)")

prop_samples_lost <- make_boxplots(samples_lost_summary_55, "Lost", "Proportion of Samples Removed",
                                   "Fig. 4. How much data was removed with Post. Est > 0.55?")

all_accuracy 

post_55

prop_samples_lost 

```

As shown in these figures, the accuracy, particularly among samples taken from regions with a medium COI as opposed to a low one, is rather low. Many of the populations have \< 75% of the samples correctly phased. This individual-level phasing accuracy requires more validation. Although adding a cutoff for posterior estimate certainty improves measures, this does lead to a loss of data, as shown in Figure 4.

# Population-Level Analyses

```{r plotting population results}
results <- full_join(sim_truth_phased, 
                     tool_results_phased, 
                     by=join_by(key==key,
                                group_id==group_id,
                                population_name==sim_id))

results <- left_join(results, sim_params, by= join_by(population_name == sim_id))
results <- results %>% mutate(
  freq = replace_na(freq, 0),
  allele_freq = replace_na(allele_freq, 0)
)

results <- results %>% left_join(country_diversity, join_by(population_name, group_id))
results_ccc_df <- results %>% mutate(group_label = paste(coi_r, group_id, pop_alpha, k_s)) 
results_ccc <-  compute_ccc(results_ccc_df %>% ungroup(), allele_freq, freq, group = c(population_name, group_id))
results_ccc <- results_ccc %>% inner_join(results_ccc_df %>% group_by(population_name, group_id) %>%
                                            summarize_all(min), 
                                          join_by(population_name == population_name, group_id ==
                                                    group_id))

results_ccc_grouped <-compute_ccc(results_ccc_df %>% ungroup(), allele_freq, freq, group = group_label)
results_ccc_grouped <- results_ccc_grouped %>% 
  inner_join(results_ccc_df %>% group_by(group_label) %>%
                                            summarize_all(min), 
                                          join_by(group_label))


make_grouped_plot(results_ccc_grouped, "freq_ccc", "CCC", "Fig. 5. Overall Plot of Concordance")

```

Now, to further visualize these concordances, we can plot them against each other. The main variable that the CCC seems to vary by is COI, so here we can see the alignment of expected and estimated allele frequencies faceted by COI.

```{r further visualizations}
results_plot <- results %>% mutate(coi_r = case_match(coi_r, 0.25 ~ "Low COI/Prop. Polyclonal",
                                                                0.5 ~ "Med COI/Prop. Polyclonal"))
test_ccc <- compute_ccc(results_plot  %>% ungroup(), allele_freq, 
                        freq, group = coi_r)
test_ccc <- test_ccc %>% mutate(label = paste0("CCC = ", round(freq_ccc, digits = 2),
                                               " [", round(freq_ccc_lower, digits = 2), "-", round(freq_ccc_upper, digits = 2), "]"))

plot <- ggplot(results_plot, aes(x = allele_freq, y = freq), color = "black") +
                    #color = ifelse(population_name %in% 
                    #                 c("simulated_Ghana_pop_043",
                    #                   "simulated_Ghana_pop_003",
                    #                   "simulated_Ghana_pop_051"), "High", "Low"))) + 
  geom_point(alpha = 0.5) + 
  ggtitle("Fig. 6. Malaria EM") + theme_bw() + 
  xlab("True Frequency") + 
  ylab("Malaria EM Frequency") +
  stat_poly_eq(group = 1) + facet_wrap(~coi_r) 

plot + geom_text(
  data    = test_ccc,
  mapping = aes(x = -Inf, y = 0.8, label = label),
  hjust   = -0.05,
  vjust   = -1
)
```

As you can see, even with a lower CCC, there is still strong agreement between the true and estimated frequencies from malaria.em.

## False Positives and False Negatives

Adapted from Bing's Code. There are some false negatives, but they are still relatively low overall. There are more false positives, but their estimated frequency is low.

```{r}
## allele freq is "real"
## freq is the estimated value

Fig7a <- results %>%
  #filter(exp_freq > 0) %>% 
  ggplot(aes(x = allele_freq, y = as.numeric(freq == 0))) +
  stat_summary_bin(    fun = "mean", geom = "bar", bins = 50, fill = "thistle", color = "white"  ) +
  labs(
    y = "Ratio (FN / Total)", # use ratio to account for base line allele frequency distribution.
    title = "Enrichment of False Negatives"
  ) + theme_bw()

Fig7b <- results %>%
  #filter(obs_freq > 0) %>% 
  ggplot(aes(x = freq, y = as.numeric(allele_freq == 0))) +
  stat_summary_bin(fun = "mean", geom = "bar", bins = 50, fill = "thistle", color = "white"  ) +
  labs(
    y = "Ratio (FP / Total)", # use ratio to account for base line allele frequency distribution.
    title = "Enrichment of False Positives"
  ) + theme_bw()

Fig7a | Fig7b
```


## Bland-Altman Plots {.tabset}

To show the bias across the range of possible allele frequencies, we generated Bland-Altman plots faceted by a variety of simulation parameters.

### By Country

```{r BA country}
make_bland_altman_plot_by_label(results %>% 
                                  mutate(Country = ifelse(grepl("Ghana", population_name), "Ghana", "Cambodia")) %>%
                                  add_diff_mean(allele_freq, freq), "freq", method = "loess",
                                linecolor = "magenta") + 
  facet_wrap(~Country, nrow = 2) +
  ggtitle("Bland-Altmans by Country")
```

### By Group ID

```{r BA Group ID}
make_bland_altman_plot_by_label(results %>%
                                  mutate(`Loci Group` = 
                                           case_match(group_id, "two_lowdiv" ~ "Two (Set A)",
                                                      "two_meddiv" ~ "Two (Set B)", "four" ~ "Four")) %>%
                                  add_diff_mean(allele_freq, freq), "freq" , method = "loess",
                                linecolor = "magenta") + 
  ggtitle("Bland-Altmans by Loci Group") +
  facet_wrap(~`Loci Group`, nrow = 2)
```

### By COI_R

```{r BA COI_r}
make_bland_altman_plot_by_label(results %>% 
                                   mutate(coi_r = case_match(coi_r, 0.25 ~ "Low COI/Prop. Polyclonal",
                                                                0.5 ~ "Med COI/Prop. Polyclonal")) %>%
                                  add_diff_mean(allele_freq, freq), "freq" , method = "loess",
                                linecolor = "magenta") +
  facet_wrap(~coi_r, nrow = 2) + ggtitle("Bland-Altmans by COI")
```

### By Relatedness

```{r BA relate}
make_bland_altman_plot_by_label(results %>% 
                                   mutate(pop_alpha = 
                                               case_match(pop_alpha, 2.33 ~ "High \nRelatedness",
                                                          9 ~ "Med. \nRelatedness",
                                                          99 ~ "Low \nRelatedness"
                                                          )) %>%
                                  add_diff_mean(allele_freq, freq), "freq" , method = "loess",
                                linecolor = "magenta") +
  facet_wrap(~pop_alpha, nrow = 2) + ggtitle("Bland-Altmans by Relatedness")
```


### By Recombination

```{r BA recomb}
make_bland_altman_plot_by_label(results %>% 
                                   mutate(k_s = case_match(k_s, 0.2 ~ "High Recombination",
                                                              0.5 ~ "Med Recombination")) %>%
                                  add_diff_mean(allele_freq, freq), "freq", method = "loess",
                                linecolor = "magenta") +
  facet_wrap(~k_s, nrow = 2) + ggtitle("Bland-Altmans by Recombination")


```

## {-}

## Summary and Next Steps

Malaria.em seems to be working! However, there are a lot of open questions around exactly how well it works on various allele sets. Additionally, the computational constraints limit wide-spread applicability and still need to be quantified for malaria.em.

* Ind level
* pop-level
* variables
* There are minimal false positives and false negatives: most false positives are "hallucinated" at very low frequencies, and the few false negative represent a very small part of the overall population.
