---
title: MalariaEM Benchmarking Results
author: Zoe Gotthold
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE, message=FALSE, echo = FALSE, error = F}
library(dplyr)
library(ggplot2)
library(jsonlite)
library(magrittr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(tidyr)
library(DT)
library(patchwork)
library(GGally)

inp <- list(
  sim_info_param_dir = "../../data/benchmarking/simulations/SpotMalaria/",
  sim_info_dir = "../../data/benchmarking/simulations/SpotMalaria/all_phased_pop_freqs_prevs_out.tsv",
  sim_info_dir_ind = "../../data/benchmarking/simulations/SpotMalaria/all_phased_sample_out.tsv",
  tool_res_dir = "../../data/benchmarking/tool_results/malariaem/",
  tool_params = "../../data/benchmarking/tool_results/malariaem.json"
)

source("../scripts/benchmarking_functions.R")
```

```{r reading in results, echo = F, warning=F, message=FALSE, error = F}
# Read all expected frequencies into one tibble ------------------------
sim_truth <- read_table(inp$sim_info_dir) %>% distinct()


# Read in all simulation parameters into one tibble --------------------
sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_param_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  dplyr::select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>%
  dplyr::select(
    -collection_year, 
    -obs_freq_monoclonal, 
    -obs_non_recombinant, 
    -mean_k, 
    -sample_count, 
    -est_between_sample_relatedness,
    -country
  ) %>%
  dplyr::rename(
    sim_id = population_name, 
    true_mean_coi = mean_coi
  )
sim_truth <- sim_truth %>%
  left_join(sim_params, by = join_by(population_name == sim_id))

country_diversity <- sim_truth %>% group_by(population_name, group_id, target) %>%
  dplyr::summarize(n_alleles = n_distinct(seq)) %>% group_by(population_name, group_id) %>%
  dplyr::summarize(max_n_alleles = max(n_alleles))

sim_truth_individual <- read_table(inp$sim_info_dir_ind)

tool_results <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "mlaf_results.tsv", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(sim_id = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  dplyr::select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(), 
        seq = col_character(), 
        target_id = col_character(), 
        gt_id = col_integer(), 
        group_id = col_character()
      ), 
      progress = FALSE
    )
  ) %>%
  dplyr::select(-tool_res_path) %>%
  unnest(tool_res)



tool_results_ind <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "phasing_results.tsv", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(sim_id = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  dplyr::select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(), 
        seq = col_character(), 
        target_id = col_character(), 
        gt_id = col_integer(), 
        group_id = col_character(),
        specimen_id = col_character()
      ), 
      progress = FALSE
    )
  ) %>%
  dplyr::select(-tool_res_path) %>%
  unnest(tool_res)

unique_seqs <- data.frame(seq = unique(c(sim_truth$seq, sim_truth_individual$seq, tool_results$seq, tool_results_ind$seq)))
unique_seqs$key <- rownames(unique_seqs)

sim_truth <- merge(sim_truth, unique_seqs, by="seq")
sim_truth_individual <- merge(sim_truth_individual, unique_seqs, by="seq")
tool_results <- merge(tool_results, unique_seqs, by="seq") %>% distinct()
tool_results_ind <- merge(tool_results_ind, unique_seqs, by="seq")

sim_truth_phased <- sim_truth %>% group_by(population_name, country, pop_simulated_gt_id, allele_freq, group_id) %>%
  arrange(target) %>% dplyr::summarize(seq = paste(seq, collapse=""),
                                target = paste(target, collapse=";"),
                                key = paste(key, collapse = ","),
                                nseqs = n())


sim_truth_phased_ind <- sim_truth_individual %>% 
  group_by(population_name, country, pop_simulated_gt_id, group_id, sample_name, simulated_within_sample_phase_hap_id) %>%
  arrange(target) %>% distinct(seq, target, key) %>% dplyr::summarize(seq = paste(seq, collapse=""),
                                target = paste(target, collapse=";"),
                                key = paste(key, collapse = ","),
                                nseqs = n()) %>%
  group_by(population_name, group_id, sample_name, target) %>%
  dplyr::summarize(key_TRUE = paste(str_sort(unique(key)), collapse="_"))


tool_results_phased <- tool_results %>% group_by(sim_id, group_id, gt_id, freq) %>% arrange(target_id) %>%
  dplyr::summarize(seq = paste(seq, collapse = ""),
            key = paste(key, collapse = ","),
            target = paste(target_id, collapse = ";")) %>% distinct()

tool_results_ind_phased <- tool_results_ind %>% group_by(sim_id, group_id, gt_id, phase_id, specimen_id) %>% 
  arrange(target_id) %>%
  dplyr::summarize(seq = paste(seq, collapse = ""),
            key = paste(key, collapse = ","),
            target = paste(target_id, collapse = ";"),
            posterior_est = min(posterior_est)) %>% distinct() %>%
  mutate(specimen_id = as.numeric(str_split_i(specimen_id, "_", 1))) %>%
  group_by(sim_id, group_id, specimen_id, target, posterior_est) %>%
  dplyr::summarize(key_TOOL = paste(str_sort(unique(key)), collapse="_"))

results_ind <- full_join(sim_truth_phased_ind, tool_results_ind_phased,
                         by = join_by(target==target,
                                      sample_name == specimen_id,
                                      group_id == group_id,
                                      population_name == sim_id)) %>%
  mutate(match = key_TRUE==key_TOOL,
         posterior_est = as.numeric(posterior_est))

results_ind <- left_join(results_ind, sim_params, by = join_by(population_name == sim_id))

# Read in tool parameters
tool_params <- jsonlite::fromJSON(inp$tool_params, flatten = TRUE) %>%
  as_tibble() 

#create_dt(tool_params)

```

# Population-Level Multi-Locus Allele Frequencies

Population-level allele frequencies were estimated for three loci groups: two
containing two loci each, and one containing four loci each. Lin's concordance
correlation coefficient (CCC) was estimated between estimated and true allele
frequencies for all loci groups and simulations. Alleles found in the estimated
but not the true set, or vice versa, were set to zero for this analysis.

```{r defining plot functions, echo = F, message=F, error = F}

make_boxplots <- function(df, response_var, ylab, title){
  df <- df %>% left_join(sim_params, by = join_by(population_name==sim_id))
  
  plot <- ggplot(df %>% mutate(pop_alpha = 
                                               case_match(pop_alpha, 2.33 ~ "High \nRelatedness",
                                                          9 ~ "Med. \nRelatedness",
                                                          99 ~ "Low \nRelatedness"
                                                          ),
                                             coi_r = case_match(coi_r, 0.25 ~ "Low COI/Prop. Polyclonal",
                                                                0.5 ~ "Med COI/Prop. Polyclonal"),
                                             k_s = case_match(k_s, 0.2 ~ "High Recombination",
                                                              0.5 ~ "Med Recombination"),
                                             `Loci Group` = case_match(group_id, "two_lowdiv" ~ "Two (Set A)",
                                                                       "two_meddiv" ~ "Two (Set B)", "four" ~ "Four")), 
         aes(x = `Loci Group`, y=!!sym(response_var), fill = `Loci Group`, color = `Loci Group`)) +
    geom_boxplot(outliers = F, alpha = 0.3) + geom_jitter() + theme_bw() + 
    ylab(ylab) + 
    ylim(c(0, 1)) + xlab(NULL) + ggtitle(title) + 
    scale_fill_manual(values = c( "goldenrod", "thistle", "orchid")) +
  scale_color_manual(values = c( "goldenrod", "thistle", "orchid")) +
  #geom_errorbar(aes(ymin = freq_ccc_lower, ymax = freq_ccc_upper), 
    #            position=position_dodge(width=0.01)) + 
  facet_grid(cols = vars(coi_r), rows = vars(k_s)) + xlab(NULL)
  return(plot)
}



make_grouped_plot <- function(df, response_var, ylab, title){
  plot <- ggplot(df %>% mutate(pop_alpha = 
                                               factor(case_match(pop_alpha, 2.33 ~ "High \nRelatedness",
                                                          9 ~ "Med. \nRelatedness",
                                                          99 ~ "Low \nRelatedness"
                                                          ), levels = c("Low \nRelatedness",
                                                                      "Med. \nRelatedness",
                                                                      "High \nRelatedness")),
                                             coi_r = case_match(coi_r, 0.25 ~ "Low COI/Prop. Polyclonal",
                                                                0.5 ~ "Med COI/Prop. Polyclonal"),
                                             k_s = case_match(k_s, 0.2 ~ "High Recombination",
                                                              0.5 ~ "Med Recombination"),
                                             `Loci Group` = case_match(group_id, "two_lowdiv" ~ "Two, Group A",
                                                                       "two_meddiv" ~ "Two, Group B", "four" ~ "Four")), 
                      aes(x = pop_alpha, y = !!sym(response_var), 
                                       fill = `Loci Group`, color = `Loci Group`)) +
  #geom_jitter() +
  geom_point(alpha = 1, position = position_dodge(width = 0.5),
             size = 3) +
  geom_errorbar(aes(ymin = freq_ccc_lower, ymax = freq_ccc_upper) , 
                position = position_dodge(width = 0.5),
                width = 0,
                linewidth = 1) + 
  scale_fill_manual(values = c( "goldenrod", "thistle", "orchid")) +
  scale_color_manual(values = c( "goldenrod", "thistle", "orchid")) +
  #geom_errorbar(aes(ymin = freq_ccc_lower, ymax = freq_ccc_upper), 
    #            position=position_dodge(width=0.01)) + 
  facet_grid(cols = vars(coi_r), rows = vars(k_s)) + xlab(NULL) + ylab(ylab) + ggtitle(title)
  return(plot)
}
  

```



```{r plotting population results, echo = F, message = F, error = F, warning = F}
results <- full_join(sim_truth_phased, 
                     tool_results_phased, 
                     by=join_by(key==key,
                                group_id==group_id,
                                population_name==sim_id))

results <- left_join(results, sim_params, by= join_by(population_name == sim_id))
results <- results %>% mutate(
  freq = replace_na(freq, 0),
  allele_freq = replace_na(allele_freq, 0)
)

results <- results %>% left_join(country_diversity, join_by(population_name, group_id))
results_ccc_df <- results %>% mutate(group_label = paste(coi_r, group_id, pop_alpha, k_s)) 
results_ccc <-  compute_ccc(results_ccc_df %>% ungroup(), allele_freq, freq, group = c(population_name, group_id))
results_ccc <- results_ccc %>% inner_join(results_ccc_df %>% group_by(population_name, group_id) %>%
                                            summarize_all(min), 
                                          join_by(population_name == population_name, group_id ==
                                                    group_id))

results_ccc_grouped <-compute_ccc(results_ccc_df %>% ungroup(), allele_freq, freq, group = group_label)
results_ccc_grouped <- results_ccc_grouped %>% 
  inner_join(results_ccc_df %>% group_by(group_label) %>%
                                            summarize_all(min), 
                                          join_by(group_label))

make_grouped_plot(results_ccc_grouped, "freq_ccc", "CCC", "Overall Plot of Concordance") +
  plot_annotation(caption = paste(strwrap("Population-level Concordance of Allele Frequencies. Point estimates with 95% confidence intervals (CIs) for each group of simulation parameters: low vs. high COI (columns), medium vs. high recombination (rows), low vs. medium vs. high relatedness (x-axis), and each loci group (colors).",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))


```

Concordances are generally high, with somewhat worse performance observed in the
four loci group and in the medium COI simulations.

Now, to further visualize these concordances, we can plot them against each other. The main simulation parameter that the CCC seems to vary by is COI, so here we can see the alignment of expected and estimated allele frequencies faceted by COI.

```{r further visualizations, echo = F, message = F, error = F, warning = F}
results_plot <- results %>% mutate(coi_r = case_match(coi_r, 0.25 ~ "Low COI/Prop. Polyclonal",
                                                                0.5 ~ "Med COI/Prop. Polyclonal"))
test_ccc <- compute_ccc(results_plot  %>% ungroup(), allele_freq, 
                        freq, group = coi_r)
test_ccc <- test_ccc %>% mutate(label = paste0("CCC = ", round(freq_ccc, digits = 2),
                                               " [", round(freq_ccc_lower, digits = 2), "-", round(freq_ccc_upper, digits = 2), "]"))

plot <- ggplot(results_plot, aes(x = allele_freq, y = freq)) +
                    #color = ifelse(population_name %in% 
                    #                 c("simulated_Ghana_pop_043",
                    #                   "simulated_Ghana_pop_003",
                    #                   "simulated_Ghana_pop_051"), "High", "Low"))) + 
  geom_point(alpha = 0.5) + 
  ggtitle("Malaria EM Population-Level Frequency (True vs. Predicted)") + theme_bw() + 
  xlab("True Frequency") + 
  ylab("Malaria EM Frequency") +
  facet_wrap(~coi_r) 

plot + geom_text(
  data    = test_ccc,
  mapping = aes(x = -Inf, y = 0.9, label = label),
  hjust   = -0.05,
  vjust   = 0
) + plot_annotation(caption = paste(strwrap("Comparison of true vs. predicted haplotype-level frequencies. Scatterplots showing the frequency of each haplotype, both true (x-axis) and estimated (y-axis). Concordance Correlation Coefficient (CCC) with 95% CIs is printed on each plot. Scatterplots are faceted by population-level COI.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))
```

Though COI of the simulated population does decrease the correlation, there is still strong agreement between the true and estimated frequencies from malaria.em.

## False Positives and False Negatives

```{r, echo = F, message=F, warning = F}
## allele freq is "real"
## freq is the estimated value

Fig7a <- results %>%
  #filter(exp_freq > 0) %>% 
  ggplot(aes(x = allele_freq, y = as.numeric(freq == 0))) +
  stat_summary_bin(    fun = "mean", geom = "bar", bins = 50, fill = "thistle", color = "white"  ) +
  labs(
    y = "Ratio (FN / Total)", # use ratio to account for base line allele frequency distribution.
    title = "Enrichment of False Negatives"
  ) + theme_bw() + coord_cartesian(ylim = c(0, 0.7)) + xlab("True frequency of missed haplotypes")

Fig7b <- results %>%
  #filter(obs_freq > 0) %>% 
  ggplot(aes(x = freq, y = as.numeric(allele_freq == 0))) +
  stat_summary_bin(fun = "mean", geom = "bar", bins = 50, fill = "thistle", color = "white"  ) +
  labs(
    y = "Ratio (FP / Total)", # use ratio to account for base line allele frequency distribution.
    title = "Enrichment of False Positives"
  ) + theme_bw() + coord_cartesian(ylim = c(0, 0.7)) + xlab("Estimated Frequency (of non-existent haplotypes)")


(Fig7a | Fig7b) + plot_annotation(caption = paste(strwrap("Left: true frequency of values that do not exist in the malaria.em estimation (i.e. those that were falsely missed), accounting for the baseline allele frequency distribution. Right: estimated frequency of values that were predicted by malaria.em despite not being present in the original sample, accounting for baseline allele frequency distribution.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))



make_false_plots <- function(cutoff){

## obviously tradeoff; depends on your use case. Don't trust low frequency haplotypes too much
Fig7a_lim <- results %>%
  #filter(exp_freq > 0) %>% 
  mutate(freq = ifelse(freq < cutoff, 0, freq)) %>%
  filter(!(freq == 0 & allele_freq == 0)) %>%
  ggplot(aes(x = allele_freq, y = as.numeric(freq == 0))) +
  stat_summary_bin(    fun = "mean", geom = "bar", bins = 50, fill = "thistle", color = "white"  ) +
  labs(
    y = "Ratio (FN / Total)", # use ratio to account for base line allele frequency distribution.
    title = "Enrichment of False Negatives"
  ) + theme_bw() + coord_cartesian(ylim = c(0, 0.7)) + xlab("True frequency of missed haplotypes")

Fig7b_lim <- results %>%
  #filter(obs_freq > 0) %>%
  mutate(freq = ifelse(freq < cutoff, 0, freq)) %>%
  filter(!(freq == 0 & allele_freq == 0)) %>%
  ggplot(aes(x = freq, y = as.numeric(allele_freq == 0))) +
  stat_summary_bin(fun = "mean", geom = "bar", bins = 50, fill = "thistle", color = "white"  ) +
  labs(
    y = "Ratio (FP / Total)", # use ratio to account for base line allele frequency distribution.
    title = "Enrichment of False Positives"
  ) + theme_bw() + coord_cartesian(ylim = c(0, 0.7)) + xlab("Estimated Frequency of false haplotypes")


  return((Fig7a_lim | Fig7b_lim) + plot_annotation(title = paste0("With a freq > ", cutoff), caption = paste(strwrap("Left: Ratio of false negative count to total allele count, binned by true frequency of missed allele. Right: Ratio of false positive count to total allele count, binned by the estimated frequency of the false positives.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11)))
  
}
```

As shown above, there are some false negatives, but they represent a relatively low proportion of the overall samples. There are more false positives, but their estimated frequency is low.

## False Positive and Negative Values with Cutoffs {.tabset}

To show how removing low-frequency alleles might impact the overall false postive and negative values, we generated those same plots for a variety of cutoffs.

### 1e-10
```{r FPN -10, echo = F, message = F, warning=FALSE}

make_false_plots(1e-10)

```

### 1e-8
```{r FPN -8, echo = F, message = F, warning=FALSE}

make_false_plots(1e-8)

```

### 1e-6
```{r FPN -6, echo = F, message = F, warning=FALSE}

make_false_plots(1e-6)

```


### 1e-4
```{r FPN 0.0001, echo = F, message = F, warning=FALSE}

make_false_plots(0.0001)

```

### 0.001
```{r FPN 0.001, echo = F, message = F, warning=FALSE}

make_false_plots(0.001)

```

### 0.01
```{r FPN 0.01, echo = F, message = F, warning=FALSE}

make_false_plots(0.01)

```


## Bland-Altman Plots {.tabset}

To show the bias across the range of possible allele frequencies, we generated Bland-Altman plots faceted by a variety of simulation parameters.

### By Country

```{r BA country, echo = F, message = F}
make_bland_altman_plot_by_label(results %>% 
                                  mutate(Country = ifelse(grepl("Ghana", population_name), "Ghana", "Cambodia")) %>%
                                  add_diff_mean(allele_freq, freq), "freq", method = "loess",
                                linecolor = "magenta") + 
  facet_wrap(~Country, nrow = 2) +
  ggtitle("Bland-Altmans by Country") + plot_annotation(caption = paste(strwrap("Bland-Altman plots separated by initial country of the simulations. Point density is shown using color, with more densely-filled segments of the graph shown in more yellow. The pink line represents the best fit through the density.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))

```

### By Group ID

```{r BA Group ID, echo = F, message = F}
make_bland_altman_plot_by_label(results %>%
                                  mutate(`Loci Group` = 
                                           case_match(group_id, "two_lowdiv" ~ "Two (Set A)",
                                                      "two_meddiv" ~ "Two (Set B)", "four" ~ "Four")) %>%
                                  add_diff_mean(allele_freq, freq), "freq" , method = "loess",
                                linecolor = "magenta") + 
  ggtitle("Bland-Altmans by Loci Group") +
  facet_wrap(~`Loci Group`, nrow = 2) + plot_annotation(caption = paste(strwrap("Bland-Altman plots separated by loci group. Point density is shown using color, with more densely-filled segments of the graph shown in more yellow. The pink line represents the best fit through the density.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))

```

### By COI_R

```{r BA COI_r, echo = F, message = F}
make_bland_altman_plot_by_label(results %>% 
                                   mutate(coi_r = case_match(coi_r, 0.25 ~ "Low COI/Prop. Polyclonal",
                                                                0.5 ~ "Med COI/Prop. Polyclonal")) %>%
                                  add_diff_mean(allele_freq, freq), "freq" , method = "loess",
                                linecolor = "magenta") +
  facet_wrap(~coi_r, nrow = 2) + ggtitle("Bland-Altmans by COI") + plot_annotation(caption = paste(strwrap("Bland-Altman plots separated by COI/proportion polyclonal of the simulations. Point density is shown using color, with more densely-filled segments of the graph shown in more yellow. The pink line represents the best fit through the density.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))

```

### By Relatedness

```{r BA relate, echo = F, message = F}
make_bland_altman_plot_by_label(results %>% 
                                   mutate(pop_alpha = 
                                               case_match(pop_alpha, 2.33 ~ "High \nRelatedness",
                                                          9 ~ "Med. \nRelatedness",
                                                          99 ~ "Low \nRelatedness"
                                                          )) %>%
                                  add_diff_mean(allele_freq, freq), "freq" , method = "loess",
                                linecolor = "magenta") +
  facet_wrap(~pop_alpha, nrow = 2) + ggtitle("Bland-Altmans by Relatedness")+ plot_annotation(caption = paste(strwrap("Bland-Altman plots separated by the relatedness of the simulations. Point density is shown using color, with more densely-filled segments of the graph shown in more yellow. The pink line represents the best fit through the density.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))

```

Generally, these Bland-Altman plots echo the high concordance observed above. In
some conditions malaria.em tends to underestimate frequencies above 0.5, as
evidenced by negative values in the y-axis of these plots.


### By Recombination

```{r BA recomb, echo = F, message = F}
make_bland_altman_plot_by_label(results %>% 
                                   mutate(k_s = case_match(k_s, 0.2 ~ "High Recombination",
                                                              0.5 ~ "Med Recombination")) %>%
                                  add_diff_mean(allele_freq, freq), "freq", method = "loess",
                                linecolor = "magenta") +
  facet_wrap(~k_s, nrow = 2) + ggtitle("Bland-Altmans by Recombination") + plot_annotation(caption = paste(strwrap("Bland-Altman plots separated by recombination level of the simulations. Point density is shown using color, with more densely-filled segments of the graph shown in more yellow. The pink line represents the best fit through the density.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))


```

## {-}

# Individual-Level Analyses

Malaria.em provides both population-level and individual results. Here, we present the individual-level results (i.e. the phasing within each individual).

```{r plotting individual results, echo = F, message=F, error = F, warning=F}
conflicted::conflicts_prefer(dplyr::filter)
results_ind <- results_ind %>% mutate(`Loci Group` = 
                                              case_match(group_id, 
                                                         "two_lowdiv" ~ "Two (Set A)",
                                                         "two_meddiv" ~ "Two (Set B)", 
                                                         "four" ~ "Four"))


results_ind_poly <- results_ind %>% filter(grepl("_", key_TOOL))

a <- ggplot(results_ind %>% filter(!is.na(posterior_est)) %>%
                                     mutate(match = factor(match, 
                                                           levels = c(TRUE, FALSE), 
                                                           labels = c("Correctly\nPhased",  "Incorrectly\nPhased"
                                                                      ))), aes(x = match, y=posterior_est)) + 
  geom_violin(alpha = 0.3, aes(fill = match)) + 
  theme_bw() + scale_fill_manual(values = c("maroon", "darkblue")) + facet_wrap(~`Loci Group`) + ggtitle("Violin Plots") + ylab("Posterior Estimate") + labs(fill = "Sample\nAccuracy") + xlab(NULL)

b <- ggplot(results_ind %>% filter(!is.na(posterior_est)) %>%
              mutate(match = factor(match, levels = c(TRUE, FALSE), 
                                    labels = c("Correctly\nPhased",  "Incorrectly\nPhased"))), 
            aes(fill = match, x = posterior_est)) + 
  geom_density(alpha = 0.3)  + theme_bw() + 
  scale_fill_manual(values = c("maroon", "darkblue")) + 
  facet_wrap(~`Loci Group`) + ggtitle("Density Plots") + 
  xlab("Posterior Estimate") + labs(fill = "Sample\nAccuracy") + ylab("Density")



a_poly <- ggplot(results_ind_poly %>% filter(!is.na(posterior_est)) %>%
                                     mutate(match = factor(match, 
                                                           levels = c(TRUE, FALSE), 
                                                           labels = c("Correctly\nPhased",  "Incorrectly\nPhased"
                                                                      ))), aes(x = match, y=posterior_est)) + 
  geom_violin(alpha = 0.3, aes(fill = match)) + 
  theme_bw() + scale_fill_manual(values = c("maroon", "darkblue")) + facet_wrap(~`Loci Group`) + ggtitle("Violin Plots") + ylab("Posterior Estimate") + labs(fill = "Sample\nAccuracy") + xlab(NULL)

b_poly <- ggplot(results_ind_poly %>% filter(!is.na(posterior_est)) %>%
              mutate(match = factor(match, levels = c(TRUE, FALSE), 
                                    labels = c("Correctly\nPhased",  "Incorrectly\nPhased"))), 
            aes(fill = match, x = posterior_est)) + 
  geom_density(alpha = 0.3)  + theme_bw() + 
  scale_fill_manual(values = c("maroon", "darkblue")) + 
  facet_wrap(~`Loci Group`) + ggtitle("Density Plots") + 
  xlab("Posterior Estimate") + labs(fill = "Sample\nAccuracy") + ylab("Density")


results_ind_summary <- results_ind %>% group_by(population_name, group_id) %>%
  dplyr::summarize(prop_perfect = sum(match)/ n(), .groups = "keep") %>% 
  mutate(country = str_split_i(population_name, "_", 2),
         id = str_split_i(population_name, "_", 4),
         short_name = paste(country, id, sep="_")) 

results_ind_summary_poly <- results_ind_poly %>% group_by(population_name, group_id) %>%
  dplyr::summarize(prop_perfect = sum(match)/ n(), .groups = "keep") %>% 
  mutate(country = str_split_i(population_name, "_", 2),
         id = str_split_i(population_name, "_", 4),
         short_name = paste(country, id, sep="_")) 

results_ind_summary_55 <- results_ind %>%
  filter(posterior_est > 0.55) %>% 
  group_by(population_name, group_id,) %>%
  dplyr::summarize(prop_perfect = sum(match)/ n(), .groups = "keep") %>% 
  mutate(country = str_split_i(population_name, "_", 2),
         id = str_split_i(population_name, "_", 4),
         short_name = paste(country, id, sep="_"))

samples_lost_summary_55 <- results_ind %>% 
  group_by(population_name, group_id) %>%
  summarize(Lost = (sum(posterior_est < 0.55) / sum(posterior_est>=0)))

all_accuracy <- make_boxplots(results_ind_summary, "prop_perfect", 
              "Proportion of Samples Correctly Phased", "Efficacy of Sample-Level Phasing")


post_55 <- make_boxplots(results_ind_summary_55, "prop_perfect", 
              "Proportion of Samples Correctly Phased", "Efficacy of Sample-Level Phasing (Post. Est > 0.55)")

prop_samples_lost <- make_boxplots(samples_lost_summary_55, "Lost", "Proportion of Samples Removed",
                                   "How much data was removed with Post. Est > 0.55?")

all_accuracy + plot_annotation(caption = 
                                 paste(strwrap("Individual-level accuracy. Each point represents a single simulated population. The y-axis shows the proportion of samples within that population that were phased completely correctly.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))
```

As shown above, the accuracy among samples taken from regions with a medium COI as opposed to a low one is rather low. Many of the populations have \< 75% of the samples correctly phased in this case.

malaria.em outputs posterior probabilities for the phasing predictions - these
may be useful as a way to filter out samples with low confidence phasing.


```{r echo = F, message=F, error = F, warning=F}
a/b + plot_annotation(title = "Comparing Posterior Estimates by Correctness of Sample") + plot_annotation(caption = 
                                 paste(strwrap("Top: Violin plots showing the posterior probabilities for phasing separated by whether the samples were correctly or incorrectly phased. Bottom: The same data as a density plot.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))
```

The posterior estimates generated for each individual were generally higher when the matching was correct. We chose a posterior estimate of > 0.55 as a cutoff for being more likely to be accurate than not in this dataset, but this value will require further validation.

```{r echo = F, message=F, error = F, warning=F}
post_55  + plot_annotation(caption = 
                                 paste(strwrap("Individual-level accuracy with posterior estimate > 0.55. Samples with lower confidence were removed. Each point represents a single simulated population, and the y-axis shows the proportion of samples within that population that were phased completely correctly.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))
```

Although adding a cutoff based on the posterior probability improves accuracy in the medium COI populations, this does lead to substantial loss of data, as shown below.

```{r echo = F, message=F, error = F, warning=F}
prop_samples_lost  + plot_annotation(caption = 
                                 paste(strwrap("Each point represents a single simulated population, and the y-axis shows the proportion of samples removed with a posterior probability cutoff of 0.55.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))

```

### Polyclonal subset

Given the clear effect of polyclonality on malaria.em performance, we further subsetted the analyses to only consider polyclonal samples (in the loci group in question).

```{r polyclonal subset, echo = F, message=F, error = F, warning=F}
a_poly / b_poly + plot_annotation(title = "Comparing Posterior Estimates by Correctness of Sample (Polyclonal)") + plot_annotation(caption = 
                                 paste(strwrap("Top: Violin plots showing the posterior probabilities for phasing separated by whether the samples were correctly or incorrectly phased. Bottom: The same data as a density plot.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))



all_accuracy_poly <- make_boxplots(results_ind_summary_poly, "prop_perfect", "Proportion of Samples Correctly", "Efficacy of Sample-Level Phasing (Polyclonal Only)")  + plot_annotation(caption = 
                                 paste(strwrap("Individual-level accuracy. Each point represents a single simulated population subsetted to only the polyclonal samples in that population, and the proportion of samples within that population that were phased completely correctly.",width = 100), collapse = "\n")) &
  theme(plot.caption = element_text(hjust = 0, size = 11))
```


## Summary and Next Steps

Malaria.em seems to be working! However, there are a lot of open questions around exactly how well it works on various allele sets. Additionally, the computational constraints limit wide-spread applicability and still need to be quantified for malaria.em.


* On a population-level, malaria.em is quite accurate. For all populations, concordance between the estimated and overall phased population-level frequencies was > 0.9.
* On an individual-level, malaria.em seems to phase samples from low-COI populations accurately. Accuracy improves upon applying a posterior-estimate filter. 
  + Our sampling showed that a posterior estimate cutoff of 0.55 seemed appropriate, but this should be further validated.
* Malaria.em appears most sensitive to population COI/ polyclonality (i.e. is generally accurate on populations with a high proportion of single infections.) 
* There are minimal false negatives, and most false positives are estimated at very low frequencies.
  + Applying a filter to exclude predicted haplotypes at a very low frequency might improve accuracy, but more validation is required to determine that cutoff. In our simulated dataset, removing samples at a predicted frequency < 0.001 appears to improve accuracy.

