---
title: Benchmarking Results - Frequency Tools Single Locus Variants
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: true
fig-dpi: 300
fig-align: center
number-sections: true
anchor-sections: true
bibliography: [references.bib]
biblio-style: apalike
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)

library(ComplexHeatmap)


get_color_scale_breaks <-function(mat_dat){
  min_val = min(mat_dat, na.rm = T)
  max_val = max(mat_dat, na.rm = T)
  
  return(c(min_val, min_val + (max_val - min_val)/2, max_val))
}
```

## Tools 

* naive single locus freq (estimating frequency based on presence/absence or weighted by within sample frequency)
* IDM[@Tsoungui_Obama2022-gz]
* estimating from microhaplotype frequencies estimated by 1 step EM algorithim from dcifer[@Gerlovina2022-ya]

## Set up 

### Libraries used in the analysis below 

```{r, echo=FALSE}
source("../tutorials/scripts/benchmarking_functions.R")
```

```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-line-numbers: true
#| filename: ../tutorials/scripts/benchmarking_functions.R
#| file: ../tutorials/scripts/benchmarking_functions.R

```

## Load in the Benchmarking Outputs

```{r}
inp <- list(
  sim_info_dir = "../data/benchmarking//simulations/SpotMalaria/", 
  madhatter_res_dir = "../data/benchmarking/tool_results/madhatter/"
)

inp_per_tools <- list(
  # single locus allele frequencies 
  "idm" = list(
    tool_res_dir = "../data/benchmarking/tool_results/IDM/", 
    tool_params = "../data/benchmarking/tool_results/idm.json"
  ), 
  "naive_slaf" = list(
    tool_res_dir = "../data/benchmarking/tool_results/naive_slaf/", 
    tool_params = "../data/benchmarking/tool_results/naive_slaf.json"
  ), 
  "dcifer_slaf_from_mhap_freqs" = list(
    tool_res_dir = "../data/benchmarking/tool_results/dcifer_slaf_from_mhap_freqs/", 
    tool_params = "../data/benchmarking/tool_results/dcifer_slaf_from_mhap_freqs.json"
  )
)



```

First, read in all of the tool results, tool parameters, expected simulated values, and simulation parameters. 

### Reading in expected truths 

Read per locus aa variant freq/prev tables 
```{r}
expected_population_aa_variants_freqs_prevs <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_aa_variants_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(),
        population_name = col_character(), 
        pop = col_character(), 
        country = col_character(), 
        GeneName = col_character(), 
        transcript_id = col_character(), 
        aa_position = col_integer(), 
        aa = col_character(), 
        is_ref = col_logical(), 
				allele_count = col_integer(),
				sample_count = col_integer(),
				total_pop_alleles = col_integer(),
				total_pop_samples = col_integer(),
				unique_alleles = col_integer(),
				freq = col_double(),
				prev  = col_double()
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  mutate(variant = paste0(transcript_id, ":", aa_position, ":", aa)) 

colnames(expected_population_aa_variants_freqs_prevs)[7:15] = paste0("expected_",
                                                                               colnames(expected_population_aa_variants_freqs_prevs)[7:15])


expected_population_aa_variants_freqs_prevs_sel = expected_population_aa_variants_freqs_prevs %>% 
  select(population_name, variant, expected_freq)
```

Read in all simulation parameters into one tibble. Encode different population parameters as descriptive factors. 

```{r}

sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>% 
  mutate(
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>% 
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  )) %>% 
  mutate(pop_cat = factor(pop_cat, levels = sort(unique((.$pop_cat)))), 
         coi_cat = factor(coi_cat, levels = sort(unique((.$coi_cat)))), 
         between_sample_related_cat = factor(between_sample_related_cat, levels = sort(unique((.$between_sample_related_cat)))), 
         nonrecombinant_cat = factor(nonrecombinant_cat, levels = sort(unique((.$nonrecombinant_cat)))), 
         meiosis_cat = factor(meiosis_cat, levels = sort(unique((.$meiosis_cat))))) %>% 
  select(-sample_count)

create_dt(sim_params)
```

## single locus variants 

naive single locus freq


### Read in naive single locus freq results

naive single locus freq outputs information on frequency and prevalence of alleles 

Getting the frequency and prevs from the `slaf.tsv.gz` files 

```{r}
naive_slaf_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$naive_slaf$tool_res_dir, 
      pattern = "^slaf.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```


The naive estimator has two modes that it can be ran in read_count_prop (weighting the frequencies by within sample allele frequencies) and presence_absence (calculating frequencies simply by their presence and absence in samples, if all samples were monocloanl this would be the same as prevalance)

```{r}
# Read in tool parameters
naive_slaf_tool_params <- jsonlite::fromJSON(inp_per_tools$naive_slaf$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_method = paramset.method,
  )

create_dt(naive_slaf_tool_params)
```



### Read in dcifer slaf from microhaplotype freqs 

dcifer slaf from microhaplotype freqs outputs information on frequency per haplotype  

Getting the frequency and prevs from the `slaf.tsv.gz` files 

```{r}
dcifer_slaf_from_mhap_freqs_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$dcifer_slaf_from_mhap_freqs$tool_res_dir, 
      pattern = "^slaf.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

Three different parameters, each set uses a different nth-max (lrank) for the COI calculation that is then used for the calculating the microhaplotype frequency with a 1 step EM algorithm. 

```{r}
# Read in tool parameters
dcifer_slaf_from_mhap_freqs_tool_params <- jsonlite::fromJSON(inp_per_tools$dcifer_slaf_from_mhap_freqs$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_lrank = paramset.lrank,
    paramset_qstart = paramset.qstart
  )

create_dt(dcifer_slaf_from_mhap_freqs_tool_params)
```


### Read in IDM results


IDM outputs information on frequency of alleles 

Getting the frequency and prevs from the `IDM_summary.tsv.gz` files 

```{r}
idm_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$idm$tool_res_dir, 
      pattern = "^IDM_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

idm 

```{r}
# Read in tool parameters
idm_tool_params <- jsonlite::fromJSON(inp_per_tools$idm$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_model = paramset.model ,
    paramset_lambda_initial = paramset.lambda_initial ,
    paramset_eps_initial = paramset.eps_initial,
  )

create_dt(idm_tool_params)
```


### Combining 

```{r}

add_expected_freqs <- function(summary_dat, expected_dat, tool_params, tool_name) {
  ## expand expected results for all parameter sets so they all get expected results
  ## fill the unmatched expected with 0's and with unmatched observed (false haps) with 0's
  summary_dat_out = summary_dat %>%
    left_join(tool_params)  %>%
    full_join(expected_dat %>%
                crossing(paramset_id = unique(sort(
                  summary_dat$paramset_id
                )))) %>%
    mutate(freq = ifelse(is.na(freq), 0, freq)) %>%
    mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq)) %>%
    mutate(tool =  tool_name) %>%
    mutate(tool_paramset_id = paste0(tool, "-", paramset_id))
  
  return(summary_dat_out)
}

naive_slaf_freq_summary_out = add_expected_freqs(naive_slaf_freq_summary, 
                                                 expected_population_aa_variants_freqs_prevs_sel,
                                                 naive_slaf_tool_params,
                                                 "naive_slaf")

dcifer_slaf_from_mhap_freqs_freq_summary_out = add_expected_freqs(dcifer_slaf_from_mhap_freqs_freq_summary,
                                                                  expected_population_aa_variants_freqs_prevs_sel, 
                                                                  dcifer_slaf_from_mhap_freqs_tool_params,
                                                                  "dcifer_slaf_from_mhap_freqs")

idm_freq_summary_out = add_expected_freqs(idm_freq_summary,
                                          expected_population_aa_variants_freqs_prevs_sel,
                                          idm_tool_params,
                                          "idm")

```



### Summarizing expected frequenies

```{r}
#bad variant positions 
#PF3D7_0731500.1:401, PF3D7_0731500.1:402, PF3D7_1343700.1:622
#PF3D7_0807800.1:380 for now, until fix for the translate of M instead of L (now fixed)

# PF3D7_0408900.2 is in calculated but not the expect, PF3D7_0408900.1 is 
# PF3D7_1133400.1:404 ama1 edge codon that is missing in some seqs cause of the edge trimming 
# PF3D7_1133400.1:439, PF3D7_1335100.1:212, PF3D7_1335900.1:90 is non-biallelic so was not included in the loci of interest column it seems 

# PF3D7_1343700.1:568 very small variation, so not included 

# PF3D7_1335100.1:231 not included in loci of interest, but is in the middle of several other ones in MSP7 so not clear 

expected_population_aa_variants_freqs_prevs_sel_filt = expected_population_aa_variants_freqs_prevs_sel %>% 
  filter(!grepl("PF3D7_0731500.1:40[1234]", variant)) %>% 
  filter(!grepl("PF3D7_1343700.1:622", variant)) %>% 
  # filter(!grepl("PF3D7_0807800.1:380", variant)) %>% 
  filter(!grepl("PF3D7_0408900.2", variant)) %>% 
  filter(!grepl("PF3D7_1133400.1:404", variant)) %>% 
  filter(!grepl("PF3D7_1133400.1:439", variant)) %>% 
  filter(!grepl("PF3D7_1335100.1:212", variant)) %>% 
  filter(!grepl("PF3D7_1335900.1:90", variant)) %>% 
  filter(!grepl("PF3D7_1343700.1:568", variant))%>% 
  filter(!grepl("PF3D7_1335100.1:231", variant))%>% 
  mutate(variant_pos = gsub(":[A-Z*]", "", variant))


# calculating he 
expected_population_aa_variants_freqs_prevs_sel_filt_he = expected_population_aa_variants_freqs_prevs_sel_filt %>% 
  group_by(population_name, variant_pos) %>% 
  summarise(he = 1 - sum(expected_freq ^2), 
            variant_n = n_distinct(variant))

expected_population_aa_variants_freqs_prevs_sel_filt_he_sum = expected_population_aa_variants_freqs_prevs_sel_filt_he %>% 
  group_by(variant_pos) %>% 
  summarise(quick_summary(he))
# filtering to just sites that have at least some variation in any of the simulated populations 

expected_population_aa_variants_freqs_prevs_sel_filt_he_sum_invariant = expected_population_aa_variants_freqs_prevs_sel_filt_he_sum %>% 
  filter(q100 == 0)
expected_population_aa_variants_freqs_prevs_sel_filt_he_sum_variant = expected_population_aa_variants_freqs_prevs_sel_filt_he_sum %>% 
  filter(q100 != 0) %>% 
  arrange(mean) %>% 
  mutate(variant_pos = factor(variant_pos, levels = unique(.$variant_pos)))

expected_population_aa_variants_freqs_prevs_sel_filt_he_variant = expected_population_aa_variants_freqs_prevs_sel_filt_he %>% 
  filter(variant_pos %!in% expected_population_aa_variants_freqs_prevs_sel_filt_he_sum_invariant$variant_pos)


expected_population_aa_variants_freqs_prevs_sel_filt_variant = expected_population_aa_variants_freqs_prevs_sel_filt %>% 
  filter(variant_pos %!in% expected_population_aa_variants_freqs_prevs_sel_filt_he_sum_invariant$variant_pos)
```

Ranges of he for loci 

```{r}
ggplot(expected_population_aa_variants_freqs_prevs_sel_filt_he_variant %>% 
         left_join(sim_params) %>% 
         mutate(variant_pos = factor(variant_pos, levels = expected_population_aa_variants_freqs_prevs_sel_filt_he_sum_variant$variant_pos))
       ) + 
  geom_boxplot(aes(x = variant_pos, y = he, color = factor(variant_n))) + 
  custom_ggplot_theme_xRotate + 
  scale_color_manual(values = colorPalette_08) + 
  facet_grid(country~.)

```

Some variants are invariant in one ancestor population but may be highly variant in the other 

```{r}
#| fig-width: 30
#| fig-height: 10
# Get ordered variant positions
variant_levels <- expected_population_aa_variants_freqs_prevs_sel_filt_he_sum_variant$variant_pos

# Create shading data frame
n_variants <- length(variant_levels)
shade_df <- data.frame(
  variant_pos = factor(variant_levels, levels = variant_levels),
  xmin = seq_along(variant_levels) - 0.5,
  xmax = seq_along(variant_levels) + 0.5,
  fill = rep(c("gray", "white"), length.out = n_variants)
)

ggplot(expected_population_aa_variants_freqs_prevs_sel_filt_he_variant %>% 
         left_join(sim_params) %>% 
         mutate(variant_pos = factor(variant_pos, levels = variant_levels))
       ) + 
  geom_rect(data = shade_df,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            inherit.aes = FALSE, alpha = 0.4) +
  scale_fill_manual(values = c("gray" = "grey50", "white" = "white"), guide = "none") +
  geom_boxplot(aes(x = variant_pos, y = he, color = factor(country))) + 
  custom_ggplot_theme_xRotate + 
  scale_color_manual(values = colorPalette_08)


```

```{r}
ggplot(expected_population_aa_variants_freqs_prevs_sel_filt_variant) + 
  geom_histogram(aes(x = expected_freq), binwidth = 0.01, boundary = 0)
```



### Analyzing results 

```{r}
combined_slfreq = bind_rows(
  naive_slaf_freq_summary_out %>% 
    select(population_name, paramset_id, variant, freq, expected_freq, tool, tool_paramset_id), 
  dcifer_slaf_from_mhap_freqs_freq_summary_out %>% 
    select(population_name, paramset_id, variant, freq, expected_freq, tool, tool_paramset_id), 
  idm_freq_summary_out %>% 
    select(population_name, paramset_id, variant, freq, expected_freq, tool, tool_paramset_id)
) %>% 
  filter(!grepl("PF3D7_0731500.1:40[1234]", variant)) %>% 
  filter(!grepl("PF3D7_1343700.1:622", variant)) %>% 
  # filter(!grepl("PF3D7_0807800.1:380", variant)) %>% 
  filter(!grepl("PF3D7_0408900.2", variant)) %>% 
  filter(!grepl("PF3D7_1133400.1:404", variant)) %>% 
  filter(!grepl("PF3D7_1133400.1:439", variant)) %>% 
  filter(!grepl("PF3D7_1335100.1:212", variant)) %>% 
  filter(!grepl("PF3D7_1335900.1:90", variant)) %>% 
  filter(!grepl("PF3D7_1343700.1:568", variant))%>% 
  filter(!grepl("PF3D7_1335100.1:231", variant))%>% 
  mutate(variant_pos = gsub(":[A-Z*]", "", variant)) %>% 
  filter(variant_pos %!in% expected_population_aa_variants_freqs_prevs_sel_filt_he_sum_invariant$variant_pos) %>% 
  filter(variant_pos %in% expected_population_aa_variants_freqs_prevs_sel_filt_he_sum_variant$variant_pos)

```


### Breakdown by variant 

```{r}
combined_slfreq_ccc_by_tool_variant_pos_params = combined_slfreq %>% 
  compute_ccc(expected_freq, freq, c(tool, variant_pos, paramset_id, tool_paramset_id))

```


```{r}
# looking at just at frequencies that aren't 1, it's important that sites with a frequency of 1 are also correct but currently 
# no tool has any issues with getting a freq of 1, having it being over representation makes evaluation biased  

combined_slfreq_ccc_by_tool_variant_pos_params_filt = combined_slfreq %>% 
  filter(expected_freq < 1) %>% 
  compute_ccc(expected_freq, freq, c(tool, variant_pos, paramset_id, tool_paramset_id))

combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt = combined_slfreq_ccc_by_tool_variant_pos_params_filt %>% 
  filter(grepl("(naive_slaf|idm_paramset20|dcifer_slaf_from_mhap_freqs_paramset)", paramset_id))

combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp = combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt %>% 
  ungroup() %>% 
  select(variant_pos, tool_paramset_id, freq_ccc) %>% 
  spread(tool_paramset_id, freq_ccc)

combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp_mat = as.matrix(combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp[,2:ncol(combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp)])
rownames(combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp_mat) = combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp$variant_pos


library(circlize)
col_fun = colorRamp2(c(min(combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp_mat, na.rm = T),
                       mean(combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp_mat, na.rm = T),1), c( "#3288bd", "#ffffbf", "#d53e4f"))

variant_meta = tibble(
  variant_pos = rownames(combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp_mat)
) %>% 
  left_join(expected_population_aa_variants_freqs_prevs_sel_filt_he_sum %>% 
              select(variant_pos, mean) %>% 
              dplyr::rename(mean_he = mean)) %>% 
  select(-variant_pos) %>% 
  as.data.frame()
row_anno = rowAnnotation(
  df = variant_meta, 
  col = list(
    mean_he = colorRamp2(c(min(variant_meta$mean_he, na.rm = T),
                       median(variant_meta$mean_he, na.rm = T),
                     max(variant_meta$mean_he, na.rm = T)),
                   c( "#ffffcc", "#feb24c", "#fc4e2a"))
  )
)

```

```{r}
#| fig-width: 8
#| fig-height: 30
combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp_mat_hp = Heatmap(combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp_mat, 
                                                                 name = "concordance", 
                                                                 cluster_rows = T, 
                                                                 row_gap = unit(0.5, "cm"),
                                                                 col = col_fun,
                                                                 rect_gp = gpar(col = "black", lwd = 1),
                                                                 left_annotation = row_anno,
                                                                 # row_split = tool_meta$tool, 
                                                                cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(
      sprintf("%.2f", combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp_mat[i, j]),
      x,
      y
      #,gp = gpar(fontsize = 8, col = "black")
    )
  })

draw(combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp_mat_hp, 
     background = "transparent", 
     padding = unit(c(100, 5, 5, 5), "mm"), 
     merge_legends = T)

cairo_pdf("combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp_mat_hp.pdf", width = 8, height = 30, bg = "transparent")
draw(combined_slfreq_ccc_by_tool_variant_pos_params_filt_filt_sp_mat_hp, background = "transparent", 
     padding = unit(c(100, 5, 5, 5), "mm"), 
     merge_legends = T)
dev.off()

```



### Overall 

```{r}
combined_slfreq_ccc_by_tool_params = combined_slfreq %>% 
  filter(expected_freq != 1) %>% 
  compute_ccc(expected_freq, freq, c(tool, paramset_id, tool_paramset_id))

combined_slfreq_ccc_by_tool_params_best = combined_slfreq_ccc_by_tool_params %>% 
  group_by(tool) %>% 
  slice_max(order_by = freq_ccc, with_ties = F)

```


The different IDM parameters of model and initial values have no effect on results. There is very little variation in the different lrank (coi estimator) for dcifer based estimate, the presence/absence naive estimator does worse than the within sample weighted frequency one. 

```{r}
combined_slfreq_ccc_by_tool_params_plot_per_paramset = ggplot(combined_slfreq_ccc_by_tool_params) + 
  geom_col(aes(x = paramset_id, y = freq_ccc, fill = tool)) + 
  geom_errorbar(aes(x = paramset_id, ymin = freq_ccc_lower, ymax = freq_ccc_upper)) + 
  scale_fill_manual(values = colorPalette_15) + 
  custom_ggplot_theme_xRotate + 
  geom_abline(slope = 0, intercept = 0.95, color = "red", linetype = "dotted")

create_dt(combined_slfreq_ccc_by_tool_params)

print(combined_slfreq_ccc_by_tool_params_plot_per_paramset)

cairo_pdf("combined_slfreq_ccc_by_tool_params_plot_per_paramset.pdf", width = 15, height = 8)
print(combined_slfreq_ccc_by_tool_params_plot_per_paramset)
dev.off()
```



### Best per population COI parameters 

The mean population COI makes a large difference to which parameter set works best for which tool. This is due to the max possible COIs in each population. 

```{r}
combined_slfreq_ccc_per_tool_per_params_per_coi_cat = combined_slfreq %>% 
  filter(grepl("(naive_slaf|idm_paramset20|dcifer_slaf_from_mhap_freqs_paramset)", paramset_id)) %>% 
  left_join(sim_params) %>% 
  filter(expected_freq != 1) %>% 
  compute_ccc(
    expected_freq, freq, 
    c(tool, paramset_id, coi_cat)
  )

combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp = combined_slfreq_ccc_per_tool_per_params_per_coi_cat %>% 
  ungroup() %>% 
  select(paramset_id, coi_cat, freq_ccc) %>% 
  spread(coi_cat, freq_ccc)

combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat = as.matrix(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp[,2:ncol(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp)])
rownames(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat) = combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp$paramset_id

tool_meta = tibble(
  paramset_id = rownames(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat)
)%>% 
  mutate(tool = gsub("_coi_(freq|variant)", "", gsub("_p.*", "", paramset_id)))


library(circlize)
col_fun = colorRamp2(get_color_scale_breaks(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat), c( "#3288bd", "#ffffbf", "#d53e4f"))

```

```{r}
#| fig-width: 8
#| fig-height: 8
combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp_hp = Heatmap(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat, 
                                                                 name = "concordance", 
                                                                 cluster_rows = F, 
                                                                 row_gap = unit(0.5, "cm"),
                                                                 col = col_fun,
                                                                 rect_gp = gpar(col = "black", lwd = 1),
                                                                 row_split = tool_meta$tool, 
                                                                   cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(
      sprintf("%.2f", combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat[i, j]),
      x,
      y
      #,gp = gpar(fontsize = 8, col = "black")
    )
  })

draw(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp_hp)

cairo_pdf("freq_ccc_per_coi_cat_per_tool_params.pdf", width = 8, height = 8, bg = "transparent")
draw(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_sp_hp, background = "transparent")
dev.off()

```




### Best per population COI parameters and relatedness 

The amount of between sample relatedness also plays a significant role in performance 

```{r}
combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness = combined_slfreq %>% 
  filter(grepl("(naive_slaf|idm_paramset20|dcifer_slaf_from_mhap_freqs_paramset)", paramset_id)) %>% 
  left_join(sim_params) %>% 
  filter(expected_freq != 1)  %>% 
  mutate(coi_cat_related_ness = paste0(coi_cat, "\n", between_sample_related_cat)) %>% 
  compute_ccc(
    expected_freq, freq, 
    c(tool, paramset_id, coi_cat_related_ness)
  )

combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp = combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness %>% 
  ungroup() %>% 
  select(paramset_id, coi_cat_related_ness, freq_ccc) %>% 
  spread(coi_cat_related_ness, freq_ccc)

combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat = as.matrix(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp[,2:ncol(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp)])
rownames(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat) = combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp$paramset_id

tool_meta = tibble(
  paramset_id = rownames(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat)
) %>% 
  mutate(tool = gsub("_coi_(freq|variant)", "", gsub("_p.*", "", paramset_id)))

library(circlize)
col_fun = colorRamp2(get_color_scale_breaks(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat), c( "#3288bd", "#ffffbf", "#d53e4f"))

```

```{r}
#| fig-width: 12
#| fig-height: 10
combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_hp = Heatmap(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat, 
                                                                 name = "concordance", 
                                                                 cluster_rows = F, 
                                                                 cluster_columns = F,
                                                                 row_gap = unit(0.5, "cm"),
                                                                 col = col_fun,
                                                                 rect_gp = gpar(col = "black", lwd = 1),
                                                                 row_split = tool_meta$tool, 
                                                                   cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(
      sprintf("%.2f", combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat[i, j]),
      x,
      y
      #,gp = gpar(fontsize = 8, col = "black")
    )
  })

draw(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_hp)

cairo_pdf("freq_ccc_per_coi_cat_related_ness_per_tool_params.pdf", width = 12, height = 10, bg = "transparent")
draw(combined_slfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_hp, background = "transparent")
dev.off()

```


### Best per freq bins 

Breaking down performance per tool per bins of frequency, this way performance can be broken down over specific frequencies 

```{r}
combined_slfreq_ccc_per_coi_bin_per_tool = combined_slfreq %>% 
  filter(grepl("(naive_slaf|idm_paramset20|dcifer_slaf_from_mhap_freqs_paramset)", paramset_id)) %>% 
  left_join(sim_params) %>% 
  filter(expected_freq != 1)  %>% 
  mutate(expected_freq_bin = cut(expected_freq, breaks =seq(0, 1, 0.2), include.lowest = T, right = F)) %>% 
  group_by(expected_freq_bin) %>% 
  filter(n_distinct(expected_freq) > 1) %>% 
  ungroup() %>% 
  compute_ccc(
    expected_freq, freq, 
    c(tool, paramset_id, expected_freq_bin)
  )


combined_slfreq_ccc_per_coi_bin_per_tool_sp = combined_slfreq_ccc_per_coi_bin_per_tool %>% 
  ungroup() %>% 
  select(paramset_id, expected_freq_bin, freq_ccc) %>% 
  spread(expected_freq_bin, freq_ccc)

combined_slfreq_ccc_per_coi_bin_per_tool_sp_mat = as.matrix(combined_slfreq_ccc_per_coi_bin_per_tool_sp[,2:ncol(combined_slfreq_ccc_per_coi_bin_per_tool_sp)])
rownames(combined_slfreq_ccc_per_coi_bin_per_tool_sp_mat) = combined_slfreq_ccc_per_coi_bin_per_tool_sp$paramset_id

tool_meta = tibble(
  paramset_id = rownames(combined_slfreq_ccc_per_coi_bin_per_tool_sp_mat)
)%>% 
  mutate(tool = gsub("_coi_(freq|variant)", "", gsub("_p.*", "", paramset_id)))


library(circlize)
col_fun = colorRamp2(get_color_scale_breaks(combined_slfreq_ccc_per_coi_bin_per_tool_sp_mat), c( "#3288bd", "#ffffbf", "#d53e4f"))

```

```{r}
#| fig-width: 16
#| fig-height: 10
combined_slfreq_ccc_per_coi_bin_per_tool_sp_hp = Heatmap(combined_slfreq_ccc_per_coi_bin_per_tool_sp_mat, 
                                                                 name = "concordance", 
                                                                 cluster_rows = F,
                                                      cluster_columns = F,
                                                                 row_gap = unit(0.5, "cm"),
                                                                 col = col_fun,
                                                                 rect_gp = gpar(col = "black", lwd = 1),
                                                                 row_split = tool_meta$tool, 
                                                                   cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(
      sprintf("%.2f", combined_slfreq_ccc_per_coi_bin_per_tool_sp_mat[i, j]),
      x,
      y
      #,gp = gpar(fontsize = 8, col = "black")
    )
  })

draw(combined_slfreq_ccc_per_coi_bin_per_tool_sp_hp)

cairo_pdf("combined_slfreq_ccc_per_coi_bin_per_tool.pdf", width = 16, height = 10, bg = "transparent")
draw(combined_slfreq_ccc_per_coi_bin_per_tool_sp_hp, background = "transparent")
dev.off()

```




Getting the absolute deviation from the expected values across different frequency bins 

```{r}
#| fig-width: 10
#| fig-height: 5


combined_slfreq_mae_by_tool_params_by_bins = combined_slfreq %>% 
  mutate(expected_bins = cut(expected_freq, seq(0, 1, by = 0.1)) ) %>% 
  compute_mae(expected_freq, freq, c(population_name, expected_bins, tool, paramset_id, tool_paramset_id) ) 


combined_slfreq_mae_by_tool_params_by_bins_sel= combined_slfreq_mae_by_tool_params_by_bins %>% 
  filter(paramset_id %in% c("dcifer_slaf_from_mhap_freqs_paramset2", "idm_paramset1", "naive_slaf_paramset1", "naive_slaf_paramset2")) %>% 
  mutate(tool_lab = case_when(
    "naive_slaf_paramset1" == paramset_id ~ "naive_slaf_wsaf_prop", 
    "naive_slaf_paramset2" == paramset_id ~ "naive_slaf_absence_presence",
    T ~ tool
  ))


combined_slfreq_mae_by_tool_params_by_bins_sel_plot = ggplot(combined_slfreq_mae_by_tool_params_by_bins_sel %>% 
         filter(!is.na(expected_bins))) + 
  geom_boxplot(aes(x = expected_bins, y = freq_mean_abs_error, color = tool_lab), position = "dodge") + 
  scale_color_manual(values = colorPalette_08) + 
  custom_ggplot_theme_xRotate

print(combined_slfreq_mae_by_tool_params_by_bins_sel_plot)

cairo_pdf("combined_slfreq_mae_by_tool_params_by_bins_sel.pdf", width = 10, height = 5)
print(combined_slfreq_mae_by_tool_params_by_bins_sel_plot)
dev.off()
```


Getting pairwise comparisons 
```{r}

combined_slfreq_sel = combined_slfreq %>% 
  filter(grepl("(naive_slaf|idm_paramset20|dcifer_slaf_from_mhap_freqs_paramset)", paramset_id))

combined_slfreq_sel_sel = combined_slfreq_sel %>% 
  ungroup() %>% 
  select(population_name, variant, variant_pos, expected_freq) %>% 
  unique() %>% 
  dplyr::rename(freq = expected_freq) %>% 
  mutate(paramset_id = "expected_freq") %>% 
  bind_rows(
    combined_slfreq_sel %>% 
  ungroup() %>% 
  select(population_name, variant, variant_pos, paramset_id, freq)
  )

combined_slfreq_sel_sel_spread = combined_slfreq_sel_sel %>% 
  spread(paramset_id, freq)


```

The different dcifer params are highly correlated with each other, showing that getting the exact COI correct isn't necessary to get to get ideal results.  

```{r}
#| fig-width: 10
#| fig-height: 10
combined_slfreq_sel_sel_spread_all_params_pairs_plot_dcifer = ggpairs(
  combined_slfreq_sel_sel_spread %>% 
    filter(expected_freq < 1),
  columns = c(
    "expected_freq",
    "dcifer_slaf_from_mhap_freqs_paramset1",
    "dcifer_slaf_from_mhap_freqs_paramset2",
    "dcifer_slaf_from_mhap_freqs_paramset3"
  ),
  lower = list(continuous = wrap(my_hex)),
  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
  upper = list(continuous = wrap(
    my_corr_ccc, digits = 3, size = 3.8
  )),
  title = "all"
)
print(combined_slfreq_sel_sel_spread_all_params_pairs_plot_dcifer)

cairo_pdf("combined_slfreq_sel_sel_spread_all_params_pairs_plot_dcifer.pdf", width = 10, height = 10)
print(combined_slfreq_sel_sel_spread_all_params_pairs_plot_dcifer)
dev.off()
```

```{r}
#| fig-width: 13
#| fig-height: 13
combined_slfreq_sel_sel_spread_all_params_pairs_plot_subset = ggpairs(
  combined_slfreq_sel_sel_spread %>% 
    filter(expected_freq < 1),
  columns = c(
    "expected_freq",
    "dcifer_slaf_from_mhap_freqs_paramset2",
    "idm_paramset20",
    "naive_slaf_paramset1",
    "naive_slaf_paramset2"
  ),
  lower = list(continuous = wrap(my_hex)),
  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
  upper = list(continuous = wrap(
    my_corr_ccc, digits = 3, size = 3.8
  )),
  title = "all"
)



cairo_pdf("combined_slfreq_sel_sel_spread_all_params_pairs_plot_subset.pdf", width = 13, height = 13)
print(combined_slfreq_sel_sel_spread_all_params_pairs_plot_subset)
dev.off()

```

```{r}
#| fig-width: 13
#| fig-height: 13
combined_slfreq_sel_sel_spread_all_params_pairs_plot_subset_low_coi = ggpairs(
  combined_slfreq_sel_sel_spread %>% 
    left_join(sim_params) %>% 
    filter(expected_freq < 1, 
           coi_cat == "MeanCOI=1.26"),
  columns = c(
    "expected_freq",
    "dcifer_slaf_from_mhap_freqs_paramset2",
    "idm_paramset20",
    "naive_slaf_paramset1",
    "naive_slaf_paramset2"
  ),
  lower = list(continuous = wrap(my_hex)),
  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
  upper = list(continuous = wrap(
    my_corr_ccc, digits = 3, size = 3.8
  )),
  title = "MeanCOI=1.26"
)

combined_slfreq_sel_sel_spread_all_params_pairs_plot_subset_high_coi = ggpairs(
  combined_slfreq_sel_sel_spread %>% 
    left_join(sim_params) %>% 
    filter(expected_freq < 1, 
           coi_cat == "MeanCOI=3.41"),
  columns = c(
    "expected_freq",
    "dcifer_slaf_from_mhap_freqs_paramset2",
    "idm_paramset20",
    "naive_slaf_paramset1",
    "naive_slaf_paramset2"
  ),
  lower = list(continuous = wrap(my_hex)),
  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
  upper = list(continuous = wrap(
    my_corr_ccc, digits = 3, size = 3.8
  )),
  title = "MeanCOI=3.41"
)

```


```{r}
#| results: asis
#| echo: false
#| fig-width: 13
#| fig-height: 13


cat(create_tabsetOfGgplotObjects(list(
  "all" = combined_slfreq_sel_sel_spread_all_params_pairs_plot_subset, 
  "low_coi" = combined_slfreq_sel_sel_spread_all_params_pairs_plot_subset_low_coi, 
  "high_coi" =combined_slfreq_sel_sel_spread_all_params_pairs_plot_subset_high_coi 
)))

```

```{r}
cairo_pdf("combined_slfreq_sel_sel_spread_all_params_pairs_plot_subset_by_mean_coi.pdf", width = 13, height = 13)
print(combined_slfreq_sel_sel_spread_all_params_pairs_plot_subset_low_coi)
print(combined_slfreq_sel_sel_spread_all_params_pairs_plot_subset_high_coi)
dev.off()

```


```{r}
combined_slfreq_best = combined_slfreq %>% 
  filter(tool_paramset_id  %in% combined_slfreq_ccc_by_tool_params_best$tool_paramset_id)
```


```{r}
#| fig-width: 5
#| fig-height: 6


combined_slfreq_best = combined_slfreq_best %>% 
  add_diff_mean(expected_freq, freq)

combined_slfreq_best_by_tool = make_diff_vs_expected_plot(combined_slfreq_best %>% 
                                                            left_join(sim_params),
                                                          expected_col = "expected_freq", 
                                                          diff_col = "freq_diff_from_exp", 
                                                          
                                                          ) + 
  facet_grid(coi_cat~tool)


combined_slfreq_best_ccc = combined_slfreq_best %>% 
  filter(expected_freq < 1) %>% 
  compute_ccc(expected_freq, freq, c(tool, population_name))

combined_slfreq_best_ccc_plot = ggplot(combined_slfreq_best_ccc) + 
  geom_boxplot(aes(x = tool, y = freq_ccc, color = tool)) + 
  theme_minimal() + 
  geom_abline(slope = 0, intercept = 0.95, color = "red", linetype = "dotted") + 
  ylim(min(combined_slfreq_best_ccc$freq_ccc), 1) + 
  scale_color_manual(values = colorPalette_08) + 
  theme(legend.position = "bottom")

print(combined_slfreq_best_ccc_plot)


cairo_pdf("combined_slfreq_best_ccc_plot.pdf", width = 5, height = 6)
print(combined_slfreq_best_ccc_plot)
dev.off()
```


