---
title: Benchmarking Results - Frequency Tools Multi locus microhaplotypes
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: true
fig-dpi: 300
fig-align: center
number-sections: true
anchor-sections: true
bibliography: [references.bib]
biblio-style: apalike
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)

library(ComplexHeatmap)


get_color_scale_breaks <-function(mat_dat){
  min_val = min(mat_dat, na.rm = T)
  max_val = max(mat_dat, na.rm = T)
  
  return(c(min_val, min_val + (max_val - min_val)/2, max_val))
}

```

## Set up 

### Libraries used in the analysis below 

```{r, echo=FALSE}
source("../tutorials/scripts/benchmarking_functions.R")
```

```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-line-numbers: true
#| filename: ../tutorials/scripts/benchmarking_functions.R
#| file: ../tutorials/scripts/benchmarking_functions.R

```

## Load in the Benchmarking Outputs

```{r}
inp <- list(
  sim_info_dir = "../data/benchmarking//simulations/SpotMalaria/", 
  madhatter_res_dir = "../data/benchmarking/tool_results/madhatter/"
)

inp_per_tools <- list(
  # multi-locus microhaplotype freq
  "malariaem" = list(
    tool_res_dir = "../data/benchmarking/tool_results/malariaem//", 
    tool_params = "../data/benchmarking/tool_results/malariaem.json"
  )
)



```



First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.

### Reading in expected truths 


Read in expected phased data 

For several haplotype groups, the expected phased haplotype frequencies are calculated below 

```{r}
all_phased_pop_freqs_prevs_out = readr::read_tsv("../data/benchmarking/simulations/SpotMalaria/all_phased_pop_freqs_prevs_out.tsv.gz") %>% 
  dplyr::rename(target_id = target) %>% 
  unique()

colnames(all_phased_pop_freqs_prevs_out)[4:9] = paste0("expected_", colnames(all_phased_pop_freqs_prevs_out[4:9]))

all_phased_sample_out = readr::read_tsv("../data/benchmarking/simulations/SpotMalaria/all_phased_sample_out.tsv.gz")%>% 
  dplyr::rename(target_id = target) %>% 
  unique()

```



Read in all simulation parameters into one tibble. Encode different population parameters as descriptive factors. 

```{r}

sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>% 
  mutate(
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>% 
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  )) %>% 
  mutate(pop_cat = factor(pop_cat, levels = sort(unique((.$pop_cat)))), 
         coi_cat = factor(coi_cat, levels = sort(unique((.$coi_cat)))), 
         between_sample_related_cat = factor(between_sample_related_cat, levels = sort(unique((.$between_sample_related_cat)))), 
         nonrecombinant_cat = factor(nonrecombinant_cat, levels = sort(unique((.$nonrecombinant_cat)))), 
         meiosis_cat = factor(meiosis_cat, levels = sort(unique((.$meiosis_cat))))) %>% 
  select(-sample_count)

create_dt(sim_params)
```



## multi-locus microhaplotype freq

### malariaem

```{r}
malariaem_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$malariaem$tool_res_dir, 
      pattern = "^mlaf_results.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

```{r}
malariaem_phased_per_sample_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$malariaem$tool_res_dir, 
      pattern = "^phasing_results.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

malariaem_tool_params 

```{r}
# Read in tool parameters
malariaem_tool_params <- jsonlite::fromJSON(inp_per_tools$malariaem$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_subset_targets = paramset.subset_targets
  )

create_dt(malariaem_tool_params)
```





####  two_meddiv


```{r}

malariaem_phased_per_sample_summary_two_meddiv = malariaem_phased_per_sample_summary %>% 
  filter(group_id == "two_meddiv")

malariaem_freq_summary_two_meddiv = malariaem_freq_summary %>% 
  filter(group_id == "two_meddiv")

all_phased_pop_freqs_prevs_out_two_meddiv = all_phased_pop_freqs_prevs_out %>% 
  filter(group_id == "two_meddiv")
all_phased_sample_out_two_meddiv = all_phased_sample_out %>% 
    filter(group_id == "two_meddiv")

two_meddiv_group_popHapId = malariaem_phased_per_sample_summary_two_meddiv %>% 
  select(target_id, seq) %>% 
  bind_rows(all_phased_sample_out_two_meddiv %>% 
              select(target_id, seq), 
            malariaem_freq_summary_two_meddiv %>% 
              select(target_id, seq)) %>% 
  group_by(target_id, seq) %>% 
  summarise(n = n()) %>%
  arrange(target_id, desc(n)) %>% 
  group_by(target_id) %>% 
  mutate(row = row_number()) %>% 
  mutate(seq_name = paste0(target_id, ".", stringr::str_pad(row, width = nchar(max(row)), pad = "0")))

two_meddiv_group_popHapId_sel = two_meddiv_group_popHapId %>% 
  ungroup() %>% 
  select(target_id, seq, seq_name)

all_phased_pop_freqs_prevs_out_two_meddiv_mod = all_phased_pop_freqs_prevs_out_two_meddiv %>% 
  left_join(two_meddiv_group_popHapId_sel)%>% 
  group_by(population_name, group_id, pop_simulated_gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_two_meddiv_mod = malariaem_freq_summary_two_meddiv %>% 
  left_join(two_meddiv_group_popHapId_sel) %>% 
  group_by(population_name, group_id, gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_two_meddiv_mod_sel = malariaem_freq_summary_two_meddiv_mod %>% 
  ungroup() %>% 
  select(population_name, paramset_id, hap_name, freq, freq_se, group_id) %>% 
  unique()

all_phased_pop_freqs_prevs_out_two_meddiv_mod_sel = all_phased_pop_freqs_prevs_out_two_meddiv_mod %>% 
  ungroup() %>% 
  select(population_name, hap_name, expected_allele_freq, group_id) %>% 
  unique()

malariaem_freq_summary_two_meddiv_mod_sel_combined = malariaem_freq_summary_two_meddiv_mod_sel %>% 
  full_join(all_phased_pop_freqs_prevs_out_two_meddiv_mod_sel %>% 
              dplyr::rename("expected_freq" = expected_allele_freq))%>%
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>%
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq))


```


####  two_lowdiv


```{r}

malariaem_phased_per_sample_summary_two_lowdiv = malariaem_phased_per_sample_summary %>% 
  filter(group_id == "two_lowdiv")

malariaem_freq_summary_two_lowdiv = malariaem_freq_summary %>% 
  filter(group_id == "two_lowdiv")

all_phased_pop_freqs_prevs_out_two_lowdiv = all_phased_pop_freqs_prevs_out %>% 
  filter(group_id == "two_lowdiv")
all_phased_sample_out_two_lowdiv = all_phased_sample_out %>% 
    filter(group_id == "two_lowdiv")

two_lowdiv_group_popHapId = malariaem_phased_per_sample_summary_two_lowdiv %>% 
  select(target_id, seq) %>% 
  bind_rows(all_phased_sample_out_two_lowdiv %>% 
              select(target_id, seq), 
            malariaem_freq_summary_two_lowdiv %>% 
              select(target_id, seq)) %>% 
  group_by(target_id, seq) %>% 
  summarise(n = n()) %>%
  arrange(target_id, desc(n)) %>% 
  group_by(target_id) %>% 
  mutate(row = row_number()) %>% 
  mutate(seq_name = paste0(target_id, ".", stringr::str_pad(row, width = nchar(max(row)), pad = "0")))

two_lowdiv_group_popHapId_sel = two_lowdiv_group_popHapId %>% 
  ungroup() %>% 
  select(target_id, seq, seq_name)

all_phased_pop_freqs_prevs_out_two_lowdiv_mod = all_phased_pop_freqs_prevs_out_two_lowdiv %>% 
  left_join(two_lowdiv_group_popHapId_sel)%>% 
  group_by(population_name, group_id, pop_simulated_gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_two_lowdiv_mod = malariaem_freq_summary_two_lowdiv %>% 
  left_join(two_lowdiv_group_popHapId_sel) %>% 
  group_by(population_name, group_id, gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_two_lowdiv_mod_sel = malariaem_freq_summary_two_lowdiv_mod %>% 
  ungroup() %>% 
  select(population_name, paramset_id, hap_name, freq, freq_se, group_id) %>% 
  unique()

all_phased_pop_freqs_prevs_out_two_lowdiv_mod_sel = all_phased_pop_freqs_prevs_out_two_lowdiv_mod %>% 
  ungroup() %>% 
  select(population_name, hap_name, expected_allele_freq, group_id) %>% 
  unique()

malariaem_freq_summary_two_lowdiv_mod_sel_combined = malariaem_freq_summary_two_lowdiv_mod_sel %>% 
  full_join(all_phased_pop_freqs_prevs_out_two_lowdiv_mod_sel %>% 
              dplyr::rename("expected_freq" = expected_allele_freq))%>%
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>%
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq))


```


####  four


```{r}

malariaem_phased_per_sample_summary_four = malariaem_phased_per_sample_summary %>% 
  filter(group_id == "four")

malariaem_freq_summary_four = malariaem_freq_summary %>% 
  filter(group_id == "four")

all_phased_pop_freqs_prevs_out_four = all_phased_pop_freqs_prevs_out %>% 
  filter(group_id == "four")
all_phased_sample_out_four = all_phased_sample_out %>% 
    filter(group_id == "four")

four_group_popHapId = malariaem_phased_per_sample_summary_four %>% 
  select(target_id, seq) %>% 
  bind_rows(all_phased_sample_out_four %>% 
              select(target_id, seq), 
            malariaem_freq_summary_four %>% 
              select(target_id, seq)) %>% 
  group_by(target_id, seq) %>% 
  summarise(n = n()) %>%
  arrange(target_id, desc(n)) %>% 
  group_by(target_id) %>% 
  mutate(row = row_number()) %>% 
  mutate(seq_name = paste0(target_id, ".", stringr::str_pad(row, width = nchar(max(row)), pad = "0")))

four_group_popHapId_sel = four_group_popHapId %>% 
  ungroup() %>% 
  select(target_id, seq, seq_name)

all_phased_pop_freqs_prevs_out_four_mod = all_phased_pop_freqs_prevs_out_four %>% 
  left_join(four_group_popHapId_sel)%>% 
  group_by(population_name, group_id, pop_simulated_gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_four_mod = malariaem_freq_summary_four %>% 
  left_join(four_group_popHapId_sel) %>% 
  group_by(population_name, group_id, gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_four_mod_sel = malariaem_freq_summary_four_mod %>% 
  ungroup() %>% 
  select(population_name, paramset_id, hap_name, freq, freq_se, group_id) %>% 
  unique()

all_phased_pop_freqs_prevs_out_four_mod_sel = all_phased_pop_freqs_prevs_out_four_mod %>% 
  ungroup() %>% 
  select(population_name, hap_name, expected_allele_freq, group_id) %>% 
  unique()

malariaem_freq_summary_four_mod_sel_combined = malariaem_freq_summary_four_mod_sel %>% 
  full_join(all_phased_pop_freqs_prevs_out_four_mod_sel %>% 
              dplyr::rename("expected_freq" = expected_allele_freq))%>%
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>%
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq))


```



```{r}

malariaem_freq_summary_all_combined = bind_rows(
  malariaem_freq_summary_two_meddiv_mod_sel_combined, 
  malariaem_freq_summary_two_lowdiv_mod_sel_combined, 
  malariaem_freq_summary_four_mod_sel_combined
) %>% 
  mutate(group_id = factor(group_id, levels = c("two_lowdiv", "two_meddiv", "four")))

malariaem_freq_summary_all_combined_ccc = compute_ccc(malariaem_freq_summary_all_combined, 
                                                                     expected_freq, freq, 
                                                                     c(group_id)
                                                                     )

malariaem_freq_summary_all_combined = malariaem_freq_summary_all_combined %>% 
  add_diff_mean(expected_freq, freq)



malariaem_freq_summary_all_combined_bland_alt = make_bland_altman_plot_by_label(malariaem_freq_summary_all_combined, "freq") + 
  facet_grid(~group_id)

cairo_pdf("malariaem_freq_summary_all_combined_bland_alt.pdf", width = 8, height = 4)
print(malariaem_freq_summary_all_combined_bland_alt)
dev.off()

malariaem_freq_summary_all_combined = malariaem_freq_summary_all_combined %>%
  mutate(TP = as.numeric(freq != 0 & expected_freq != 0)) %>% 
  mutate(FP = as.numeric(freq != 0 & expected_freq == 0)) %>% 
  mutate(FN = as.numeric(freq == 0 & expected_freq != 0))


malariaem_freq_summary_all_combined_contig = malariaem_freq_summary_all_combined %>% 
  group_by(group_id) %>% 
  add_contigency_metrics_no_TN(T)

malariaem_freq_summary_all_combined_contig_precission_recall = malariaem_freq_summary_all_combined_contig %>% 
  gather(measure, value, 7, 9)


malariaem_freq_summary_all_combined_contig_precission_recall_plot = ggplot(malariaem_freq_summary_all_combined_contig_precission_recall) + 
  geom_col(aes(x = group_id, y = value, fill = measure), color = "black", position = "dodge") + 
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values =colorPalette_08)

cairo_pdf("malariaem_freq_summary_all_combined_contig_precission_recall_plot.pdf", width = 4, height = 4)
print(malariaem_freq_summary_all_combined_contig_precission_recall_plot)
dev.off()


malariaem_freq_summary_all_combined_fp = malariaem_freq_summary_all_combined %>% 
  filter(FP == 1)


malariaem_freq_summary_all_combined_false_positive_freq_plot = ggplot(malariaem_freq_summary_all_combined_fp) + 
  geom_histogram(aes(x = freq)) +
  scale_y_log10() + 
  facet_grid(~group_id) + 
  custom_ggplot_theme + 
  labs(y = "count (log10)")

cairo_pdf("malariaem_freq_summary_all_combined_false_positive_freq_plot.pdf", width = 6, height = 4)
print(malariaem_freq_summary_all_combined_false_positive_freq_plot)
dev.off()
```


