---
title: Benchmarking Results - Frequency Tools Single Microhaplotype
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: true
fig-dpi: 300
fig-align: center
number-sections: true
anchor-sections: true
bibliography: [references.bib]
biblio-style: apalike
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)

library(ComplexHeatmap)

get_color_scale_breaks <-function(mat_dat){
  min_val = min(mat_dat, na.rm = T)
  max_val = max(mat_dat, na.rm = T)
  
  return(c(min_val, min_val + (max_val - min_val)/2, max_val))
}
```

## Tools 

* MOIRE[@Murphy2024-or]
* 1 step EM algorithim from dcifer[@Gerlovina2022-ya]  
* naive estimator (estimating frequency based on presence/absence or weighted by within sample frequency)

## Set up 

### Libraries used in the analysis below 

```{r, echo=FALSE}
source("../tutorials/scripts/benchmarking_functions.R")
```

```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-line-numbers: true
#| filename: ../tutorials/scripts/benchmarking_functions.R
#| file: ../tutorials/scripts/benchmarking_functions.R

```

## Load in the Benchmarking Outputs

```{r}
inp <- list(
  sim_info_dir = "../data/benchmarking//simulations/SpotMalaria/", 
  madhatter_res_dir = "../data/benchmarking/tool_results/madhatter/"
)

inp_per_tools <- list(
  # single allele frequency tools
  "moire" = list(
      tool_res_dir = "../data/benchmarking/tool_results/MOIRE/", 
      tool_params = "../data/benchmarking/tool_results/moire.json"
  ), 
  "naive_mhap_freqs" = list(
      tool_res_dir = "../data/benchmarking/tool_results/naive_mhap_freqs/", 
      tool_params = "../data/benchmarking/tool_results/naive_mhap_freqs.json"
  ),
  "dcifer_mhap_freqs" = list(
      tool_res_dir = "../data/benchmarking/tool_results/dcifer_mhap_freqs/", 
      tool_params = "../data/benchmarking/tool_results/dcifer_mhap_freqs.json"
  )
)



```



First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.

### Reading in expected truths 


Read per in expected allele frequencies tables population level 
```{r}
population_trimmed_mhap_freqs_prevs <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_trimmed_mhap_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(),
        population_name = col_character(), 
        pop = col_character(), 
        country = col_character(), 
        target = col_character(), 
        seq = col_character()
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(target_id = target)

# select only what's needed 
population_trimmed_mhap_freqs_prevs_sel = population_trimmed_mhap_freqs_prevs %>% 
  select(population_name, target_id, seq, freq) %>% 
  dplyr::rename(expected_freq = freq)

```


Read in all simulation parameters into one tibble. Encode different population parameters as descriptive factors. 

```{r}

sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>% 
  mutate(
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>% 
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  )) %>% 
  mutate(pop_cat = factor(pop_cat, levels = sort(unique((.$pop_cat)))), 
         coi_cat = factor(coi_cat, levels = sort(unique((.$coi_cat)))), 
         between_sample_related_cat = factor(between_sample_related_cat, levels = sort(unique((.$between_sample_related_cat)))), 
         nonrecombinant_cat = factor(nonrecombinant_cat, levels = sort(unique((.$nonrecombinant_cat)))), 
         meiosis_cat = factor(meiosis_cat, levels = sort(unique((.$meiosis_cat))))) %>% 
  select(-sample_count)

create_dt(sim_params)
```

Read in madhatter pipeline allele tables 

```{r}
madhatter_allele_data <- tibble(
    allele_data_path = list.files(
      inp$madhatter_res_dir, 
      pattern = "allele_data.txt.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    allele_data = map(
      allele_data_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  mutate(tool_res_dir = dirname(allele_data_path)) %>%
  mutate(population_name = gsub("_mh_results", "", basename(tool_res_dir))) %>%
  select(-allele_data_path, -tool_res_dir) %>%
  unnest(allele_data) %>% 
  mutate(SampleID = gsub("_S[0-9]+$", "", SampleID)) %>% 
  dplyr::rename(library_sample_name = SampleID, 
                target_id = Locus, 
                seq = ASV, 
                reads = Reads) 
```

## single allele microhaplotype frequencies 

### MOIRE  

Getting the allele frequencies data in the `allele_freq_summary.tsv.gz` files 

```{r}
moire_allele_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$moire$tool_res_dir, 
      pattern = "allele_freq_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 
```

Getting what tool parameters were used. For MOIRE benchmarking we set the priors on the assumed amount of relatedness with moire_paramset1 assuming uniform relatedness and then each moire_paramset 2 and 3 increasing the amount of assumed relatedness. moire_paramsets3-6 are the same prior relatedness as 1-3 but have more parallel burin to explore more pace for convergence.   

```{r}
# Read in tool parameters
moire_tool_params <- jsonlite::fromJSON(inp_per_tools$moire$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    r_alpha = paramset.r_alpha,
    r_beta = paramset.r_beta, 
    burnin = paramset.burnin,
    pt_chains = paramset.pt_chains,
    pt_num_threads = paramset.pt_num_threads,
    pt_grad_lower = paramset.pt_grad_lower,
    max_runtime = paramset.max_runtime
  )

create_dt(moire_tool_params)
```


### dcifer mhaps freqs 

Read in dcifer_mhap_freqs results

Loading results for dcifer's microhaplotype single locus allele frequency estimator, a 1 step EM algorithm which utilizes COI to help inform allele estimates.  

```{r}
dcifer_mhap_freqs_allele_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$dcifer_mhap_freqs$tool_res_dir, 
      pattern = "mhaps_freqs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

Getting what tool parameters were used. For dcifer_mhap_freqs benchmarking we used dcifer's COI tool for choosing COI based on a 1st, 2nd, and 3rd max, we also had just a general set COI of 1, 2, and 5 for all sample to examine differences in estimated versus just one COI. Also tried qstart (frequency start point for EM) to 0.01, 0.5 (default) and (0.99). 

```{r}
# Read in tool parameters
dcifer_mhap_freqs_tool_params <- jsonlite::fromJSON(inp_per_tools$dcifer_mhap_freqs$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    lrank = paramset.lrank,
    coi = paramset.coi,
    qstart = paramset.qstart
  )

create_dt(dcifer_mhap_freqs_tool_params)
```


### Naive mhaps freqs 

```{r}
naive_mhap_freqs_allele_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$naive_mhap_freqs$tool_res_dir, 
      pattern = "mhaps_freqs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

Naive microhaplotype frequency estimator has two modes, read_count_prop (allele frequencies normalized by within sample allele frequency as defined by the read counts) and presence_absence, calculated simply by how many samples each allele appears in 

```{r}
# Read in tool parameters
naive_mhap_freqs_tool_params <- jsonlite::fromJSON(inp_per_tools$naive_mhap_freqs$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    method = paramset.method
  )

create_dt(naive_mhap_freqs_tool_params)
```


### Combining results into one 

```{r}


add_expected_freqs <- function(summary_dat,
                               expected_dat,
                               tool_params,
                               tool_name) {
  ## expand expected results for all parameter sets so they all get expected results
  ## fill the unmatched expected with 0's and with unmatched observed (false haps) with 0's
  summary_dat_out = summary_dat %>%
    left_join(tool_params)  %>%
    full_join(expected_dat %>%
                crossing(paramset_id = unique(sort(
                  summary_dat$paramset_id
                )))) %>%
    mutate(freq = ifelse(is.na(freq), 0, freq)) %>%
    mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq)) %>%
    mutate(tool =  tool_name) %>%
    mutate(tool_paramset_id = paste0(tool, "-", paramset_id))
  
  return(summary_dat_out)
}

```



```{r}

# moire 
moire_allele_freq_summary_out = add_expected_freqs(moire_allele_freq_summary, 
                                                   population_trimmed_mhap_freqs_prevs_sel, 
                                                   moire_tool_params, 
                                                   "moire")

# dcifer_mhap_freqs
dcifer_mhap_freqs_allele_freq_summary_out = add_expected_freqs(dcifer_mhap_freqs_allele_freq_summary, 
                                                   population_trimmed_mhap_freqs_prevs_sel, 
                                                   dcifer_mhap_freqs_tool_params, 
                                                   "dcifer")

# naive_mhap_freqs
naive_mhap_freqs_allele_freq_summary_out = add_expected_freqs(naive_mhap_freqs_allele_freq_summary, 
                                                   population_trimmed_mhap_freqs_prevs_sel, 
                                                   naive_mhap_freqs_tool_params, 
                                                   "naive")

# combine 
combined_mhaps_freqs = bind_rows(
  moire_allele_freq_summary_out %>% 
    select(population_name, tool, paramset_id, tool_paramset_id, target_id, seq, freq, expected_freq), 
  dcifer_mhap_freqs_allele_freq_summary_out %>% 
    select(population_name, tool, paramset_id, tool_paramset_id, target_id, seq, freq, expected_freq), 
  naive_mhap_freqs_allele_freq_summary_out %>% 
    select(population_name, tool, paramset_id, tool_paramset_id, target_id, seq, freq, expected_freq), 
) %>% 
  left_join(sim_params)

```


### Summarizing expected frequencies 

```{r}
# calculating he 
population_trimmed_mhap_freqs_prevs_sel_he = population_trimmed_mhap_freqs_prevs_sel %>% 
  group_by(population_name, target_id) %>% 
  summarise(he = 1 - sum(expected_freq ^2), 
            seq_n = n_distinct(seq))

population_trimmed_mhap_freqs_prevs_sel_he_sum = population_trimmed_mhap_freqs_prevs_sel_he %>% 
  group_by(target_id) %>% 
  summarise(quick_summary(he))
# filtering to just sites that have at least some variation in any of the simulated populations 

population_trimmed_mhap_freqs_prevs_sel_he_sum_invariant = population_trimmed_mhap_freqs_prevs_sel_he_sum %>% 
  filter(q100 == 0)
population_trimmed_mhap_freqs_prevs_sel_he_sum_variant = population_trimmed_mhap_freqs_prevs_sel_he_sum %>% 
  filter(q100 != 0) %>% 
  arrange(mean) %>% 
  mutate(target_id = factor(target_id, levels = unique(.$target_id)))

population_trimmed_mhap_freqs_prevs_sel_he_variant = population_trimmed_mhap_freqs_prevs_sel_he %>% 
  filter(target_id %!in% population_trimmed_mhap_freqs_prevs_sel_he_sum_invariant$target_id)


population_trimmed_mhap_freqs_prevs_sel_variant = population_trimmed_mhap_freqs_prevs_sel %>% 
  filter(target_id %!in% population_trimmed_mhap_freqs_prevs_sel_he_sum_invariant$target_id)
```

Ranges of he for loci 

```{r}
ggplot(population_trimmed_mhap_freqs_prevs_sel_he_variant %>% 
         left_join(sim_params) %>% 
         mutate(target_id = factor(target_id, levels = population_trimmed_mhap_freqs_prevs_sel_he_sum_variant$target_id))
       ) + 
  geom_boxplot(aes(x = target_id, y = he)) + 
  custom_ggplot_theme_xRotate + 
  scale_color_manual(values = colorPalette_08) + 
  facet_grid(country~.)

```

Some variants are invariant in one ancestor population but may be highly variant in the other 

```{r}
#| fig-width: 30
#| fig-height: 10
variant_levels <- population_trimmed_mhap_freqs_prevs_sel_he_sum_variant$target_id

# Create shading data frame
n_variants <- length(variant_levels)
shade_df <- data.frame(
  target_id = factor(variant_levels, levels = variant_levels),
  xmin = seq_along(variant_levels) - 0.5,
  xmax = seq_along(variant_levels) + 0.5,
  fill = rep(c("gray", "white"), length.out = n_variants)
)

ggplot(population_trimmed_mhap_freqs_prevs_sel_he_variant %>% 
         left_join(sim_params) %>% 
         mutate(target_id = factor(target_id, levels = variant_levels))
       ) + 
  geom_rect(data = shade_df,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            inherit.aes = FALSE, alpha = 0.4) +
  scale_fill_manual(values = c("gray" = "grey50", "white" = "white"), guide = "none") +
  geom_boxplot(aes(x = target_id, y = he, color = factor(country))) + 
  custom_ggplot_theme_xRotate + 
  scale_color_manual(values = colorPalette_08)


```

```{r}
ggplot(population_trimmed_mhap_freqs_prevs_sel_variant) + 
  geom_histogram(aes(x = expected_freq), binwidth = 0.01, boundary = 0)
```


Filter out always invariant targets 

```{r}
combined_mhaps_freqs = combined_mhaps_freqs %>% 
  filter(target_id  %!in%  population_trimmed_mhap_freqs_prevs_sel_he_sum_invariant$target_id)
```


### Breakdown by target_id 


```{r}
combined_mhaps_freqs_ccc_by_tool_target_id_params = combined_mhaps_freqs %>% 
  compute_ccc(expected_freq, freq, c(tool, target_id, paramset_id, tool_paramset_id))
```

```{r}
# looking at just at frequencies that aren't 1, it's important that sites with a frequency of 1 are also correct but currently 
# no tool has any issues with getting a freq of 1, having it being over representation makes evaluation biased  

combined_mhaps_freqs_ccc_by_tool_target_id_params_filt = combined_mhaps_freqs %>% 
  filter(expected_freq < 1) %>% 
  compute_ccc(expected_freq, freq, c(tool, target_id, paramset_id, tool_paramset_id))

combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt = combined_mhaps_freqs_ccc_by_tool_target_id_params_filt %>% 
  filter(grepl("(naive_mhap|moire|dcifer_mhap_freqs_paramset0[123])", paramset_id))

combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp = combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt %>% 
  ungroup() %>% 
  select(target_id, tool_paramset_id, freq_ccc) %>% 
  spread(tool_paramset_id, freq_ccc)

combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp_mat = as.matrix(combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp[,2:ncol(combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp)])
rownames(combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp_mat) = combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp$target_id


library(circlize)
col_fun = colorRamp2(get_color_scale_breaks(combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp_mat), c( "#3288bd", "#ffffbf", "#d53e4f"))

target_id_meta = tibble(
  target_id = rownames(combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp_mat)
) %>% 
  left_join(population_trimmed_mhap_freqs_prevs_sel_he_sum %>% 
              select(target_id, mean) %>% 
              dplyr::rename(mean_he = mean)) %>% 
  select(-target_id) %>% 
  as.data.frame()
row_anno = rowAnnotation(
  df = target_id_meta, 
  col = list(
    mean_he = colorRamp2(c(min(target_id_meta$mean_he, na.rm = T),
                       median(target_id_meta$mean_he, na.rm = T),
                     max(target_id_meta$mean_he, na.rm = T)),
                   c( "#ffffcc", "#feb24c", "#fc4e2a"))
  )
)

```

```{r}
#| fig-width: 15
#| fig-height: 30
combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp_mat_hp = Heatmap(combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp_mat, 
                                                                 name = "concordance", 
                                                                 cluster_rows = T, 
                                                                 row_gap = unit(0.5, "cm"),
                                                                 col = col_fun,
                                                                 rect_gp = gpar(col = "black", lwd = 1),
                                                                 left_annotation = row_anno,
                                                                 # row_split = tool_meta$tool, 
                                                                cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(
      sprintf("%.2f", combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp_mat[i, j]),
      x,
      y
      #,gp = gpar(fontsize = 8, col = "black")
    )
  })

draw(combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp_mat_hp, 
     background = "transparent", 
     padding = unit(c(50, 5, 5, 5), "mm"), 
     merge_legends = T)

cairo_pdf("combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp_mat_hp.pdf", width = 15, height = 30, bg = "transparent")
draw(combined_mhaps_freqs_ccc_by_tool_target_id_params_filt_filt_sp_mat_hp, background = "transparent", 
     padding = unit(c(50, 5, 5, 5), "mm"), 
     merge_legends = T)
dev.off()

```



### Overall 

```{r}

combined_mhaps_freqs_ccc_by_tool_params = combined_mhaps_freqs %>% 
  filter(expected_freq != 1) %>% 
  compute_ccc(expected_freq, freq, c(tool, paramset_id, tool_paramset_id))

combined_mhaps_freqs_ccc_by_tool_params_best = combined_mhaps_freqs_ccc_by_tool_params %>% 
  group_by(tool) %>% 
  slice_max(order_by = freq_ccc, with_ties = F)

```


The different IDM parameters of model and initial values have no effect on results. There is very little variation in the different lrank (coi estimator) for dcifer based estimate, the presence/absence naive estimator does worse than the within sample weighted frequency one. 

```{r}
#| fig-height: 8
#| fig-width: 30


combined_mhaps_freqs_ccc_by_tool_params_plot_per_paramset = ggplot(combined_mhaps_freqs_ccc_by_tool_params) + 
  geom_col(aes(x = paramset_id, y = freq_ccc, fill = tool)) + 
  geom_errorbar(aes(x = paramset_id, ymin = freq_ccc_lower, ymax = freq_ccc_upper)) + 
  scale_fill_manual(values = colorPalette_15) + 
  custom_ggplot_theme_xRotate + 
  geom_abline(slope = 0, intercept = 0.95, color = "red", linetype = "dotted")

create_dt(combined_mhaps_freqs_ccc_by_tool_params)

print(combined_mhaps_freqs_ccc_by_tool_params_plot_per_paramset)

cairo_pdf("combined_mhaps_freqs_ccc_by_tool_params_plot_per_paramset.pdf", width = 30, height = 8)
print(combined_mhaps_freqs_ccc_by_tool_params_plot_per_paramset)
dev.off()
```



### Best per population COI parameters 

The mean population COI makes a large difference to which parameter set works best for which tool. This is due to the max possible COIs in each population. 

```{r}
combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat = combined_mhaps_freqs %>% 
  filter(grepl("(naive_mhap|moire|dcifer_mhap_freqs_paramset0[123])", paramset_id)) %>% 
  left_join(sim_params) %>% 
  filter(expected_freq != 1) %>% 
  compute_ccc(
    expected_freq, freq, 
    c(tool, paramset_id, coi_cat)
  )

combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp = combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat %>% 
  ungroup() %>% 
  select(paramset_id, coi_cat, freq_ccc) %>% 
  spread(coi_cat, freq_ccc)

combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp_mat = as.matrix(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp[,2:ncol(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp)])
rownames(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp_mat) = combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp$paramset_id

tool_meta = tibble(
  paramset_id = rownames(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp_mat)
)%>% 
  mutate(tool = gsub("_coi_(freq|variant)", "", gsub("_p.*", "", paramset_id)))


library(circlize)
col_fun = colorRamp2(get_color_scale_breaks(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp_mat), c( "#3288bd", "#ffffbf", "#d53e4f"))

```

```{r}
#| fig-width: 8
#| fig-height: 8
combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp_hp = Heatmap(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp_mat, 
                                                                 name = "concordance", 
                                                                 cluster_rows = F, 
                                                                 row_gap = unit(0.5, "cm"),
                                                                 col = col_fun,
                                                                 rect_gp = gpar(col = "black", lwd = 1),
                                                                 row_split = tool_meta$tool, 
                                                                   cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(
      sprintf("%.2f", combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp_mat[i, j]),
      x,
      y
      #,gp = gpar(fontsize = 8, col = "black")
    )
  })

draw(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp_hp)

cairo_pdf("freq_ccc_per_coi_cat_per_tool_params.pdf", width = 8, height = 8, bg = "transparent")
draw(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_sp_hp, background = "transparent")
dev.off()

```

### Best per population COI parameters and relatedness 

The amount of between sample relatedness also plays a significant role in performance 

```{r}
combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness = combined_mhaps_freqs %>% 
  filter(grepl("(naive_mhap|moire|dcifer_mhap_freqs_paramset0[123])", paramset_id)) %>% 
  left_join(sim_params) %>% 
  filter(expected_freq != 1)  %>% 
  mutate(coi_cat_related_ness = paste0(coi_cat, "\n", between_sample_related_cat)) %>% 
  compute_ccc(
    expected_freq, freq, 
    c(tool, paramset_id, coi_cat_related_ness)
  )

combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp = combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness %>% 
  ungroup() %>% 
  select(paramset_id, coi_cat_related_ness, freq_ccc) %>% 
  spread(coi_cat_related_ness, freq_ccc)

combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat = as.matrix(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp[,2:ncol(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp)])
rownames(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat) = combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp$paramset_id

tool_meta = tibble(
  paramset_id = rownames(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat)
) %>% 
  mutate(tool = gsub("_coi_(freq|variant)", "", gsub("_p.*", "", paramset_id)))

library(circlize)
col_fun = colorRamp2(get_color_scale_breaks(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat), c( "#3288bd", "#ffffbf", "#d53e4f"))

```

```{r}
#| fig-width: 12
#| fig-height: 10
combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_hp = Heatmap(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat, 
                                                                 name = "concordance", 
                                                                 cluster_rows = F, 
                                                                 cluster_columns = F,
                                                                 row_gap = unit(0.5, "cm"),
                                                                 col = col_fun,
                                                                 rect_gp = gpar(col = "black", lwd = 1),
                                                                 row_split = tool_meta$tool, 
                                                                   cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(
      sprintf("%.2f", combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat[i, j]),
      x,
      y
      #,gp = gpar(fontsize = 8, col = "black")
    )
  })

draw(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_hp)

cairo_pdf("freq_ccc_per_coi_cat_related_ness_per_tool_params.pdf", width = 12, height = 10, bg = "transparent")
draw(combined_mhaps_freqs_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_hp, background = "transparent")
dev.off()

```


### Best per freq bins 

Breaking down performance per tool per bins of frequency, this way performance can be broken down over specific frequencies 

```{r}
combined_mhaps_freqs_ccc_per_coi_bin_per_tool = combined_mhaps_freqs %>% 
  filter(grepl("(naive_mhap|moire|dcifer_mhap_freqs_paramset0[123])", paramset_id)) %>% 
  left_join(sim_params) %>% 
  filter(expected_freq != 1)  %>% 
  mutate(expected_freq_bin = cut(expected_freq, breaks =seq(0, 1, 0.2), include.lowest = T, right = F)) %>% 
  group_by(expected_freq_bin) %>% 
  filter(n_distinct(expected_freq) > 1) %>% 
  ungroup() %>% 
  compute_ccc(
    expected_freq, freq, 
    c(tool, paramset_id, expected_freq_bin)
  )


combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp = combined_mhaps_freqs_ccc_per_coi_bin_per_tool %>% 
  ungroup() %>% 
  select(paramset_id, expected_freq_bin, freq_ccc) %>% 
  spread(expected_freq_bin, freq_ccc)

combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp_mat = as.matrix(combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp[,2:ncol(combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp)])
rownames(combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp_mat) = combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp$paramset_id

tool_meta = tibble(
  paramset_id = rownames(combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp_mat)
)%>% 
  mutate(tool = gsub("_coi_(freq|variant)", "", gsub("_p.*", "", paramset_id)))


library(circlize)
col_fun = colorRamp2(get_color_scale_breaks(combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp_mat), c( "#3288bd", "#ffffbf", "#d53e4f"))

```

```{r}
#| fig-width: 16
#| fig-height: 10
combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp_hp = Heatmap(combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp_mat, 
                                                                 name = "concordance", 
                                                                 cluster_rows = F,
                                                      cluster_columns = F,
                                                                 row_gap = unit(0.5, "cm"),
                                                                 col = col_fun,
                                                                 rect_gp = gpar(col = "black", lwd = 1),
                                                                 row_split = tool_meta$tool, 
                                                                   cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(
      sprintf("%.2f", combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp_mat[i, j]),
      x,
      y
      #,gp = gpar(fontsize = 8, col = "black")
    )
  })

draw(combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp_hp)

cairo_pdf("combined_mhaps_freqs_ccc_per_coi_bin_per_tool.pdf", width = 16, height = 10, bg = "transparent")
draw(combined_mhaps_freqs_ccc_per_coi_bin_per_tool_sp_hp, background = "transparent")
dev.off()

```




Getting the absolute deviation from the expected values across different frequency bins 

```{r}
#| fig-width: 10
#| fig-height: 5


combined_mhaps_freqs_mae_by_tool_params_by_bins = combined_mhaps_freqs %>% 
  mutate(expected_bins = cut(expected_freq, seq(0, 1, by = 0.1)) ) %>% 
  compute_mae(expected_freq, freq, c(population_name, expected_bins, tool, paramset_id, tool_paramset_id) ) 


combined_mhaps_freqs_mae_by_tool_params_by_bins_sel= combined_mhaps_freqs_mae_by_tool_params_by_bins %>% 
  filter(paramset_id %in% c(combined_mhaps_freqs_ccc_by_tool_params_best$paramset_id, 
                            "naive_mhap_freqs_paramset1", 
                            "naive_mhap_freqs_paramset2")) %>% 
  mutate(tool_lab = case_when(
    "naive_mhap_freqs_paramset2" == paramset_id ~ "naive_mhap_freqs_wsaf_prop", 
    "naive_mhap_freqs_paramset1" == paramset_id ~ "naive_mhap_freqs_absence_presence",
    T ~ tool
  ))


combined_mhaps_freqs_mae_by_tool_params_by_bins_sel_plot = ggplot(combined_mhaps_freqs_mae_by_tool_params_by_bins_sel %>% 
         filter(!is.na(expected_bins))) + 
  geom_boxplot(aes(x = expected_bins, y = freq_mean_abs_error, color = tool_lab), position = "dodge") + 
  scale_color_manual(values = colorPalette_08) + 
  custom_ggplot_theme_xRotate

print(combined_mhaps_freqs_mae_by_tool_params_by_bins_sel_plot)

cairo_pdf("combined_mhaps_freqs_mae_by_tool_params_by_bins_sel.pdf", width = 10, height = 5)
print(combined_mhaps_freqs_mae_by_tool_params_by_bins_sel_plot)
dev.off()
```


Getting pairwise comparisons 

```{r}

combined_mhaps_freqs_sel = combined_mhaps_freqs %>% 
  filter(grepl("(naive_mhap|moire|dcifer_mhap_freqs_paramset0[123])", paramset_id))

combined_mhaps_freqs_sel_sel = combined_mhaps_freqs_sel %>% 
  ungroup() %>% 
  select(population_name, seq, target_id, expected_freq) %>% 
  unique() %>% 
  dplyr::rename(freq = expected_freq) %>% 
  mutate(paramset_id = "expected_freq") %>% 
  bind_rows(
    combined_mhaps_freqs_sel %>% 
  ungroup() %>% 
  select(population_name, seq, target_id, paramset_id, freq)
  )

combined_mhaps_freqs_sel_sel_spread = combined_mhaps_freqs_sel_sel %>% 
  spread(paramset_id, freq)


```

The different dcifer params are highly correlated with each other, showing that getting the exact COI correct isn't necessary to get to get ideal results.  

```{r}
#| fig-width: 10
#| fig-height: 10
combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_dcifer = ggpairs(
  combined_mhaps_freqs_sel_sel_spread %>% 
    filter(expected_freq < 1),
  columns = c(
    "expected_freq",
    "dcifer_mhap_freqs_paramset01",
    "dcifer_mhap_freqs_paramset02",
    "dcifer_mhap_freqs_paramset03"
  ),
  lower = list(continuous = wrap(my_hex)),
  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
  upper = list(continuous = wrap(
    my_corr_ccc, digits = 3, size = 3.8
  )),
  title = "all"
)
print(combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_dcifer)

cairo_pdf("combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_dcifer.pdf", width = 10, height = 10)
print(combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_dcifer)
dev.off()
```

```{r}
#| fig-width: 13
#| fig-height: 13
combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_subset = ggpairs(
  combined_mhaps_freqs_sel_sel_spread %>% 
    filter(expected_freq < 1),
  columns =
    unique(c("expected_freq", 
      combined_mhaps_freqs_ccc_by_tool_params_best$paramset_id, 
                            "naive_mhap_freqs_paramset1", 
                            "naive_mhap_freqs_paramset2")),
  lower = list(continuous = wrap(my_hex)),
  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
  upper = list(continuous = wrap(
    my_corr_ccc, digits = 3, size = 3.8
  )),
  title = "all"
)


cairo_pdf("combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_subset.pdf", width = 13, height = 13)
print(combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_subset)
dev.off()

```

```{r}
#| fig-width: 13
#| fig-height: 13
combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_subset_low_coi = ggpairs(
  combined_mhaps_freqs_sel_sel_spread %>% 
    left_join(sim_params) %>% 
    filter(expected_freq < 1, 
           coi_cat == "MeanCOI=1.26"),
  columns =
    unique(c("expected_freq", 
      combined_mhaps_freqs_ccc_by_tool_params_best$paramset_id, 
                            "naive_mhap_freqs_paramset1", 
                            "naive_mhap_freqs_paramset2")),
  lower = list(continuous = wrap(my_hex)),
  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
  upper = list(continuous = wrap(
    my_corr_ccc, digits = 3, size = 3.8
  )),
  title = "MeanCOI=1.26"
)

combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_subset_high_coi = ggpairs(
  combined_mhaps_freqs_sel_sel_spread %>% 
    left_join(sim_params) %>% 
    filter(expected_freq < 1, 
           coi_cat == "MeanCOI=3.41"),
  columns =
    unique(c("expected_freq", 
      combined_mhaps_freqs_ccc_by_tool_params_best$paramset_id, 
                            "naive_mhap_freqs_paramset1", 
                            "naive_mhap_freqs_paramset2")),
  lower = list(continuous = wrap(my_hex)),
  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
  upper = list(continuous = wrap(
    my_corr_ccc, digits = 3, size = 3.8
  )),
  title = "MeanCOI=3.41"
)

```


```{r}
#| results: asis
#| echo: false
#| fig-width: 13
#| fig-height: 13


cat(create_tabsetOfGgplotObjects(list(
  "all" = combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_subset, 
  "low_coi" = combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_subset_low_coi, 
  "high_coi" =combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_subset_high_coi 
)))

```

```{r}
cairo_pdf("combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_subset_by_mean_coi.pdf", width = 13, height = 13)
print(combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_subset_low_coi)
print(combined_mhaps_freqs_sel_sel_spread_all_params_pairs_plot_subset_high_coi)
dev.off()

```


```{r}
combined_mhaps_freqs_best = combined_mhaps_freqs %>% 
  filter(tool_paramset_id  %in% combined_mhaps_freqs_ccc_by_tool_params_best$tool_paramset_id)
```


```{r}
#| fig-width: 8
#| fig-height: 3


combined_mhaps_freqs_best = combined_mhaps_freqs_best %>% 
  add_diff_mean(expected_freq, freq)

combined_mhaps_freqs_best_by_tool = make_diff_vs_expected_plot(combined_mhaps_freqs_best %>% 
                                                            left_join(sim_params),
                                                          expected_col = "expected_freq", 
                                                          diff_col = "freq_diff_from_exp", 
                                                          
                                                          ) + 
  facet_grid(coi_cat~tool)

print(combined_mhaps_freqs_best_by_tool)

cairo_pdf("combined_mhaps_freqs_best_by_tool.pdf", width = 8, height = 3)
print(combined_mhaps_freqs_best_by_tool)
dev.off()


```


Though the naive method performs better, all methods perform well above the 0.95 cut off

```{r}
#| fig-width: 5
#| fig-height: 6


combined_mhaps_freqs_best_ccc = combined_mhaps_freqs_best %>% 
  filter(expected_freq < 1) %>% 
  compute_ccc(expected_freq, freq, c(tool, population_name))

combined_mhaps_freqs_best_ccc_plot = ggplot(combined_mhaps_freqs_best_ccc) + 
  geom_boxplot(aes(x = tool, y = freq_ccc, color = tool)) + 
  theme_minimal() + 
  geom_abline(slope = 0, intercept = 0.95, color = "red", linetype = "dotted") + 
  ylim(min(combined_mhaps_freqs_best_ccc$freq_ccc), 1) + 
  scale_color_manual(values = colorPalette_08) + 
  theme(legend.position = "bottom") + 
  ylim(min(0.94, combined_mhaps_freqs_best_ccc$freq_ccc), 1)

print(combined_mhaps_freqs_best_ccc_plot)


cairo_pdf("combined_mhaps_freqs_best_ccc_plot.pdf", width = 5, height = 6)
print(combined_mhaps_freqs_best_ccc_plot)
dev.off()
```





<!-- ```{r} -->
<!-- all_mhaps_freqs_summary = all_mhaps_freqs_summary%>%  -->
<!--   add_diff_mean(expected_freq, freq) -->

<!-- all_mhaps_freqs_summary_bland_altman = make_bland_altman_plot_by_label(all_mhaps_freqs_summary %>%  -->
<!--                                                                              filter(freq != 0, expected_freq != 0 ), "freq") +  -->
<!--   facet_wrap(tool_paramset_id ~ .) -->

<!-- cairo_pdf("all_mhaps_freqs_summary_bland_altman.pdf", width = 25, height = 25) -->
<!-- print(all_mhaps_freqs_summary_bland_altman) -->
<!-- dev.off() -->



<!-- all_mhaps_freqs_summary_bland_altman_filt = make_bland_altman_plot_by_label(all_mhaps_freqs_summary %>%  -->
<!--                                                                              filter(freq != 0, expected_freq != 0 ) %>%  -->
<!--                                                                               filter(expected_freq > 0.05,  -->
<!--                                                                                      expected_freq < 0.95), "freq") +  -->
<!--   facet_wrap(tool_paramset_id ~ .) -->

<!-- cairo_pdf("all_mhaps_freqs_summary_bland_altman_filt.pdf", width = 25, height = 25) -->
<!-- print(all_mhaps_freqs_summary_bland_altman_filt) -->
<!-- dev.off() -->

<!-- ``` -->

<!-- ```{r} -->
<!-- all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat_between_sample_related_cat_filt = compute_ccc(all_mhaps_freqs_summary %>%  -->
<!--                                                                 filter(expected_freq > 0.05,  -->
<!--                                                                        expected_freq < 0.95), -->
<!--                                                                          expected_freq, freq,  -->
<!--                                                                 c(tool, paramset_id, coi_cat, between_sample_related_cat)) %>%  -->
<!--   mutate(tool_paramset_id = paste0(tool, paramset_id)) -->

<!-- all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat_between_sample_related_cat_filt_best = all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat_between_sample_related_cat_filt %>%  -->
<!--   group_by(tool, coi_cat, between_sample_related_cat) %>%  -->
<!--   slice_max(n = 1, order_by = freq_ccc)  %>%  -->
<!--   arrange(coi_cat, between_sample_related_cat) -->

<!-- ``` -->



<!-- ```{r} -->
<!-- all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat_filt = compute_ccc(all_mhaps_freqs_summary %>%  -->
<!--                                                                 filter(expected_freq > 0.05,  -->
<!--                                                                        expected_freq < 0.95), -->
<!--                                                                          expected_freq, freq,  -->
<!--                                                                 c(tool, paramset_id, coi_cat)) %>%  -->
<!--   mutate(tool_paramset_id = paste0(tool, paramset_id)) -->

<!-- all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat_filt_best = all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat_filt %>%  -->
<!--   group_by(tool, coi_cat) %>%  -->
<!--   slice_max(n = 1, order_by = freq_ccc)  %>%  -->
<!--   arrange(coi_cat) -->


<!-- all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat = compute_ccc(all_mhaps_freqs_summary, expected_freq, freq,  -->
<!--                                                                 c(tool, paramset_id, coi_cat)) %>%  -->
<!--   mutate(tool_paramset_id = paste0(tool, paramset_id)) -->

<!-- all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat_best = all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat %>%  -->
<!--   group_by(tool, coi_cat) %>%  -->
<!--   slice_max(n = 1, order_by = freq_ccc)  %>%  -->
<!--   arrange(coi_cat) -->

<!-- all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat_plot = ggplot(all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat) + -->
<!--   geom_boxplot(aes(x = tool_paramset_id, y = freq_ccc, color = tool)) + -->
<!--   custom_ggplot_theme_xRotate +  -->
<!--   scale_color_manual(values = colorPalette_08) -->

<!-- cairo_pdf("all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat_plot.pdf", width = 10, height = 8) -->
<!-- print(all_mhaps_freqs_summary_ccc_by_tool_params_by_coi_cat_plot) -->
<!-- dev.off() -->


<!-- ``` -->

<!-- ```{r} -->
<!-- all_mhaps_freqs_summary_ccc_by_tool_params_by_target_id = compute_ccc(all_mhaps_freqs_summary, expected_freq, freq, c(tool, paramset_id, target_id)) %>%  -->
<!--   mutate(tool_paramset_id = paste0(tool, paramset_id)) -->

<!-- all_mhaps_freqs_summary_ccc_by_tool_params_by_target_id_best = all_mhaps_freqs_summary_ccc_by_tool_params_by_target_id %>%  -->
<!--   group_by(tool, target_id) %>%  -->
<!--   slice_max(n = 1, order_by = freq_ccc)   -->

<!-- all_mhaps_freqs_summary_ccc_by_tool_params_by_target_id_plot = ggplot(all_mhaps_freqs_summary_ccc_by_tool_params_by_target_id) + -->
<!--   geom_boxplot(aes(x = tool_paramset_id, y = freq_ccc, color = tool)) + -->
<!--   custom_ggplot_theme_xRotate +  -->
<!--   scale_fill_manual(values = colorPalette_08) -->

<!-- cairo_pdf("all_mhaps_freqs_summary_ccc_by_tool_params_by_target_id_plot.pdf", width = 10, height = 8) -->
<!-- print(all_mhaps_freqs_summary_ccc_by_tool_params_by_target_id_plot) -->
<!-- dev.off() -->

<!-- ``` -->


<!-- ```{r} -->
<!-- all_mhaps_freqs_summary_ccc_by_tool_params = compute_ccc(all_mhaps_freqs_summary, expected_freq, freq, c(tool, paramset_id)) -->

<!-- all_mhaps_freqs_summary_ccc_by_tool_params_best = all_mhaps_freqs_summary_ccc_by_tool_params %>%  -->
<!--   group_by(tool) %>%  -->
<!--   slice_max(n = 1, order_by = freq_ccc)  %>%  -->
<!--   mutate(tool_paramset_id = paste0(tool, paramset_id)) -->

<!-- all_mhaps_freqs_summary_ccc_by_tool_params_filt = compute_ccc(all_mhaps_freqs_summary %>%  -->
<!--                                                                 filter(expected_freq > 0.05,  -->
<!--                                                                        expected_freq < 0.95), -->
<!--                                                               expected_freq, freq, c(tool, paramset_id)) -->

<!-- all_mhaps_freqs_summary_ccc_by_tool_params_filt_best = all_mhaps_freqs_summary_ccc_by_tool_params_filt %>%  -->
<!--   group_by(tool) %>%  -->
<!--   slice_max(n = 1, order_by = freq_ccc)  %>%  -->
<!--   mutate(tool_paramset_id = paste0(tool, paramset_id)) -->

<!-- all_mhaps_freqs_summary_best = all_mhaps_freqs_summary %>%  -->
<!--   filter(tool_paramset_id %in% all_mhaps_freqs_summary_ccc_by_tool_params_best$tool_paramset_id)  -->
<!-- ``` -->



<!-- ```{r} -->
<!-- all_mhaps_freqs_summary_best_sp = all_mhaps_freqs_summary_best %>% -->
<!--   ungroup() %>% -->
<!--   select(population_name, target_id, seq, tool, freq)  %>% -->
<!--   bind_rows( -->
<!--     all_mhaps_freqs_summary_best %>% -->
<!--       ungroup()  %>% -->
<!--       select(population_name, target_id, seq, expected_freq) %>% -->
<!--       dplyr::rename(freq = expected_freq) %>% -->
<!--       mutate(tool = "expected_freq") %>%  -->
<!--       unique() -->
<!--   ) %>% -->
<!--   spread(tool, freq) -->


<!-- all_mhaps_freqs_summary_best_sp_pair_plot =  ggpairs(all_mhaps_freqs_summary_best_sp, -->
<!--                                                   columns = c("dcifer_mhap_freqs", "moire", "naive_mhap_freqs", "expected_freq"), -->
<!--                                                   lower = list(continuous = wrap(my_hex)), -->
<!--                                                   diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)), -->
<!--                                                   upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)), -->
<!--                                                   title = "all") -->


<!-- all_mhaps_freqs_summary_best_sp_pair_plot_filt =  ggpairs(all_mhaps_freqs_summary_best_sp %>%  -->
<!--                                                        filter(expected_freq < 0.95,  -->
<!--                                                               expected_freq > 0.05), -->
<!--                                                   columns = c("dcifer_mhap_freqs", "moire", "naive_mhap_freqs", "expected_freq"), -->
<!--                                                   lower = list(continuous = wrap(my_hex)), -->
<!--                                                   diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)), -->
<!--                                                   upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)), -->
<!--                                                   title = "all") -->


<!-- cairo_pdf("all_mhaps_freqs_summary_best_sp_pair_plot_filt.pdf", width = 10, height = 10) -->
<!-- print(all_mhaps_freqs_summary_best_sp_pair_plot_filt) -->
<!-- dev.off() -->
<!-- ``` -->









