---
title: Field Samples Results - Single Locus frequency Tools
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: true
fig-dpi: 300
fig-align: center
number-sections: true
anchor-sections: true
bibliography: [references.bib]
biblio-style: apalike
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)


```

## Set up 

### Libraries used in the analysis below 

```{r, echo=FALSE}
source("../tutorials/scripts/benchmarking_functions.R")
```

```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-line-numbers: true
#| filename: ../tutorials/scripts/benchmarking_functions.R
#| file: ../tutorials/scripts/benchmarking_functions.R

```



## Load in the Benchmarking Outputs


```{r}
inp_per_tools <- list(
  # single locus allele frequencies 
  "idm" = list(
    tool_res_dir = "../data/benchmarking/field_samples/immrse_field/full/IDM/", 
    tool_params = "../data/benchmarking/field_samples/immrse_field/full/IDM/tool_params.json"
  ), 
  "naive_slaf" = list(
    tool_res_dir = "../data/benchmarking/field_samples/immrse_field/full/naive_slaf/", 
    tool_params = "../data/benchmarking/field_samples/immrse_field/full/naive_slaf/tool_params.json"
  ), 
  "dcifer_slaf_from_mhap_freqs" = list(
    tool_res_dir = "../data/benchmarking/field_samples/immrse_field/full/dcifer_slaf_from_mhap_freqs/", 
    tool_params = "../data/benchmarking/field_samples/immrse_field/full/dcifer_slaf_from_mhap_freqs/tool_params.json"
  )
)

```



## single locus variants 

naive single locus freq


### Read in naive single locus freq results

naive single locus freq outputs information on frequency and prevalence of alleles 

Getting the frequency and prevs from the `slaf.tsv.gz` files 

```{r}
naive_slaf_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$naive_slaf$tool_res_dir, 
      pattern = "^slaf.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res)  %>% 
  mutate(tool = "naive_slaf")

```


The naive estimator has two modes that it can be ran in read_count_prop (weighting the frequencies by within sample allele frequencies) and presence_absence (calculating frequencies simply by their presence and absence in samples, if all samples were monocloanl this would be the same as prevalance)

```{r}
# Read in tool parameters
naive_slaf_tool_params <- jsonlite::fromJSON(inp_per_tools$naive_slaf$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_method = paramset.method,
  )

create_dt(naive_slaf_tool_params)
```



### Read in dcifer slaf from microhaplotype freqs 

dcifer slaf from microhaplotype freqs outputs information on frequency per haplotype  

Getting the frequency and prevs from the `slaf.tsv.gz` files 

```{r}
dcifer_slaf_from_mhap_freqs_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$dcifer_slaf_from_mhap_freqs$tool_res_dir, 
      pattern = "^slaf.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res)  %>% 
  mutate(tool = "dcifer_slaf_from_mhap_freqs")

```

Three different parameters, each set uses a different nth-max (lrank) for the COI calculation that is then used for the calculating the microhaplotype frequency with a 1 step EM algorithm. 

```{r}
# Read in tool parameters
dcifer_slaf_from_mhap_freqs_tool_params <- jsonlite::fromJSON(inp_per_tools$dcifer_slaf_from_mhap_freqs$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_coi_lrank = paramset.coi_lrank,
    paramset_qstart = paramset.qstart
  )

create_dt(dcifer_slaf_from_mhap_freqs_tool_params)
```


### Read in IDM results


IDM outputs information on frequency of alleles 

Getting the frequency and prevs from the `IDM_summary.tsv.gz` files 

```{r}
idm_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$idm$tool_res_dir, 
      pattern = "^IDM_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(tool = "idm")

```

idm 

```{r}
# Read in tool parameters
idm_tool_params <- jsonlite::fromJSON(inp_per_tools$idm$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_model = paramset.model ,
    paramset_lambda_initial = paramset.lambda_initial ,
    paramset_eps_initial = paramset.eps_initial,
  )

create_dt(idm_tool_params)
```

## combine 


## combine 

```{r}
#| fig-width: 18
#| fig-height: 18

combined_slaf_freqs = bind_rows(
  naive_slaf_freq_summary                   %>% select(population_name, paramset_id, freq, variant, tool), 
  dcifer_slaf_from_mhap_freqs_freq_summary  %>% select(population_name, paramset_id, freq, variant, tool), 
  idm_freq_summary                          %>% select(population_name, paramset_id, freq, variant, tool)
) %>% 
  mutate(variant_pos = gsub(":[A-Z*]", "", variant))

combined_slaf_freqs_sp = combined_slaf_freqs %>% 
  ungroup() %>% 
  select(-tool) %>% 
  spread(paramset_id, freq)


```


### IDM parameters 

All IDM results are exactly the same no matter the parameters 

```{r}
combined_slaf_freqs_sp_pairwise_ccc_idm = compute_pairwise_ccc(combined_slaf_freqs_sp, cols = colnames(combined_slaf_freqs_sp)[grepl("idm", colnames(combined_slaf_freqs_sp))])

create_dt(combined_slaf_freqs_sp_pairwise_ccc_idm)

```


### naive parameters 


```{r}
#| fig-width: 8
#| fig-height: 8

combined_slaf_freqs_sp_naive_pair_plot =  ggpairs(combined_slaf_freqs_sp,
                                                  columns = colnames(combined_slaf_freqs_sp)[grepl("naive", colnames(combined_slaf_freqs_sp))],
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "slaf allele frequencies")

print(combined_slaf_freqs_sp_naive_pair_plot)


cairo_pdf("field_samples_combined_slaf_freqs_sp_naive_pair_plot.pdf", width = 8, height = 8)
print(combined_slaf_freqs_sp_naive_pair_plot)
dev.off()


```



### dcifer parameters 

Very minimumal difference between parameters of auto determining the COI by nth rank for dcifer allele frequency estimate. 

```{r}
#| fig-width: 8
#| fig-height: 8

combined_slaf_freqs_sp_dcifer_pair_plot =  ggpairs(combined_slaf_freqs_sp,
                                                  columns = colnames(combined_slaf_freqs_sp)[grepl("dcifer", colnames(combined_slaf_freqs_sp))],
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "slaf allele frequencies")

print(combined_slaf_freqs_sp_dcifer_pair_plot)


cairo_pdf("field_samples_combined_slaf_freqs_sp_dcifer_pair_plot.pdf", width = 8, height = 8)
print(combined_slaf_freqs_sp_dcifer_pair_plot)
dev.off()


```


### all tools 

Seeing similar patterns to simulation results in terms of the agreement between tools 

```{r}
#| fig-width: 15
#| fig-height: 15

combined_slaf_freqs_sp_all_pair_plot =  ggpairs(combined_slaf_freqs_sp,
                                                  columns = colnames(combined_slaf_freqs_sp)[grepl("(dcifer|naive|idm_paramset20)", colnames(combined_slaf_freqs_sp))],
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "slaf allele frequencies")

print(combined_slaf_freqs_sp_all_pair_plot)


cairo_pdf("field_samples_combined_slaf_freqs_sp_all_pair_plot.pdf", width = 15, height = 15)
print(combined_slaf_freqs_sp_all_pair_plot)
dev.off()


```
