---
title: Field Samples Results - COI Tools
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: true
fig-dpi: 300
fig-align: center
number-sections: true
anchor-sections: true
bibliography: [references.bib]
biblio-style: apalike
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)


```

## Set up 

### Libraries used in the analysis below 

```{r, echo=FALSE}
source("../tutorials/scripts/benchmarking_functions.R")
```

```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-line-numbers: true
#| filename: ../tutorials/scripts/benchmarking_functions.R
#| file: ../tutorials/scripts/benchmarking_functions.R

```



## Load in the Benchmarking Outputs

```{r}
inp_per_tools <- list(
  # coi tools 
  "therealmccoil" = list(
      tool_res_dir = "../data/benchmarking/field_samples/immrse_field/full/THEREALMcCOIL/", 
      tool_params = "../data/benchmarking/field_samples/immrse_field/full/THEREALMcCOIL/tool_params.json"
  ), 
  # also does allele frequencies 
  "moire" = list(
      tool_res_dir = "../data/benchmarking/field_samples/immrse_field/full/MOIRE/", 
      tool_params = "../data/benchmarking/field_samples/immrse_field/full/MOIRE/tool_params.json"
  ), 
  "naive_coi" = list(
      tool_res_dir = "../data/benchmarking/field_samples/immrse_field/full/naive_coi/", 
      tool_params = "../data/benchmarking/field_samples/immrse_field/full/naive_coi/tool_params.json"
  ), 
  "coiaf" = list(
      tool_res_dir = "../data/benchmarking/field_samples/immrse_field/full/coiaf/", 
      tool_params = "../data/benchmarking/field_samples/immrse_field/full/coiaf/tool_params.json"
  )
)



```


## THEREALMcCOI

```{r}
therealmccoil_coi_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$therealmccoil$tool_res_dir, 
      pattern = "^coi.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name)) %>% 
  mutate(tool = "therealmccoil")

```

THEREALMcCOI can run in two different models, categorical vs proportional

```{r}
# Read in tool parameters
therealmccoil_tool_params <- jsonlite::fromJSON(inp_per_tools$therealmccoil$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_model = paramset.model,
  ) 

create_dt(therealmccoil_tool_params)
```


### Read in MOIRE results

MOIRE outputs information on complexity of infection (COI) and on population allele frequencies, will read in only the COIs results for the below analysis

Getting the COI from the `coi_summary.tsv.gz` files 

```{r}
moire_coi_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$moire$tool_res_dir, 
      pattern = "^coi_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name)) %>% 
  mutate(tool = "moire-coi")

```


Getting what tool parameters were used. For MOIRE benchmarking we set the priors on the assumed amount of relatedness with moire_paramset1 assuming uniform relatedness and then each moire_paramset 2 and 3 increasing the amount of assumed relatedness. moire_paramsets3-6 are the same prior relatedness as 1-3 but have more parallel burin to explore more pace for convergence.   

```{r}
# Read in tool parameters
moire_tool_params <- jsonlite::fromJSON(inp_per_tools$moire$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    r_alpha = paramset.r_alpha,
    r_beta = paramset.r_beta
  )

create_dt(moire_tool_params)
```


### Read in COIAF results

Getting the COI from the `coi.tsv.gz` files 

```{r}
coiaf_coi_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$coiaf$tool_res_dir, 
      pattern = "^coi.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))%>% 
  mutate(tool = "coiaf")


# when the COI for the sample is 1, coi_freq returns NA 
coiaf_coi_summary = coiaf_coi_summary %>% 
  mutate(coi_freq = ifelse(is.na(coi_freq), 1, coi_freq))

coiaf_coi_summary_gat = coiaf_coi_summary %>% 
  gather(measure, coi, 4:5) %>% 
  rowwise() %>% 
  mutate(paramset_id = gsub("coiaf_", paste0("coiaf_", measure, "_"), paramset_id))


```

coiaf can run with different assumptions of sequencing error and different maxs set on COI 

```{r}
# Read in tool parameters
coiaf_tool_params <- jsonlite::fromJSON(inp_per_tools$coiaf$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_seq_error = paramset.seq_error,
    paramset_max_coi = paramset.max_coi
  ) 

create_dt(coiaf_tool_params)
```


### Read in naive COI results

Getting naive COI outputs from the `coi.tsv.gz` files 

```{r}
naive_coi_coi_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$naive_coi$tool_res_dir, 
      pattern = "^coi.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name)) %>% 
  mutate(tool = "naive_coi")

```

naive_coi can run with different assumptions of sequencing error and different maxs set on COI 

```{r}
# Read in tool parameters
naive_coi_tool_params <- jsonlite::fromJSON(inp_per_tools$naive_coi$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_method = paramset.method,
    paramset_quantile_threshold = paramset.quantile_threshold,
    paramset_integer_threshold = paramset.integer_threshold
  ) %>% 
  mutate(paramset_threshold = ifelse(is.na(paramset_integer_threshold), paramset_quantile_threshold, paramset_integer_threshold)) %>% 
  select(-paramset_quantile_threshold, -paramset_integer_threshold)

create_dt(naive_coi_tool_params)
```



## Combine results 


```{r}

combined_COI_tools_results = bind_rows(
  therealmccoil_coi_summary %>% select(population_name, paramset_id, library_sample_name, coi, tool), 
  moire_coi_summary %>% select(population_name, paramset_id, library_sample_name, coi, tool), 
  coiaf_coi_summary_gat %>% select(population_name, paramset_id, library_sample_name, coi, tool), 
  naive_coi_coi_summary %>% select(population_name, paramset_id, library_sample_name, coi, tool)
)

combined_COI_tools_results_sp = combined_COI_tools_results %>% 
  ungroup() %>% 
  select(-tool) %>% 
  spread(paramset_id, coi)


# getting the parameter set that was found to be the "best" out of the simulations 

best_coi_tool_paramset_ids = readr::read_tsv("best_coi_tool_paramset_ids.txt", col_names = "paramset_id")
combined_COI_tools_results_best_paramset = combined_COI_tools_results %>% 
  filter(paramset_id %in% best_coi_tool_paramset_ids$paramset_id)

combined_COI_tools_results_best_paramset_sp = combined_COI_tools_results_best_paramset %>% 
  ungroup() %>% 
  select(-tool) %>% 
  spread(paramset_id, coi)

```


Plotting pairwise comparisons between the parameter sets that were fond to be "best" for the simulated data  

```{r}
#| fig-height: 13
#| fig-width: 13


combined_COI_tools_results_best_paramset_sp_all_pair_plot =  ggpairs(combined_COI_tools_results_best_paramset_sp,
                                                  columns = rev(colnames(combined_COI_tools_results_best_paramset_sp)[grepl("paramset", colnames(combined_COI_tools_results_best_paramset_sp))]),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "coi")

print(combined_COI_tools_results_best_paramset_sp_all_pair_plot)

cairo_pdf("combined_field_samples_COI_tools_results_best_paramset_sp_all_pair_plot.pdf", width = 13, height = 13)
print(combined_COI_tools_results_best_paramset_sp_all_pair_plot)
dev.off()


```

```{r}

# combined_COI_tools_results_sp_all_pair_plot =  ggpairs(combined_COI_tools_results_sp,
#                                                   columns = colnames(combined_COI_tools_results_sp)[grepl("paramset", colnames(combined_COI_tools_results_sp))],
#                                                   lower = list(continuous = wrap(my_hex)),
#                                                   diag  = list(continuous = wrap("barDiag", binwidth = 1)),
#                                                   upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
#                                                   title = "coi")
# 
# cairo_pdf("combined_field_samples_COI_tools_results_sp_all_pair_plot.pdf", width = 75, height = 75)
# print(combined_COI_tools_results_sp_all_pair_plot)
# dev.off()
# 

```


```{r}
#| fig-height: 18
#| fig-width: 18

subset_coi_columns = colnames(combined_COI_tools_results_sp)[grepl("paramset", colnames(combined_COI_tools_results_sp))]
# remove ones that clearly don't work 
subset_coi_columns = subset_coi_columns[!grepl("coiaf", subset_coi_columns)]
subset_coi_columns = subset_coi_columns[!grepl("therealmccoil_paramset2", subset_coi_columns)]

subset_coi_columns = rev(subset_coi_columns)

combined_COI_tools_results_sp_all_pair_subset_plot =  ggpairs(combined_COI_tools_results_sp,
                                                  columns = subset_coi_columns,
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "coi")

print(combined_COI_tools_results_sp_all_pair_subset_plot)

cairo_pdf("combined_field_samples_COI_tools_results_sp_all_pair_subset_plot.pdf", width = 17, height = 17)
print(combined_COI_tools_results_sp_all_pair_subset_plot)
dev.off()

```

```{r}
#| fig-height: 7
#| fig-width: 7


combined_COI_tools_results_sp_all_pair_moire_subset_plot =  ggpairs(combined_COI_tools_results_sp,
                                                  columns = subset_coi_columns[grepl("moire", subset_coi_columns)],
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "coi")

print(combined_COI_tools_results_sp_all_pair_moire_subset_plot)

cairo_pdf("field_samples_combined_COI_tools_results_sp_all_pair_moire_subset_plot.pdf", width = 7, height = 7)
print(combined_COI_tools_results_sp_all_pair_moire_subset_plot)
dev.off()

```



