---
title: Field Samples Results - Microhaplotype frequency Tools
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: true
fig-dpi: 300
fig-align: center
number-sections: true
anchor-sections: true
bibliography: [references.bib]
biblio-style: apalike
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)


```

## Set up 

### Libraries used in the analysis below 

```{r, echo=FALSE}
source("../tutorials/scripts/benchmarking_functions.R")
```

```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-line-numbers: true
#| filename: ../tutorials/scripts/benchmarking_functions.R
#| file: ../tutorials/scripts/benchmarking_functions.R

```



## Load in the Benchmarking Outputs

```{r}
inp_per_tools <- list(
  "naive_mhap_freqs" = list(
      tool_res_dir = "../data/benchmarking/field_samples/immrse_field/full/naive_mhap_freqs/", 
      tool_params = "../data/benchmarking/field_samples/immrse_field/full/naive_mhap_freqs/tool_params.json"
  ), 
  "moire" = list(
      tool_res_dir = "../data/benchmarking/field_samples/immrse_field/full/MOIRE/", 
      tool_params = "../data/benchmarking/field_samples/immrse_field/full/MOIRE/tool_params.json"
  ), 
  "dcifer_mhap_freqs" = list(
      tool_res_dir = "../data/benchmarking/field_samples/immrse_field/full/dcifer_mhap_freqs//", 
      tool_params = "../data/benchmarking/field_samples/immrse_field/full/dcifer_mhap_freqs/tool_params.json"
  )
)

```


### Naive mhaps freqs 

```{r}
naive_mhap_freqs_allele_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$naive_mhap_freqs$tool_res_dir, 
      pattern = "mhaps_freqs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(tool = "naive_mhap_freqs")

```

Naive microhaplotype frequency estimator has two modes, read_count_prop (allele frequencies normalized by within sample allele frequency as defined by the read counts) and presence_absence, calculated simply by how many samples each allele appears in 

```{r}
# Read in tool parameters
naive_mhap_freqs_tool_params <- jsonlite::fromJSON(inp_per_tools$naive_mhap_freqs$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    method = paramset.method
  )

create_dt(naive_mhap_freqs_tool_params)
```


### dcifer mhaps freqs 

```{r}
dcifer_mhap_freqs_allele_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$dcifer_mhap_freqs$tool_res_dir, 
      pattern = "mhaps_freqs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(tool = "dcifer_mhap_freqs")

```


dcifer has a single step allele frequency estimator that is a single stemp EM alogorithm while accounting for COI. coi_lrank is the nth rank to estimate COI, was run with setting 1sth, 2nd and 5th ranks. 

```{r}
# Read in tool parameters
dcifer_mhap_freqs_tool_params <- jsonlite::fromJSON(inp_per_tools$dcifer_mhap_freqs$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    coi_lrank = paramset.coi_lrank, 
    qstart = paramset.qstart
  )

create_dt(dcifer_mhap_freqs_tool_params)
```

### MOIRE  

Getting the allele frequencies data in the `allele_freq_summary.tsv.gz` files 

```{r}
moire_allele_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$moire$tool_res_dir, 
      pattern = "allele_freq_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(tool = "moire")
```

Getting what tool parameters were used. For MOIRE benchmarking we set the priors on the assumed amount of relatedness with moire_paramset1 assuming uniform relatedness and then each moire_paramset 2 and 3 increasing the amount of assumed relatedness.  

```{r}
# Read in tool parameters
moire_tool_params <- jsonlite::fromJSON(inp_per_tools$moire$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    r_alpha = paramset.r_alpha,
    r_beta = paramset.r_beta
    #, 
    # burnin = paramset.burnin,
    # pt_chains = paramset.pt_chains,
    # pt_num_threads = paramset.pt_num_threads,
    # pt_grad_lower = paramset.pt_grad_lower,
    # max_runtime = paramset.max_runtime
  )

create_dt(moire_tool_params)
```


## combine 

```{r}
#| fig-width: 18
#| fig-height: 18


combined_allele_mhaps_freqs = bind_rows(
  moire_allele_freq_summary             %>% select(population_name, paramset_id, freq, target_id, seq, tool), 
  naive_mhap_freqs_allele_freq_summary  %>% select(population_name, paramset_id, freq, target_id, seq, tool), 
  dcifer_mhap_freqs_allele_freq_summary %>% select(population_name, paramset_id, freq, target_id, seq, tool)
)


combined_allele_mhaps_freqs_sp = combined_allele_mhaps_freqs %>% 
  ungroup() %>% 
  select(-tool) %>% 
  spread(paramset_id, freq)


combined_allele_mhaps_freqs_sp_all_pair_plot =  ggpairs(combined_allele_mhaps_freqs_sp,
                                                  columns = colnames(combined_allele_mhaps_freqs_sp)[grepl("paramset", colnames(combined_allele_mhaps_freqs_sp))],
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 0.01, boundary = 0)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "microhaplotype allele frequencies")

print(combined_allele_mhaps_freqs_sp_all_pair_plot)


cairo_pdf("field_samples_combined_allele_mhaps_freqs_sp_all_pair_plot.pdf", width = 18, height = 18)
print(combined_allele_mhaps_freqs_sp_all_pair_plot)
dev.off()
```


