---
title: Frequency tools Benchmarking Results
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)


```


## Set up 

### Libraries used in the analysis below 

```{r, echo=FALSE}
source("../tutorials/scripts/benchmarking_functions.R")

# define once
quick_summary <- function(x, probs = c(0, 0.25, 0.5, 0.75, 1)) {
  q <- quantile(x, probs = probs, na.rm = TRUE)
  tibble(
    n = length(x[!is.na(x)]), 
    sd = sd(x, na.rm = TRUE),
    mean = mean(x, na.rm = TRUE),
    min  = min(x, na.rm = TRUE),
    max  = max(x, na.rm = TRUE),
    !!!setNames(as.list(q), paste0("q", probs*100))
  )
}
```

```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-line-numbers: true
#| filename: ../tutorials/scripts/benchmarking_functions.R
#| file: ../tutorials/scripts/benchmarking_functions.R

```

## Load in the Benchmarking Outputs

```{r}
inp <- list(
  sim_info_dir = "../data/benchmarking//simulations/SpotMalaria/", 
  madhatter_res_dir = "../data/benchmarking/tool_results/madhatter/"
)

inp_per_tools <- list(
  "therealmccoil" = list(
      tool_res_dir = "../data/benchmarking/tool_results/THEREALMcCOIL/", 
      tool_params = "../data/benchmarking/tool_results/therealmccoil.json"
  ), 
  "moire" = list(
      tool_res_dir = "../data/benchmarking/tool_results/MOIRE/", 
      tool_params = "../data/benchmarking/tool_results/moire.json"
  ), 
  "naive_coi" = list(
      tool_res_dir = "../data/benchmarking/tool_results/naive_coi/", 
      tool_params = "../data/benchmarking/tool_results/naive_coi.json"
  ),
  "coiaf" = list(
      tool_res_dir = "../data/benchmarking/tool_results/coiaf//", 
      tool_params = "../data/benchmarking/tool_results/coiaf.json"
  ), 
  # multi-locus tools
  "fem" = list(
    tool_res_dir = "../data/benchmarking/tool_results/FEM/", 
    tool_params = "../data/benchmarking/tool_results/fem.json"
  ), 
  "mlbm" = list(
    tool_res_dir = "../data/benchmarking/tool_results/MLBM//", 
    tool_params = "../data/benchmarking/tool_results/mlbm.json"
  ), 
  "naive_mlafp" = list(
    tool_res_dir = "../data/benchmarking/tool_results/naive_mlafp/", 
    tool_params = "../data/benchmarking/tool_results/naive_mlafp.json"
  ), 
  # single locus allele frequencies 
  "idm" = list(
    tool_res_dir = "../data/benchmarking/tool_results/IDM/", 
    tool_params = "../data/benchmarking/tool_results/idm.json"
  ), 
  "naive_slaf" = list(
    tool_res_dir = "../data/benchmarking/tool_results/naive_slaf/", 
    tool_params = "../data/benchmarking/tool_results/naive_slaf.json"
  ), 
  # multi-locus microhaplotype freq

  "malariaem" = list(
    tool_res_dir = "../data/benchmarking/tool_results/malariaem//", 
    tool_params = "../data/benchmarking/tool_results/malariaem.json"
  )
)



```



First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.

### Reading in expected truths 

Read per in expected microhaplotype frequencies per sample tables to get true COIs

```{r}
expected_simulated_frequencies_per_sample <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "trimmed_expected_simulated_frequencies_per_sample.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(library_sample_name = libary_sample_name)

expected_simulated_frequencies_per_sample_COIs = expected_simulated_frequencies_per_sample %>% 
  select(panel, population_name, country, library_sample_name, COI) %>% 
  distinct() %>% 
  dplyr::rename(expected_COI = COI) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```


Read in expected phased data 

For several haplotype groups, the expected phased haplotype frequencies are calculated below 

```{r}
all_phased_pop_freqs_prevs_out = readr::read_tsv("../data/benchmarking/simulations/SpotMalaria/all_phased_pop_freqs_prevs_out.tsv.gz") %>% 
  dplyr::rename(target_id = target) %>% 
  unique()

colnames(all_phased_pop_freqs_prevs_out)[4:9] = paste0("expected_", colnames(all_phased_pop_freqs_prevs_out[4:9]))

all_phased_sample_out = readr::read_tsv("../data/benchmarking/simulations/SpotMalaria/all_phased_sample_out.tsv.gz")%>% 
  dplyr::rename(target_id = target) %>% 
  unique()

```


Read per in expected aa multilocus freq/prev tables 


```{r}
expected_population_aa_multiloci_variants_freqs_prevs <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_aa_multiloci_variants_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(),
        population_name = col_character(), 
        pop = col_character(), 
        country = col_character(), 
        variant = col_character(), 
        aas_in_variant = col_integer(),
				allele_count = col_integer(),
				sample_count = col_integer(),
				total_samples = col_integer(),
				total_alleles = col_integer(),
				freq = col_double(),
				prev  = col_double(), 
				group_id = col_character()
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth)

colnames(expected_population_aa_multiloci_variants_freqs_prevs)[5:11] = paste0("expected_",
                                                                               colnames(expected_population_aa_multiloci_variants_freqs_prevs)[5:11])


```



Read per locus aa variant freq/prev tables 
```{r}
expected_population_aa_variants_freqs_prevs <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_aa_variants_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(),
        population_name = col_character(), 
        pop = col_character(), 
        country = col_character(), 
        GeneName = col_character(), 
        transcript_id = col_character(), 
        aa_position = col_integer(), 
        aa = col_character(), 
        is_ref = col_logical(), 
				allele_count = col_integer(),
				sample_count = col_integer(),
				total_pop_alleles = col_integer(),
				total_pop_samples = col_integer(),
				unique_alleles = col_integer(),
				freq = col_double(),
				prev  = col_double()
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  mutate(variant = paste0(transcript_id, ":", aa_position, ":", aa)) 

colnames(expected_population_aa_variants_freqs_prevs)[7:15] = paste0("expected_",
                                                                               colnames(expected_population_aa_variants_freqs_prevs)[7:15])


expected_population_aa_variants_freqs_prevs_sel = expected_population_aa_variants_freqs_prevs %>% 
  select(population_name, variant, expected_freq)
```






Read in all simulation parameters into one tibble. Encode different population parameters as descriptive factors. 

```{r}

sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>% 
  mutate(
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>% 
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  )) %>% 
  mutate(pop_cat = factor(pop_cat, levels = sort(unique((.$pop_cat)))), 
         coi_cat = factor(coi_cat, levels = sort(unique((.$coi_cat)))), 
         between_sample_related_cat = factor(between_sample_related_cat, levels = sort(unique((.$between_sample_related_cat)))), 
         nonrecombinant_cat = factor(nonrecombinant_cat, levels = sort(unique((.$nonrecombinant_cat)))), 
         meiosis_cat = factor(meiosis_cat, levels = sort(unique((.$meiosis_cat))))) %>% 
  select(-sample_count)

create_dt(sim_params)
```

Read in madhatter pipeline allele tables 

```{r}
madhatter_allele_data <- tibble(
    allele_data_path = list.files(
      inp$madhatter_res_dir, 
      pattern = "allele_data.txt.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    allele_data = map(
      allele_data_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  mutate(tool_res_dir = dirname(allele_data_path)) %>%
  mutate(population_name = gsub("_mh_results", "", basename(tool_res_dir))) %>%
  select(-allele_data_path, -tool_res_dir) %>%
  unnest(allele_data) %>% 
  mutate(SampleID = gsub("_S[0-9]+$", "", SampleID)) %>% 
  dplyr::rename(library_sample_name = SampleID, 
                target_id = Locus, 
                seq = ASV, 
                reads = Reads) 
```



## multi-locus variants 

*  FreqEstimationModel
*  MLBM
*  naive multi locus freq estimator



### Read in FreqEstimationModel results

FreqEstimationModel outputs information on frequency and prevalence of alleles 

Getting the frequency and prevs from the `FEM_summary.tsv.gz` files 

```{r}
fem_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$fem$tool_res_dir, 
      pattern = "^FEM_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

fem can take a COI argument 

```{r}
# Read in tool parameters
fem_tool_params <- jsonlite::fromJSON(inp_per_tools$fem$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_coi = paramset.coi,
  )

create_dt(fem_tool_params)
```



### Read in MLBM results

MLBM outputs information on frequency but not prevalence of alleles 

Getting the frequency from the `MLBM_summary.tsv.gz` files 

```{r}
mlbm_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$mlbm$tool_res_dir, 
      pattern = "^MLBM_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

mlbm has no additional parameters  

```{r}
# Read in tool parameters
mlbm_tool_params <- jsonlite::fromJSON(inp_per_tools$mlbm$tool_params, flatten = TRUE) %>%
  as_tibble()

create_dt(mlbm_tool_params)
```



### Read in naive multi locus freq estimator results

naive multi locus freq estimator outputs information on frequency but not prevalence of alleles 

Getting the freqs and prevs from the `mlafp.tsv.gz` files 

```{r}
naive_mlafp_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$naive_mlafp$tool_res_dir, 
      pattern = "^mlafp.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

naive_mlafp has no additional parameters  

```{r}
# Read in tool parameters
naive_mlafp_tool_params <- jsonlite::fromJSON(inp_per_tools$naive_mlafp$tool_params, flatten = TRUE) %>%
  as_tibble()

create_dt(naive_mlafp_tool_params)
```


### combining freqs 

```{r}
fem_freq_summary_out = fem_freq_summary %>% 
  left_join(fem_tool_params)  %>% 
  full_join(bind_rows(
    expected_population_aa_multiloci_variants_freqs_prevs %>% mutate(paramset_id = "fem_paramset1"), 
    expected_population_aa_multiloci_variants_freqs_prevs %>% mutate(paramset_id = "fem_paramset2"), 
    expected_population_aa_multiloci_variants_freqs_prevs %>% mutate(paramset_id = "fem_paramset3")
  )) %>% 
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>% 
  mutate(prev = ifelse(is.na(prev), 0, prev)) %>% 
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq)) %>% 
  mutate(expected_prev = ifelse(is.na(expected_prev), 0, expected_prev)) %>% 
  mutate(tool =  "fem") 

mlbm_freq_summary_out = mlbm_freq_summary %>% 
  left_join(mlbm_tool_params)  %>% 
  full_join(expected_population_aa_multiloci_variants_freqs_prevs %>% mutate(paramset_id = "mlbm_paramset1")) %>% 
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>% 
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq))%>% 
  mutate(tool =  "mlbm") 


naive_mlafp_freq_summary_variant_info = naive_mlafp_freq_summary %>% 
  select(variant) %>% 
  unique() %>% 
  rowwise() %>% 
  mutate(observed_aas_in_variant = variantstring::variant_to_long(variant)[[1]] %>% select(gene, pos) %>% unique() %>% nrow())

# filter to just the full called haplotype, the wrapper for the naive caller will generate all possible haplotypes and sub haplotypes it can (e.g. if 6 loci are being called, it will call for for 5,4,3,2 loci ) 
naive_mlafp_freq_summary_out = naive_mlafp_freq_summary %>% 
  select(-sample_total) %>% 
  left_join(naive_mlafp_tool_params %>% 
              select(-loci_groups)) %>% 
  left_join(expected_population_aa_multiloci_variants_freqs_prevs %>% 
              ungroup() %>% 
              select(group_id, expected_aas_in_variant) %>% 
              unique()) %>% 
  left_join(naive_mlafp_freq_summary_variant_info)  %>% 
  full_join(expected_population_aa_multiloci_variants_freqs_prevs %>% 
              select(-expected_aas_in_variant)
            %>% mutate(paramset_id = "naive_mlafp_paramset1"), by = c("population_name", "paramset_id", "group_id",  "variant")) 
  
naive_mlafp_freq_summary_out_filt = naive_mlafp_freq_summary_out%>% 
  filter(observed_aas_in_variant == expected_aas_in_variant)
  
naive_mlafp_freq_summary_out_filt = naive_mlafp_freq_summary_out_filt %>% 
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>% 
  mutate(prev = ifelse(is.na(prev), 0, prev)) %>% 
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq)) %>% 
  mutate(expected_prev = ifelse(is.na(expected_prev), 0, expected_prev)) %>% 
  mutate(tool =  "naive_mlafp")



combined_mlfreq = bind_rows(
  fem_freq_summary_out %>% 
    select(-sequence, -median_freq, 
           -CI_2.5, -CI_97.5), 
  mlbm_freq_summary_out, 
  naive_mlafp_freq_summary_out_filt
) %>% 
  mutate(tool_paramset_id = paste0(tool, paramset_id))



```

```{r}
combined_mlfreq_ccc_by_tool_params = compute_ccc(combined_mlfreq, expected_freq, freq, c(tool, paramset_id))

combined_mlfreq_ccc_by_tool_params_best = combined_mlfreq_ccc_by_tool_params %>% 
  group_by(tool) %>% 
  slice_max(n = 1, order_by = freq_ccc)  %>% 
  mutate(tool_paramset_id = paste0(tool, paramset_id))


combined_mlfreq_best = combined_mlfreq %>% 
  filter(tool_paramset_id %in% combined_mlfreq_ccc_by_tool_params_best$tool_paramset_id) %>% 
  mutate(group_id = factor(group_id, levels = c("two_group1", "pfdhfr", "pfdhps", "four_group1", "six_low_div", "pfdhfr_pfdhps", "six_med_div")))
```

```{r}

combined_mlfreq_best = combined_mlfreq_best %>% 
  add_diff_mean(expected_freq, freq)

combined_mlfreq_best_altman_tool_groupid = make_bland_altman_plot_by_label(combined_mlfreq_best %>% 
                                                                             filter(freq != 0, expected_freq != 0 ), "freq") + 
  facet_grid(tool ~ group_id)


cairo_pdf("combined_mlfreq_best_altman_tool_groupid.pdf", width = 12, height = 6)
print(combined_mlfreq_best_altman_tool_groupid)
dev.off()
```


```{r}
combined_mlfreq_best_CCC = combined_mlfreq_best %>% 
  compute_ccc(expected_freq, freq, c(tool, group_id, population_name)) %>% 
  left_join(sim_params)


cairo_pdf("combined_mlfreq_best_CCC_groupid_tool_ccc_boxplot.pdf", width = 10, height = 5)
ggplot(combined_mlfreq_best_CCC) + 
  geom_boxplot(aes(x = group_id, y =  freq_ccc, fill = tool)) + 
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted") + 
  facet_grid(coi_cat~.)
dev.off()


```

```{r}
combined_mlfreq_best_mod = combined_mlfreq_best %>% 
  mutate(TP = as.numeric(freq != 0 & expected_freq != 0))%>% 
  mutate(FP = as.numeric(freq != 0 & expected_freq == 0))%>% 
  mutate(FN = as.numeric(freq == 0 & expected_freq != 0))




combined_mlfreq_best_mod_contig = combined_mlfreq_best_mod %>% 
  group_by(tool, group_id) %>% 
  add_contigency_metrics_no_TN(T)


combined_mlfreq_best_mod_contig_recall_precision = combined_mlfreq_best_mod_contig %>% 
  select(tool, group_id, recall, precision) %>% 
  gather(measure, value, 3:4)


cairo_pdf("combined_mlfreq_best_mod_contig_recall_precision.pdf", width = 8, height = 6)
print(ggplot(combined_mlfreq_best_mod_contig_recall_precision) + 
  geom_col(aes(x = group_id, y = value, fill = tool), color = "black", position = "dodge") + 
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) +
  facet_wrap(~measure))
dev.off()


combined_mlfreq_best_mod_contig_by_coi_cat = combined_mlfreq_best_mod %>% 
  left_join(sim_params, by = "population_name") %>% 
  group_by(tool, group_id, coi_cat) %>% 
  add_contigency_metrics_no_TN(T)


combined_mlfreq_best_mod_contig_by_coi_cat_recall_precision = combined_mlfreq_best_mod_contig_by_coi_cat %>% 
  select(tool, group_id, coi_cat, recall, precision) %>% 
  gather(measure, value,4:5)


cairo_pdf("combined_mlfreq_best_mod_contig_coi_cat_recall_precision.pdf", width = 8, height = 6)
print(ggplot(combined_mlfreq_best_mod_contig_by_coi_cat_recall_precision) + 
  geom_col(aes(x = group_id, y = value, fill = tool), color = "black", position = "dodge") + 
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) +
  facet_wrap(coi_cat~measure))
dev.off()


```


```{r}
combined_mlfreq_best_mod_fp = combined_mlfreq_best_mod %>% 
  filter(FP == 1)


ggplot(combined_mlfreq_best_mod_fp) + 
  geom_histogram(aes(x = freq)) + 
  facet_grid(group_id ~ tool, scales = "free")
```



```{r}

fem_freq_summary_out_mod = fem_freq_summary_out %>% 
  add_diff_mean(expected_freq, freq)

fem_freq_summary_out_mod_bland_altman_paramset_id_groupid = make_bland_altman_plot_by_label(fem_freq_summary_out_mod, "freq") + 
  facet_grid( paramset_id ~ group_id)

fem_freq_summary_out_ccc = fem_freq_summary_out %>% 
  compute_ccc(expected_freq, freq, c(paramset_id, group_id, population_name)) %>% 
  left_join(sim_params)


```



## single locus variants 

naive single locus freq


### Read in naive single locus freq results

naive single locus freq outputs information on frequency and prevalence of alleles 

Getting the frequency and prevs from the `slaf.tsv.gz` files 

```{r}
naive_slaf_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$naive_slaf$tool_res_dir, 
      pattern = "^slaf.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```


```{r}
# Read in tool parameters
naive_slaf_tool_params <- jsonlite::fromJSON(inp_per_tools$naive_slaf$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_method = paramset.method,
  )

create_dt(naive_slaf_tool_params)
```


### Read in IDM results


IDM outputs information on frequency of alleles 

Getting the frequency and prevs from the `IDM_summary.tsv.gz` files 

```{r}
idm_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$idm$tool_res_dir, 
      pattern = "^IDM_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

idm 

```{r}
# Read in tool parameters
idm_tool_params <- jsonlite::fromJSON(inp_per_tools$idm$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_model = paramset.model ,
    paramset_lambda_initial = paramset.lambda_initial ,
    paramset_eps_initial = paramset.eps_initial,
  )

create_dt(idm_tool_params)
```


### Combining 

```{r}

naive_slaf_freq_summary_out = naive_slaf_freq_summary %>% 
  left_join(naive_slaf_tool_params)  %>% 
  full_join(bind_rows(
    expected_population_aa_variants_freqs_prevs_sel %>% mutate(paramset_id = "naive_slaf_paramset1"), 
    expected_population_aa_variants_freqs_prevs_sel %>% mutate(paramset_id = "naive_slaf_paramset2")
  )) %>% 
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>% 
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq)) %>% 
  mutate(tool =  "naive_slaf") %>% 
  mutate(tool_paramset_id = paste0(tool, "-", paramset_id))




idm_freq_summary_out = idm_freq_summary %>%
  left_join(idm_tool_params)  %>%
  full_join(
    map_dfr(
      unique(idm_freq_summary$paramset_id),
      ~
        expected_population_aa_variants_freqs_prevs_sel %>%
        mutate(paramset_id = .x)
    )
  ) %>%
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>%
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq)) %>%
  mutate(tool =  "idm") %>%
  mutate(tool_paramset_id = paste0(tool, "-", paramset_id))


```


```{r}
combined_slfreq = bind_rows(
  naive_slaf_freq_summary_out %>% 
    select(population_name, paramset_id, variant, freq, expected_freq, tool, tool_paramset_id), 
  idm_freq_summary_out %>% 
    select(population_name, paramset_id, variant, freq, expected_freq, tool, tool_paramset_id)
)

combined_slfreq_ccc_by_tool_params = combined_slfreq %>% 
  compute_ccc(expected_freq, freq, c(tool, paramset_id, tool_paramset_id))

combined_slfreq_ccc_by_tool_params_best = combined_slfreq_ccc_by_tool_params %>% 
  group_by(tool) %>% 
  slice_max(order_by = freq_ccc, with_ties = F)

combined_slfreq_summary  = combined_slfreq %>% 
  group_by(tool, paramset_id) %>% 
  summarise(quick_summary(freq))

```

```{r}
combined_slfreq_best = combined_slfreq %>% 
  filter(tool_paramset_id  %in% combined_slfreq_ccc_by_tool_params_best$tool_paramset_id)
```


```{r}
combined_slfreq_best = combined_slfreq_best %>% 
  add_diff_mean(expected_freq, freq)
combined_slfreq_best_by_tool = make_bland_altman_plot_by_label(combined_slfreq_best, "freq") + 
  facet_wrap(~tool)


combined_slfreq_best_ccc = combined_slfreq_best %>% 
  compute_ccc(expected_freq, freq, c(tool, population_name))

ggplot(combined_slfreq_best_ccc) + 
  geom_boxplot(aes(x = tool, y = freq_ccc))
```


```{r}
combined_slfreq_best_mod =  combined_slfreq_best %>%
  mutate(TP = as.numeric(freq != 0 & expected_freq != 0))%>% 
  mutate(FP = as.numeric(freq != 0 & expected_freq == 0))%>% 
  mutate(FN = as.numeric(freq == 0 & expected_freq != 0))




combined_slfreq_best_mod_contig = combined_slfreq_best_mod %>% 
  group_by(tool) %>% 
  add_contigency_metrics_no_TN(T)

```



## multi-locus microhaplotype freq

### malariaem

```{r}
malariaem_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$malariaem$tool_res_dir, 
      pattern = "^mlaf_results.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

```{r}
malariaem_phased_per_sample_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$malariaem$tool_res_dir, 
      pattern = "^phasing_results.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

malariaem_tool_params 

```{r}
# Read in tool parameters
malariaem_tool_params <- jsonlite::fromJSON(inp_per_tools$malariaem$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_subset_targets = paramset.subset_targets
  )

create_dt(malariaem_tool_params)
```





####  two_meddiv


```{r}

malariaem_phased_per_sample_summary_two_meddiv = malariaem_phased_per_sample_summary %>% 
  filter(group_id == "two_meddiv")

malariaem_freq_summary_two_meddiv = malariaem_freq_summary %>% 
  filter(group_id == "two_meddiv")

all_phased_pop_freqs_prevs_out_two_meddiv = all_phased_pop_freqs_prevs_out %>% 
  filter(group_id == "two_meddiv")
all_phased_sample_out_two_meddiv = all_phased_sample_out %>% 
    filter(group_id == "two_meddiv")

two_meddiv_group_popHapId = malariaem_phased_per_sample_summary_two_meddiv %>% 
  select(target_id, seq) %>% 
  bind_rows(all_phased_sample_out_two_meddiv %>% 
              select(target_id, seq), 
            malariaem_freq_summary_two_meddiv %>% 
              select(target_id, seq)) %>% 
  group_by(target_id, seq) %>% 
  summarise(n = n()) %>%
  arrange(target_id, desc(n)) %>% 
  group_by(target_id) %>% 
  mutate(row = row_number()) %>% 
  mutate(seq_name = paste0(target_id, ".", stringr::str_pad(row, width = nchar(max(row)), pad = "0")))

two_meddiv_group_popHapId_sel = two_meddiv_group_popHapId %>% 
  ungroup() %>% 
  select(target_id, seq, seq_name)

all_phased_pop_freqs_prevs_out_two_meddiv_mod = all_phased_pop_freqs_prevs_out_two_meddiv %>% 
  left_join(two_meddiv_group_popHapId_sel)%>% 
  group_by(population_name, group_id, pop_simulated_gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_two_meddiv_mod = malariaem_freq_summary_two_meddiv %>% 
  left_join(two_meddiv_group_popHapId_sel) %>% 
  group_by(population_name, group_id, gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_two_meddiv_mod_sel = malariaem_freq_summary_two_meddiv_mod %>% 
  ungroup() %>% 
  select(population_name, paramset_id, hap_name, freq, freq_se, group_id) %>% 
  unique()

all_phased_pop_freqs_prevs_out_two_meddiv_mod_sel = all_phased_pop_freqs_prevs_out_two_meddiv_mod %>% 
  ungroup() %>% 
  select(population_name, hap_name, expected_allele_freq, group_id) %>% 
  unique()

malariaem_freq_summary_two_meddiv_mod_sel_combined = malariaem_freq_summary_two_meddiv_mod_sel %>% 
  full_join(all_phased_pop_freqs_prevs_out_two_meddiv_mod_sel %>% 
              dplyr::rename("expected_freq" = expected_allele_freq))%>%
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>%
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq))


```


####  two_lowdiv


```{r}

malariaem_phased_per_sample_summary_two_lowdiv = malariaem_phased_per_sample_summary %>% 
  filter(group_id == "two_lowdiv")

malariaem_freq_summary_two_lowdiv = malariaem_freq_summary %>% 
  filter(group_id == "two_lowdiv")

all_phased_pop_freqs_prevs_out_two_lowdiv = all_phased_pop_freqs_prevs_out %>% 
  filter(group_id == "two_lowdiv")
all_phased_sample_out_two_lowdiv = all_phased_sample_out %>% 
    filter(group_id == "two_lowdiv")

two_lowdiv_group_popHapId = malariaem_phased_per_sample_summary_two_lowdiv %>% 
  select(target_id, seq) %>% 
  bind_rows(all_phased_sample_out_two_lowdiv %>% 
              select(target_id, seq), 
            malariaem_freq_summary_two_lowdiv %>% 
              select(target_id, seq)) %>% 
  group_by(target_id, seq) %>% 
  summarise(n = n()) %>%
  arrange(target_id, desc(n)) %>% 
  group_by(target_id) %>% 
  mutate(row = row_number()) %>% 
  mutate(seq_name = paste0(target_id, ".", stringr::str_pad(row, width = nchar(max(row)), pad = "0")))

two_lowdiv_group_popHapId_sel = two_lowdiv_group_popHapId %>% 
  ungroup() %>% 
  select(target_id, seq, seq_name)

all_phased_pop_freqs_prevs_out_two_lowdiv_mod = all_phased_pop_freqs_prevs_out_two_lowdiv %>% 
  left_join(two_lowdiv_group_popHapId_sel)%>% 
  group_by(population_name, group_id, pop_simulated_gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_two_lowdiv_mod = malariaem_freq_summary_two_lowdiv %>% 
  left_join(two_lowdiv_group_popHapId_sel) %>% 
  group_by(population_name, group_id, gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_two_lowdiv_mod_sel = malariaem_freq_summary_two_lowdiv_mod %>% 
  ungroup() %>% 
  select(population_name, paramset_id, hap_name, freq, freq_se, group_id) %>% 
  unique()

all_phased_pop_freqs_prevs_out_two_lowdiv_mod_sel = all_phased_pop_freqs_prevs_out_two_lowdiv_mod %>% 
  ungroup() %>% 
  select(population_name, hap_name, expected_allele_freq, group_id) %>% 
  unique()

malariaem_freq_summary_two_lowdiv_mod_sel_combined = malariaem_freq_summary_two_lowdiv_mod_sel %>% 
  full_join(all_phased_pop_freqs_prevs_out_two_lowdiv_mod_sel %>% 
              dplyr::rename("expected_freq" = expected_allele_freq))%>%
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>%
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq))


```


####  four


```{r}

malariaem_phased_per_sample_summary_four = malariaem_phased_per_sample_summary %>% 
  filter(group_id == "four")

malariaem_freq_summary_four = malariaem_freq_summary %>% 
  filter(group_id == "four")

all_phased_pop_freqs_prevs_out_four = all_phased_pop_freqs_prevs_out %>% 
  filter(group_id == "four")
all_phased_sample_out_four = all_phased_sample_out %>% 
    filter(group_id == "four")

four_group_popHapId = malariaem_phased_per_sample_summary_four %>% 
  select(target_id, seq) %>% 
  bind_rows(all_phased_sample_out_four %>% 
              select(target_id, seq), 
            malariaem_freq_summary_four %>% 
              select(target_id, seq)) %>% 
  group_by(target_id, seq) %>% 
  summarise(n = n()) %>%
  arrange(target_id, desc(n)) %>% 
  group_by(target_id) %>% 
  mutate(row = row_number()) %>% 
  mutate(seq_name = paste0(target_id, ".", stringr::str_pad(row, width = nchar(max(row)), pad = "0")))

four_group_popHapId_sel = four_group_popHapId %>% 
  ungroup() %>% 
  select(target_id, seq, seq_name)

all_phased_pop_freqs_prevs_out_four_mod = all_phased_pop_freqs_prevs_out_four %>% 
  left_join(four_group_popHapId_sel)%>% 
  group_by(population_name, group_id, pop_simulated_gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_four_mod = malariaem_freq_summary_four %>% 
  left_join(four_group_popHapId_sel) %>% 
  group_by(population_name, group_id, gt_id) %>% 
  arrange(target_id) %>% 
  mutate(hap_name = paste0(seq_name, collapse = "__"))

malariaem_freq_summary_four_mod_sel = malariaem_freq_summary_four_mod %>% 
  ungroup() %>% 
  select(population_name, paramset_id, hap_name, freq, freq_se, group_id) %>% 
  unique()

all_phased_pop_freqs_prevs_out_four_mod_sel = all_phased_pop_freqs_prevs_out_four_mod %>% 
  ungroup() %>% 
  select(population_name, hap_name, expected_allele_freq, group_id) %>% 
  unique()

malariaem_freq_summary_four_mod_sel_combined = malariaem_freq_summary_four_mod_sel %>% 
  full_join(all_phased_pop_freqs_prevs_out_four_mod_sel %>% 
              dplyr::rename("expected_freq" = expected_allele_freq))%>%
  mutate(freq = ifelse(is.na(freq), 0, freq)) %>%
  mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq))


```



```{r}

malariaem_freq_summary_all_combined = bind_rows(
  malariaem_freq_summary_two_meddiv_mod_sel_combined, 
  malariaem_freq_summary_two_lowdiv_mod_sel_combined, 
  malariaem_freq_summary_four_mod_sel_combined
) %>% 
  mutate(group_id = factor(group_id, levels = c("two_lowdiv", "two_meddiv", "four")))

malariaem_freq_summary_all_combined_ccc = compute_ccc(malariaem_freq_summary_all_combined, 
                                                                     expected_freq, freq, 
                                                                     c(group_id)
                                                                     )

malariaem_freq_summary_all_combined = malariaem_freq_summary_all_combined %>% 
  add_diff_mean(expected_freq, freq)



malariaem_freq_summary_all_combined_bland_alt = make_bland_altman_plot_by_label(malariaem_freq_summary_all_combined, "freq") + 
  facet_grid(~group_id)

cairo_pdf("malariaem_freq_summary_all_combined_bland_alt.pdf", width = 8, height = 4)
print(malariaem_freq_summary_all_combined_bland_alt)
dev.off()

malariaem_freq_summary_all_combined = malariaem_freq_summary_all_combined %>%
  mutate(TP = as.numeric(freq != 0 & expected_freq != 0)) %>% 
  mutate(FP = as.numeric(freq != 0 & expected_freq == 0)) %>% 
  mutate(FN = as.numeric(freq == 0 & expected_freq != 0))


malariaem_freq_summary_all_combined_contig = malariaem_freq_summary_all_combined %>% 
  group_by(group_id) %>% 
  add_contigency_metrics_no_TN(T)

malariaem_freq_summary_all_combined_contig_precission_recall = malariaem_freq_summary_all_combined_contig %>% 
  gather(measure, value, 7, 9)


malariaem_freq_summary_all_combined_contig_precission_recall_plot = ggplot(malariaem_freq_summary_all_combined_contig_precission_recall) + 
  geom_col(aes(x = group_id, y = value, fill = measure), color = "black", position = "dodge") + 
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values =colorPalette_08)

cairo_pdf("malariaem_freq_summary_all_combined_contig_precission_recall_plot.pdf", width = 4, height = 4)
print(malariaem_freq_summary_all_combined_contig_precission_recall_plot)
dev.off()


malariaem_freq_summary_all_combined_fp = malariaem_freq_summary_all_combined %>% 
  filter(FP == 1)


malariaem_freq_summary_all_combined_false_positive_freq_plot = ggplot(malariaem_freq_summary_all_combined_fp) + 
  geom_histogram(aes(x = freq)) +
  scale_y_log10() + 
  facet_grid(~group_id) + 
  custom_ggplot_theme + 
  labs(y = "count (log10)")

cairo_pdf("malariaem_freq_summary_all_combined_false_positive_freq_plot.pdf", width = 6, height = 4)
print(malariaem_freq_summary_all_combined_false_positive_freq_plot)
dev.off()
```


