---
title: Processing all COI tools Benchmarking Results
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: true
fig-dpi: 300
fig-align: center
number-sections: true
anchor-sections: true
bibliography: [references.bib]
biblio-style: apalike
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)
```

## Intro

COI tools were evaluated on their ability to accurately estimate COI for amplicon data from simulated data, see [here](Benchmarking_Overview.qmd) for details. 

COI tools evaluated: 

*  THEREALMcCOIL[@Chang2017-mq]  
*  MOIRE[@Murphy2024-or]  
*  coiaf[@Paschalidis2023-fv]  
*  naive estimators (bespoke code)  


## Metrics 

The following metrics were calculated on the results. Several tools have several parameters that can be modified and so have several results over different parameter sets. For final comparison between tools, the best parameter set was choosen. 


*  Root mean squared error (RMSE) and Mean absolute error (MAE)[@Chai2014-cn]
    *  Calculated by taking the square root of the mean sum of squared difference between two metrics, not bounded but greater than or equal to 0 
*  [Bland-altman plots](https://en.wikipedia.org/wiki/Bland%E2%80%93Altman_plot) 
    *  Plotting the observed minus the expected on the y and on the x-aixs the mean between the two measures, can illustrate where and by how measures differ 
*  [Concoordance](https://www.rdocumentation.org/packages/DescTools/versions/0.99.60/topics/CCC)
    *  R function documentation from library DescTools(Signorell, 2025) can be found here: <https://www.rdocumentation.org/packages/DescTools/versions/0.99.60/topics/CCC>
    *  Description: "Computes Lin's (Lin, 1989; Steichen & Cox, 2002) concordance correlation coefficient for agreement on a continuous measure obtained by two methods. The concordance correlation coefficient combines measures of both precision and accuracy to determine how far the observed data deviate from the line of perfect concordance (that is, the line at 45 degrees on a square scatter plot). Lin's coefficient increases in value as a function of the nearness of the data's reduced major axis to the line of perfect concordance (the accuracy of the data) and of the tightness of the data about its reduced major axis (the precision of the data)."


## Set up 

### Libraries used in the analysis below 

```{r, echo=FALSE}
source("../tutorials/scripts/benchmarking_functions.R")

```

```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-line-numbers: true
#| filename: ../tutorials/scripts/benchmarking_functions.R
#| file: ../tutorials/scripts/benchmarking_functions.R

```

## Load in the Benchmarking Outputs

```{r}
inp <- list(
  sim_info_dir = "../data/benchmarking//simulations/SpotMalaria/", 
  madhatter_res_dir = "../data/benchmarking/tool_results/madhatter/"
)

inp_per_tools <- list(
  "therealmccoil" = list(
      tool_res_dir = "../data/benchmarking/tool_results/THEREALMcCOIL/", 
      tool_params = "../data/benchmarking/tool_results/therealmccoil.json"
  ), 
  "moire" = list(
      tool_res_dir = "../data/benchmarking/tool_results/MOIRE/", 
      tool_params = "../data/benchmarking/tool_results/moire.json"
  ), 
  "naive_coi" = list(
      tool_res_dir = "../data/benchmarking/tool_results/naive_coi/", 
      tool_params = "../data/benchmarking/tool_results/naive_coi.json"
  ),
  "coiaf" = list(
      tool_res_dir = "../data/benchmarking/tool_results/coiaf//", 
      tool_params = "../data/benchmarking/tool_results/coiaf.json"
  ) 
)
```



First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.

### Reading in expected truths 

Read per in expected allele frequencies per sample tables to get true COIs

```{r}
expected_simulated_frequencies_per_sample <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "trimmed_expected_simulated_frequencies_per_sample.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(library_sample_name = libary_sample_name)

expected_simulated_frequencies_per_sample_COIs = expected_simulated_frequencies_per_sample %>% 
  select(panel, population_name, country, library_sample_name, COI) %>% 
  distinct() %>% 
  dplyr::rename(expected_COI = COI) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name)) %>% 
  mutate(library_sample_name = as.character(library_sample_name))

```

Read per in expected allele frequencies tables population level 
```{r}
population_trimmed_mhap_freqs_prevs <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_trimmed_mhap_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(),
        population_name = col_character(), 
        pop = col_character(), 
        country = col_character(), 
        target = col_character(), 
        seq = col_character()
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(target_id = target)
```


Read in all simulation parameters into one tibble. Encode different population parameters as descriptive factors. 

```{r}

sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>% 
  mutate(
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>% 
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  )) %>% 
  mutate(pop_cat = factor(pop_cat, levels = sort(unique((.$pop_cat)))), 
         coi_cat = factor(coi_cat, levels = sort(unique((.$coi_cat)))), 
         between_sample_related_cat = factor(between_sample_related_cat, levels = sort(unique((.$between_sample_related_cat)))), 
         nonrecombinant_cat = factor(nonrecombinant_cat, levels = sort(unique((.$nonrecombinant_cat)))), 
         meiosis_cat = factor(meiosis_cat, levels = sort(unique((.$meiosis_cat)))))

create_dt(sim_params)
```

Read in madhatter pipeline allele tables 

```{r}
madhatter_allele_data <- tibble(
    allele_data_path = list.files(
      inp$madhatter_res_dir, 
      pattern = "allele_data.txt.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    allele_data = map(
      allele_data_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  mutate(tool_res_dir = dirname(allele_data_path)) %>%
  mutate(population_name = gsub("_mh_results", "", basename(tool_res_dir))) %>%
  select(-allele_data_path, -tool_res_dir) %>%
  unnest(allele_data) %>% 
  mutate(SampleID = gsub("_S[0-9]+$", "", SampleID)) %>% 
  dplyr::rename(library_sample_name = SampleID, 
                target_id = Locus, 
                seq = ASV, 
                reads = Reads) 
```



### Read in THEREALMcCOI results

THEREALMcCOI outputs information on complexity of infection (COI) and on population allele frequencies and several different outputs need to be read in. 

Getting the complexity of infection (COI) from the `coi.tsv.gz` files 

```{r}
therealmccoil_coi_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$therealmccoil$tool_res_dir, 
      pattern = "^coi.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id)) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

THEREALMcCOI can run in two different models, categorical vs proportional

```{r}
# Read in tool parameters
therealmccoil_tool_params <- jsonlite::fromJSON(inp_per_tools$therealmccoil$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_model = paramset.model,
  ) 

create_dt(therealmccoil_tool_params)
```


### Read in MOIRE results

MOIRE outputs information on complexity of infection (COI) and on population allele frequencies, will read in only the COIs results for the below analysis, see [here](benchmarking_freq_tools.qmd) for freq results

Getting the complexity of infection (COI) from the `coi_summary.tsv.gz` files 

```{r}
moire_coi_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$moire$tool_res_dir, 
      pattern = "^coi_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id)) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Getting the effective COI from the `effective_coi_summary.tsv.gz` files. 

```{r}
moire_effective_coi_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$moire$tool_res_dir, 
      pattern = "effective_coi_summary.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id))%>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Getting what tool parameters were used. For MOIRE benchmarking we set the priors on the assumed amount of relatedness with moire_paramset1 assuming uniform relatedness and then each moire_paramset 2 and 3 increasing the amount of assumed relatedness. moire_paramsets3-6 are the same prior relatedness as 1-3 but have more parallel burin to explore more pace for convergence.   

```{r}
# Read in tool parameters
moire_tool_params <- jsonlite::fromJSON(inp_per_tools$moire$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    r_alpha = paramset.r_alpha,
    r_beta = paramset.r_beta
  )

create_dt(moire_tool_params)
```


### Read in COIAF results

COIAF outputs information on complexity of infection (COI) 

Getting the complexity of infection (COI) from the `coi.tsv.gz` files 

```{r}
coiaf_coi_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$coiaf$tool_res_dir, 
      pattern = "^coi.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id)) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))


coiaf_coi_summary = coiaf_coi_summary %>% 
  mutate(coi_freq = ifelse(is.na(coi_freq), 1, coi_freq))
```

coiaf can run with different assumptions of sequencing error and different maxs set on COI 

```{r}
# Read in tool parameters
coiaf_tool_params <- jsonlite::fromJSON(inp_per_tools$coiaf$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_seq_error = paramset.seq_error,
    paramset_max_coi = paramset.max_coi
  ) 

create_dt(coiaf_tool_params)
```


### Read in naive COI results

naive COI outputs information on complexity of infection (COI)  

Getting the complexity of infection (COI) from the `coi.tsv.gz` files 

```{r}
naive_coi_coi_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$naive_coi$tool_res_dir, 
      pattern = "^coi.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id)) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

naive_coi can run with different assumptions of sequencing error and different maxs set on COI 

```{r}
# Read in tool parameters
naive_coi_tool_params <- jsonlite::fromJSON(inp_per_tools$naive_coi$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_method = paramset.method,
    paramset_quantile_threshold = paramset.quantile_threshold,
    paramset_integer_threshold = paramset.integer_threshold
  ) %>% 
  mutate(paramset_threshold = ifelse(is.na(paramset_integer_threshold), paramset_quantile_threshold, paramset_integer_threshold)) %>% 
  select(-paramset_quantile_threshold, -paramset_integer_threshold)

create_dt(naive_coi_tool_params)
```



## Benchmarking 

### COI  

#### Metrics 

*  Plotting Bland-Altman plots (plotting the mean of expected and obs vs difference) 
*  Calculating the RMSE (taking the square root of the mean squared difference between expected and observed)  
*  Calculating [Lin's concordance correlation coefficient](https://www.rdocumentation.org/packages/DescTools/versions/0.99.60/topics/CCC) for agreement on a continuous measure 
*  Calculating contingency table of determining a sample as monoclonal vs polyclonal 

#### Results 

```{r}
therealmccoil_coi_summary_out = therealmccoil_coi_summary %>%
    mutate(tool = "therealmccoil")%>% 
    select(population_name, paramset_id, library_sample_name, tool, coi)%>% 
  left_join(expected_simulated_frequencies_per_sample_COIs %>% 
              mutate(library_sample_name = as.character(library_sample_name))) %>% 
  left_join(sim_params)

moire_coi_summary_out = moire_coi_summary %>%
    mutate(tool = "moire") %>% 
    select(population_name, paramset_id, library_sample_name, tool, coi, naive_coi, offset_naive_coi) %>% 
  left_join(expected_simulated_frequencies_per_sample_COIs %>% 
              mutate(library_sample_name = as.character(library_sample_name))) %>% 
  left_join(sim_params)
  
moire_effective_coi_summary_out = moire_effective_coi_summary %>%
    mutate(tool = "moire")%>% 
    select(population_name, paramset_id, library_sample_name, tool, ecoi)%>% 
  left_join(expected_simulated_frequencies_per_sample_COIs %>% 
              mutate(library_sample_name = as.character(library_sample_name))) %>% 
  left_join(sim_params)

coiaf_coi_summary_out =  coiaf_coi_summary %>%
    mutate(tool = "coiaf") %>% 
    select(population_name, paramset_id, library_sample_name, tool, coi_freq, coi_variant) %>% 
  left_join(expected_simulated_frequencies_per_sample_COIs %>% 
              mutate(library_sample_name = as.character(library_sample_name))) %>% 
  left_join(sim_params)
  
naive_coi_coi_summary_out = naive_coi_coi_summary %>%
    mutate(tool = "naive_coi") %>% 
  left_join(expected_simulated_frequencies_per_sample_COIs %>% 
              mutate(library_sample_name = as.character(library_sample_name))) %>% 
  left_join(sim_params)


```

```{r, echo = F, eval = F}

coiaf_coi_summary_out_ccc = coiaf_coi_summary_out %>% 
  compute_ccc(expected_COI, c(coi_freq, coi_variant), c(paramset_id, population_name)) %>% 
  left_join(coiaf_tool_params) %>% 
  left_join(sim_params)

coiaf_coi_summary_out_rmse = coiaf_coi_summary_out %>% 
  compute_rmse(expected_COI, c(coi_freq, coi_variant), c(paramset_id, population_name)) %>% 
  left_join(coiaf_tool_params) %>% 
  left_join(sim_params)

coiaf_coi_summary_out = coiaf_coi_summary_out %>% 
  add_diff_mean(expected_COI, coi_freq)%>% 
  add_diff_mean(expected_COI, coi_variant)

make_bland_altman_plot_by_label(coiaf_coi_summary_out, label = "coi_freq") + 
  facet_grid(coi_cat~paramset_id)

make_bland_altman_plot_by_label(coiaf_coi_summary_out, label = "coi_variant") + 
  facet_grid(coi_cat~paramset_id)
```

```{r, echo = F, eval = F}

coiaf_coi_summary_out_sp_coi_freq_paramset_id = coiaf_coi_summary_out %>% 
  select(population_name, paramset_id, library_sample_name, tool, coi_freq, expected_COI) %>% 
  spread(paramset_id, coi_freq)

coiaf_coi_summary_out_sp_coi_freq_paramset_id_plot =  ggpairs(coiaf_coi_summary_out_sp_coi_freq_paramset_id,
                                                  columns = c("coiaf_paramset1", "coiaf_paramset2", "coiaf_paramset3", "coiaf_paramset4", "expected_COI"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "all")

coiaf_coi_summary_out_sp_coi_variant_paramset_id = coiaf_coi_summary_out %>% 
  select(population_name, paramset_id, library_sample_name, tool, coi_variant, expected_COI) %>% 
  spread(paramset_id, coi_variant)

coiaf_coi_summary_out_sp_coi_variant_paramset_id_plot =  ggpairs(coiaf_coi_summary_out_sp_coi_variant_paramset_id,
                                                  columns = c("coiaf_paramset1", "coiaf_paramset2", "coiaf_paramset3", "coiaf_paramset4", "expected_COI"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "all")

```


```{r, echo = F, eval = F}

naive_coi_coi_summary_out_ccc = naive_coi_coi_summary_out %>% 
  compute_ccc(expected_COI, coi, c(paramset_id, population_name)) %>% 
  left_join(naive_coi_tool_params) %>% 
  left_join(sim_params)

naive_coi_coi_summary_out = naive_coi_coi_summary_out %>% 
  add_diff_mean(expected_COI, coi)

make_bland_altman_plot_by_label(naive_coi_coi_summary_out, label = "coi") + 
  facet_grid(coi_cat~paramset_id)
```



## Combining and getting best parameter set per program 

Summarizing results over each program's various parameters to determine it's best peforming parameter to include for the final comparison 


```{r}
coiaf_coi_summary_out_gat = coiaf_coi_summary_out %>% 
  gather(measure, coi, 5:6) %>% 
  mutate(paramset_id = paste0(paramset_id, "-", measure))
moire_coi_summary_out_mod = moire_coi_summary_out %>% 
  mutate(tool = paste0(tool, "-coi"))
moire_coi_summary_out_mod_offset_naive_coi = moire_coi_summary_out %>% 
  select(-coi) %>% 
  dplyr::rename(coi = offset_naive_coi) %>% 
  mutate(tool = "offset_naive")
moire_effective_coi_summary_out_mod =  moire_effective_coi_summary_out %>% 
  mutate(tool = paste0(tool, "-ecoi")) %>% 
  dplyr::rename(coi = ecoi)


all_combined = bind_rows(
  coiaf_coi_summary_out_gat %>% 
    select(population_name, tool, paramset_id, library_sample_name, coi, population_name_library_sample_name, panel, country, expected_COI),
  moire_coi_summary_out_mod %>% 
    select(population_name, tool, paramset_id, library_sample_name, coi, population_name_library_sample_name, panel, country, expected_COI), 
  moire_effective_coi_summary_out_mod%>% 
    select(population_name, tool, paramset_id, library_sample_name, coi, population_name_library_sample_name, panel, country, expected_COI),  
  therealmccoil_coi_summary_out%>% 
    select(population_name, tool, paramset_id, library_sample_name, coi, population_name_library_sample_name, panel, country, expected_COI),  
  naive_coi_coi_summary_out %>% 
    select(population_name, tool, paramset_id, library_sample_name, coi, population_name_library_sample_name, panel, country, expected_COI), 
  
  # moire_coi_summary_out_mod_offset_naive_coi %>% 
  #   select(population_name, tool, paramset_id, library_sample_name, coi, population_name_library_sample_name, panel, country, expected_COI)
) %>% 
  mutate(tool_paramset_id = paste0(tool, paramset_id))

all_combined_ccc_per_tool_per_params = all_combined %>% 
  compute_ccc(
    expected_COI, coi, 
    c(tool, paramset_id)
  )

all_combined_ccc_per_tool_per_params_best = all_combined_ccc_per_tool_per_params %>% 
  group_by(tool) %>% 
  slice_max(order_by = coi_ccc, n = 1, with_ties = F) %>% 
  mutate(tool_paramset_id = paste0(tool, paramset_id))


all_combined_best = all_combined %>% 
  filter(tool_paramset_id %in% all_combined_ccc_per_tool_per_params_best$tool_paramset_id)%>% 
  add_diff_mean(expected_COI, coi) %>% 
  left_join(sim_params)
```


```{r}

all_combined_best_ccc = all_combined_best %>% 
  compute_ccc(
    expected_COI, coi, 
    c(tool, population_name)
  ) %>% 
  left_join(sim_params)

all_combined_best_ccc_plot = ggplot(all_combined_best_ccc) + 
  geom_boxplot(aes(y = coi_ccc, fill = tool)) + 
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08, 
                    guide = guide_legend(ncol = 2) 
                    ) + 
  theme(axis.text.x = element_blank(), 
        axis.ticks.x = element_blank(), 
        axis.title.x = element_blank()) + 
  geom_hline(linetype = "dotted", color = "red", yintercept = 0.95)

```

```{r}
#| fig-width: 4
#| fig-height: 4

print(all_combined_best_ccc_plot)
```

```{r}

cairo_pdf("all_combined_best_ccc_plot.pdf", width = 4, height = 4)
print(all_combined_best_ccc_plot)
dev.off()
```

```{r}
all_combined_best_by_tool = make_bland_altman_plot_by_label(all_combined_best, label = "coi") + 
  facet_grid(~tool)

```

```{r}
#| fig-width: 10
#| fig-height: 5

print(all_combined_best_by_tool)
```

```{r}
cairo_pdf("all_coi_tools_combined_best_bland_altman.pdf", width = 10, height = 5)
print(all_combined_best_by_tool)
dev.off()

```


```{r}
all_combined_best_sp = all_combined_best %>% 
  select(-paramset_id, -tool_paramset_id, -coi_diff_from_exp, -coi_mean_with_exp) %>% 
  mutate(tool = gsub("-", "_", tool)) %>% 
  spread(tool, coi)

all_combined_best_sp_tool_plot =  ggpairs(all_combined_best_sp,
                                                  columns = c("coiaf", "moire_coi", "moire_ecoi", "naive_coi", "therealmccoil", "expected_COI"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "coi")

```

```{r}
#| fig-width: 13
#| fig-height: 13

print(all_combined_best_sp_tool_plot)
```

```{r}
cairo_pdf("all_combined_best_sp_tool_plot.pdf", width = 13, height = 13)
print(all_combined_best_sp_tool_plot)
dev.off()

all_combined_best_sp_tool_plot_high_coi =  ggpairs(all_combined_best_sp %>% 
                                            filter(coi_cat != "MeanCOI=1.26"),
                                                  columns = c("coiaf", "moire_coi", "moire_ecoi", "naive_coi", "therealmccoil", "expected_COI"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "coi")


```

```{r}
#| fig-width: 13
#| fig-height: 13

print(all_combined_best_sp_tool_plot_high_coi)
```

```{r}
cairo_pdf("all_combined_best_sp_tool_plot_high_coi.pdf", width = 13, height = 13)
print(all_combined_best_sp_tool_plot_high_coi)
dev.off()
```





