---
title: Benchmarking Results - Frequency Tools Multi locus Variants
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: true
fig-dpi: 300
fig-align: center
number-sections: true
anchor-sections: true
bibliography: [references.bib]
biblio-style: apalike
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)

library(ComplexHeatmap)


get_color_scale_breaks <-function(mat_dat){
  min_val = min(mat_dat, na.rm = T)
  max_val = max(mat_dat, na.rm = T)
  
  return(c(min_val, min_val + (max_val - min_val)/2, max_val))
}

```

## Set up 

### Libraries used in the analysis below 

```{r, echo=FALSE}
source("../tutorials/scripts/benchmarking_functions.R")
```

```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-line-numbers: true
#| filename: ../tutorials/scripts/benchmarking_functions.R
#| file: ../tutorials/scripts/benchmarking_functions.R

```

## Load in the Benchmarking Outputs

```{r}
inp <- list(
  sim_info_dir = "../data/benchmarking//simulations/SpotMalaria/", 
  madhatter_res_dir = "../data/benchmarking/tool_results/madhatter/"
)

inp_per_tools <- list(
  # multi-locus tools
  "fem" = list(
    tool_res_dir = "../data/benchmarking/tool_results/FEM/", 
    tool_params = "../data/benchmarking/tool_results/fem.json"
  ), 
  "mlbm" = list(
    tool_res_dir = "../data/benchmarking/tool_results/MLBM//", 
    tool_params = "../data/benchmarking/tool_results/mlbm.json"
  ), 
  "naive_mlafp" = list(
    tool_res_dir = "../data/benchmarking/tool_results/naive_mlafp/", 
    tool_params = "../data/benchmarking/tool_results/naive_mlafp.json"
  ), 
  "naive_mlafp_njh" = list(
    tool_res_dir = "../data/benchmarking/tool_results/naive_mlafp_njh/", 
    tool_params = "../data/benchmarking/tool_results/naive_mlafp_njh.json"
  ), 
  "snpslice" = list(
    tool_res_dir = "../data/benchmarking/tool_results/SNPSlice/", 
    tool_params = "../data/benchmarking/tool_results/snpslice.json"
  )
)



```



First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.

### Reading in expected truths 

Read per in expected aa multilocus freq/prev tables 


```{r}
expected_population_aa_multiloci_variants_freqs_prevs <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_aa_multiloci_variants_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(),
        population_name = col_character(), 
        pop = col_character(), 
        country = col_character(), 
        variant = col_character(), 
        aas_in_variant = col_integer(),
				allele_count = col_integer(),
				sample_count = col_integer(),
				total_samples = col_integer(),
				total_alleles = col_integer(),
				freq = col_double(),
				prev  = col_double(), 
				group_id = col_character()
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth)

colnames(expected_population_aa_multiloci_variants_freqs_prevs)[5:11] = paste0("expected_",
                                                                               colnames(expected_population_aa_multiloci_variants_freqs_prevs)[5:11])


```

Read in all simulation parameters into one tibble. Encode different population parameters as descriptive factors. 

```{r}

sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>% 
  mutate(
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>% 
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  )) %>% 
  mutate(pop_cat = factor(pop_cat, levels = sort(unique((.$pop_cat)))), 
         coi_cat = factor(coi_cat, levels = sort(unique((.$coi_cat)))), 
         between_sample_related_cat = factor(between_sample_related_cat, levels = sort(unique((.$between_sample_related_cat)))), 
         nonrecombinant_cat = factor(nonrecombinant_cat, levels = sort(unique((.$nonrecombinant_cat)))), 
         meiosis_cat = factor(meiosis_cat, levels = sort(unique((.$meiosis_cat))))) %>% 
  select(-sample_count)

create_dt(sim_params)
```


## multi-locus variants 


* FreqEstimationModel(FEM)[@Taylor2014-hk]
* MultiLociBiallelicModel(MLBM)[@Tsoungui_Obama2022-gz]
* snp-slice[@Ju2024-fh]
* naive multi locus freq estimator variant_string
* naive multi locus freq estimator njh



### Read in FreqEstimationModel results

FreqEstimationModel outputs information on frequency and prevalence of alleles 

Getting the frequency and prevs from the `FEM_summary.tsv.gz` files 

```{r}
fem_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$fem$tool_res_dir, 
      pattern = "^FEM_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

fem can take a COI argument 

```{r}
# Read in tool parameters
fem_tool_params <- jsonlite::fromJSON(inp_per_tools$fem$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_coi = paramset.coi,
  )

create_dt(fem_tool_params)
```



### Read in MLBM results

MLBM outputs information on frequency but not prevalence of alleles 

Getting the frequency from the `MLBM_summary.tsv.gz` files 

```{r}
mlbm_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$mlbm$tool_res_dir, 
      pattern = "^MLBM_summary.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

mlbm has no additional parameters  

```{r}
# Read in tool parameters
mlbm_tool_params <- jsonlite::fromJSON(inp_per_tools$mlbm$tool_params, flatten = TRUE) %>%
  as_tibble()

create_dt(mlbm_tool_params)
```



### Read in naive multi locus freq estimator results

naive multi locus freq estimator outputs information on frequency but not prevalence of alleles 

Getting the freqs and prevs from the `mlafp.tsv.gz` files 

```{r}
naive_mlafp_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$naive_mlafp$tool_res_dir, 
      pattern = "^mlafp.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

naive_mlafp has no additional parameters  

```{r}
# Read in tool parameters
naive_mlafp_tool_params <- jsonlite::fromJSON(inp_per_tools$naive_mlafp$tool_params, flatten = TRUE) %>%
  as_tibble()

create_dt(naive_mlafp_tool_params)
```



### Read in SNPSLICE multi locus freq estimator results

SNPSLICE multi locus freq estimator  

Getting the freqs and prevs from the `mlaf.tsv.gz` files 

```{r}
snpslice_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$snpslice$tool_res_dir, 
      pattern = "^mlaf.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

snpslice   

```{r}
# Read in tool parameters
snpslice_tool_params <- jsonlite::fromJSON(inp_per_tools$snpslice$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_model = paramset.model,
    paramset_loci_limit = paramset.loci_limit,
    paramset_alpha = paramset.alpha,
    paramset_gap = paramset.gap,
    paramset_use_mcmc_for_af_and_coi = paramset.use_mcmc_for_af_and_coi
  ) 

create_dt(snpslice_tool_params)
```



### Read in naive multi locus freq estimator results njh

naive multi locus freq estimator outputs information on frequency but not prevalence of alleles 

Getting the freqs and prevs from the `mlafp.tsv.gz` files 

```{r}
naive_mlafp_njh_freq_summary <- tibble(
    tool_res_path = list.files(
      inp_per_tools$naive_mlafp_njh$tool_res_dir, 
      pattern = "^mlafp.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 

```

naive_mlafp_njh has no additional parameters  

```{r}
# Read in tool parameters
naive_mlafp_njh_tool_params <- jsonlite::fromJSON(inp_per_tools$naive_mlafp_njh$tool_params, flatten = TRUE) %>%
  as_tibble()

create_dt(naive_mlafp_njh_tool_params)
```


### Summarizing expected frequenies

```{r}
# calculating he 
expected_population_aa_multiloci_variants_freqs_prevs_he = expected_population_aa_multiloci_variants_freqs_prevs %>% 
  group_by(population_name, group_id) %>% 
  summarise(he = 1 - sum(expected_freq ^2), 
            variant_n = n_distinct(variant))

expected_population_aa_multiloci_variants_freqs_prevs_he_sum = expected_population_aa_multiloci_variants_freqs_prevs_he %>% 
  group_by(group_id) %>% 
  summarise(quick_summary(he))
# filtering to just sites that have at least some variation in any of the simulated populations 

expected_population_aa_multiloci_variants_freqs_prevs_he_sum_invariant = expected_population_aa_multiloci_variants_freqs_prevs_he_sum %>% 
  filter(q100 == 0)
expected_population_aa_multiloci_variants_freqs_prevs_he_sum_variant = expected_population_aa_multiloci_variants_freqs_prevs_he_sum %>% 
  filter(q100 != 0) %>% 
  arrange(mean) %>% 
  mutate(group_id = factor(group_id, levels = unique(.$group_id)))

expected_population_aa_multiloci_variants_freqs_prevs_he_variant = expected_population_aa_multiloci_variants_freqs_prevs_he %>% 
  filter(group_id %!in% expected_population_aa_multiloci_variants_freqs_prevs_he_sum_invariant$group_id)


expected_population_aa_multiloci_variants_freqs_prevs_variant = expected_population_aa_multiloci_variants_freqs_prevs %>% 
  filter(group_id %!in% expected_population_aa_multiloci_variants_freqs_prevs_he_sum_invariant$group_id)
```

Ranges of he for loci 

```{r}
ggplot(expected_population_aa_multiloci_variants_freqs_prevs_he_variant %>% 
         left_join(sim_params) %>% 
         mutate(group_id = factor(group_id, levels = expected_population_aa_multiloci_variants_freqs_prevs_he_sum_variant$group_id))
       ) + 
  geom_boxplot(aes(x = group_id, y = he)) + 
  custom_ggplot_theme_xRotate + 
  scale_color_manual(values = colorPalette_08) + 
  facet_grid(country~.)

```

Some variants are invariant in one ancestor population but may be highly variant in the other 

```{r}
#| fig-width: 30
#| fig-height: 10
# Get ordered variant positions
variant_levels <- expected_population_aa_multiloci_variants_freqs_prevs_he_sum_variant$group_id

# Create shading data frame
n_variants <- length(variant_levels)
shade_df <- data.frame(
  group_id = factor(variant_levels, levels = variant_levels),
  xmin = seq_along(variant_levels) - 0.5,
  xmax = seq_along(variant_levels) + 0.5,
  fill = rep(c("gray", "white"), length.out = n_variants)
)

ggplot(expected_population_aa_multiloci_variants_freqs_prevs_he_variant %>% 
         left_join(sim_params) %>% 
         mutate(group_id = factor(group_id, levels = variant_levels))
       ) + 
  geom_rect(data = shade_df,
            aes(xmin = xmin, xmax = xmax, ymin = -Inf, ymax = Inf, fill = fill),
            inherit.aes = FALSE, alpha = 0.4) +
  scale_fill_manual(values = c("gray" = "grey50", "white" = "white"), guide = "none") +
  geom_boxplot(aes(x = group_id, y = he, color = factor(country))) + 
  custom_ggplot_theme_xRotate + 
  scale_color_manual(values = colorPalette_08)


```


So unlike the single locus variants, these groups rarely have allele frequencies of 1 and majority of frequencies are low level frequencies. 

```{r}
ggplot(expected_population_aa_multiloci_variants_freqs_prevs_variant) + 
  geom_histogram(aes(x = expected_freq), binwidth = 0.01)
```



### combining freqs 

```{r}
add_expected_freqs <- function(summary_dat, expected_dat, tool_params, tool_name) {
  
  # some tools  
  summary_dat_variant_info = summary_dat %>% 
    select(variant) %>% 
    unique() %>% 
    rowwise() %>% 
    mutate(observed_aas_in_variant = variantstring::variant_to_long(variant)[[1]] %>%
             select(gene, pos) %>%
             unique() %>%
             nrow())

  ## expand expected results for all parameter sets so they all get expected results
  ## fill the unmatched expected with 0's and with unmatched observed (false haps) with 0's
  summary_dat_out = summary_dat %>%
    left_join(tool_params%>% 
              select(-loci_groups)) %>% 
    left_join(expected_dat %>% 
                ungroup() %>% 
                select(group_id, expected_aas_in_variant) %>% 
                unique()) %>% 
    left_join(summary_dat_variant_info) %>%
    full_join(expected_dat %>%
                crossing(paramset_id = unique(sort(
                  summary_dat$paramset_id
                )))%>% 
              select(-expected_aas_in_variant), 
              by = c("population_name", "paramset_id", "group_id",  "variant")) %>%
    mutate(freq = ifelse(is.na(freq), 0, freq)) %>%
    mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq)) %>%
    mutate(tool =  tool_name) %>%
    mutate(tool_paramset_id = paste0(tool, "-", paramset_id))
  
  # not all tools output prev
  if ("prev" %in% colnames(summary_dat_out)) {
    summary_dat_out = summary_dat_out %>%
      mutate(prev = ifelse(is.na(prev), 0, prev)) %>%
      mutate(expected_freq = ifelse(is.na(expected_freq), 0, expected_freq)) %>%
      mutate(expected_prev = ifelse(is.na(expected_prev), 0, expected_prev))
  }
  
  return(summary_dat_out)
}

```


```{r}
fem_freq_summary_out = add_expected_freqs(fem_freq_summary, 
                                          expected_population_aa_multiloci_variants_freqs_prevs, 
                                          fem_tool_params, 
                                          "fem")

mlbm_freq_summary_out = add_expected_freqs(mlbm_freq_summary, 
                                          expected_population_aa_multiloci_variants_freqs_prevs, 
                                          mlbm_tool_params, 
                                          "mlbm")

snpslice_freq_summary_out = add_expected_freqs(snpslice_freq_summary, 
                                          expected_population_aa_multiloci_variants_freqs_prevs, 
                                          snpslice_tool_params, 
                                          "snpslice")

naive_mlafp_freq_summary_out = add_expected_freqs(naive_mlafp_freq_summary, 
                                          expected_population_aa_multiloci_variants_freqs_prevs, 
                                          naive_mlafp_tool_params, 
                                          "naive_mlafp")

naive_mlafp_njh_freq_summary_out = add_expected_freqs(naive_mlafp_njh_freq_summary, 
                                          expected_population_aa_multiloci_variants_freqs_prevs, 
                                          naive_mlafp_njh_tool_params, 
                                          "naive_mlafp_njh")

```

```{r}
combined_mlfreq = 
  bind_rows(
    fem_freq_summary_out %>%
      select(population_name, tool, paramset_id, tool_paramset_id, group_id, variant, freq, expected_freq, observed_aas_in_variant, expected_aas_in_variant), 
    mlbm_freq_summary_out %>%
      select(population_name, tool, paramset_id, tool_paramset_id, group_id, variant, freq, expected_freq, observed_aas_in_variant, expected_aas_in_variant), 
    snpslice_freq_summary_out %>%
      select(population_name, tool, paramset_id, tool_paramset_id, group_id, variant, freq, expected_freq, observed_aas_in_variant, expected_aas_in_variant), 
    naive_mlafp_freq_summary_out %>%
      select(population_name, tool, paramset_id, tool_paramset_id, group_id, variant, freq, expected_freq, observed_aas_in_variant, expected_aas_in_variant), 
    naive_mlafp_njh_freq_summary_out %>%
      select(population_name, tool, paramset_id, tool_paramset_id, group_id, variant, freq, expected_freq, observed_aas_in_variant, expected_aas_in_variant)
  ) %>%  # remove output for tools that have output multi-locus haplotypes that aren't the full group
  filter(is.na(observed_aas_in_variant) | observed_aas_in_variant == expected_aas_in_variant)

```





```{r}
# looking at just at frequencies that aren't 1, it's important that sites with a frequency of 1 are also correct but currently 
# no tool has any issues with getting a freq of 1, having it being over representation makes evaluation biased  

combined_mlfreq_ccc_by_tool_group_id_params_filt = combined_mlfreq %>% 
  compute_ccc(expected_freq, freq, c(tool, group_id, paramset_id, tool_paramset_id))

combined_mlfreq_ccc_by_tool_group_id_params_filt_filt = combined_mlfreq_ccc_by_tool_group_id_params_filt 

combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp = combined_mlfreq_ccc_by_tool_group_id_params_filt_filt %>% 
  ungroup() %>% 
  select(group_id, tool_paramset_id, freq_ccc) %>% 
  spread(tool_paramset_id, freq_ccc)

combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp_mat = as.matrix(combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp[,2:ncol(combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp)])
rownames(combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp_mat) = combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp$group_id


library(circlize)
col_fun = colorRamp2(get_color_scale_breaks(combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp_mat), c( "#3288bd", "#ffffbf", "#d53e4f"))

variant_meta = tibble(
  group_id = rownames(combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp_mat)
) %>% 
  left_join(expected_population_aa_multiloci_variants_freqs_prevs_he_sum %>% 
              select(group_id, mean) %>% 
              dplyr::rename(mean_he = mean)) %>% 
  select(-group_id) %>% 
  as.data.frame()
row_anno = rowAnnotation(
  df = variant_meta, 
  col = list(
    mean_he = colorRamp2(c(min(variant_meta$mean_he, na.rm = T),
                       median(variant_meta$mean_he, na.rm = T),
                     max(variant_meta$mean_he, na.rm = T)),
                   c( "#ffffcc", "#feb24c", "#fc4e2a"))
  )
)

```

```{r}
#| fig-width: 15
#| fig-height: 10
combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp_mat_hp = Heatmap(combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp_mat, 
                                                                 name = "concordance", 
                                                                 cluster_rows = T,
                                                                 cluster_columns = F,
                                                                 row_gap = unit(0.5, "cm"),
                                                                 col = col_fun,
                                                                 rect_gp = gpar(col = "black", lwd = 1),
                                                                 left_annotation = row_anno,
                                                                 # row_split = tool_meta$tool, 
                                                                cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(
      sprintf("%.2f", combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp_mat[i, j]),
      x,
      y
      #,gp = gpar(fontsize = 8, col = "black")
    )
  })

draw(combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp_mat_hp, 
     background = "transparent", 
     padding = unit(c(50, 5, 5, 5), "mm"), 
     merge_legends = T)

cairo_pdf("combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp_mat_hp.pdf", width = 15, height = 10, bg = "transparent")
draw(combined_mlfreq_ccc_by_tool_group_id_params_filt_filt_sp_mat_hp, background = "transparent", 
     padding = unit(c(50, 5, 5, 5), "mm"), 
     merge_legends = T)
dev.off()

```



### Best per population COI parameters 

The mean population COI makes a large difference to which parameter set works best for which tool. This is due to the max possible COIs in each population. 

```{r}
combined_mlfreq_ccc_per_tool_per_params_per_coi_cat = combined_mlfreq %>% 
  left_join(sim_params) %>% 
  compute_ccc(
    expected_freq, freq, 
    c(tool, paramset_id, coi_cat, group_id)
  )

combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp = combined_mlfreq_ccc_per_tool_per_params_per_coi_cat %>% 
  ungroup() %>% 
  mutate(group_id_coi_cat = paste0(group_id, "\n", coi_cat)) %>% 
  select(paramset_id, group_id_coi_cat, freq_ccc) %>% 
  spread(group_id_coi_cat, freq_ccc)

combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat = as.matrix(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp[,2:ncol(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp)])
rownames(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat) = combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp$paramset_id

tool_meta = tibble(
  paramset_id = rownames(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat)
)%>% 
  mutate(tool = gsub("_coi_(freq|variant)", "", gsub("_p.*", "", paramset_id)))

col_meta = tibble(
  group_id_coi_cat = colnames(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat)
) %>% 
  mutate(group_id = gsub("\\n.*", "", group_id_coi_cat)) %>% 
  mutate(group_id = factor(group_id,
                           levels = c("two_group1",
                                      "pfdhfr",
                                      "pfdhps",
                                      "four_group1",
                                      "six_low_div",
                                      "pfdhfr_pfdhps",
                                      "six_med_div")))

library(circlize)
col_fun = colorRamp2(get_color_scale_breaks(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat), c( "#3288bd", "#ffffbf", "#d53e4f"))

```


```{r}
#| fig-width: 20
#| fig-height: 8
combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp_hp = Heatmap(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat, 
                                                                 name = "concordance", 
                                                                 cluster_rows = F, 
                                                                 row_gap = unit(0.5, "cm"),
                                                                 col = col_fun,
                                                                 rect_gp = gpar(col = "black", lwd = 1),
                                                                 row_split = tool_meta$tool, 
                                                                 row_title = NULL,
                                                                 column_split = col_meta$group_id, 
                                                                 column_gap = unit(0.5, "cm"),
                                                                 column_title_rot = -45,
                                                                 cluster_column_slices = F,
                                                                   cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(
      sprintf("%.2f", combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp_mat[i, j]),
      x,
      y
      #,gp = gpar(fontsize = 8, col = "black")
    )
  })

draw(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp_hp)

cairo_pdf("multilocus_freq_ccc_per_coi_cat_per_tool_params.pdf", width = 20, height = 8, bg = "transparent")
draw(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_sp_hp, background = "transparent")
dev.off()

```



### Best per population COI parameters and relatedness 

The amount of between sample relatedness also plays a significant role in performance 

```{r}
combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness = combined_mlfreq %>% 
  left_join(sim_params) %>% 
  mutate(coi_cat_related_ness = paste0(coi_cat, "\n", between_sample_related_cat)) %>% 
  compute_ccc(
    expected_freq, freq, 
    c(tool, paramset_id, coi_cat_related_ness, group_id)
  )

combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp = combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness %>% 
  ungroup() %>% 
  mutate(group_id_coi_cat_related_ness = paste0(group_id, "\n", coi_cat_related_ness)) %>% 
  select(paramset_id, group_id_coi_cat_related_ness, freq_ccc) %>% 
  spread(group_id_coi_cat_related_ness, freq_ccc)

combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat = as.matrix(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp[,2:ncol(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp)])
rownames(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat) = combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp$paramset_id

tool_meta = tibble(
  paramset_id = rownames(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat)
) %>% 
  mutate(tool = gsub("_coi_(freq|variant)", "", gsub("_p.*", "", paramset_id)))

col_meta = tibble(
  group_id_coi_cat = colnames(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat)
) %>% 
  mutate(group_id = gsub("\\n.*", "", group_id_coi_cat)) %>% 
  mutate(group_id = factor(group_id,
                           levels = c("two_group1",
                                      "pfdhfr",
                                      "pfdhps",
                                      "four_group1",
                                      "six_low_div",
                                      "pfdhfr_pfdhps",
                                      "six_med_div"))) %>% 
  mutate(new_col_name = gsub(".*\\nMean", "Mean", group_id_coi_cat))

colnames(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat) = col_meta$new_col_name

library(circlize)
col_fun = colorRamp2(get_color_scale_breaks(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat), c( "#3288bd", "#ffffbf", "#d53e4f"))

```

```{r}
#| fig-width: 24
#| fig-height: 10
combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_hp = Heatmap(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat, 
                                                                 name = "concordance", 
                                                                 cluster_rows = F, 
                                                                 cluster_columns = F,
                                                                 row_gap = unit(0.5, "cm"),
                                                                 col = col_fun,
                                                                 rect_gp = gpar(col = "black", lwd = 1),
                                                                 row_split = tool_meta$tool, 
                                                                 row_title = NULL,
                                                                 column_split = col_meta$group_id, 
                                                                 column_gap = unit(0.5, "cm"),
                                                                 column_title_rot = -45,
                                                                 cluster_column_slices = F, 
                                                                   cell_fun = function(j, i, x, y, width, height, fill) {
    grid.text(
      sprintf("%.2f", combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_mat[i, j]),
      x,
      y
      #,gp = gpar(fontsize = 8, col = "black")
    )
  })

draw(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_hp)

cairo_pdf("multilocus_freq_ccc_per_coi_cat_related_ness_per_tool_params.pdf", width = 24, height = 10, bg = "transparent")
draw(combined_mlfreq_ccc_per_tool_per_params_per_coi_cat_related_ness_sp_hp, background = "transparent")
dev.off()

```




### Overall 

```{r}
combined_mlfreq_ccc_by_tool_params = combined_mlfreq %>% 
  compute_ccc(expected_freq, freq, c(tool, paramset_id, tool_paramset_id))

combined_mlfreq_ccc_by_tool_params_best = combined_mlfreq_ccc_by_tool_params %>% 
  group_by(tool) %>% 
  slice_max(order_by = freq_ccc, with_ties = F)

```


Overall and broken down by group there is frequently is just small changes in performance based on parameters. 

```{r}
combined_mlfreq_ccc_by_tool_params_plot_per_paramset = ggplot(combined_mlfreq_ccc_by_tool_params) + 
  geom_col(aes(x = paramset_id, y = freq_ccc, fill = tool)) + 
  geom_errorbar(aes(x = paramset_id, ymin = freq_ccc_lower, ymax = freq_ccc_upper)) + 
  scale_fill_manual(values = colorPalette_15) + 
  custom_ggplot_theme_xRotate + 
  geom_abline(slope = 0, intercept = 0.95, color = "red", linetype = "dotted")

create_dt(combined_mlfreq_ccc_by_tool_params)

print(combined_mlfreq_ccc_by_tool_params_plot_per_paramset)

cairo_pdf("combined_mlfreq_ccc_by_tool_params_plot_per_paramset.pdf", width = 15, height = 8)
print(combined_mlfreq_ccc_by_tool_params_plot_per_paramset)
dev.off()


```




```{r}
combined_mlfreq_ccc_by_tool_params = compute_ccc(combined_mlfreq, expected_freq, freq, c(tool, paramset_id))

combined_mlfreq_ccc_by_tool_params_best = combined_mlfreq_ccc_by_tool_params %>%
  group_by(tool) %>%
  slice_max(n = 1, order_by = freq_ccc)  %>%
  mutate(tool_paramset_id = paste0(tool, "-", paramset_id))

combined_mlfreq_best = combined_mlfreq %>%
  filter(tool_paramset_id %in% combined_mlfreq_ccc_by_tool_params_best$tool_paramset_id) %>%
  mutate(group_id = factor(group_id,
                           levels = c("two_group1",
                                      "pfdhfr",
                                      "pfdhps",
                                      "four_group1",
                                      "six_low_div",
                                      "pfdhfr_pfdhps",
                                      "six_med_div")))
```

```{r}
#| fig-width: 15
#| fig-height: 8


combined_mlfreq_best = combined_mlfreq_best %>%
  add_diff_mean(expected_freq, freq)

combined_mlfreq_best_altman_tool_groupid = make_diff_vs_expected_plot(combined_mlfreq_best,
                                                                      expected_col = "expected_freq", 
                                                                      diff_col = "freq_diff_from_exp",
                                                                      "freq") +
  facet_grid(tool ~ group_id)

print(combined_mlfreq_best_altman_tool_groupid)


cairo_pdf("combined_mlfreq_best_altman_tool_groupid.pdf", width = 15, height = 8)
print(combined_mlfreq_best_altman_tool_groupid)
dev.off()
```


```{r}
#| fig-width: 8
#| fig-height: 6


combined_mlfreq_best_mod = combined_mlfreq_best %>%
  mutate(TP = as.numeric(freq != 0 & expected_freq != 0))%>%
  mutate(FP = as.numeric(freq != 0 & expected_freq == 0))%>%
  mutate(FN = as.numeric(freq == 0 & expected_freq != 0))

combined_mlfreq_best_mod_contig = combined_mlfreq_best_mod %>%
  group_by(tool, group_id) %>%
  add_contigency_metrics_no_TN(T)


combined_mlfreq_best_mod_contig_recall_precision = combined_mlfreq_best_mod_contig %>%
  select(tool, group_id, recall, precision) %>%
  gather(measure, value, 3:4)

combined_mlfreq_best_mod_contig_recall_precision_plot = ggplot(combined_mlfreq_best_mod_contig_recall_precision) +
  geom_col(aes(x = group_id, y = value, fill = tool), color = "black", position = "dodge") +
  custom_ggplot_theme_xRotate +
  scale_fill_manual(values = colorPalette_08) +
  facet_wrap(~measure)

print(combined_mlfreq_best_mod_contig_recall_precision_plot)

cairo_pdf("combined_mlfreq_best_mod_contig_recall_precision.pdf", width = 8, height = 6)
print(combined_mlfreq_best_mod_contig_recall_precision_plot)
dev.off()

```

Models FEM, mlbm, snpslice have a better recall for the higher diversity group of six but this comes at the cost of precision 

```{r}
#| fig-width: 8
#| fig-height: 6


combined_mlfreq_best_mod_contig_by_coi_cat = combined_mlfreq_best_mod %>%
  left_join(sim_params, by = "population_name") %>%
  group_by(tool, group_id, coi_cat) %>%
  add_contigency_metrics_no_TN(T)

combined_mlfreq_best_mod_contig_by_coi_cat_recall_precision = combined_mlfreq_best_mod_contig_by_coi_cat %>%
  select(tool, group_id, coi_cat, recall, precision) %>%
  gather(measure, value,4:5)


combined_mlfreq_best_mod_contig_by_coi_cat_recall_precision_plot = ggplot(combined_mlfreq_best_mod_contig_by_coi_cat_recall_precision) +
  geom_col(aes(x = group_id, y = value, fill = tool), color = "black", position = "dodge") +
  custom_ggplot_theme_xRotate +
  scale_fill_manual(values = colorPalette_08) +
  facet_wrap(coi_cat~measure)


print(combined_mlfreq_best_mod_contig_by_coi_cat_recall_precision_plot)

cairo_pdf("combined_mlfreq_best_mod_contig_coi_cat_recall_precision.pdf", width = 8, height = 6)
print(combined_mlfreq_best_mod_contig_by_coi_cat_recall_precision_plot)
dev.off()


```


The false positives are at low frequencies for the models but the majority of haplotypes are at the same frequencies so a cut off would not solve the false positive problem. In addition, multi-locus that may not exist ever (like biological incompatiable combinations) may be called by these models.  

```{r}
#| fig-width: 10
#| fig-height: 8


combined_mlfreq_best_mod_FP = combined_mlfreq_best_mod %>%
  filter(FP == 1)

combined_mlfreq_best_mod_FP_plot = ggplot(combined_mlfreq_best_mod_FP) +
  geom_histogram(aes(x = freq), binwidth = 0.02, boundary = 0) +
  facet_grid(group_id ~ tool, scales = "free") +
  custom_ggplot_theme + 
  xlim(0,1)

cairo_pdf("combined_mlfreq_best_mod_false_positives.pdf", width = 10, height = 8)
print(combined_mlfreq_best_mod_FP_plot)
dev.off()

combined_mlfreq_best_mod_TP = combined_mlfreq_best_mod %>%
  filter(TP == 1)

combined_mlfreq_best_mod_TP_plot = ggplot(combined_mlfreq_best_mod_TP) +
  geom_histogram(aes(x = freq), binwidth = 0.02, boundary = 0) +
  facet_grid(group_id ~ tool, scales = "free") +
  custom_ggplot_theme + 
  xlim(0,1)

cairo_pdf("combined_mlfreq_best_mod_true_positives.pdf", width = 10, height = 8)
print(combined_mlfreq_best_mod_TP_plot)
dev.off()

```

 

```{r}
#| results: asis
#| echo: false
#| fig-width: 10
#| fig-height: 8

cat(create_tabsetOfGgplotObjects(
  list("true_positive" = combined_mlfreq_best_mod_TP_plot,
       "false_positive" = combined_mlfreq_best_mod_FP_plot)
))

```



