---
title: Frequency tools Benchmarking Results
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)
```


## Set up 

### Libraries used in the analysis below 

```{r, echo=FALSE}
source("../tutorials/scripts/benchmarking_functions.R")
```

```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-line-numbers: true
#| filename: ../scripts/benchmarking_functions.R
#| file: ../scripts/benchmarking_functions.R

```

## Load in the Benchmarking Outputs

```{r}
inp <- list(
  sim_info_dir = "../data/benchmarking//simulations/SpotMalaria/", 
  madhatter_res_dir = "../data/benchmarking/tool_results/madhatter/"
)

inp_per_tools <- list(
  "therealmccoil" = list(
      tool_res_dir = "../data/benchmarking/tool_results/THEREALMcCOIL/", 
      tool_params = "../data/benchmarking/tool_results/therealmccoil.json"
  ), 
  "moire" = list(
      tool_res_dir = "../data/benchmarking/tool_results/MOIRE/", 
      tool_params = "../data/benchmarking/tool_results/moire.json"
  ), 
  "naive_coi" = list(
      tool_res_dir = "../data/benchmarking/tool_results/naive_coi/", 
      tool_params = "../data/benchmarking/tool_results/naive_coi.json"
  ),
  "coiaf" = list(
      tool_res_dir = "../data/benchmarking/tool_results/coiaf//", 
      tool_params = "../data/benchmarking/tool_results/coiaf.json"
  ) 
)
```



First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.

### Reading in expected truths 

Read per in expected allele frequencies per sample tables to get true COIs

```{r}
expected_simulated_frequencies_per_sample <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "trimmed_expected_simulated_frequencies_per_sample.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(library_sample_name = libary_sample_name)

expected_simulated_frequencies_per_sample_COIs = expected_simulated_frequencies_per_sample %>% 
  select(panel, population_name, country, library_sample_name, COI) %>% 
  distinct() %>% 
  dplyr::rename(expected_COI = COI) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Read per in expected allele frequencies tables population level 
```{r}
population_trimmed_mhap_freqs_prevs <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_trimmed_mhap_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(),
        population_name = col_character(), 
        pop = col_character(), 
        country = col_character(), 
        target = col_character(), 
        seq = col_character()
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(target_id = target)
```


Read in all simulation parameters into one tibble. Encode different population parameters as descriptive factors. 

```{r}

sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>% 
  mutate(
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>% 
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  )) %>% 
  mutate(pop_cat = factor(pop_cat, levels = sort(unique((.$pop_cat)))), 
         coi_cat = factor(coi_cat, levels = sort(unique((.$coi_cat)))), 
         between_sample_related_cat = factor(between_sample_related_cat, levels = sort(unique((.$between_sample_related_cat)))), 
         nonrecombinant_cat = factor(nonrecombinant_cat, levels = sort(unique((.$nonrecombinant_cat)))), 
         meiosis_cat = factor(meiosis_cat, levels = sort(unique((.$meiosis_cat)))))

create_dt(sim_params)
```

Read in madhatter pipeline allele tables 

```{r}
madhatter_allele_data <- tibble(
    allele_data_path = list.files(
      inp$madhatter_res_dir, 
      pattern = "allele_data.txt.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    allele_data = map(
      allele_data_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  mutate(tool_res_dir = dirname(allele_data_path)) %>%
  mutate(population_name = gsub("_mh_results", "", basename(tool_res_dir))) %>%
  select(-allele_data_path, -tool_res_dir) %>%
  unnest(allele_data) %>% 
  mutate(SampleID = gsub("_S[0-9]+$", "", SampleID)) %>% 
  dplyr::rename(library_sample_name = SampleID, 
                target_id = Locus, 
                seq = ASV, 
                reads = Reads) 
```

