---
title: "MALECOT"
output: html_document
author: "Bob Verity"
date: "11-Dec-2023"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose

The MALECOT R package is intended for the inference of population structure in malaria data. It operates similar to STRUCTURE by dividing the dataset in K subpopulations. Unlike STRUCTURE, it considers that each sample may have multiple genotypes due to complexity of infection (COI).

The tools uses Bayesian Markov Monte Carlo to estimate:

1. Allele frequencies in each subpopulation.
2. Group allocation (i.e., it allocates each sample to a group).
3. Complexity of Infection - COI (The number of genotypes represented in each sample).

## Existing resources

The [MALECOT website](https://bobverity.github.io/MALECOT/index.html) provides installation intructions, vignettes with different example workflows, recommendations to increase statistical power and performance, and documentation on the functions included in the package.

## Citation

Please use the following citation:

```
@Manual{,
    title = {MALECOT: Joint estimation of COI and population structure for malaria
genetic data},
    author = {Bob Verity},
    year = {2023},
    note = {R package version 0.1.1},
  }
```
