---
title: "MLMOI"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose

The purpose of MLMOI is to estimate complexity of infection with various
corrections for unobservable states. The program also generates simulated
datasets. The approach seems to involve simulating datasets and checking to see
if the observed data correlates well with the simulated datasets derived by a
given model. We were unable to install the package as it has been removed from
the R-CRAN library, and had difficulty following the vignettes in the archived
package. More specifically, We had difficulty interpreting unitless numbers
associated with a given locus for a given sample, and understanding why some
missing samples (no sample name) had filled in values and other named samples
had missing values). Manual installation (with no package available) was also
difficult to follow.

## Existing resources

- Any existing online tutorials?
- Any important papers?

## Citation

BibTeX style citation. For an R package, you can get this using `citation(package = "name")`:

Here is an example for DRpower, using `citation(package = "DRpower")`:

```
@Manual{,
    title = {DRpower: Study design and analysis for pfhrp2/3 deletion prevalence studies},
    author = {Bob Verity and Shazia Ruybal},
    note = {R package version 1.0.2},
  }
```