---
title: "hmmibdr"
output: html_document
author: "Sophie Berube"
date: "13-12-2023"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose
This package is an Rcpp wrapper for the hmmibd software. Currently, there are several c and python scripts (within the hmmibd software) required to convert files from VCF format to a text format that can then be used by either hmmibd or hmmibdr to perform further analysis. However, these scripts require the user to run c and python code; these are the same requirements as those for performing analysis using only the hmmibd software . Therefore, we recommend the use of the hmmibd software directly for the entire analysis. 

Please refer to the hmmibd tutorial for a complete demonstration of the tool. 

## Existing resources

- A short tutorial for hmmibdr is located on the [Github Page](https://github.com/OJWatson/hmmibdr/tree/main)
- See the hmmibd for further information.

## Citation

```
@Manual{,
    title = {hmmibdr: HMM Identity by Descent},
    author = {OJ Watson},
    year = {2023},
    note = {R package version 0.2.0},
  }

```