---
title: "Analysis tutorial"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

First we will load the libraries we will need:

```{r, warning=FALSE, message=FALSE, error=FALSE}
library(hierfstat)
library(vcfR)
```

We're going to be using WGS data from Vietnam which was generated for the 
[Pf3k Project](https://www.malariagen.net/parasite/pf3k). First load in the metadata containing 
the sample names and information on where the samples were collected.

```{r, warning=FALSE, message=FALSE, error=FALSE}
meta<-read.csv("../../data/wgs/pf3k/Vietnam/pf3k.metadata.Vietnam.csv")
```

Then we can load in the genetic data from VCF format.

```{r, warning=FALSE, message=FALSE, error=FALSE, results = 'hide'}
vcf<-read.vcfR("../../data/wgs/pf3k/Vietnam/SNP_INDEL_Pf3D7_ALL_v3.combined.filtered.vqslod6.biallelic_snp.Vietnam.vcf.gz")
```

Next we extract the genotypes and convert them into a dosage format.

```{r, warning=FALSE, message=FALSE, error=FALSE, results = 'hide'}
vcf_gts<-extract.gt(vcf)

recode_gt<-function(x){
  x<-gsub("[|]","/",x)
  x<-gsub("0/0",0,x)
  x<-gsub("0/1",1,x)
  x<-gsub("1/1",2,x)
  as.numeric(x)
}

raw_gts<-apply(vcf_gts,MARGIN = 2,function(x){recode_gt(x)})

```


Now we can remove loci which are monomorphic.
```{r, warning=FALSE, message=FALSE, error=FALSE}
calculate_af<-function(x){
  sum(x,na.rm = T)/length(x)*2
}
af <- apply(raw_gts,MARGIN = 1,calculate_af)
transposed_gts<-as.data.frame(t(raw_gts[which(af>0 & af<1),]))

```

We now have a matrix where the samples are rows and the columns are loci. Hierfstat requires you to also
provide population assignments for each sample by adding it to the data as the first column.

```{r, warning=FALSE, message=FALSE, error=FALSE}
dat<-cbind(meta$site,transposed_gts)

```

Now we are ready to calculate some statistics!

```{r, warning=FALSE, message=FALSE, error=FALSE}
results<-basic.stats(dat)
results$overall
```

The `results$overall` table contains basic statistics averaged over loci. The 
statistics presented are defined in eq.7.38– 7.43 pp.164–5 of Nei (1987).


## Summary

We loaded in data from VCF format, converted this to dosage and finally calculated basic statistics.