---
title: THEREALMcCOI Benchmarking Results
author: Nicholas Hathaway
format: html
toc-depth: 5
code-fold: true
fig-width: 15
fig-height: 15
embed-resources: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, comment = "", cache = F, fig.align = "center")
options(width = 200)
options(readr.show_col_types = FALSE)
```


## Set up 

### Libraries used in the analysis below 

```{r, echo=FALSE}
source("../scripts/benchmarking_functions.R")
```

```{r}
#' @title Concordance / Error Metrics (single or multiple observed columns)
#' @keywords CCC RMSE MAE COI
#'
#' @description
#' Compute metrics between an expected column and one or more observed columns.
#' Each function accepts an optional set of grouping columns. If omitted, an
#' overall (single-row) summary is returned.
#'
#' @param df A data frame.
#' @param expected_col Unquoted column name of expected values.
#' @param observed_cols One symbol or \code{c(sym1, sym2, ...)} of observed columns.
#' @param group_cols Optional grouping columns supplied like
#'   \code{c(population_name, paramset_id)}. If omitted or \code{NULL},
#'   computes overall metrics (no grouping).
#'
#' @return A tibble with group columns (if any) and metric columns per observed variable.
#'
#' @importFrom dplyr group_by summarise distinct left_join bind_cols
#' @importFrom rlang enquo enexpr is_call call_args as_quosure caller_env as_name
#' @importFrom purrr map map_dbl
#' @name vectorized_metrics


# ---- CCC (DescTools::CCC) ----------------------------------------------------

#' @rdname vectorized_metrics
#' @export
compute_ccc <- function(df, expected_col, observed_cols, group_cols = NULL) {
  stopifnot(is.data.frame(df))
  expected_col <- rlang::enquo(expected_col)

  # capture observed columns (symbol or c(...))
  obs_expr <- rlang::enexpr(observed_cols)
  obs_quos <- if (rlang::is_call(obs_expr, "c")) {
    purrr::map(rlang::call_args(obs_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(obs_expr, env = rlang::caller_env()))
  }

  # capture grouping columns (NULL/omitted => no groups)
  grp_expr <- rlang::enexpr(group_cols)
  grp_quos <- if (missing(group_cols) || is.null(grp_expr)) {
    list()
  } else if (rlang::is_call(grp_expr, "c")) {
    purrr::map(rlang::call_args(grp_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(grp_expr, env = rlang::caller_env()))
  }

  if (length(grp_quos) > 0) {
    out <- df %>% dplyr::distinct(!!!grp_quos)
    join_by_cols <- names(out)
    for (q in obs_quos) {
      obs_name <- rlang::as_name(q)
      one <- df %>%
        dplyr::group_by(!!!grp_quos) %>%
        dplyr::summarise(.ccc = list(DescTools::CCC(!!expected_col, !!q)), .groups = "drop") %>%
        dplyr::mutate(
          !!paste0(obs_name, "_ccc")       := purrr::map_dbl(.ccc, ~ .x$rho.c$est),
          !!paste0(obs_name, "_ccc_lower") := purrr::map_dbl(.ccc, ~ .x$rho.c$lwr.ci),
          !!paste0(obs_name, "_ccc_upper") := purrr::map_dbl(.ccc, ~ .x$rho.c$upr.ci)
        ) %>%
        dplyr::select(-.ccc)
      out <- dplyr::left_join(out, one, by = join_by_cols)
    }
  } else {
    pieces <- purrr::map(obs_quos, function(q) {
      obs_name <- rlang::as_name(q)
      tmp <- df %>% dplyr::summarise(.ccc = list(DescTools::CCC(!!expected_col, !!q)))
      tibble::tibble(
        !!paste0(obs_name, "_ccc")       := purrr::map_dbl(tmp$.ccc, ~ .x$rho.c$est),
        !!paste0(obs_name, "_ccc_lower") := purrr::map_dbl(tmp$.ccc, ~ .x$rho.c$lwr.ci),
        !!paste0(obs_name, "_ccc_upper") := purrr::map_dbl(tmp$.ccc, ~ .x$rho.c$upr.ci)
      )
    })
    out <- dplyr::bind_cols(pieces)
  }

  out
}

# ---- RMSE --------------------------------------------------------------------

#' @rdname vectorized_metrics
#' @export
compute_rmse <- function(df, expected_col, observed_cols, group_cols = NULL) {
  stopifnot(is.data.frame(df))
  expected_col <- rlang::enquo(expected_col)

  obs_expr <- rlang::enexpr(observed_cols)
  obs_quos <- if (rlang::is_call(obs_expr, "c")) {
    purrr::map(rlang::call_args(obs_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(obs_expr, env = rlang::caller_env()))
  }

  grp_expr <- rlang::enexpr(group_cols)
  grp_quos <- if (missing(group_cols) || is.null(grp_expr)) {
    list()
  } else if (rlang::is_call(grp_expr, "c")) {
    purrr::map(rlang::call_args(grp_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(grp_expr, env = rlang::caller_env()))
  }

  metric_fun <- function(q) {
    obs_name <- rlang::as_name(q)
    if (length(grp_quos) > 0) {
      df %>%
        dplyr::group_by(!!!grp_quos) %>%
        dplyr::summarise(
          !!paste0(obs_name, "_rmse") := {
            x <- !!q; y <- !!expected_col
            ok <- !is.na(x) & !is.na(y)
            n  <- sum(ok)
            if (n == 0) NA_real_ else sqrt(sum((x[ok] - y[ok])^2) / n)
          },
          .groups = "drop"
        )
    } else {
      dplyr::summarise(df,
        !!paste0(obs_name, "_rmse") := {
          x <- !!q; y <- !!expected_col
          ok <- !is.na(x) & !is.na(y)
          n  <- sum(ok)
          if (n == 0) NA_real_ else sqrt(sum((x[ok] - y[ok])^2) / n)
        }
      )
    }
  }

  if (length(grp_quos) > 0) {
    out <- df %>% dplyr::distinct(!!!grp_quos)
    join_by_cols <- names(out)
    for (q in obs_quos) {
      out <- dplyr::left_join(out, metric_fun(q), by = join_by_cols)
    }
  } else {
    out <- dplyr::bind_cols(purrr::map(obs_quos, metric_fun))
  }

  out
}

# ---- MAE ---------------------------------------------------------------------

#' @rdname vectorized_metrics
#' @export
compute_mae <- function(df, expected_col, observed_cols, group_cols = NULL) {
  stopifnot(is.data.frame(df))
  expected_col <- rlang::enquo(expected_col)

  obs_expr <- rlang::enexpr(observed_cols)
  obs_quos <- if (rlang::is_call(obs_expr, "c")) {
    purrr::map(rlang::call_args(obs_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(obs_expr, env = rlang::caller_env()))
  }

  grp_expr <- rlang::enexpr(group_cols)
  grp_quos <- if (missing(group_cols) || is.null(grp_expr)) {
    list()
  } else if (rlang::is_call(grp_expr, "c")) {
    purrr::map(rlang::call_args(grp_expr), rlang::as_quosure, env = rlang::caller_env())
  } else {
    list(rlang::as_quosure(grp_expr, env = rlang::caller_env()))
  }

  metric_fun <- function(q) {
    obs_name <- rlang::as_name(q)
    if (length(grp_quos) > 0) {
      df %>%
        dplyr::group_by(!!!grp_quos) %>%
        dplyr::summarise(
          !!paste0(obs_name, "_mean_abs_error") := {
            x <- !!q; y <- !!expected_col
            ok <- !is.na(x) & !is.na(y)
            if (!any(ok)) NA_real_ else mean(abs(x[ok] - y[ok]))
          },
          .groups = "drop"
        )
    } else {
      dplyr::summarise(df,
        !!paste0(obs_name, "_mean_abs_error") := {
          x <- !!q; y <- !!expected_col
          ok <- !is.na(x) & !is.na(y)
          if (!any(ok)) NA_real_ else mean(abs(x[ok] - y[ok]))
        }
      )
    }
  }

  if (length(grp_quos) > 0) {
    out <- df %>% dplyr::distinct(!!!grp_quos)
    join_by_cols <- names(out)
    for (q in obs_quos) {
      out <- dplyr::left_join(out, metric_fun(q), by = join_by_cols)
    }
  } else {
    out <- dplyr::bind_cols(purrr::map(obs_quos, metric_fun))
  }

  out
}

```



```{r}
#| echo: true
#| eval: false
#| code-fold: true
#| code-line-numbers: true
#| filename: ../scripts/benchmarking_functions.R
#| file: ../scripts/benchmarking_functions.R

```

## Load in the Benchmarking Outputs

```{r}
inp <- list(
  sim_info_dir = "../../data/benchmarking//simulations/SpotMalaria", 
  tool_res_dir = "../../data/benchmarking/tool_results/THEREALMcCOIL/", 
  madhatter_res_dir = "../../data/benchmarking/tool_results/madhatter/",
  tool_params = "../../data/benchmarking/tool_results/therealmccoil.json"
)
```



First, all of the tool results, tool parameters, expected simulated values, and simulation parameters are read into a tibble.

### Reading in expected truths 

Read per in expected allele frequencies per sample tables to get true COIs

```{r}
expected_simulated_frequencies_per_sample <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "trimmed_expected_simulated_frequencies_per_sample.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(library_sample_name = libary_sample_name)

expected_simulated_frequencies_per_sample_COIs = expected_simulated_frequencies_per_sample %>% 
  select(panel, population_name, country, library_sample_name, COI) %>% 
  distinct() %>% 
  dplyr::rename(expected_COI = COI) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

Read per in expected allele frequencies tables population level 
```{r}
population_trimmed_mhap_freqs_prevs <- tibble(
    sim_truth_path = list.files(
      inp$sim_info_dir, 
      pattern = "population_trimmed_mhap_freqs_prevs.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_truth = map(
      sim_truth_path, 
      read_tsv, 
      col_types = cols(
        .default = col_double(),
        population_name = col_character(), 
        pop = col_character(), 
        country = col_character(), 
        target = col_character(), 
        seq = col_character()
      ),
      progress = FALSE
    )
  ) %>%
  select(-sim_truth_path) %>%
  unnest(sim_truth) %>% 
  dplyr::rename(target_id = target)
```


Read in all simulation parameters into one tibble. Encode different population parameters as descriptive factors. 

```{r}

sim_params <- tibble(
    sim_params_path = list.files(
      inp$sim_info_dir, 
      pattern = "simulation_parameters.json", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    sim_params = map(
      sim_params_path, 
      jsonlite::fromJSON
    )
  ) %>%
  select(-sim_params_path) %>%
  mutate(sim_params = map(sim_params, as_tibble)) %>%
  unnest(sim_params) %>% 
  mutate(
    coi_cat = case_when(
      coi_r == 0.25 & coi_p == 0.7 ~"MeanCOI=1.26",
      coi_r == 0.5  & coi_p == 0.2 ~"MeanCOI=3.41"
    ), 
    between_sample_related_cat = case_when(
      pop_alpha == 99 ~ "_1%_related", 
      pop_alpha == 9 ~ "10%_related", 
      pop_alpha == 2.33 ~ "30%_related"
    ), 
    nonrecombinant_cat = case_when(
      k_s == 0.5 ~ "50%_non-recombinant", 
      k_s == 0.2 ~ "20%_non-recombinant", 
    ), 
    meiosis_cat = case_when(
      k_s == 0.5 ~ "mean_meiosis=1.0", 
      k_s == 0.2 ~ "mean_meiosis=3.5", 
    )
  ) %>% 
  mutate(pop_cat = paste0(
    between_sample_related_cat, "__", 
    coi_cat, "__", 
    nonrecombinant_cat
  )) %>% 
  mutate(pop_cat = factor(pop_cat, levels = sort(unique((.$pop_cat)))), 
         coi_cat = factor(coi_cat, levels = sort(unique((.$coi_cat)))), 
         between_sample_related_cat = factor(between_sample_related_cat, levels = sort(unique((.$between_sample_related_cat)))), 
         nonrecombinant_cat = factor(nonrecombinant_cat, levels = sort(unique((.$nonrecombinant_cat)))), 
         meiosis_cat = factor(meiosis_cat, levels = sort(unique((.$meiosis_cat)))))

create_dt(sim_params)
```

Read in madhatter pipeline allele tables 

```{r}
madhatter_allele_data <- tibble(
    allele_data_path = list.files(
      inp$madhatter_res_dir, 
      pattern = "allele_data.txt.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(
    allele_data = map(
      allele_data_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  mutate(tool_res_dir = dirname(allele_data_path)) %>%
  mutate(population_name = gsub("_mh_results", "", basename(tool_res_dir))) %>%
  select(-allele_data_path, -tool_res_dir) %>%
  unnest(allele_data) %>% 
  mutate(SampleID = gsub("_S[0-9]+$", "", SampleID)) %>% 
  dplyr::rename(library_sample_name = SampleID, 
                target_id = Locus, 
                seq = ASV, 
                reads = Reads) 
```



### Read in THEREALMcCOI results

THEREALMcCOI outputs information on complexity of infection (COI) and on population allele frequencies and several different outputs need to be read in. 

Getting the allele frequencies data in the `slaf.tsv.gz` files 

```{r}
therealmccoil_single_variant_freq_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "slaf.tsv.gz", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) 
```

Getting the complexity of infection (COI) from the `coi.tsv.gz` files 

```{r}
therealmccoil_coi_summary <- tibble(
    tool_res_path = list.files(
      inp$tool_res_dir, 
      pattern = "^coi.tsv.gz$", 
      full.names = TRUE, 
      recursive = TRUE
    )
  ) %>%
  mutate(tool_res_dir = dirname(tool_res_path)) %>%
  mutate(population_name = basename(dirname(tool_res_dir))) %>%
  mutate(paramset_id = basename(tool_res_dir)) %>%
  select(-tool_res_dir) %>%
  mutate(
    tool_res = map(
      tool_res_path, 
      read_tsv, 
      progress = FALSE
    )
  ) %>%
  select(-tool_res_path) %>%
  unnest(tool_res) %>% 
  mutate(specimen_id = gsub("_S[0-9]+$", "", specimen_id)) %>% 
  dplyr::rename(library_sample_name = specimen_id) %>% 
  mutate(population_name_library_sample_name = paste0(population_name, "-", library_sample_name))

```

THEREALMcCOI can run in two different models, categorical vs proportional

```{r}
# Read in tool parameters
tool_params <- jsonlite::fromJSON(inp$tool_params, flatten = TRUE) %>%
  as_tibble() %>%
  rename(
    paramset_model = paramset.model,
  ) 

create_dt(tool_params)
```



## Benchmarking 

### COI  

#### Metrics 

*  Plotting Bland-Altman plots (plotting the mean of expected and obs vs difference) 
*  Calculating the RMSE (taking the square root of the mean squared difference between expected and observed)  
*  Calculating [Lin's concordance correlation coefficient](https://www.rdocumentation.org/packages/DescTools/versions/0.99.60/topics/CCC) for agreement on a continuous measure 
*  Calculating contingency table of determining a sample as monoclonal vs polyclonal 

#### Results 


```{r}
# combine the COI summaries with the expected outputs for COI
therealmccoil_all_coi_summary = therealmccoil_coi_summary %>% 
  left_join(expected_simulated_frequencies_per_sample_COIs %>% 
              mutate(library_sample_name = as.character(library_sample_name))) %>% 
                                                        left_join(sim_params) %>% 
  left_join(tool_params)

# calculate the means and differences for Bland-Altman plots

therealmccoil_all_coi_summary = therealmccoil_all_coi_summary %>% 
  add_diff_mean(expected_COI, coi)

```


##### Bland-Altman COI plots 

Over both models 

```{r}
therealmccoil_all_coi_summary_overall_coi_bland_altman_plot = make_bland_altman_plot_by_label(therealmccoil_all_coi_summary, "coi", expected_label = "Expected COI")

print(therealmccoil_all_coi_summary_overall_coi_bland_altman_plot)
```

Showing the different parameter models 

```{r}
therealmccoil_all_coi_summary_coi_bland_altman_plot = make_bland_altman_plot_by_label(therealmccoil_all_coi_summary, "coi") + 
  facet_wrap(~paramset_model, ncol = 1)

print(therealmccoil_all_coi_summary_coi_bland_altman_plot)
```


```{r}
therealmccoil_all_coi_summary_coi_by_pop_bland_altman_plot = make_bland_altman_plot_by_label(therealmccoil_all_coi_summary, "coi") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)

print(therealmccoil_all_coi_summary_coi_by_pop_bland_altman_plot)
```


```{r}
therealmccoil_all_coi_summary_coi_by_pop_bland_altman_plot_categorical = make_bland_altman_plot_by_label(therealmccoil_all_coi_summary %>% 
                                                                                                            filter(paramset_model == "categorical"), "coi") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)

therealmccoil_all_coi_summary_coi_by_pop_bland_altman_plot_proportional = make_bland_altman_plot_by_label(therealmccoil_all_coi_summary %>% 
                                                                                                            filter(paramset_model == "proportional"), "coi") + 
  facet_grid(interaction(coi_cat, nonrecombinant_cat, sep = "\n")~between_sample_related_cat)


```


```{r}
#| results: asis
#| echo: false

cat(create_tabsetOfGgplotObjects(
  list("categorical" = therealmccoil_all_coi_summary_coi_by_pop_bland_altman_plot_categorical,
       "proportional" = therealmccoil_all_coi_summary_coi_by_pop_bland_altman_plot_proportional)
))
```




##### Concordance 

Plotting concordance between all COI metrics reported by therealmccoil.

Overall 

```{r}
therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_all = ggpairs(therealmccoil_all_coi_summary,
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "all")

```

```{r}
print(therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_all)
```


Plotting difference parameter sets  

```{r}
therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_categorical = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(paramset_model == "categorical"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "categorical")

therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_proportional = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(paramset_model == "proportional"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "proportional")

```


```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("categorical" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_categorical,
                                      "proportional" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_proportional
                                      )))
```

Different population metrics 


COI categories 

```{r}
therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI1.26 = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(coi_cat == "MeanCOI=1.26"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=1.26")

therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI3.41 = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(coi_cat == "MeanCOI=3.41"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=3.41")

therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI1.26_proportional = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(coi_cat == "MeanCOI=1.26",
                                                           paramset_model == "proportional"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=1.26 proportional model")

therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI3.41_proportional = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(coi_cat == "MeanCOI=3.41",
                                                           paramset_model == "proportional"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=3.41 proportional model")


therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI1.26_categorical = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(coi_cat == "MeanCOI=1.26",
                                                           paramset_model == "categorical"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=1.26 categorical model")

therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI3.41_categorical = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(coi_cat == "MeanCOI=3.41",
                                                           paramset_model == "categorical"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "MeanCOI=3.41  categorical model")

```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("MeanCOI_1.26" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI1.26,
                                      "MeanCOI_3.41" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI3.41, 
                                      "MeanCOI_1.26_categorical" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI1.26_categorical,
                                      "MeanCOI_3.41_categorical" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI3.41_categorical, 
                                      "MeanCOI_1.26_proportional" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI1.26_proportional,
                                      "MeanCOI_3.41_proportional" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_MeanCOI3.41_proportional
                                      )))
```


between sample related categories 

```{r}
therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_1perc_related = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(between_sample_related_cat == "_1%_related"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "_1%_related")

therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_10perc_related = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(between_sample_related_cat == "10%_related"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "10%_related")

therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_30perc_related = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(between_sample_related_cat == "30%_related"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "30%_related")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("_1perc_between_sample_related" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_1perc_related,
                                      "10perc_between_sample_related" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_10perc_related, 
                                      "30perc_between_sample_related" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_30perc_related
                                      )))
```

Recombination categories 
```{r}
therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_50perc_non_recomb = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(nonrecombinant_cat == "50%_non-recombinant"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "50%_non-recombinant")

therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_20perc_non_recomb = ggpairs(therealmccoil_all_coi_summary %>%
                                                    filter(nonrecombinant_cat == "20%_non-recombinant"),
                                                  columns = c("expected_COI", "coi"),
                                                  lower = list(continuous = wrap(my_hex)),
                                                  diag  = list(continuous = wrap("barDiag", binwidth = 1)),
                                                  upper = list(continuous = wrap(my_corr_ccc, digits = 3, size = 3.8)),
                                                  title = "20%_non-recombinant")
```

```{r}
#| results: asis
#| echo: false
#| fig-column: screen-inset

cat(create_tabsetOfGgplotObjects(list("20perc_non_recomb" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_20perc_non_recomb,
                                      "50perc_non_recomb" = therealmccoil_all_coi_summary_coi_measures_plot_therealmccoil_50perc_non_recomb
                                      )))
```



```{r}
# calculate Concordance Correlation Coefficient (CCC) 
therealmccoil_all_coi_summary_ccc_by_pop <- therealmccoil_all_coi_summary %>%
  compute_ccc(
    expected_col  = expected_COI,
    observed_cols = c(coi),
    group_cols    = c(population_name, paramset_id)
  )

# Gather into long format, left join in the simulation parameters
therealmccoil_all_coi_summary_ccc_by_pop_gat = therealmccoil_all_coi_summary_ccc_by_pop %>%
  select(population_name, paramset_id, ends_with("_ccc")) %>% 
  gather(coi_estimate, ccc, 3:ncol(.)) %>%
  mutate(coi_estimate = gsub("_ccc", "", coi_estimate))%>%
  left_join(sim_params) %>% 
  left_join(tool_params)
```

CCC for all runs 

```{r}
therealmccoil_all_coi_summary_ccc_plot_just_estimators = ggplot(therealmccoil_all_coi_summary_ccc_by_pop_gat %>%
                                                            mutate(paramset_model =
                                                                     factor(paramset_model,
                                                                            level = c("categorical",
                                                                                      "proportional"
                                                                                      )))) +
  geom_boxplot(
    aes(y = ccc, x = coi_estimate, fill = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

ggplotly(therealmccoil_all_coi_summary_ccc_plot_just_estimators) %>% layout(boxmode = "group")
```

ccc over all comparing parameter sets

```{r}
therealmccoil_all_coi_summary_ccc_plot_fill_paramset_id = ggplot(therealmccoil_all_coi_summary_ccc_by_pop_gat %>%
                                                            mutate(paramset_model =
                                                                     factor(paramset_model,
                                                                            level = c("categorical",
                                                                                      "proportional"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_model, y = ccc, x = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")
```

```{r}
ggplotly(therealmccoil_all_coi_summary_ccc_plot_fill_paramset_id) %>% layout(boxmode = "group")
```

ccc on the different simulation population characteristics 

```{r}
therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_by_country = ggplot(therealmccoil_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country = ggplot(therealmccoil_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country = ggplot(therealmccoil_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat_by_country = ggplot(therealmccoil_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc,
        fill = country)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08) + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot = ggplot(therealmccoil_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = ccc)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat = ggplot(therealmccoil_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = ccc)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat = ggplot(therealmccoil_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = ccc)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat = ggplot(therealmccoil_all_coi_summary_ccc_by_pop_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = ccc)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  geom_hline(yintercept = 0.95, color = "red", linetype = "dotted")

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot,
  "est_between_sample_related" = therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat,
  "coi_cat" = therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat,
  "nonrecombinant" = therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = therealmccoil_all_coi_summary_ccc_by_pop_gat_large_plot_nonrecombinant_cat_by_country
)))
```

##### RMSE   

```{r}
# calculate RMSE 
therealmccoil_all_coi_summary_rmse <- therealmccoil_all_coi_summary %>%
  compute_rmse(
    expected_col  = expected_COI,
    observed_cols = c(coi),
    group_cols    = c(population_name, paramset_model)
  )

# Gather into long format, left join in the simulation parameters
therealmccoil_all_coi_summary_rmse_gat = therealmccoil_all_coi_summary_rmse %>%
  gather(coi_estimate, rmse, 3:ncol(.)) %>%
  mutate(coi_estimate = gsub("_rmse", "", coi_estimate))%>%
  left_join(sim_params) %>% 
  left_join(tool_params)

```

RMSE across all parameter sets for all 4 estimators 

```{r}
therealmccoil_all_coi_summary_rmse_plot_just_estimators = ggplot(therealmccoil_all_coi_summary_rmse_gat %>%
                                                            mutate(paramset_model =
                                                                     factor(paramset_model,
                                                                            level = c("categorical",
                                                                                      "proportional"
                                                                                      )))) +
  geom_boxplot(
    aes(y = rmse, x = coi_estimate, fill = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

ggplotly(therealmccoil_all_coi_summary_rmse_plot_just_estimators)
```

RMSE over all comparing parameter sets for each estimate

```{r}
therealmccoil_all_coi_summary_rmse_plot_fill_paramset_model = ggplot(therealmccoil_all_coi_summary_rmse_gat %>%
                                                            mutate(paramset_model =
                                                                     factor(paramset_model,
                                                                            level = c("categorical",
                                                                                      "proportional"
                                                                                      )))) +
  geom_boxplot(
    aes(fill = paramset_model, y = rmse, x = coi_estimate)
  ) +
  theme_minimal() +
  scale_fill_manual(values = colorPalette_08)

```

```{r}
#| results: asis
#| echo: false

ggplotly(therealmccoil_all_coi_summary_rmse_plot_fill_paramset_model) %>% layout(boxmode = "group")
```

RMSE on the different simulation population characteristics 

```{r}
therealmccoil_all_coi_summary_rmse_gat_large_plot_by_country = ggplot(therealmccoil_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

therealmccoil_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat_by_country = ggplot(therealmccoil_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

therealmccoil_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat_by_country = ggplot(therealmccoil_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

therealmccoil_all_coi_summary_rmse_gat_large_plot_coi_cat_by_country = ggplot(therealmccoil_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse,
        fill = country)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate + 
  scale_fill_manual(values = colorPalette_08)

therealmccoil_all_coi_summary_rmse_gat_large_plot = ggplot(therealmccoil_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = pop_cat,
        y = rmse)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate

therealmccoil_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat = ggplot(therealmccoil_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = between_sample_related_cat,
        y = rmse)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate

therealmccoil_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat = ggplot(therealmccoil_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = nonrecombinant_cat,
        y = rmse)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate

therealmccoil_all_coi_summary_rmse_gat_large_plot_coi_cat = ggplot(therealmccoil_all_coi_summary_rmse_gat) +
  geom_boxplot(
    aes(x = coi_cat,
        y = rmse)
  ) +
  facet_grid(paramset_model~coi_estimate) +
  custom_ggplot_theme_xRotate

```

```{r}
#| echo: false
#| results: asis

cat(create_tabsetOfGgplotObjects(list(
  "all_intersection" = therealmccoil_all_coi_summary_rmse_gat_large_plot,
  "est_between_sample_related" = therealmccoil_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat,
  "coi_cat" = therealmccoil_all_coi_summary_rmse_gat_large_plot_coi_cat,
  "nonrecombinant" = therealmccoil_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat,
  "all_intersection_by_country" = therealmccoil_all_coi_summary_rmse_gat_large_plot_by_country,
  "est_between_sample_related_by_country" = therealmccoil_all_coi_summary_rmse_gat_large_plot_between_sample_related_cat_by_country,
  "coi_by_country" = therealmccoil_all_coi_summary_rmse_gat_large_plot_coi_cat_by_country,
  "nonrecombinant_by_country" = therealmccoil_all_coi_summary_rmse_gat_large_plot_nonrecombinant_cat_by_country
)))
```



