---
title: "Tool landscaping"
format: html
---

## Overview
During the [2023 RADISH23 hackathon](website_docs/radish23.qmd), we conducted a landscaping exercise of available tools that are commonly applied to *Plasmodium* genetic data. In order to limit the scope of this exercise we applied the following constraints:

- Focus on downstream analysis tools. This includes tools whose primary goal is to extract signal from pre-processed data, but does not include tools that are primarily used within upstream bioinformatic steps, such as variant callers and quality filters.
- Focus on *Plasmodium* genetics, including both *P. falciparum* and *P. vivax*. We did not consider applications to mosquito genetics, despite some tools and techniques being applicable to both organisms.
- Focus on tools that are specifically engineered for *Plasmodium*. While there are many generic methods that can be applied to any type of genetic data (e.g. F-statistics and other population genetic metrics), we did not seek to evaluate these tools.

MORE DETAILS HERE

## Landscaping matrix
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(DT)
```

```{r echo=F, include=F}
#Read csv data into R
matrix <- read.csv("tables/MMS_software_landscaping.csv")
```

```{r echo=F}
matrix_dt <- matrix %>% 
  select(Software_name, Software_authors, Theme, Stage, Implementation, Functionalities, Tutorial_status, Tutorial_by, Reason_relegated_or_superseded) %>% 
  rename("Software name" = "Software_name",
         "Software authors" = "Software_authors",
         "Tutorial status" = "Tutorial_status",
         "Tutorial by" = "Tutorial_by",
         "Reason relegated or superseded" = "Reason_relegated_or_superseded") 

datatable(matrix_dt,
          extensions = c('FixedColumns', 'FixedHeader'),
          options = list(
            pageLength = nrow(matrix_dt),
            fixedHeader = T,
            scrollY = "600px",
            scrollX = T,
            fixedColumns = list(leftColumns = 2),
            dom = "ft",
            autoWidth = T
          ),
          caption = htmltools::tags$caption(style = "caption-side: bottom; text-align: left; font-style: italic",
                                            "Last updated: 15 Dec 2023"),
          class = 'white-space: nowrap'
          )
```

