---
title: "Insert tool title"
output: html_document
author: "Insert name"
date: "Insert todays date"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

<br/>
<br/>

## Summary sheet

```{r, echo=FALSE, warning=FALSE}
# read in summary sheet from file
df_summary <- read.csv("summary_sheet.csv", header = FALSE)

# produce formatted table
df_summary |>
  kbl(format = "html", col.names = NULL, table.attr = "style='width:50%;'", escape = FALSE) |>
  kable_styling(position = "left", bootstrap_options = c("striped", "hover", "condensed"))
```

## Purpose

The moimix package uses whole genome/exome data to estimate the following: 

1.MOI using a binomial mixture model and Fws
2.within host allele frequencies
3.minor allele frequencies
4.

## Existing resources

- Any existing online tutorials?
- Any important papers?

## Citation

BibTeX style citation. For an R package, you can get this using `citation(package = "name")`:

Here is an example for DRpower, using `citation(package = "DRpower")`:

```
@Manual{,
    title = {DRpower: Study design and analysis for pfhrp2/3 deletion prevalence studies},
    author = {Bob Verity and Shazia Ruybal},
    note = {R package version 1.0.2},
  }
```